### 2019-10-10 draft.tex
Elsa Template added

\documentclass[review]{elsarticle}

\usepackage{lineno,hyperref}

\modulolinenumbers[5]

\journal{Future Generation Computer Systems}
\begin{document}

\begin{frontmatter}

\title{Here goes the title}

%% Group authors per affiliation:
\author{Elsevier\fnref{myfootnote}}
\address{Radarweg 29, Amsterdam}
%\fntext[myfootnote]{Since 1880.}

%% or include affiliations in footnotes:
%\author[mymainaddress,mysecondaryaddress]{Elsevier Inc}
%\ead[url]{www.elsevier.com}

%\author[mysecondaryaddress]{Global Customer Service\corref{mycorrespondingauthor}}
%\cortext[mycorrespondingauthor]{Corresponding author}
%\ead{support@elsevier.com}

%\address[mymainaddress]{1600 John F Kennedy Boulevard, Philadelphia}
%\address[mysecondaryaddress]{360 Park Avenue South, New York}

\begin{abstract}
This template helps you to create a properly formatted \LaTeX\ manuscript.
\end{abstract}

\begin{keyword}
\texttt{elsarticle.cls}\sep \LaTeX\sep Elsevier \sep template
\MSC[2010] 00-01\sep  99-00
\end{keyword}

\end{frontmatter}

\linenumbers

\section{The Elsevier article class}

\paragraph{Installation} If the document class \emph{elsarticle} is not available on your computer, you can download and install the system package \emph{texlive-publishers} (Linux) or install the \LaTeX\ package \emph{elsarticle} using the package manager of your \TeX\ installation, which is typically \TeX\ Live or Mik\TeX.

\paragraph{Usage} Once the package is properly installed, you can use the document class \emph{elsarticle} to create a manuscript. Please make sure that your manuscript follows the guidelines in the Guide for Authors of the relevant journal. It is not necessary to typeset your manuscript in exactly the same way as an article, unless you are submitting to a camera-ready copy (CRC) journal.

\paragraph{Functionality} The Elsevier article class is based on the standard article class and supports almost all of the functionality of that class. In addition, it features commands and options to format the
\begin{itemize}
\item document style
\item baselineskip
\item front matter
\item keywords and MSC codes
\item theorems, definitions and proofs
\item lables of enumerations
\item citation style and labeling.
\end{itemize}

\section{Front matter}

The author names and affiliations could be formatted in two ways:
\begin{enumerate}[(1)]
\item Group the authors per affiliation.
\item Use footnotes to indicate the affiliations.
\end{enumerate}
See the front matter of this document for examples. You are recommended to conform your choice to the journal you are submitting to.

\section{Bibliography styles}

There are various bibliography styles available. You can select the style of your choice in the preamble of this document. These styles are Elsevier styles based on standard styles like Harvard and Vancouver. Please use Bib\TeX\ to generate your bibliography and include DOIs whenever available.

Here are two sample references: \cite{Feynman1963118,Dirac1953888}.

\bibliographystyle{elsarticle-num}
\bibliography{mybibfile}

\end{document}

### 2020-01-09 draft.tex
ignore
title authors
draft abstract

@@
 
-\title{Here goes the title}
+\title{An Event-based Architecture for Multi-population Optimization Algorithms}
 
 %% Group authors per affiliation:
-\author{Elsevier\fnref{myfootnote}}
-\address{Radarweg 29, Amsterdam}
+%\author{Mario García Valdez}
+%\address{Instituto Tecnol\'ogico de Tijuana, Tijuana BC, Mexico}
+%\ead{mario@tectijuana.edu.mx}
+
+%\author{Juan J. Merelo Guerv\'os}
+%\address{Universidad de Granada, Granada, Spain}
+%\ead{jmerelo@geneura.ugr.es}
+
 %\fntext[myfootnote]{Since 1880.}
@@
 %% or include affiliations in footnotes:
-%\author[mymainaddress,mysecondaryaddress]{Elsevier Inc}
-%\ead[url]{www.elsevier.com}
+\author[itt]{Mario Garci\'a Valdez \corref{mycorrespondingauthor}}
+\cortext[mycorrespondingauthor]{Corresponding author}
+\ead{mario@tectijuana.edu.mx}
 
-%\author[mysecondaryaddress]{Global Customer Service\corref{mycorrespondingauthor}}
-%\cortext[mycorrespondingauthor]{Corresponding author}
-%\ead{support@elsevier.com}
+\author[granada]{Juan J. Merelo Guerv\'os}
+\ead{jmerelo@geneura.ugr.es}
 
-%\address[mymainaddress]{1600 John F Kennedy Boulevard, Philadelphia}
-%\address[mysecondaryaddress]{360 Park Avenue South, New York}
+\address[itt]{Department of Graduate Studies, Instituto Tecnol\'ogico de Tijuana, Tijuana BC, Mexico}
+\address[granada]{Universidad de Granada, Granada, Spain}
 
 \begin{abstract}
-This template helps you to create a properly formatted \LaTeX\ manuscript.
+    Multi-population based methods have been used extensively in recent years to
+    improve the performance of nature-inspired optimization algorithms.
+    Researchers try to avoid premature convergence and maintain population
+    diversity by dividing the original population into multiple small
+    subpopulations. Each subpopulation then evolves independently using a
+    specified search strategy. Finally, these subpopulations communicate with
+    each other to exchange individuals. Multi-population based methods are
+    parallel and asynchronous by nature; researchers take advantage of this to
+    scale algorithms with distributed, multi-threaded, and parallel
+    implementations. The design of parallel and distributed versions of
+    multi-population algorithms for their execution in cloud infrastructures is
+    not a trivial task. Designers need to adapt the algorithms to leverage the
+    benefits of scalability, elasticity, fault-tolerance, reproducibility,  and
+    cost-effectiveness of cloud-based systems. In this work, we present the
+    design and implementation of a novel event-driven architecture, designed to
+    distribute the processing of population-based algorithms asynchronously. The
+    system provides researchers a serverless strategy for creating multiple
+    populations with different parameters of execution, allowing the
+    implementation of multiple algorithms. We provide a container-based platform
+    where researchers can deploy experiments by just specifying the resources
+    and parameters of the algorithm. The performance of the platform and the
+    benefits of a multi-population approach has been evaluated using an ensemble
+    of Genetic  Algorithms (GAs) and Particle Swarm Optimization (PSO).
+    Experiments show that the framework allows the combined algorithms to
+    outperform, with a high probability, single-algorithm versions. The
+    framework we provide also has few parameters to tune since single-algorithm
+    parameters are selected randomly. The architecture and the implemented
+    platform is an excellent alternative for locally or a cloud-based
+    deployment.
 \end{abstract}
@@
 \begin{keyword}
-\texttt{elsarticle.cls}\sep \LaTeX\sep Elsevier \sep template
-\MSC[2010] 00-01\sep  99-00
+Multi-population \sep Nature-inspired algorithm \sep Parallel Genetic Algorithms \sep Cloud-Computing
+\sep Event-driven architecture 
 \end{keyword}
@@
 \item keywords and MSC codes
-\item theorems, definitions and proofs
-\item lables of enumerations
-\item citation style and labeling.
+\item theorems definitions and proofs
+\item citation style and labeling
 \end{itemize}

### 2020-01-09 draft.tex
removed example chapters

@@
 
-%% Group authors per affiliation:
-%\author{Mario García Valdez}
-%\address{Instituto Tecnol\'ogico de Tijuana, Tijuana BC, Mexico}
-%\ead{mario@tectijuana.edu.mx}
-
-%\author{Juan J. Merelo Guerv\'os}
-%\address{Universidad de Granada, Granada, Spain}
-%\ead{jmerelo@geneura.ugr.es}
-
-%\fntext[myfootnote]{Since 1880.}
-
-%% or include affiliations in footnotes:
 \author[itt]{Mario Garci\'a Valdez \corref{mycorrespondingauthor}}
@@
 
-\section{The Elsevier article class}
-
-\paragraph{Installation} If the document class \emph{elsarticle} is not available on your computer, you can download and install the system package \emph{texlive-publishers} (Linux) or install the \LaTeX\ package \emph{elsarticle} using the package manager of your \TeX\ installation, which is typically \TeX\ Live or Mik\TeX.
-
-\paragraph{Usage} Once the package is properly installed, you can use the document class \emph{elsarticle} to create a manuscript.
-Please make sure that your manuscript follows the guidelines in the Guide for Authors of the relevant journal.
-It is not necessary to typeset your manuscript in exactly the same way as an article, unless you are submitting to a camera-ready copy (CRC) journal.
-
-\paragraph{Functionality} The Elsevier article class is based on the standard article class and supports almost all of the functionality of that class.
-In addition, it features commands and options to format the
-\begin{itemize}
-\item document style
-\item baselineskip
-\item front matter
-\item keywords and MSC codes
-\item theorems definitions and proofs
-\item citation style and labeling
-\end{itemize}
-
-\section{Front matter}
-
-The author names and affiliations could be formatted in two ways:
-\begin{enumerate}[(1)]
-\item Group the authors per affiliation.
-\item Use footnotes to indicate the affiliations.
-\end{enumerate}
-See the front matter of this document for examples. You are recommended to conform your choice to the journal you are submitting to.
+\section{Introduction}
 

### 2020-01-10 draft.tex
abstract

@@
 \begin{abstract}
-    Multi-population based methods have been used extensively in recent years to
-    improve the performance of nature-inspired optimization algorithms.
-    Researchers try to avoid premature convergence and maintain population
-    diversity by dividing the original population into multiple small
-    subpopulations. Each subpopulation then evolves independently using a
-    specified search strategy. Finally, these subpopulations communicate with
-    each other to exchange individuals. Multi-population based methods are
-    parallel and asynchronous by nature; researchers take advantage of this to
-    scale algorithms with distributed, multi-threaded, and parallel
-    implementations. The design of parallel and distributed versions of
-    multi-population algorithms for their execution in cloud infrastructures is
-    not a trivial task. Designers need to adapt the algorithms to leverage the
-    benefits of scalability, elasticity, fault-tolerance, reproducibility,  and
-    cost-effectiveness of cloud-based systems. In this work, we present the
-    design and implementation of a novel event-driven architecture, designed to
-    distribute the processing of population-based algorithms asynchronously. The
-    system provides researchers a serverless strategy for creating multiple
-    populations with different parameters of execution, allowing the
-    implementation of multiple algorithms. We provide a container-based platform
-    where researchers can deploy experiments by just specifying the resources
-    and parameters of the algorithm. The performance of the platform and the
-    benefits of a multi-population approach has been evaluated using an ensemble
-    of Genetic  Algorithms (GAs) and Particle Swarm Optimization (PSO).
-    Experiments show that the framework allows the combined algorithms to
-    outperform, with a high probability, single-algorithm versions. The
-    framework we provide also has few parameters to tune since single-algorithm
-    parameters are selected randomly. The architecture and the implemented
-    platform is an excellent alternative for locally or a cloud-based
-    deployment.
+    Multi-population based optimization methods are parallel and asynchronous by
+    nature; researchers take advantage of this to scale algorithms with
+    distributed, multi-threaded, and parallel implementations. However, the design of
+    parallel and distributed versions of multi-population algorithms for their
+    execution in cloud infrastructures is not a trivial task. Designers need to
+    adapt the algorithms to leverage the benefits of scalability, elasticity,
+    fault-tolerance, reproducibility, and cost-effectiveness of reactive
+    cloud-based systems. In this work, we present the design and
+    implementation of a reactive container-based application for the execution
+    of multi-population algorithms. The application can be used to execute
+    multiple experiments, both locally or in a cloud environment. Researchers
+    can deploy experiments by specifying the infrastructure as part of an
+    experiment definition so that it is much easier to reproduce results. The
+    performance of the platform and the benefits of a multi-population approach
+    has been evaluated using an ensemble of Genetic  Algorithms (GAs) and
+    Particle Swarm Optimization (PSO). Experiments show that the framework
+    allows the combined algorithms to outperform, with a high probability,
+    single-algorithm versions. The framework we provide also has fewer
+    parameters to tune since single-algorithm parameters are selected randomly.
+    The architecture and the implemented platform is an excellent alternative
+    for locally or a cloud-based deployment.
 \end{abstract}
@@
 \section{Introduction}
+
+In the last decades, nature-inspired optimization algorithms have been succesfully
+applied to solve complex real-world problems \cite{yang2014nature}. Algorithms
+inspired in natural processes include evolutionary algorithms (EAs)
+\cite{back1996evolutionary} and swarm intelligence (SI) \cite{kennedy2006swarm},
+among others. These population-based algorithms share the common characteristic
+of using an initial set of random candidate solutions that are later used to
+generate a new set of candidates, using a nature-inspired heuristic. Popular EAs
+are Genetic Algorithms (GAs), Genetic Programming (GP), grey wolf optimization
+(GWO) and Differential Evolution (DE), while examples of (SI) are particle swarm
+optimization (PSO) and ant colony algorithms (ACO).
+
+As in nature, population-based algorithms are intrinsically parallel and
+asynchronous. The fitness of each individual can be evaluated independently of
+others, and similarly, each population could evolve in isolation. Since 
+earlier works, researchers have been proposing some form of parallelization
+\cite{muhlenbein1988evolution} to increase the scalability of these algorithms.
+The island model was one of the first techniques proposed for parallelization,
+which lead to an increased performance
+\cite{gorges1990explicit,grosso1985computer}. The concept was to divide a large
+population into communicated subpopulations. Since then, other population-based
+algorithms have adopted the concept, and researchers have found other benefits
+besides the execution speed; these include avoiding a premature convergence and
+maintaining the diversity of the global population. Researchers use the term
+multi-population based methods when generally referring to techniques using
+subpopulations as part of their optimization strategy.
+
+When designing efficient multi-population algorithms, researchers need to
+consider additional issues. These include the number and size of subpopulations,
+the interaction between them, the search area of subpopulations, and the search
+strategy and parametrization of each subpopulation. We must take into account
+these high-level issues when designing a parallel architecture.
+
+The parallel execution of multi-population algorithms has been the focus of many
+works,  going from earlier hardware-based works using transputers,
+multi-threaded, and multi-core systems. Recently the trend has been on
+exploiting a higher number of processing units by using GPUs, or distributed
+systems: client-server, web-based, map-reduce,  grid, and cloud computing. Cloud
+computing is becoming the standard way of running enterprise applications, not
+only because of the convenience of the pay-as-you-go model or the non-existent
+cost of administration, but it also offers a way of describing the
+infrastructure as part of the code, so that it is much easier to reproduce
+results and has been a boon for scientific computing.  However,  cloud computing
+has also been evolving, going from monolithic applications, that is,
+applications built on a single stack of services that communicate by layers, to
+microservice and serverless architectures that favor the parallel and
+asynchronous communication of heterogeneous resources. In these architectures,
+services or even functions are single processing points, that depart from the
+monolithic or even distributed paradigm to become a loose collection of {\em
+microservices} \cite{microservices}, which in many cases are stateless, that
+react to events and they only exist while they are doing some processing.
+
+Based on these considerations, in this work, we present the design and
+implementation of a reactive multi-population based platform extending the
+architecture proposed in [Anonymous]; this is a so-called {\em serverless}
+architecture that asynchronously processes isolated and heterogeneous
+subpopulations. In the proposed reactive system, each subpopulation is trated as
+a message that is pushed asynchronously into a message queue. Messages trigger
+stateless functions that receive the subpopulation and proceed to run an
+algorithm, using the parameters and population included in the message. After
+the specified number of iterations, each stateless function returns the evolved
+subpopulation by again pushing a message to another queue, used for receiving
+the resulting subpopulations. Subpopulations are received from the queue by a
+controller that is responsible for mixing the individuals from different
+subpopulations and producing new subpopulations. These new subpopulations are
+pushed again by the controller into the message queue, creating a loop. The
+cycle stops when the controller receives a subpopulation containing a candidate
+solution that satisfies a particular condition, or a maximum number of messages
+were received.
+
+To evaluate the capability of a cross-breed multi-population solution,
+we conducted several experiments using different benchmark functions, comparing the
+results of single versus cross-breed multi-population algorithms. For the experiments, we choose to
+compare the PSO and GA algorithms, as they are well understood, and there are
+several implementations in the literature. We implemented both algorithms as
+stateless functions, and more algorithms can be added in the same way.
+
+The rest of the paper is organized as follows
+
+\section{Multi-population design}
+
+Multi-population based methods divide the original population into
+smaller subpopulations or islands, with every subpopulation carrying out the
+algorithm independently, with synchronous or asynchronous communication with the
+rest of the islands. This relative isolation helps in maintaining an overall
+diversity since each subpopulation will search in a particular area, at least
+between communications. The recombination mentioned above (mixing) or migration
+between subpopulations is needed to avoid a premature convergence of candidate
+solutions since smaller populations are known to perform better for a given
+problem than bigger populations. However, it gives them the added advantage of
+parallel operation. Additionally, and in some cases, multi-population algorithms
+scale better than expected due to the interaction between the algorithm and the
+parallelism of the operation \cite{ALBA20027}.
+
+However, in most cases, algorithms applied to each subpopulation are
+homogeneous, or at any rate, the same variant of the algorithm. As long as this
+parallel operation is not synchronous, other population-based algorithms, or, as
+a matter of fact, any algorithm, could be easily integrated. That is why several
+works based on multi-population are heterogeneous, integrating various
+optimization algorithms, and often performing better than single-population or
+homogeneous optimization algorithms \cite{wu2016differential,nseef2016adaptive}.
+
+Heterogeneous algorithms add another degree of freedom to the problem of finding
+the correct parameter settings for an algorithm; because some parameters affect
+the accuracy of the solution and the convergence speed of the algorithms as they
+tip the balance between exploration and exploitation of the search space. On the
+other hand, current studies show that by having a high number of subpopulations
+interacting in parallel, the effect of the individual parameters of each
+subpopulation is compensated by those selected in other subpopulations. In this
+work, we will use random settings within a specific range as results have shown
+this is a valid solution to this problem.
+
+Some parameters, specially the population size, are
+kept fixed in order to control more easily the execution of the algorithm. For
+instance, by having the size of subpopulations fixed, it is easier to control
+the number of evaluations and the communication costs, when the algorithm is in
+operation.
+
+Combining multiple algorithms, with different parameters, interacting with each
+other at the same time, can benefit from the strengths of each. For instance, a
+genetic algorithm could find a promising global solution that is not optimal
+while another algorithm, more suitable for a local search, finds the global
+optimum. This approach has been followed extensively in recent years, with
+success. Moreover, there is a need for frameworks, architecture, and
+implementation models that can allow researchers the development of new
+parallel, asynchronous, heterogeneous, and parameter-free algorithms in a scalable way.
 

### 2020-01-10 draft.tex
Update abstract
This is a shorter version

@@
 \begin{abstract} 
-    Multi-population based optimization methods are parallel and asynchronous by
-    nature; researchers take advantage of this to scale algorithms with
-    distributed, multi-threaded, and parallel implementations. However, the design of
-    parallel and distributed versions of multi-population algorithms for their
-    execution in cloud infrastructures is not a trivial task. Designers need to
-    adapt the algorithms to leverage the benefits of scalability, elasticity,
-    fault-tolerance, reproducibility, and cost-effectiveness of reactive
-    cloud-based systems. In this work, we present the design and
-    implementation of a reactive container-based application for the execution
-    of multi-population algorithms. The application can be used to execute
-    multiple experiments, both locally or in a cloud environment. Researchers
-    can deploy experiments by specifying the infrastructure as part of an
-    experiment definition so that it is much easier to reproduce results. The
-    performance of the platform and the benefits of a multi-population approach
-    has been evaluated using an ensemble of Genetic  Algorithms (GAs) and
-    Particle Swarm Optimization (PSO). Experiments show that the framework
-    allows the combined algorithms to outperform, with a high probability,
-    single-algorithm versions. The framework we provide also has fewer
-    parameters to tune since single-algorithm parameters are selected randomly.
-    The architecture and the implemented platform is an excellent alternative
-    for locally or a cloud-based deployment.
+Multi-population based methods have been used extensively in
+recent years to improve the performance of nature-inspired optimization
+algorithms, these methods are parallel and asynchronous by nature; researchers
+take advantage of this to scale algorithms with distributed, multi-threaded,
+parallel and cloud-based implementations. However, the design of distributed
+versions of multi-population algorithms for their execution in cloud
+infrastructures is not a trivial task. Designers need to adapt the algorithms to
+leverage the benefits of scalability, elasticity, fault-tolerance,
+reproducibility, and cost-effectiveness of reactive cloud-based systems. In this
+work, we present the design and implementation of a reactive container-based
+application for the execution of multi-population algorithms. Researchers can
+use the application to execute multiple experiments, both locally or in a cloud
+environment and deploy experiments by specifying the infrastructure as part of
+an experiment definition so that it is much easier to reproduce results. We
+evaluated the performance of the platform and the benefits of a multi-population
+approach by using an ensemble of Genetic Algorithms (GAs) and Particle Swarm
+Optimization (PSO) in a benchmark for the optimization of continuous functions.
+Experiments show that the framework allows the ensemble to outperform
+single-algorithm versions. The framework we provide also has fewer parameters to
+tune since subpopulation parameters are selected randomly. The architecture and
+the implemented platform is an excellent alternative for locally or a
+cloud-based deployment. 
 \end{abstract}

### 2020-01-10 draft.tex
Fix  LaTeX warning
Intro draft

@@
 
-\author[itt]{Mario Garci\'a Valdez \corref{mycorrespondingauthor}}
+\author[itt]{Mario Garci\'a Valdez}\corref{mycorrespondingauthor}
 \cortext[mycorrespondingauthor]{Corresponding author}
@@
 
-As in nature, population-based algorithms are intrinsically parallel and
+Population-based algorithms are intrinsically parallel and
 asynchronous. The fitness of each individual can be evaluated independently of
@@
 The island model was one of the first techniques proposed for parallelization,
-which lead to an increased performance
-\cite{gorges1990explicit,grosso1985computer}. The concept was to divide a large
-population into communicated subpopulations. Since then, other population-based
-algorithms have adopted the concept, and researchers have found other benefits
+which lead to an increased performance \cite{gorges1990explicit,grosso1985computer}. 
+The concept was to divide a large population into communicated subpopulations. 
+Since then, other population-based algorithms have adopted the concept, 
+and researchers have found additional benefits
 besides the execution speed; these include avoiding a premature convergence and
@@
 The parallel execution of multi-population algorithms has been the focus of many
-works,  going from earlier hardware-based works using transputers,
+works,  going from earlier hardware-based implementations using transputers,
 multi-threaded, and multi-core systems. Recently the trend has been on
@@
 infrastructure as part of the code, so that it is much easier to reproduce
-results and has been a boon for scientific computing.  However,  cloud computing
+results and this has been a boon for scientific computing.  However,  cloud computing
 has also been evolving, going from monolithic applications, that is,
@@
 asynchronous communication of heterogeneous resources. In these architectures,
-services or even functions are single processing points, that depart from the
-monolithic or even distributed paradigm to become a loose collection of {\em
-microservices} \cite{microservices}, which in many cases are stateless, that
-react to events and they only exist while they are doing some processing.
+services or even functions are seen as single processing points, that departing
+from the monolithic and even distributed paradigms, become a loose collection of 
+{\em stateless functions}, that react to events and they only exist while they 
+are doing some processing.
 
-Based on these considerations, in this work, we present the design and
-implementation of a reactive multi-population based platform extending the
-architecture proposed in [Anonymous]; this is a so-called {\em serverless}
-architecture that asynchronously processes isolated and heterogeneous
-subpopulations. In the proposed reactive system, each subpopulation is trated as
-a message that is pushed asynchronously into a message queue. Messages trigger
-stateless functions that receive the subpopulation and proceed to run an
-algorithm, using the parameters and population included in the message. After
-the specified number of iterations, each stateless function returns the evolved
-subpopulation by again pushing a message to another queue, used for receiving
-the resulting subpopulations. Subpopulations are received from the queue by a
-controller that is responsible for mixing the individuals from different
-subpopulations and producing new subpopulations. These new subpopulations are
-pushed again by the controller into the message queue, creating a loop. The
-cycle stops when the controller receives a subpopulation containing a candidate
-solution that satisfies a particular condition, or a maximum number of messages
-were received.
+Based on a reactive architecture \cite{guervos2018introducing} , each subpopulation is represented as an event message that is pushed asynchronously into a message queue.
+Incoming messages trigger stateless functions that receive the subpopulation and proceed to run an algorithm, using the parameters and population included in the message.
+After the specified number of iterations, each stateless function returns the evolved subpopulation by again pushing a message to another queue, used for receiving the resulting subpopulations.
+Subpopulations are received from the queue by a controller that is responsible for mixing the individuals from different subpopulations and producing new subpopulations.
+These new subpopulations are pushed again by the controller into the message queue, creating a loop.
+The cycle stops when the controller receives a subpopulation containing a candidate solution that satisfies a particular condition, or a maximum number of messages were received.
 
-To evaluate the capability of a cross-breed multi-population solution,
-we conducted several experiments using different benchmark functions, comparing the
-results of single versus cross-breed multi-population algorithms. For the experiments, we choose to
-compare the PSO and GA algorithms, as they are well understood, and there are
-several implementations in the literature. We implemented both algorithms as
-stateless functions, and more algorithms can be added in the same way.
+Taking these factors into account, in this work, we present Swarm, a  multi-container Docker application that reactively processes isolated and heterogeneous subpopulations.
+The platform uses the Docker Compose tool for defining and running experiments as multi-container applications.
+Docker Container images host the stateless functions responsible for running algorithms on subpopulations.
+Researchers then define the configuration of services used by a multi-population algorithm using a YAML file.
+Other researchers can deploy and start all the services from this configuration file with a single command.
+Once the services are up, multiple instances of an experiment can be sent to a queue for their execution.
+The platform includes containerized services that are compatible with other container orchestration technologies like Kubernetes and Docker Swarm.
+The services included are a Redis Server used for the Message Queue implementation and logging and an Experiment Controller implemented in python.
+Developers of sub-population processing functions have the freedom to implement new population-based algorithms or operators in the language they choose inside a container, as they only need to subscribe to a channel of the Message Queue and have the ability to read the population and parameters as JSON file.
 
-The rest of the paper is organized as follows
+This paper extends our earlier publication on the topic \cite{guervos2018introducing}, and we highlight the main contributions as follows:
+
+\begin{itemize}
+	\item First, we present the design and implementation of a reactive container-based application for the asynchronous execution of multi-population algorithms.
+The source code and example container definitions are publicly available.
+	\item Second, we propose a method for the deployment and execution of multiple experiments by specifying the infrastructure as part of an experiment definition in both local or cloud environments,  facilitating the reproduction of experimental results.
+	\item Third, the application is compatible with the COCO benchmark framework, allowing researchers to compare the performance of their algorithms with other works.
+	\item Lastly, we present an empirical study to validate the applicability of our application,  measuring the execution time, speedup, and a comparison between homogeneous and an ensemble of multi-populations, using Genetic Algorithms (GAs) and Particle Swarm Optimization (PSO) in a benchmark for the optimization of continuous functions.
+\end{itemize}
 
@@
 
-\section{Bibliography styles}
-
-There are various bibliography styles available.
-You can select the style of your choice in the preamble of this document.
-These styles are Elsevier styles based on standard styles like Harvard and Vancouver.
-Please use Bib\TeX\ to generate your bibliography and include DOIs whenever available.
-
-Here are two sample references: \cite{Feynman1963118,Dirac1953888}.
-
 \bibliographystyle{elsarticle-num}
-\bibliography{mybibfile}
+\bibliography{multipopulation, serverless}
 

### 2020-01-10 draft.tex
reflow
Fix error bibliography
Some rephrasing

@@
 recent years to improve the performance of nature-inspired optimization
-algorithms, these methods are parallel and asynchronous by nature; researchers
-take advantage of this to scale algorithms with distributed, multi-threaded,
-parallel and cloud-based implementations. However, the design of distributed
-versions of multi-population algorithms for their execution in cloud
-infrastructures is not a trivial task. Designers need to adapt the algorithms to
-leverage the benefits of scalability, elasticity, fault-tolerance,
-reproducibility, and cost-effectiveness of reactive cloud-based systems. In this
+algorithms, these methods are parallel and asynchronous by nature and researchers
+leverage this to scale algorithms % scale algorithms or implementations? - JJ
+with distributed, multi-threaded,
+parallel and cloud-based implementations.
+However, the design  of cloud-native,  distributed, multi-population algorithms is not a trivial task.
+Designers need to adapt the algorithms, while keeping its nature, to
+leverage the scalability, elasticity, fault-tolerance,
+reproducibility, and cost-effectiveness of reactive cloud-based systems.
+% reactive is mentioned here for the first time.
+Possibly you should just refer to "cloud-native" and then specify that it implies asynchrony, reactivity, and so on - JJ
+In this
 work, we present the design and implementation of a reactive container-based
@@
 environment and deploy experiments by specifying the infrastructure as part of
-an experiment definition so that it is much easier to reproduce results. We
+an experiment definition so that results are easily reproducible. We
 evaluated the performance of the platform and the benefits of a multi-population
@@
 
-Based on a reactive architecture \cite{guervos2018introducing} , each subpopulation is represented as an event message that is pushed asynchronously into a message queue.
-Incoming messages trigger stateless functions that receive the subpopulation and proceed to run an algorithm, using the parameters and population included in the message.
-After the specified number of iterations, each stateless function returns the evolved subpopulation by again pushing a message to another queue, used for receiving the resulting subpopulations.
-Subpopulations are received from the queue by a controller that is responsible for mixing the individuals from different subpopulations and producing new subpopulations.
-These new subpopulations are pushed again by the controller into the message queue, creating a loop.
-The cycle stops when the controller receives a subpopulation containing a candidate solution that satisfies a particular condition, or a maximum number of messages were received.
+Based on a reactive architecture \cite{guervos2018introducing} , each
+subpopulation is represented as an event message that is pushed asynchronously
+into a message queue. Incoming messages trigger stateless functions that receive
+the subpopulation and proceed to run an algorithm, using the parameters and
+population included in the message. After the specified number of iterations,
+each stateless function returns the evolved subpopulation by again pushing a
+message to another queue, used for receiving the resulting subpopulations.
+Subpopulations are received from the queue by a controller that is responsible
+for mixing the individuals from different subpopulations and producing new
+subpopulations. These new subpopulations are pushed again by the controller into
+the message queue, creating a loop. The cycle stops when the controller receives
+a subpopulation containing a candidate solution that satisfies a particular
+condition, or a maximum number of messages were received.
 
-Taking these factors into account, in this work, we present Swarm, a  multi-container Docker application that reactively processes isolated and heterogeneous subpopulations.
-The platform uses the Docker Compose tool for defining and running experiments as multi-container applications.
-Docker Container images host the stateless functions responsible for running algorithms on subpopulations.
-Researchers then define the configuration of services used by a multi-population algorithm using a YAML file.
-Other researchers can deploy and start all the services from this configuration file with a single command.
-Once the services are up, multiple instances of an experiment can be sent to a queue for their execution.
-The platform includes containerized services that are compatible with other container orchestration technologies like Kubernetes and Docker Swarm.
-The services included are a Redis Server used for the Message Queue implementation and logging and an Experiment Controller implemented in python.
-Developers of sub-population processing functions have the freedom to implement new population-based algorithms or operators in the language they choose inside a container, as they only need to subscribe to a channel of the Message Queue and have the ability to read the population and parameters as JSON file.
+Taking these factors into account, in this work, we present Swarm, a
+multi-container Docker application that reactively processes isolated and
+heterogeneous subpopulations. The platform uses the Docker Compose tool for
+defining and running experiments as multi-container applications. Docker
+Container images host the stateless functions responsible for running algorithms
+on subpopulations. Researchers then define the configuration of services used by
+a multi-population algorithm using a YAML file. Other researchers can deploy and
+start all the services from this configuration file with a single command. Once
+the services are up, multiple instances of an experiment can be sent to a queue
+for their execution. The platform includes containerized services that are
+compatible with other container orchestration technologies like Kubernetes and
+Docker Swarm. The services included are a Redis Server used for the Message
+Queue implementation and logging and an Experiment Controller implemented in
+python. Developers of sub-population processing functions have the freedom to
+implement new population-based algorithms or operators in the language they
+choose inside a container, as they only need to subscribe to a channel of the
+Message Queue and have the ability to read the population and parameters as JSON  
+file.
 
-This paper extends our earlier publication on the topic \cite{guervos2018introducing}, and we highlight the main contributions as follows:
+This paper extends our earlier publication on the topic
+\cite{guervos2018introducing}, and we highlight the main contributions as
+follows:
 
 \begin{itemize}
-	\item First, we present the design and implementation of a reactive container-based application for the asynchronous execution of multi-population algorithms.
-The source code and example container definitions are publicly available.
-	\item Second, we propose a method for the deployment and execution of multiple experiments by specifying the infrastructure as part of an experiment definition in both local or cloud environments,  facilitating the reproduction of experimental results.
-	\item Third, the application is compatible with the COCO benchmark framework, allowing researchers to compare the performance of their algorithms with other works.
-	\item Lastly, we present an empirical study to validate the applicability of our application,  measuring the execution time, speedup, and a comparison between homogeneous and an ensemble of multi-populations, using Genetic Algorithms (GAs) and Particle Swarm Optimization (PSO) in a benchmark for the optimization of continuous functions.
+    \item First, we present the design and implementation of a reactive 
+    container-based application for the asynchronous execution of multi-population 
+    algorithms. The source code and example container definitions are publicly available.
+    \item Second, we propose a method for the deployment and execution of 
+    multiple experiments by specifying the infrastructure as part of an 
+    experiment definition in both local or cloud environments,
+    facilitating the reproduction of experimental results.
+    \item Third, the application is compatible with the COCO benchmark 
+    framework, allowing researchers to compare the performance of their 
+    algorithms with other works.
+    \item Lastly, we present an empirical study to validate the 
+    applicability of our application,  measuring the execution time, 
+    speedup, and a comparison between homogeneous and an ensemble of 
+    multi-populations, using Genetic Algorithms (GAs) and Particle 
+    Swarm Optimization (PSO) in a benchmark for the optimization of 
+    continuous functions.
 \end{itemize}
@@
 \bibliographystyle{elsarticle-num}
-\bibliography{multipopulation, serverless}
+\bibliography{multipopulation,serverless}
 

### 2020-01-10 draft.tex
Some commments to the introduction

@@
 
-\author[itt]{Mario Garci\'a Valdez}\corref{mycorrespondingauthor}
+\author[itt]{Mario Garc\'ia Valdez}\corref{mycorrespondingauthor}
 \cortext[mycorrespondingauthor]{Corresponding author}
@@
 applied to solve complex real-world problems \cite{yang2014nature}. Algorithms
-inspired in natural processes include evolutionary algorithms (EAs)
+inspired by natural processes include evolutionary algorithms (EAs)
 \cite{back1996evolutionary} and swarm intelligence (SI) \cite{kennedy2006swarm},
@@
 
-Population-based algorithms are intrinsically parallel and
-asynchronous. The fitness of each individual can be evaluated independently of
+Population-based algorithms are intrinsically parallel, and can be implemented in an asynchronous. % They needn't be intrinsically asynchronous, I think. Most are implemented synchronouslu - JJ
+The fitness of each individual can be evaluated independently of
 others, and similarly, each population could evolve in isolation. Since 
@@
 multi-population based methods when generally referring to techniques using
-subpopulations as part of their optimization strategy.
+subpopulations as part of their optimization strategy. % Some comment about asynchrony would be in order - JJ
 
@@
 works,  going from earlier hardware-based implementations using transputers,
-multi-threaded, and multi-core systems. Recently the trend has been on
+multi-threaded, and multi-core systems. % many citations here - JJ
+Recently the trend has been on
 exploiting a higher number of processing units by using GPUs, or distributed
-systems: client-server, web-based, map-reduce,  grid, and cloud computing. Cloud
-computing is becoming the standard way of running enterprise applications, not
+systems; techniques used to achieve this include client-server, web-based, map-reduce,  grid, and cloud computing. % many references here too - JJ
+Cloud
+computing has become the standard way of running enterprise applications, not
 only because of the convenience of the pay-as-you-go model or the non-existent
-cost of administration, but it also offers a way of describing the
+cost of administration, but also because it offers a way of describing the
 infrastructure as part of the code, so that it is much easier to reproduce
 results and this has been a boon for scientific computing.  However,  cloud computing
-has also been evolving, going from monolithic applications, that is,
+has also been evolving, going from simply putting in the cloud old-style monolithic applications, that is,
 applications built on a single stack of services that communicate by layers, to
 microservice and serverless architectures that favor the parallel and
-asynchronous communication of heterogeneous resources. In these architectures,
-services or even functions are seen as single processing points, that departing
-from the monolithic and even distributed paradigms, become a loose collection of 
-{\em stateless functions}, that react to events and they only exist while they 
-are doing some processing.
+asynchronous communication of heterogeneous resources; in the process, a new process for designing {\em cloud native applications} has been created, with brand new methodologies and techniques.
+In these architectures,
+services or even functions are seen as single processing points,  departing
+from the monolithic and even distributed paradigms, become a loosely coupled collection of 
+{\em stateless functions}, that react to events and only come into existence while they 
+are doing some processing. % many references more here - JJ
 
+% Again, you have to justify the need for a reactive architecture and define it before this paragraph - JJ
 Based on a reactive architecture \cite{guervos2018introducing} , each
@@
 
-Taking these factors into account, in this work, we present Swarm, a
+% There is a big gap here where you go from generic considerations
+% about cloud computing to the presentation of this method. In the
+% middle, there should be motivation for it, what are you trying to
+% address precisely, and how you intend to address it. "Approaching
+% multi-population algorithms in cloud computing" is not nearly
+% enough. There are several points here, and every one of them needs
+% to be addressed and justified
+% * Using containers as opposed to a serverless architecture or a
+% virtual-machine based one.
+% * Combining several metaheuristics instead of using a single one.
+% * Reducing the amount of parameters.
+% There needs to be a motivation for every one, relate them to the
+% objectives of the paper (which need to be clarified) and eventually
+% prove the contribution of every one of them to the final result - JJ
+Taking these factors into account, in this work, we present Swarm, % maybe a less generic name? - JJ
+a
 multi-container Docker application that reactively processes isolated and
 heterogeneous subpopulations. The platform uses the Docker Compose tool for
-defining and running experiments as multi-container applications. Docker
+defining and running experiments as multi-container applications.
+% Docker compose is not really for production, more for development environments, you have to justify this.
+Also, it does not handle scaling so need to justify it too - JJ
+Docker
 Container images host the stateless functions responsible for running algorithms
 on subpopulations. Researchers then define the configuration of services used by
-a multi-population algorithm using a YAML file. Other researchers can deploy and
+a multi-population algorithm using a YAML file. Any researcher can deploy and
 start all the services from this configuration file with a single command. Once
@@
 Queue implementation and logging and an Experiment Controller implemented in
-python. Developers of sub-population processing functions have the freedom to
+Python. Developers of sub-population processing functions have the freedom to
 implement new population-based algorithms or operators in the language they

### 2020-01-10 draft.tex
Comments to contributions
Adds paper on superlinear speedups for swarm, refs #2
Revision completed

@@
     container-based application for the asynchronous execution of multi-population 
-    algorithms. The source code and example container definitions are publicly available.
+    algorithms. The source code and example container definitions are
+    publicly available. % Not really a contribution. It's the mean to
+                        % achieve the contribution - JJ
     \item Second, we propose a method for the deployment and execution of 
@@
     experiment definition in both local or cloud environments,
-    facilitating the reproduction of experimental results.
+    facilitating the reproduction of experimental results. % Not clear
+                                % what the contribution is. If this
+                                % has been used for the first time in
+                                % an EA environment, it should be
+                                % stated that way. If not, it's also a
+                                % mean to achieve the contribution - JJ
     \item Third, the application is compatible with the COCO benchmark 
     framework, allowing researchers to compare the performance of their 
-    algorithms with other works.
+    algorithms with other works. % Not really a contribution, but an
+                                % implementation detail. Convenient,
+                                % but does not advance the state of
+                                % the art - JJ
     \item Lastly, we present an empirical study to validate the 
@@
     Swarm Optimization (PSO) in a benchmark for the optimization of 
-    continuous functions.
+    continuous functions. % "Validate the applicability" falls short
+                          % of being also a contribution. It must
+                          % advance the state of the art, in which way
+                          % does it advance? Speed (needs to be
+                          % proved) Convenience (needs to be proved
+                          % too) Scalability? (needs to be proved
+                          % also). - JJ
 \end{itemize}
 
-\section{Multi-population design}
+% The rest of the paper is organized as follows...
+
+% A state of the art should be included here too - JJ
+
+\section{Multi-population design} % Is this part of the state of the
+                                % art? - JJ
 
@@
 algorithm independently, with synchronous or asynchronous communication with the
-rest of the islands. This relative isolation helps in maintaining an overall
+rest of the islands. % exemplify how this is done in different
+                     % metaheuristics, especially those we will be
+                     % using here - JJ
+This relative isolation helps in maintaining an overall
 diversity since each subpopulation will search in a particular area, at least
@@
 solutions since smaller populations are known to perform better for a given
-problem than bigger populations. However, it gives them the added advantage of
+problem than bigger populations. % Smaller populations converge
+                                % prematurely or perform better? - JJ
+However, it gives them the added advantage of
 parallel operation. Additionally, and in some cases, multi-population algorithms
 scale better than expected due to the interaction between the algorithm and the
-parallelism of the operation \cite{ALBA20027}.
+parallelism of the operation \cite{ALBA20027}. % Need to clarify this
+                                % and if it extends to Swarm
+                                % algorithms too
+                                % Also, is your objective to achieve this superlinear speedup? - JJ
 
@@
 interacting in parallel, the effect of the individual parameters of each
-subpopulation is compensated by those selected in other subpopulations. In this
+subpopulation is compensated by those selected in other
+subpopulations. % reference - JJ
+In this
 work, we will use random settings within a specific range as results have shown
-this is a valid solution to this problem.
+this is a valid solution to this problem. % This should have been
+                                % mentioned in the introduction - JJ
 
@@
 the number of evaluations and the communication costs, when the algorithm is in
-operation.
+operation. % And to keep them more or less synchronized, which can be
+           % interesting, too... But this should go to the
+           % introduction also - JJ
 
@@
 optimum. This approach has been followed extensively in recent years, with
-success. Moreover, there is a need for frameworks, architecture, and
+success. % References - JJ
+Moreover, there is a need for frameworks, architecture, and
 implementation models that can allow researchers the development of new

### 2020-01-10 draft.tex
Addressing comment about async
Some comment about asynchrony would be in order

@@
 algorithms, these methods are parallel and asynchronous by nature and researchers
-leverage this to scale algorithms % scale algorithms or implementations? - JJ
+leverage this to scale implementations % scale algorithms or implementations? - JJ   
+%I think they are both, but in this case, implementation is more appropriate - Mario  
 with distributed, multi-threaded,
-parallel and cloud-based implementations.
-However, the design  of cloud-native,  distributed, multi-population algorithms is not a trivial task.
-Designers need to adapt the algorithms, while keeping its nature, to
+parallel and cloud-based implementations. However, the design  of cloud-native,  
+distributed, multi-population algorithms is not a trivial task. Designers need to adapt the algorithms, while keeping its nature, to
 leverage the scalability, elasticity, fault-tolerance,
-reproducibility, and cost-effectiveness of reactive cloud-based systems.
-% reactive is mentioned here for the first time.
-Possibly you should just refer to "cloud-native" and then specify that it implies asynchrony, reactivity, and so on - JJ
+reproducibility, and cost-effectiveness of cloud-native systems. 
+% reactive is mentioned here for the first time. Possibly you should just refer to "cloud-native" and then specify that it implies asynchrony, reactivity, and so on - JJ
+% OK - Mario
 In this
@@
 
-Population-based algorithms are intrinsically parallel, and can be implemented in an asynchronous. % They needn't be intrinsically asynchronous, I think. Most are implemented synchronouslu - JJ
+Population-based algorithms are intrinsically parallel, and can be implemented to run asynchronously. % They needn't be intrinsically asynchronous, I think. Most are implemented synchronouslu - JJ  
+%OK - M
 The fitness of each individual can be evaluated independently of 
@@
 subpopulations as part of their optimization strategy. % Some comment about asynchrony would be in order - JJ
+% Comments below - Mario
+Moreover, parallel implementations can run in two ways: synchronous and
+asynchronous. When operations run in parallel but maintain synchronism with
+other operations,  all operations need to finish to continue to the next step.
+For instance, in a  parent/worker model, the parent needs to wait for all
+workers to finish their operations before moving to the next iteration.  In
+contrast, in an asynchronous execution,  operations are not synchronized with
+each other. In the previous, the parent operation can continue to the next
+iteration, even if a worker has not finished its work.  Several works have
+proposed asynchronous algorithms, reporting benefits in execution time, and
+scalability. In these works, sub-populations
+are processed independently and communicate asynchronously through a central
+repository or message queue [refs needed to our previous work].
 

### 2020-01-11 draft.tex
some references added

@@
 generate a new set of candidates, using a nature-inspired heuristic. Popular EAs
-are Genetic Algorithms (GAs), Genetic Programming (GP), grey wolf optimization
+are Genetic Algorithms (GAs) \cite{holland1992adaptation,eiben2003genetic}, 
+Genetic Programming (GP) \cite{back1996evolutionary}, Grey Wolf Optimization
 (GWO) and Differential Evolution (DE), while examples of (SI) are particle swarm
@@
 iteration, even if a worker has not finished its work.  Several works have
-proposed asynchronous algorithms, reporting benefits in execution time, and
-scalability. In these works, sub-populations
+proposed asynchronous algorithms \cite{coleman89,baugh2003asynchronous}, reporting benefits in execution time, and
+scalability. In other works, sub-populations
 are processed independently and communicate asynchronously through a central
-repository or message queue [refs needed to our previous work].
+repository \cite{sofea:cec2012,agajaj} or message queues \cite{salza2019speed, guervos2018introducing} .
 
@@
 The parallel execution of multi-population algorithms has been the focus of many
-works,  going from earlier hardware-based implementations using transputers,
-multi-threaded, and multi-core systems. % many citations here - JJ
-Recently the trend has been on
-exploiting a higher number of processing units by using GPUs, or distributed
-systems; techniques used to achieve this include client-server, web-based, map-reduce,  grid, and cloud computing. % many references here too - JJ
-Cloud
-computing has become the standard way of running enterprise applications, not
+works,  going from earlier hardware-based implementations using transputers ,
+multi-threaded \cite{merelo2019scaling}, and multi-core systems \cite{Serrano2008,lai2019adaptive}. % many citations here - JJ
+Recently the trend has been on exploiting a higher number of processing units 
+by using GPUs \cite{tan2015survey,li2007efficient}, or distributed systems; techniques used to achieve this include 
+client-server , web-based \cite{JSON}, map-reduce \cite{fazenda2012},  grid \cite{munawar2010design,Gonzalez09}, 
+voluntary computing \cite{MilkyWay}, and cloud computing 
+\cite{GValdez2015,salza2019speed,valenzuela2015implementing,Sherry2012}. % many references here too - JJ
+Cloud computing has become the standard way of running enterprise applications, not
 only because of the convenience of the pay-as-you-go model or the non-existent
@@
 applications built on a single stack of services that communicate by layers, to
-microservice and serverless architectures that favor the parallel and
-asynchronous communication of heterogeneous resources; in the process, a new process for designing {\em cloud native applications} has been created, with brand new methodologies and techniques.
-In these architectures,
+microservice and serverless architectures \cite{varghese2018next} that favor the parallel and
+asynchronous communication of heterogeneous resources; in the process, 
+a new process for designing {\em cloud native applications} has been created, 
+with brand new methodologies and techniques. In these architectures,
 services or even functions are seen as single processing points,  departing
@@
 Taking these factors into account, in this work, we present Swarm, % maybe a less generic name? - JJ
-a
-multi-container Docker application that reactively processes isolated and
+% agree - Mario
+a multi-container Docker application that reactively processes isolated and
 heterogeneous subpopulations. The platform uses the Docker Compose tool for
 defining and running experiments as multi-container applications. 
-% Docker compose is not really for production, more for development environments, you have to justify this.
-Also, it does not handle scaling so need to justify it too - JJ
-Docker
-Container images host the stateless functions responsible for running algorithms
+% Docker compose is not really for production, more for development environments, 
+% you have to justify this. Also, it does not handle scaling so need to justify it too - JJ
+Docker Container images host the stateless functions responsible for running algorithms
 on subpopulations. Researchers then define the configuration of services used by

### 2020-01-13 draft.tex
Revised abstract

@@
 recent years to improve the performance of nature-inspired optimization
-algorithms, these methods are parallel and asynchronous by nature and researchers
+algorithms. These methods are parallel and asynchronous by nature and researchers
 leverage this to scale implementations % scale algorithms or implementations? - JJ   
-%I think they are both, but in this case, implementation is more appropriate - Mario  
+% I think they are both, but in this case, implementation is more appropriate - Mario
 with distributed, multi-threaded,
-parallel and cloud-based implementations. However, the design  of cloud-native,  
-distributed, multi-population algorithms is not a trivial task. Designers need to adapt the algorithms, while keeping its nature, to
+parallel and cloud-based implementations. % scale implementations with ... implementations. Please rephrase - JJ
+However, the design  of these cloud-native,  
+distributed, multi-population algorithms is not a trivial task. Designers need to adapt the algorithms,
+% design of algorithms is not trivial... designer need to adapt the algorithms... Maybe rephrase to avoid repetition - JJ
+while keeping its nature, to
 leverage the scalability, elasticity, fault-tolerance,
 reproducibility, and cost-effectiveness of cloud-native systems. 
-% reactive is mentioned here for the first time. Possibly you should just refer to "cloud-native" and then specify that it implies asynchrony, reactivity, and so on - JJ
-% OK - Mario
+% Something about "this is the problem we want to address... " to link the previous sentence with this one - JJ
 In this
 work, we present the design and implementation of a reactive container-based
+% again maybe it's better cloud-native instead of reactive - JJ
 application for the execution of multi-population algorithms. Researchers can
@@
 environment and deploy experiments by specifying the infrastructure as part of
-an experiment definition so that results are easily reproducible. We
+an experiment definition so that results are easily reproducible. % A sentence on how this implementation is better than the state of the art - JJ
+% For instance "Generally the current implementations include a single algorithm, which is bad because... " - JJ
+% Also "Current implementations are more or less functional equivalent to parallel, including the fact that they run copies of the same, parametrized, algorithm...
+% If this is too long, maybe just some reference about "current implementation do not leverage the fact that (you can use different algorithms, and differently parametrized...) and then qualify this in the introduction - JJ
+We
 evaluated the performance of the platform and the benefits of a multi-population
+% performance evaluation always needs to have an objective. What is the main objective of this performance evaluation? - JJ
 approach by using an ensemble of Genetic Algorithms (GAs) and Particle Swarm
@@
 Experiments show that the framework allows the ensemble to outperform
-single-algorithm versions. The framework we provide also has fewer parameters to
-tune since subpopulation parameters are selected randomly. The architecture and
+single-algorithm versions. % If this is the objective, then
+% 1) say that the state of the art is using single-algorithm implementations and this might not be the best
+% 2) state it clearly as the objective of the performance evaluation - JJ
+The framework we provide also has fewer parameters to
+tune since subpopulation parameters are selected randomly. % Again, this can't be a result if it's not a research question. - JJ
+% It needs to be a consequence of using a cloud-native architecture - JJ
+% Before this, "experiments prove that... "
+The architecture and
 the implemented platform is an excellent alternative for locally or a
-cloud-based deployment.
+cloud-based deployment, % thus proving that cloud-native bioinspired algorithms perform better in their "natural" environment that other kind of algorithms - JJ
 \end{abstract}

### 2020-01-13 draft.tex
Fixing references

--- 2020-01-11 draft.tex
@@
 The parallel execution of multi-population algorithms has been the focus of many
-works,  going from earlier hardware-based implementations using transputers ,
+works,  going from earlier hardware-based implementations using transputers \cite{gorges1990explicit},
 multi-threaded \cite{merelo2019scaling}, and multi-core systems \cite{Serrano2008,lai2019adaptive}. % many citations here - JJ
@@
 by using GPUs \cite{tan2015survey,li2007efficient}, or distributed systems; techniques used to achieve this include 
-client-server , web-based \cite{JSON}, map-reduce \cite{fazenda2012},  grid \cite{munawar2010design,Gonzalez09}, 
+web-based \cite{JSON}, map-reduce \cite{fazenda2012},  grid \cite{munawar2010design,Gonzalez09}, 
 voluntary computing \cite{MilkyWay}, and cloud computing 
-\cite{GValdez2015,salza2019speed,valenzuela2015implementing,Sherry2012}. % many references here too - JJ
+\cite{GValdez2015,salza2019speed,valenzuela2015implementing,Sherry2012,FlexGP}. % many references here too - JJ
 Cloud computing has become the standard way of running enterprise applications, not
@@
 applications built on a single stack of services that communicate by layers, to
-microservice and serverless architectures \cite{varghese2018next} that favor the parallel and
+microservices \cite{microservices} and serverless architectures \cite{varghese2018next, Varghese2018849} that favor the parallel and
 asynchronous communication of heterogeneous resources; in the process, 
 a new process for designing {\em cloud native applications} has been created, 
-with brand new methodologies and techniques. In these architectures,
+with brand new methodologies and techniques \cite{Baldini2016287}. In these architectures,
 services or even functions are seen as single processing points,  departing
 from the monolithic and even distributed paradigms, become a loosely coupled collection of 
-{\em stateless functions}, that react to events and only come into existence while they 
+{\em stateless functions} \cite{malawski2017serverless}, that react to events and only come into existence while they 
 are doing some processing. % many references more here - JJ
@@
 % prove the contribution of every one of them to the final result - JJ
-Taking these factors into account, in this work, we present Swarm, % maybe a less generic name? - JJ
+Taking these factors into account, in this work, we present EcoSwarm, % maybe a less generic name? - JJ
 % agree - Mario
@@
 \bibliographystyle{elsarticle-num}
-\bibliography{multipopulation,serverless}
+\bibliography{multipopulation}
 

### 2020-01-13 draft.tex
references fixed
no url, and no  list of editors, and no duplicates

@@
 are processed independently and communicate asynchronously through a central
-repository \cite{sofea:cec2012,agajaj} or message queues \cite{salza2019speed, guervos2018introducing} .
+repository \cite{sofea:cec2012,JSON} or message queues \cite{salza2019speed, guervos2018introducing} .
 
@@
 voluntary computing \cite{MilkyWay}, and cloud computing 
-\cite{GValdez2015,salza2019speed,valenzuela2015implementing,Sherry2012,FlexGP}. % many references here too - JJ
+\cite{GValdez2015,salza2019speed,valenzuela2015implementing,FlexGP}. % many references here too - JJ
 Cloud computing has become the standard way of running enterprise applications, not

### 2020-01-13 draft.tex
Addressing the first comments to the abstract

@@
 algorithms. These methods are parallel and asynchronous by nature and researchers
-leverage this to scale implementations % scale algorithms or implementations? - JJ   
-% I think they are both, but in this case, implementation is more appropriate - Mario
-with distributed, multi-threaded,
-parallel and cloud-based implementations. % scale implementations with ... implementations. Please rephrase - JJ
+leverage this to scale implementations 
+applying distributed, multi-threaded, parallel, and cloud-based computing. 
+% scale implementations with ... implementations. Please rephrase - JJ 
+% done - Mario
 However, the design  of these cloud-native,  
-distributed, multi-population algorithms is not a trivial task. Designers need to adapt the algorithms,
+distributed, multi-population algorithms is not a trivial task. Designers need to adapt their solutions,
 % design of algorithms is not trivial... designer need to adapt the algorithms... Maybe rephrase to avoid repetition - JJ
-while keeping its nature, to
-leverage the scalability, elasticity, fault-tolerance,
-reproducibility, and cost-effectiveness of cloud-native systems. 
+to leverage the scalability, elasticity, fault-tolerance,
+reproducibility, and cost-effectiveness of cloud-native systems while conserving the intended functionality.
 % Something about "this is the problem we want to address... " to link the previous sentence with this one - JJ

### 2020-01-14 draft.tex
Changes to abstarct

@@
 However, the design  of these cloud-native,  
-distributed, multi-population algorithms is not a trivial task. Designers need to adapt their solutions,
-% design of algorithms is not trivial... designer need to adapt the algorithms... Maybe rephrase to avoid repetition - JJ
-to leverage the scalability, elasticity, fault-tolerance,
-reproducibility, and cost-effectiveness of cloud-native systems while conserving the intended functionality.
+distributed, multi-population algorithms is not a trivial task. Designers need
+to adapt their solutions, 
+% design of algorithms is not trivial... designer need to adapt the algorithms... Maybe rephrase to avoid repetition - JJ to leverage
+the scalability, elasticity, fault-tolerance, reproducibility, and
+cost-effectiveness of cloud systems while conserving the intended functionality.
+% I am removing the native part above, because is also very repetetive - Mario 
 % Something about "this is the problem we want to address... " to link the previous sentence with this one - JJ 
-In this
-work, we present the design and implementation of a reactive container-based
-% again maybe it's better cloud-native instead of reactive - JJ
-application for the execution of multi-population algorithms. Researchers can
+Furthermore, currently, there are no
+cloud-native solutions that support basic algorithm design practices that can make
+multi-population methods more efficient: having multiple search strategies and
+varying the number and size of subpopulations. We address these problems in this work,
+by presenting the design and implementation of a cloud-native 
+container-based % again maybe it's better cloud-native instead of reactive - JJ
+application for the execution of ensemble multi-population algorithms. Researchers can
 use the application to execute multiple experiments, both locally or in a cloud
 environment and deploy experiments by specifying the infrastructure as part of
-an experiment definition so that results are easily reproducible. % A sentence on how this implementation is better than the state of the art - JJ
+an experiment definition so that results are easily reproducible.
+% A sentence on how this implementation is better than the state of the art - JJ 
+% Added a comment above - Mario, 
+% But maybe we need to say it again? - Mario
 % For instance "Generally the current implementations include a single algorithm, which is bad because... " - JJ 
-% Also "Current implementations are more or less functional equivalent to parallel, including the fact that they run copies of the same, parametrized, algorithm...
-% If this is too long, maybe just some reference about "current implementation do not leverage the fact that (you can use different algorithms, and differently parametrized...) and then qualify this in the introduction - JJ
-We
-evaluated the performance of the platform and the benefits of a multi-population
-% performance evaluation always needs to have an objective. What is the main objective of this performance evaluation? - JJ
-approach by using an ensemble of Genetic Algorithms (GAs) and Particle Swarm
-Optimization (PSO) in a benchmark for the optimization of continuous functions.
-Experiments show that the framework allows the ensemble to outperform
-single-algorithm versions. % If this is the objective, then
-% 1) say that the state of the art is using single-algorithm implementations and this might not be the best
-% 2) state it clearly as the objective of the performance evaluation - JJ
-The framework we provide also has fewer parameters to
-tune since subpopulation parameters are selected randomly. % Again, this can't be a result if it's not a research question. - JJ
-% It needs to be a consequence of using a cloud-native architecture - JJ
-% Before this, "experiments prove that... "
-The architecture and
-the implemented platform is an excellent alternative for locally or a
-cloud-based deployment, % thus proving that cloud-native bioinspired algorithms perform better in their "natural" environment that other kind of algorithms - JJ
+% Also "Current implementations are more or less functional equivalent 
+%to parallel, including the fact that they run copies of the same, parametrized, algorithm... 
+% If this is too long, maybe just some reference about 
+%"current implementation do not leverage the fact that (you can use different algorithms,
+% and differently parametrized...) and then qualify this in the introduction - JJ
+State of the art approaches use a single-algorithm implementation, and this might
+not be the best solution. We compared the speedup, execution time, and the number
+of evaluations between a single and an ensemble approach by using Genetic Algorithms (GAs)
+and Particle Swarm Optimization (PSO) algorithms against a benchmark for the optimization
+of continuous functions. Experiments show that the framework allows the ensemble 
+to outperform single-algorithm versions,  
+% We also run experiments to see the effect of population size for seveal problem sizes  
+% Against NUMBER OF WORKERS (this concept is not mentioned yet)
+% We can add something like:      
+% Moreover, we evaluated the effect other 
+% design parameters had on the performance; in particular, the number and the size of 
+% populations concerning the problem size. - Mario
+thus proving that cloud-native bioinspired algorithms perform better 
+in their "natural" environment than other kinds of algorithms.
+
+%The architecture and the implemented platform is an excellent alternative for
+%locally or a cloud-based deployment, 
+% thus proving that cloud-native bioinspired algorithms perform better in their "natural" environment that other kind of algorithms - JJ 
 \end{abstract}

### 2020-01-14 draft.tex
More additions to intro

@@
 % We can add something like:      
-% Moreover, we evaluated the effect other 
-% design parameters had on the performance; in particular, the number and the size of 
-% populations concerning the problem size. - Mario
-thus proving that cloud-native bioinspired algorithms perform better 
-in their "natural" environment than other kinds of algorithms.
-
-%The architecture and the implemented platform is an excellent alternative for
-%locally or a cloud-based deployment, 
-% thus proving that cloud-native bioinspired algorithms perform better in their "natural" environment that other kind of algorithms - JJ 
+Moreover, we evaluated the effect other design parameters had on the
+performance; in particular, the number and the size of populations concerning
+the problem size. The architecture and the implemented platform is an excellent
+alternative for locally or in the cloud, thus proving that cloud-native
+bioinspired algorithms perform better in their "natural" environment than other
+kinds of algorithms.
 \end{abstract}
@@
 Genetic Programming (GP) \cite{back1996evolutionary}, Grey Wolf Optimization
-(GWO) and Differential Evolution (DE), while examples of (SI) are particle swarm
-optimization (PSO) and ant colony algorithms (ACO).
+(GWO) \cite{mirjalili2014grey} and Differential Evolution (DE) \cite{karabouga2004simple},
+while examples of (SI) \cite{kennedy2006swarm} are particle swarm
+optimization (PSO) \cite{clerc2010particle} and ant colony algorithms (ACO) \cite{dorigo1999ant}.
 
-Population-based algorithms are intrinsically parallel, and can be implemented to run asynchronously. % They needn't be intrinsically asynchronous, I think. Most are implemented synchronouslu - JJ  
-%OK - M
+Population-based algorithms are intrinsically parallel, and can be implemented to run asynchronously.
 The fitness of each individual can be evaluated independently of 
@@
 besides the execution speed; these include avoiding a premature convergence and
-maintaining the diversity of the global population. Researchers use the term
+maintaining the diversity of the global population \cite{li2015multi}. Researchers use the term
 multi-population based methods when generally referring to techniques using
@@
 Moreover, parallel implementations can run in two ways: synchronous and
-asynchronous. When operations run in parallel but maintain synchronism with
-other operations,  all operations need to finish to continue to the next step.
-For instance, in a  parent/worker model, the parent needs to wait for all
+asynchronous. When operations run in parallel but must maintain synchronism with
+the other operations,  all operations need to finish to continue to the next
+step. For instance, in a  parent/worker model, the parent needs to wait for all
 workers to finish their operations before moving to the next iteration.  In
 contrast, in an asynchronous execution,  operations are not synchronized with
-each other. In the previous, the parent operation can continue to the next
-iteration, even if a worker has not finished its work.  Several works have
-proposed asynchronous algorithms \cite{coleman89,baugh2003asynchronous}, reporting benefits in execution time, and
-scalability. In other works, sub-populations
-are processed independently and communicate asynchronously through a central
-repository \cite{sofea:cec2012,JSON} or message queues \cite{salza2019speed, guervos2018introducing} .
-
+each other. Following the previous example, now the parent operation can
+continue to the next iteration, even if a worker has not finished its work.  In
+the literature, we can find many asynchronous algorithms
+\cite{coleman89,baugh2003asynchronous}, reporting benefits in execution time,
+and scalability. In the particular case of asynchronous and cloud-native
+multi-population algorithms, recent works propose an asynchronous communication
+through a central repository \cite{sofea:cec2012, JSON} or message queues
+\cite{salza2019speed, guervos2018introducing}, in this work we follow this
+practice.
 When designing efficient multi-population algorithms, researchers need to
-consider additional issues. These include the number and size of subpopulations,
+consider additional issues \cite{Ma2019}. These include the number and size of subpopulations,
 the interaction between them, the search area of subpopulations, and the search
@@
 
-The parallel execution of multi-population algorithms has been the focus of many
-works,  going from earlier hardware-based implementations using transputers \cite{gorges1990explicit},
+The trend concerning the parallel execution multi-population algorithms  
+go from earlier hardware-based implementations using transputers \cite{gorges1990explicit},
 multi-threaded \cite{merelo2019scaling}, and multi-core systems \cite{Serrano2008,lai2019adaptive}. % many citations here - JJ
-Recently the trend has been on exploiting a higher number of processing units 
+And recently the focus has been on exploiting a higher number of processing units 
 by using GPUs \cite{tan2015survey,li2007efficient}, or distributed systems; techniques used to achieve this include 

### 2020-01-14 draft.tex
More intro

@@
 
-The trend concerning the parallel execution multi-population algorithms  
-go from earlier hardware-based implementations using transputers \cite{gorges1990explicit},
-multi-threaded \cite{merelo2019scaling}, and multi-core systems \cite{Serrano2008,lai2019adaptive}. % many citations here - JJ
-And recently the focus has been on exploiting a higher number of processing units 
-by using GPUs \cite{tan2015survey,li2007efficient}, or distributed systems; techniques used to achieve this include 
-web-based \cite{JSON}, map-reduce \cite{fazenda2012},  grid \cite{munawar2010design,Gonzalez09}, 
+The trend concerning the parallel execution multi-population algorithms go from
+earlier hardware-based implementations using transputers
+\cite{gorges1990explicit}, multi-threaded \cite{merelo2019scaling}, and
+multi-core systems \cite{Serrano2008,lai2019adaptive}. % many citations here - JJ 
+And recently the focus has been on exploiting a higher number of processing
+units by using GPUs \cite{tan2015survey,li2007efficient}, or distributed
+systems; techniques used to achieve this include web-based \cite{JSON},
+map-reduce \cite{fazenda2012},  grid \cite{munawar2010design,Gonzalez09},
 voluntary computing \cite{MilkyWay}, and cloud computing
 \cite{GValdez2015,salza2019speed,valenzuela2015implementing,FlexGP}. % many references here too - JJ 
-Cloud computing has become the standard way of running enterprise applications, not
-only because of the convenience of the pay-as-you-go model or the non-existent
-cost of administration, but also because it offers a way of describing the
-infrastructure as part of the code, so that it is much easier to reproduce
-results and this has been a boon for scientific computing.  However,  cloud computing
-has also been evolving, going from simply putting in the cloud old-style monolithic applications, that is,
-applications built on a single stack of services that communicate by layers, to
-microservices \cite{microservices} and serverless architectures \cite{varghese2018next, Varghese2018849} that favor the parallel and
-asynchronous communication of heterogeneous resources; in the process, 
-a new process for designing {\em cloud native applications} has been created, 
-with brand new methodologies and techniques \cite{Baldini2016287}. In these architectures,
-services or even functions are seen as single processing points,  departing
-from the monolithic and even distributed paradigms, become a loosely coupled collection of 
-{\em stateless functions} \cite{malawski2017serverless}, that react to events and only come into existence while they 
-are doing some processing. % many references more here - JJ
+Cloud computing has become the standard way of running
+enterprise applications, not only because of the convenience of the
+pay-as-you-go model or the non-existent cost of administration, but also because
+it offers a way of describing the infrastructure as part of the code, so that it
+is much easier to reproduce results and this has been a boon for scientific
+computing.  However,  cloud computing has also been evolving, going from simply
+putting in the cloud old-style monolithic applications, that is, applications
+built on a single stack of services that communicate by layers, to microservices
+\cite{microservices} and serverless architectures \cite{varghese2018next,
+Varghese2018849} that favor the parallel and asynchronous communication of
+heterogeneous resources; in the process, a new process for designing {\em cloud
+native applications} has been created, with brand new methodologies and
+techniques \cite{Baldini2016287}. In these architectures, services or even
+functions are seen as single processing points, departing from the monolithic
+and even distributed paradigms, become a loosely coupled collection of {\em
+stateless functions} \cite{malawski2017serverless}, that react to events and
+only come into existence while they are doing some processing. % many references more here - JJ 
+Systems developed with the patterns outlined above have the
+properties of a reactive system \cite{boner2014reactive} and they are commonly
+more flexible, loosely-coupled, and scalable. % Some properies and reference - Mario  
+% Again, you have to justify the need for a reactive architecture and define it before this paragraph - JJ
 
-% Again, you have to justify the need for a reactive architecture and define it before this paragraph - JJ
-Based on a reactive architecture \cite{guervos2018introducing} , each
-subpopulation is represented as an event message that is pushed asynchronously
-into a message queue. Incoming messages trigger stateless functions that receive
-the subpopulation and proceed to run an algorithm, using the parameters and
-population included in the message. After the specified number of iterations,
-each stateless function returns the evolved subpopulation by again pushing a
-message to another queue, used for receiving the resulting subpopulations.
-Subpopulations are received from the queue by a controller that is responsible
-for mixing the individuals from different subpopulations and producing new
-subpopulations. These new subpopulations are pushed again by the controller into
-the message queue, creating a loop. The cycle stops when the controller receives
-a subpopulation containing a candidate solution that satisfies a particular
-condition, or a maximum number of messages were received.
+% This paragraph will be moved to the Proposed Method section - Mario
+% Based on a reactive architecture \cite{guervos2018introducing} , each
+% subpopulation is represented as an event message that is pushed asynchronously
+% into a message queue. Incoming messages trigger stateless functions that receive
+% the subpopulation and proceed to run an algorithm, using the parameters and
+% population included in the message. After the specified number of iterations,
+% each stateless function returns the evolved subpopulation by again pushing a
+% message to another queue, used for receiving the resulting subpopulations.
+% Subpopulations are received from the queue by a controller that is responsible
+% for mixing the individuals from different subpopulations and producing new
+% subpopulations. These new subpopulations are pushed again by the controller into
+% the message queue, creating a loop. The cycle stops when the controller receives
+% a subpopulation containing a candidate solution that satisfies a particular
+% condition, or a maximum number of messages were received.
 
@@
 % prove the contribution of every one of them to the final result - JJ
+
 Taking these factors into account, in this work, we present EcoSwarm, % maybe a less generic name? - JJ

### 2020-01-14 draft.tex
Motivations

@@
 distributed, multi-population algorithms is not a trivial task. Designers need
-to adapt their solutions, 
-% design of algorithms is not trivial... designer need to adapt the algorithms... Maybe rephrase to avoid repetition - JJ to leverage
+to adapt their solutions 
+% design of algorithms is not trivial... designer need to adapt the algorithms... Maybe rephrase to avoid repetition - JJ 
+to leverage
 the scalability, elasticity, fault-tolerance, reproducibility, and
@@
 cloud-native solutions that support basic algorithm design practices that can make
-multi-population methods more efficient: having multiple search strategies and
+multi-population methods more efficient: having multiple search strategies, and
 varying the number and size of subpopulations. We address these problems in this work,
@@
 use the application to execute multiple experiments, both locally or in a cloud
-environment and deploy experiments by specifying the infrastructure as part of
+environment, and deploy experiments by specifying the infrastructure as part of
 an experiment definition so that results are easily reproducible.
@@
 of continuous functions. Experiments show that the framework allows the ensemble 
-to outperform single-algorithm versions,  
+to outperform single-algorithm versions.  
 % We also run experiments to see the effect of population size for seveal problem sizes  
@@
 Moreover, we evaluated the effect other design parameters had on the
-performance; in particular, the number and the size of populations concerning
+performance, in particular, the number and the size of populations concerning
 the problem size. The architecture and the implemented platform is an excellent
@@
 
+Researchers and algorithm designers willing to  adapt to the cloud their
+multi-population solutions face the following challenges:
+\begin{itemize}
+    \item They need to change their current solutions to a reactive architecture in which processes
+    communicate with each other by exchanging messages asynchronously and react
+    to a continuous stream of data. One of the options is to consider
+    sub-populations as messages to be modified asynchronously by functions.
+
+    \item To scale the system, when it is possible, they must use stateless functions for
+    processing.
+
+    \item They need to establish a workflow of local development and prototyping while having the 
+    option of deployment in a cloud provider. It is desirable to use modern development methodologies and technologies.
+
+    \item They must log experimental results and be able to replicate the experiments.
+
+    \item They must consider additional parameters that can affect the monetary cost or the performance of the
+    execution — for instance, the format and size of messages, the number of
+    worker processes and their capabilities.  
+\end{itemize}
+
 Taking these factors into account, in this work, we present EcoSwarm, % maybe a less generic name? - JJ
@@
 % you have to justify this. Also, it does not handle scaling so need to justify it too - JJ
+% justification: - Mario
+We preferred the use of a container-based application for this prototype because
+it is more suitable for development and replication without the need for
+additional components. Nevertheless, the application is compatible with other
+orchestration and deployment technologies if more scalabilty like Docker Swarm
+and Kubernetes with additional configuration.  
 Docker Container images host the stateless functions responsible for running algorithms
-on subpopulations. Researchers then define the configuration of services used by
+on subpopulations(these can be used directly in a FaaS implementation).
+% Need to improve the next paragraph or maybe move it to another place - Mario
+Researchers then define the configuration of services used by
 a multi-population algorithm using a YAML file. Any researcher can deploy and

### 2020-01-15 draft.tex
Minor change
Details added to Intro

@@
 practice.
-When designing efficient multi-population algorithms, researchers need to
-consider additional issues \cite{Ma2019}. These include the number and size of subpopulations,
-the interaction between them, the search area of subpopulations, and the search
-strategy and parametrization of each subpopulation. We must take into account
-these high-level issues when designing a parallel architecture.
 
@@
 
-Researchers and algorithm designers willing to  adapt to the cloud their
-multi-population solutions face the following challenges:
+When designing efficient multi-population algorithms, researchers need to
+consider additional issues \cite{Ma2019}. These include the number and size of subpopulations,
+the interaction between them, the search area of subpopulations, and the search
+strategy and parametrization of each subpopulation. We must take into account
+these high-level issues when designing a parallel architecture. Researchers 
+and algorithm designers willing to  adapt to the cloud their
+multi-population solutions face the additional challenges:
 \begin{itemize}
@@
 additional components. Nevertheless, the application is compatible with other
-orchestration and deployment technologies if more scalabilty like Docker Swarm
-and Kubernetes with additional configuration.  
-Docker Container images host the stateless functions responsible for running algorithms
-on subpopulations(these can be used directly in a FaaS implementation).
+orchestration and deployment technologies if more scalabilty is needed. 
+With additional configuration settings, applications can be deployed 
+to a Docker Swarm or a Kubernetes system.
+
+Docker Container images host the stateless functions responsible for running 
+a search strategy on subpopulations, these functions could also be ported
+directly to a Function as a Service (FaaS) \cite{Roberts2016} implementation.
 % Need to improve the next paragraph or maybe move it to another place - Mario
@@
                                 % art? - JJ
-
+% Also, multipopulation design of what?x
 Multi-population based methods divide the original population into

### 2020-01-15 draft.tex
Paper Organization

@@
 
-% The rest of the paper is organized as follows...
+The organization of the paper is as follows. First, in Section \ref{multi}, we present a
+background of the fundamental issues of integrating nature-inspired optimization
+and multi-population methods. Section \ref{soa} presents state of the art relevant to
+our work. In Section \ref{method}, we present the proposed method.  Section \ref{setup} describes
+the design of the empirical evaluation we designed to assess the effectiveness
+of the method, and in Section \ref{results}, we report and discuss the results. Finally, we
+offer the conclusions of this paper and suggestions on future work in Section \ref{conclusions}.
 
@@
 
-\section{Multi-population design} % Is this part of the state of the
+\section{Multi-population methods} % Is this part of the state of the
+\label{multi}
                                 % art? - JJ
+% This the Background for the design decisions, but I think is also needed
+% to better understand the state of the art.
+
 % Also, multipopulation design of what?x
@@
 
+\section{State of the Art} 
+\label{soa}
+
+\section{Proposed Method} 
+\label{method}
+
+\section{Experimental Setup} 
+\label{setup}
+
+\section{Experimental Results} 
+\label{results}
+
+\section{Conclusions} 
+\label{conclusions}
+
 \bibliographystyle{elsarticle-num}

### 2020-01-15 draft.tex
SoA Outline
Only cloud-container methos

@@
 \label{soa}
+Hadoop
+    \cite{fazenda2012} 
+    FlexGP \cite{FlexGP}, 
+    elephant56 \cite{salza2016elephant56,ferrucci2018using}, 
+Pool
+    SofEA \cite{merelo2012sofea,sofea:cec2012} , 
+    EvoSpace \cite{garcia2015evospace},
+Volunteer
+    evospace-js \cite{garcia2017evospace}
+Queue
+    Containers \cite{salza2016approach}
+    Seamless local and Cloud \cite{leclerc2016seamless}
+    Serverless KafKEO \cite{guervos2018introducing}, 
+    Docker Orchestation \cite{dziurzanski2020scalable,zhao2019cloud}
 

### 2020-01-15 draft.tex
Outline SOA

--- 2020-01-15 draft.tex
@@
 \label{soa}
+Hadoop
+    Hadoop\cite{fazenda2012} 
+    FlexGP \cite{FlexGP}, 
+    elephant56 \cite{salza2016elephant56,ferrucci2018using}, 
+Pool
+    SofEA \cite{merelo2012sofea,sofea:cec2012} , 
+    EvoSpace \cite{garcia2015evospace},
+Volunteer
+    evospace-js \cite{garcia2017evospace}
+Queue
+    Containers \cite{valenzuela2015implementing} \cite{salza2016approach} 
+    Seamless local and Cloud \cite{leclerc2016seamless}
+    Serverless KafKEO \cite{guervos2018introducing}, 
+    Docker Orchestation \cite{dziurzanski2020scalable,zhao2019cloud}
 

### 2020-01-16 draft.tex
KafkEO image

--- 2020-01-15 draft.tex
@@
 \label{soa}
+Hadoop 
+    Comparasion \cite{ferrucci2018using}
+    Hadoop\cite{fazenda2012} 
+    FlexGP \cite{FlexGP}, 
+    elephant56 \cite{salza2016elephant56},
+
+Pool
+    SofEA \cite{merelo2012sofea,sofea:cec2012} , 
+    EvoSpace \cite{garcia2015evospace},
+Volunteer
+    evospace-js \cite{garcia2017evospace}
+Queue
+    Containers \cite{valenzuela2015implementing} \cite{salza2016approach} 
+    Seamless local and Cloud \cite{leclerc2016seamless}
+    Serverless KafKEO \cite{guervos2018introducing}, 
+    Docker Orchestation \cite{dziurzanski2020scalable,zhao2019cloud}
+PSO 
+    PSO AWS \cite{li2015parallel}
 

### 2020-01-17 draft.tex
Intro to Proposed method

@@
 
+In this section, we present the design
+and a container-based implementation of the model we propose for the execution
+of multi-population based algorithms. We follow the requirements outlined in
+Section \ref{multi} and the general design principles of cloud-native
+applications we have mention earlier. As a first step, we describe the general
+architecture first, and then we describe each component in detail. The primary
+data element used throughout the system is the subpopulation, and at a higher
+level, we can view the system as a continuous stream of subpopulations flowing
+from one process to the other. By "continuous stream," we mean that ideally,
+everything must happen asynchronously without components needing to wait for
+others. In this kind of system, this streaming functionality is normally
+implemented by using a message queue system. For instance, if one process needs
+to communicate a subpopulation to another, it must pack the sub-population into
+a message, and then push the message asynchronously to the queuing system; this
+means that after the process pushes the message, it continues its execution
+without waiting for a result. On the other end, a recipient subscribes to the
+message queue by defining an event handler method that is triggered when a
+message is pushed to the queue. The two components at each end of the queue are
+Producers that push messages to the queue and Consumers that pull them. This
+pattern is an essential component of a reactive architecture because it is
+highly scalable; one of the reasons for this are stateless functions. We can
+define stateless functions as methods that do not have side effects, meaning
+they do not need to read, keep, or modify data outside of the method. If
+Consumers are implemented using stateless functions, then there is no difference
+in having one or many copies of the function pulling work from the queue at the
+same time, because no side effects are resulting from the concurrency.  
+Based on a reactive architecture we proposed in a previous work
+\cite{guervos2018introducing}, we now describe the general architecture shown in
+Figure .
+
 \section{Experimental Setup} 

### 2020-01-17 draft.tex
images

--- 2020-01-16 draft.tex
@@
 \usepackage{lineno,hyperref}
+\graphicspath{ {./images/} }
 
@@
 % Again, you have to justify the need for a reactive architecture and define it before this paragraph - JJ
-
-% This paragraph will be moved to the Proposed Method section - Mario
-% Based on a reactive architecture \cite{guervos2018introducing} , each
-% subpopulation is represented as an event message that is pushed asynchronously
-% into a message queue. Incoming messages trigger stateless functions that receive
-% the subpopulation and proceed to run an algorithm, using the parameters and
-% population included in the message. After the specified number of iterations,
-% each stateless function returns the evolved subpopulation by again pushing a
-% message to another queue, used for receiving the resulting subpopulations.
-% Subpopulations are received from the queue by a controller that is responsible
-% for mixing the individuals from different subpopulations and producing new
-% subpopulations. These new subpopulations are pushed again by the controller into
-% the message queue, creating a loop. The cycle stops when the controller receives
-% a subpopulation containing a candidate solution that satisfies a particular
-% condition, or a maximum number of messages were received.
 
@@
 \label{method}
+This paragraph will be moved to the Proposed Method section - Mario
+Based on a reactive architecture \cite{guervos2018introducing} , each
+subpopulation is represented as an event message that is pushed asynchronously
+into a message queue. Incoming messages trigger stateless functions that receive
+the subpopulation and proceed to run an algorithm, using the parameters and
+population included in the message. After the specified number of iterations,
+each stateless function returns the evolved subpopulation by again pushing a
+message to another queue, used for receiving the resulting subpopulations.
+Subpopulations are received from the queue by a controller that is responsible
+for mixing the individuals from different subpopulations and producing new
+subpopulations. These new subpopulations are pushed again by the controller into
+the message queue, creating a loop. The cycle stops when the controller receives
+a subpopulation containing a candidate solution that satisfies a particular
+condition, or a maximum number of messages were received.
+
+\subsection{Event Driven Model} 
+\label{edm}
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\textwidth]{kafkEO}
+    \caption{Example of a parametric plot ($\sin (x), \cos(x), x$)}
+\end{figure}
+
+\subsection{Workers} 
+\label{edm}
+
+\subsection{Controller} 
+\label{edm}
+
+\subsection{Experimental Process} 
+\label{edm}
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\textwidth]{experiment_flow}
+    \caption{Example of a parametric plot ($\sin (x), \cos(x), x$)}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\textwidth]{worker}
+    \caption{Example of a parametric plot ($\sin (x), \cos(x), x$)}
+\end{figure}
 

### 2020-01-17 draft.tex

@@
 \label{method}
-This paragraph will be moved to the Proposed Method section - Mario
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\textwidth]{kafkEO}
+    \caption{Example of a parametric plot ($\sin (x), \cos(x), x$)}
+    \label{fig:kafkEO}
+\end{figure}
+
+In this section, we present the design
+and a container-based implementation of the model we propose for the execution
+of multi-population based algorithms. We follow the requirements outlined in
+Section \ref{multi} and the general design principles of cloud-native
+applications we have mention earlier. As a first step, we describe the general
+architecture first, and then we describe each component in detail.
+
+The primary data element used throughout the system is the subpopulation, and at a higher
+level, we can view the system as a continuous stream of subpopulations flowing
+from one process to the other. By "continuous stream," we mean that ideally,
+everything must happen asynchronously without components needing to wait for
+others. In this kind of system, this streaming functionality is normally
+implemented by using a message queue system. For instance, if one process needs
+to communicate a subpopulation to another, it must pack the sub-population into
+a message, and then push the message asynchronously to the queuing system; this
+means that after the process pushes the message, it continues its execution
+without waiting for a result. On the other end, the recipient subscribes to the
+message queue by defining an event handler method that is going to be triggered when a
+message is pushed to the message queue. The two components at each end of the queue are
+Producers that push messages to the queue and Consumers that pull them. This
+pattern is an essential component of a reactive architecture because it is
+highly scalable; one of the reasons for this are stateless functions. We can
+define stateless functions as methods that do not have side effects, meaning
+they do not need to read, keep, or modify data outside of the method. If
+Consumers are implemented using stateless functions, then there is no difference
+netween having one or many copies of the same function pulling work from the queue all at the
+same time, because no side effects are going to result from the concurrency.
+Based un these general principles we define next the proposed model.
+
+\subsection{Event Driven Model} 
+\label{edm}
+Based on a reactive architecture we proposed in a previous work
+\cite{guervos2018introducing}, we now describe the general architecture shown in
+Figure \ref{fig:kafkEO}. Again, we can explain the general model using the analogy
+of Producers and Consumers of messages using a queue. First, we can see that
+there are two queues, one labeled Input, and the other one Output. In the diagram, 
+push operations on a queue are represented by solid arrows connecting to the left side
+of the box, and pull or pop operations as solid arrows leaving from the right side.
+
 Based on a reactive architecture \cite{guervos2018introducing} , each
@@
 condition, or a maximum number of messages were received.
-
-\subsection{Event Driven Model} 
-\label{edm}
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\textwidth]{kafkEO}
-    \caption{Example of a parametric plot ($\sin (x), \cos(x), x$)}
-\end{figure}
 

### 2020-01-18 draft.tex
first draft of the proposed method  intro

@@
     \centering
-    \includegraphics[width=\textwidth]{kafkEO}
+    \includegraphics[width=\textwidth]{kafkEOsmall}
     \caption{Example of a parametric plot ($\sin (x), \cos(x), x$)}
@@
 \cite{guervos2018introducing}, we now describe the general architecture shown in
-Figure \ref{fig:kafkEO}. Again, we can explain the general model using the analogy
-of Producers and Consumers of messages using a queue. First, we can see that
+Figure \ref{fig:kafkEO}. Again, we can explain the model using the analogy
+of Producers and Consumers of messages. First, we can see that
 there are two queues, one labeled Input, and the other one Output. In the diagram, 
@@
 of the box, and pull or pop operations as solid arrows leaving from the right side.
+The algorithm starts with the Setup process pulling a configuration message from
+the experiements queue, this is not shown in the diagram. The configuration messsage 
+includes all the parameters needed to execute the algorithm, among other things,
+the number of subpopulations, the number of individuals, and the number of iterations
+of the algorithm. We give more details about the configuration data structure
+in Section []. Once the configuration is read, we can follow the path of 
+messagess as follows:
+
+\begin{enumerate}
+\item In this step, the specified number of subpopulations is created acording to the 
+parameters found in the configuration structure. The population at this moment is just
+static data, including each individual inside. Each subpopulation includes a metadata
+section where its algorithm and execution parameters are specified. For instance, 
+for a GA, the mutation rate, type of crossover operator, and other values are indicated.
+Each subpopulation is then pushed to the Input queue, so they can be consumed 
+by Stateless functions responsible for the execution of the search.
+
+\item One or more Stateless functions are constantly pulling subpopulation messages
+from the Input queue. Taking the population data and metadata as parameters, these 
+functions are responsible for the actual execution of the optimization algorithm. 
+They take the current state of the population and run the algorithm for a certain 
+number of iterations. Once they finish the execution, the population state
+is packed again along with aditional metadata about the execution of the algorithm.
+The resulting populations is now pushed to the Output queue. Once finished, another 
+subpopulation is pulled from the queue.
+
+\item The Controller process is responsible for keeping track of the progress of 
+the search. It pulls current subpopulations from the Output queue, inspects the metadata 
+and if an optimal solution has been found or the maximum number of iteration has been
+reached it signals the stop of the execution. Otherwise, it passes the stream of messages 
+to a migration process, where subpopulations are mixed with each other. New populations 
+are generated from this migration, and they are again pushed to the Input queue to continue 
+in a loop. The Controller is also responsible of logging the metadata recived with the
+messages.
+
+\end{enumerate}
 

### 2020-01-18 draft.tex
More description and images

@@
 
+% Based on a reactive architecture \cite{guervos2018introducing} , each
+% subpopulation is represented as an event message that is pushed asynchronously
+% into a message queue. Incoming messages trigger stateless functions that receive
+% the subpopulation and proceed to run an algorithm, using the parameters and
+% population included in the message. After the specified number of iterations,
+% each stateless function returns the evolved subpopulation by again pushing a
+% message to another queue, used for receiving the resulting subpopulations.
+% Subpopulations are received from the queue by a controller that is responsible
+% for mixing the individuals from different subpopulations and producing new
+% subpopulations. These new subpopulations are pushed again by the controller into
+% the message queue, creating a loop. The cycle stops when the controller receives
+% a subpopulation containing a candidate solution that satisfies a particular
+% condition, or a maximum number of messages were received.
+
 In this section, we present the design
@@
 messages.  
-
 \end{enumerate}
 
-Based on a reactive architecture \cite{guervos2018introducing} , each
-subpopulation is represented as an event message that is pushed asynchronously
-into a message queue. Incoming messages trigger stateless functions that receive
-the subpopulation and proceed to run an algorithm, using the parameters and
-population included in the message. After the specified number of iterations,
-each stateless function returns the evolved subpopulation by again pushing a
-message to another queue, used for receiving the resulting subpopulations.
-Subpopulations are received from the queue by a controller that is responsible
-for mixing the individuals from different subpopulations and producing new
-subpopulations. These new subpopulations are pushed again by the controller into
-the message queue, creating a loop. The cycle stops when the controller receives
-a subpopulation containing a candidate solution that satisfies a particular
-condition, or a maximum number of messages were received.
+This reactive architecture has the following advantages:
+
+An important aspect of proposal is the decoupling of the population and 
+the population-based algorithm. It is common that in a classic island-based algorithm
+each island is executed in a separate processing node, i.e. in a virtual machine, cpu,
+or thread. In this case we can have just one processing node, or many nodes, each 
+running a different search strategy, or using its own parameters.
+
+Also, the reactive controller, gives designers more control over the milti-population algorithm.
+In this process, designers can dynamicly change the number of populations, population parameters,
+and migration details on-the-fly.
+
+Another advantege is that algorithm designers have many options for implementing this simple architecure. 
+The same basic components can be implmented as a single multithreaded program, or 
+as a highly scalable serverless cloud application. Most modern languages include 
+constructs for asynchronous programming using queues or channels, for a multi-threaded
+execution.
+
+On the other hand, there are several caveats as well:
+
+It is more costly to move entire subpopulations as messages than only passing certain individuals 
+from a process to another. Designers must conmsider this cost when working with large
+individuals, and if possible send sub-populations using several messages, use compression or 
+adjust the size of subpopulations. Pool-based algorithms also suffer from this drawback, but 
+web based immplementations have been working with continous optimization problems of 1000 dimensions.
+But, in other cases this is not a viable approach.
 
@@
 
-\subsection{Controller} 
+\subsection{Comparasion} 
 \label{edm}
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\textwidth]{classicisland}
+    \caption{Example of a parametric plot ($\sin (x), \cos(x), x$)}
+    \label{fig:kafkEO}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\textwidth]{pool_island}
+    \caption{Example of a parametric plot ($\sin (x), \cos(x), x$)}
+    \label{fig:kafkEO}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\textwidth]{population_pool}
+    \caption{Example of a parametric plot ($\sin (x), \cos(x), x$)}
+    \label{fig:kafkEO}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\textwidth]{population_message}
+    \caption{Example of a parametric plot ($\sin (x), \cos(x), x$)}
+    \label{fig:kafkEO}
+\end{figure}
 

### 2020-01-19 draft.tex
Rewrites abstract, reducing it

@@
 \begin{abstract} 
-Multi-population based methods have been used extensively in
+Splitting a population into multiple instances is a technique that has been used extensively in
 recent years to improve the performance of nature-inspired optimization
-algorithms. These methods are parallel and asynchronous by nature and researchers
-leverage this to scale implementations 
-applying distributed, multi-threaded, parallel, and cloud-based computing. 
-% scale implementations with ... implementations. Please rephrase - JJ 
-% done - Mario
-However, the design  of these cloud-native,  
-distributed, multi-population algorithms is not a trivial task. Designers need
-to adapt their solutions 
-% design of algorithms is not trivial... designer need to adapt the algorithms... Maybe rephrase to avoid repetition - JJ 
-to leverage
+algorithms. Work on several populations can be done in parallel, and
+also asynchronously, which can be leveraged to create scalable
+implementations based on, among other methods, distributed, multi-threaded, parallel, and cloud-based computing. 
+% I have rephrased all this. Please check it out - JJ
+The design of these cloud-native, distributed, multi-population
+algorithms is not a trivial task. Coming from monolithic
+(single-instance) solutions, adaptations at several levels, from the
+algorithmic to the functional, must be made to leverage
 the scalability, elasticity, fault-tolerance, reproducibility, and
-cost-effectiveness of cloud systems while conserving the intended functionality.
-% I am removing the native part above, because is also very repetetive - Mario 
-% Something about "this is the problem we want to address... " to link the previous sentence with this one - JJ 
-Furthermore, currently, there are no
-cloud-native solutions that support basic algorithm design practices that can make
-multi-population methods more efficient: having multiple search strategies, and
-varying the number and size of subpopulations. We address these problems in this work,
-by presenting the design and implementation of a cloud-native 
-container-based % again maybe it's better cloud-native instead of reactive - JJ
-application for the execution of ensemble multi-population algorithms. Researchers can
-use the application to execute multiple experiments, both locally or in a cloud
-environment, and deploy experiments by specifying the infrastructure as part of
-an experiment definition so that results are easily reproducible.
-% A sentence on how this implementation is better than the state of the art - JJ 
-% Added a comment above - Mario, 
-% But maybe we need to say it again? - Mario
-% For instance "Generally the current implementations include a single algorithm, which is bad because... " - JJ 
-% Also "Current implementations are more or less functional equivalent 
-%to parallel, including the fact that they run copies of the same, parametrized, algorithm... 
-% If this is too long, maybe just some reference about 
-%"current implementation do not leverage the fact that (you can use different algorithms,
-% and differently parametrized...) and then qualify this in the introduction - JJ
-State of the art approaches use a single-algorithm implementation, and this might
-not be the best solution. We compared the speedup, execution time, and the number
-of evaluations between a single and an ensemble approach by using Genetic Algorithms (GAs)
-and Particle Swarm Optimization (PSO) algorithms against a benchmark for the optimization
-of continuous functions. Experiments show that the framework allows the ensemble 
-to outperform single-algorithm versions.  
-% We also run experiments to see the effect of population size for seveal problem sizes  
-% Against NUMBER OF WORKERS (this concept is not mentioned yet)
-% We can add something like:      
-Moreover, we evaluated the effect other design parameters had on the
-performance, in particular, the number and the size of populations concerning
-the problem size. The architecture and the implemented platform is an excellent
+cost-effectiveness of cloud systems while, at the same time,
+conserving the intended functionality. Instead of an evolutive
+approach, in this paper we propose to create from scratch a
+cloud-native optimization framework that is able to include multiple
+algorithms and use few parameters to work. This solution goes beyond
+current state of the art via a framework that is able to support
+different algorithms at the same time, work asynchronously, and also
+be easily deployable to any cloud platform. We evaluate the
+performance and scalability of this solution, together with the effect
+other design parameters had on the 
+performance, in particular, the number and the size of populations
+with respect to problem size. The architecture and the implemented platform is an excellent
 alternative for locally or in the cloud, thus proving that cloud-native
 bioinspired algorithms perform better in their "natural" environment than other
-kinds of algorithms.
+kinds of algorithms, and setting a new baseline for scaling and
+performance in the cloud.
 \end{abstract}
@@
 Hadoop 
-    Comparasion \cite{ferrucci2018using}
+    Comparison \cite{ferrucci2018using}
     Hadoop\cite{fazenda2012} 

### 2020-01-19 draft.tex
Some comments and reformatting

@@
 performance in the cloud.
+% Still a big longish. We need another iteration to reduce it - JJ
 \end{abstract}
@@
 \label{soa}
+% a lot of this has been done in the previous method... Maybe there's
+% not that much to do - JJ
 Hadoop 
@@
     \includegraphics[width=\textwidth]{kafkEOsmall}
-    \caption{Example of a parametric plot ($\sin (x), \cos(x), x$)}
+    \caption{Example of a parametric plot ($\sin (x), \cos(x), x$)} % This caption? - JJ
     \label{fig:kafkEO}
@@
 The primary data element used throughout the system is the subpopulation, and at a higher
-level, we can view the system as a continuous stream of subpopulations flowing
+level, we can view the system as a continuous stream of subpopulations
+flowing
+% We talk about "multi-population" elsewhere, including the
+% abstract. We should unify terminology - JJ
 from one process to the other. By "continuous stream," we mean that ideally,
@@
 to communicate a subpopulation to another, it must pack the sub-population into
-a message, and then push the message asynchronously to the queuing system; this
+a message, and then send the message asynchronously % I don't know if
+                                % asynchronously applies here. It's
+                                % simply sent - JJ
+to the queuing system; this
 means that after the process pushes the message, it continues its execution
@@
 Consumers are implemented using stateless functions, then there is no difference
-netween having one or many copies of the same function pulling work from the queue all at the
-same time, because no side effects are going to result from the concurrency.
-Based un these general principles we define next the proposed model.
+between having one or many copies of the same function pulling work from the queue all at the
+same time, because no side effects are going to result from the
+concurrency.
+% This is important and we should probably mention that in the
+% introduction - JJ 
+Based on these general principles we define next the proposed model.
 
@@
 Each subpopulation is then pushed to the Input queue, so they can be consumed 
-by Stateless functions responsible for the execution of the search.
+by Stateless functions responsible for the execution of the search.  % Why the caps in Stateless? - JJ
 
@@
 
-This reactive architecture has the following advantages:
-
+This reactive architecture has the following advantages:\begin{itemize}
+\item 
 An important aspect of proposal is the decoupling of the population and 
@@
 running a different search strategy, or using its own parameters. 
-
+\item 
 Also, the reactive controller, gives designers more control over the milti-population algorithm.
@@
 and migration details on-the-fly.
-
+\item 
 Another advantege is that algorithm designers have many options for implementing this simple architecure. 
@@
 execution.
+\end{itemize}
 
 On the other hand, there are several caveats as well:
-
 It is more costly to move entire subpopulations as messages than only passing certain individuals 

### 2020-01-19 draft.tex
Revised and fixed some things

@@
 \item 
-Also, the reactive controller, gives designers more control over the milti-population algorithm.
+Also, the reactive controller gives designers more control over the multi-population algorithm.
 In this process, designers can dynamicly change the number of populations, population parameters,
@@
 \item 
-Another advantege is that algorithm designers have many options for implementing this simple architecure. 
+Another advantege is that algorithm designers have many options for implementing this simple architecture. 
 The same basic components can be implmented as a single multithreaded program, or 
@@
 
-\subsection{Comparasion} 
+\subsection{Comparison} 
 \label{edm}
@@
     \includegraphics[width=\textwidth]{population_message}
-    \caption{Example of a parametric plot ($\sin (x), \cos(x), x$)}
+    \caption{Example of a parametric plot ($\sin (x), \cos(x), x$)} % Er, same caption. Better insert it now, or we might roftet... Also, same label. - JJ
     \label{fig:kafkEO}

### 2020-01-20 draft.tex
Adressing comments

@@
 with respect to problem size. The architecture and the implemented platform is an excellent
-alternative for locally or in the cloud, thus proving that cloud-native
+alternative for running locally or in the cloud, thus proving that cloud-native
 bioinspired algorithms perform better in their "natural" environment than other
@@
 
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\textwidth]{kafkEOsmall}
-    \caption{Example of a parametric plot ($\sin (x), \cos(x), x$)} % This caption? - JJ
-    \label{fig:kafkEO}
-\end{figure}
-
 % Based on a reactive architecture \cite{guervos2018introducing} , each
@@
 % abstract. We should unify terminology - JJ
+% "multi-population" is the whole thing, a subpopulation is a member of that.
+% "A multi-population has many subpopulations or should we call them just populations?"
+% I will add a comment about this in the Multi-population section - Mario
 from one process to the other. By "continuous stream," we mean that ideally,
@@
                                 % simply sent - JJ 
+                                % Meaning it does not wait for a respone, it continues doing other stuff - Mario 
 to the queuing system; this
@@
 
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\textwidth]{kafkEOsmall}
+    \caption{The proposed general architecture, 
+     showing each process in a swimlane, message-dataflow,
+     message queues, and high-level tasks for each process.} % This caption? - JJ
+    \label{fig:kafkEO}
+\end{figure}
+
 \subsection{Event Driven Model} 
@@
 of the box, and pull or pop operations as solid arrows leaving from the right side.
+Also, the architecture has at least four processes indicated in the diagram as
+swimlanes: First, the Setup process, responsible for the reading a configuration
+file, and creating the initial subpopulations. Second, a Controller process,
+responsible for the migration between subpopulations, and keeping track of the
+iterations of the algorithm. The Message Queue process runs the Input and Output
+queues. Finally, there is at least one Stateless Function process responsible
+for running the isolated algorithms. In the example, two processes PSO and GA,
+are shown.
+
 The algorithm starts with the Setup process pulling a configuration message from
@@
 for a GA, the mutation rate, type of crossover operator, and other values are indicated.
-Each subpopulation is then pushed to the Input queue, so they can be consumed 
+
+\item Each subpopulation is then pushed to the Input queue, so they can be consumed 
 by Stateless functions responsible for the execution of the search.  % Why the caps in Stateless? - JJ  
+% It refers to the swimlines in the diagram, maybe I should point it out  - Mario
+% Added a comment above. - Mario
 
@@
 They take the current state of the population and run the algorithm for a certain 
-number of iterations. Once they finish the execution, the population state
+number of iterations.
+
+\item Once they finish the execution, the population state
 is packed again along with aditional metadata about the execution of the algorithm.
-The resulting populations is now pushed to the Output queue. Once finished, another 
+The resulting populations are now pushed to the Output queue. Once finished, another 
 subpopulation is pulled from the queue.
@@
 and if an optimal solution has been found or the maximum number of iteration has been
-reached it signals the stop of the execution. Otherwise, it passes the stream of messages 
+reached it signals the stop of the execution.
+
+\item Otherwise, it passes the stream of messages 
 to a migration process, where subpopulations are mixed with each other. New populations 
@@
 Also, the reactive controller gives designers more control over the multi-population algorithm.
-In this process, designers can dynamicly change the number of populations, population parameters,
+In this process, designers can dynamically change the number of populations, population parameters,
 and migration details on-the-fly.

### 2020-01-20 draft.tex
Captions added

@@
     \includegraphics[width=\textwidth]{classicisland}
-    \caption{Example of a parametric plot ($\sin (x), \cos(x), x$)}
-    \label{fig:kafkEO}
+    \caption{ Population State and dataflow between processing nodes of a classic Island-based multi-population algorithm. }
+    \label{fig:classicisland}
 \end{figure}
@@
     \includegraphics[width=\textwidth]{pool_island}
-    \caption{Example of a parametric plot ($\sin (x), \cos(x), x$)}
-    \label{fig:kafkEO}
+    \caption{Population State and dataflow between processing nodes of a Pool-based multi-population algorithm.}
+    \label{fig:pool_island}
 \end{figure}
@@
     \includegraphics[width=\textwidth]{population_pool}
-    \caption{Example of a parametric plot ($\sin (x), \cos(x), x$)}
-    \label{fig:kafkEO}
+    \caption{Population State and dataflow between processing nodes of the EvoSpace algorithm.}
+    \label{fig:population_pool}
 \end{figure}
@@
     \includegraphics[width=\textwidth]{population_message}
-    \caption{Example of a parametric plot ($\sin (x), \cos(x), x$)} % Er, same caption. Better insert it now, or we might roftet... Also, same label. - JJ
-    \label{fig:kafkEO}
+    \caption{Population State and dataflow between processing nodes of the EvoSpace algorithm.} % Er, same caption. Better insert it now, or we might roftet... Also, same label. - JJ
+    % Ok, it was a "temporal" Copy-Paste solution - Mario
+    \label{fig:population_message}
 \end{figure}
@@
     \includegraphics[width=\textwidth]{experiment_flow}
-    \caption{Example of a parametric plot ($\sin (x), \cos(x), x$)}
+    \caption{EvoSwarm Workflow}
 \end{figure}
@@
     \includegraphics[width=\textwidth]{worker}
-    \caption{Example of a parametric plot ($\sin (x), \cos(x), x$)}
+    \caption{Serverless function implementation details. Showing the worker algorithm in each container}
 \end{figure}

### 2020-01-21 draft.tex
Bigger Figures
New Figuares can be seen when printed. Using 12 points  font.
Comp iniciated

@@
 \usepackage{lineno,hyperref}
+\usepackage{rotating}
 \graphicspath{ {./images/} }
@@
 
-\begin{figure}[h]
+\begin{sidewaysfigure}
     \centering
@@
     \label{fig:kafkEO}
-\end{figure}
+\end{sidewaysfigure}
 
@@
 
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=\textwidth]{population_message}
+    \caption{Population State and dataflow between processing nodes of a message-based algorithm.}
+    \label{fig:population_message}
+\end{figure}
+
 On the other hand, there are several caveats as well:
@@
 
-\subsection{Workers} 
-\label{edm}
-
-\subsection{Comparison} 
-\label{edm}
-
-\begin{figure}[h]
+\begin{figure}[ht]
     \centering
@@
 
-\begin{figure}[h]
+\subsection{Comparison} 
+\label{comparison}
+
+Next, we compare the message-driven architecture we presented against four other
+proposals for multi-population algorithms found in the literature. We center the
+comparison on the coupling and communication between the main components:
+Subpopulations, Processing Nodes, and Algorithms.  In Figure \ref{fig:population_message} we show the
+main components of a message-driven architecture. We can see that the algorithm
+and the subpopulations are separated, and the Population Queue keeps the state
+of subpopulations. While algorithms only need the subpopulations as parameters
+for their execution.  In contrast, in the classic Island model shown in Figure
+\ref{fig:classicisland}, algorithms are coupled with a subpopulation, and the same processing
+node keeps both the algorithm process and the state of the subpopulation.  Other
+processing nodes follow the same configuration and form of execution, and they
+only pass specific individuals between them.  In this model, adding additional
+processing nodes, on-runtime, can be more difficult, because they need to have
+an addressing mechanism. % JJ, help I don't know if this is clearly understood, a mean like a DNS - Mario
+
+\begin{figure}[ht]
     \centering
@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
     \centering
@@
 
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\textwidth]{population_message}
-    \caption{Population State and dataflow between processing nodes of the EvoSpace algorithm.} % Er, same caption. Better insert it now, or we might roftet... Also, same label. - JJ
-    % Ok, it was a "temporal" Copy-Paste solution - Mario
-    \label{fig:population_message}
-\end{figure}
+\subsection{Workers} 
+\label{edm}
 
@@
     \caption{EvoSwarm Workflow}
-\end{figure}
+\end{sidewaysfigure}
 
-\begin{figure}[h]
+\begin{figure}[ht]
     \centering

### 2020-01-21 draft.tex
New section

@@
 and multi-population methods. Section \ref{soa} presents state of the art relevant to
-our work. In Section \ref{method}, we present the proposed method.  Section \ref{setup} describes
+our work. In Section \ref{method}, we present the proposed method and the container-based
+application in Section \ref{docker} .  Section \ref{setup} describes
 the design of the empirical evaluation we designed to assess the effectiveness
@@
 
-\subsection{Comparison} 
+\subsection{Comparison with other cloud-based works} 
 \label{comparison}
@@
 of subpopulations. While algorithms only need the subpopulations as parameters
-for their execution.  In contrast, in the classic Island model shown in Figure
+for their execution.
+
+% We need to add references to the respective papers - Mario
+
+In contrast, in the classic Island model shown in Figure
 \ref{fig:classicisland}, algorithms are coupled with a subpopulation, and the same processing
@@
 
+A typical design pattern used to alleviate the above drawbacks is the use of a
+central repository or pool of individuals that is available to all processing
+nodes.  In Figure \ref{fig:pool_island} we show the main components of a pool-based
+multi-population algorithm. Although algorithms and subpopulation state remains
+coupled, now, processing nodes do not communicate directly with each other.
+Instead, they interchange individuals with the pool. Communication between nodes
+is not affected if the system adds or removes nodes because they do not know
+about each other.  Since the state of the global population is stored in
+external processing nodes, the system needs additional communication and
+processing for keeping track of each subpopulation.
+
+We show another pool-based approach in Figure  \ref{fig:population_pool}. In
+this design, the global population is stored in a central repository, while
+isolated algorithms take random samples of the global population and use this
+temporal subpopulation as parameters. An advantage of this design is that the
+sampling provides a type of migration between isolated subpopulations. A problem
+found is that when a processing node returns a subpopulation to the pool, the
+state of the subpopulation is lost. But having the algorithm decoupled opens the
+possibility of having a serverless-function implementation.
+
 \begin{figure}[ht]
@@
 
-\subsection{Workers} 
+\section{Container-based application} 
+\label{docker}
+
+We have mentioned several advantages of a cloud-native application for running
+multi-population based algorithms, but there are in our experience several
+practical issues of this approach. In general, cloud platforms have free-tiers
+for their services, and many offer academic discounts to students. But once we
+pass to the pay-as-you-go tier, computing resources need to be managed.  First,
+in many cases, there is a credit-card need to guarantee the payment of the
+resources consumed and someone responsible for assigning quotas and, more
+importantly, special keys for using the cloud's API.  If these keys are
+compromised or shared by their holders,  additional costs or administrative
+procedures are in order. For those cases when a prototype platform is needed,
+for instance, students developing a new algorithm variant or experimenting with
+a published work, a container-based platform can be more practical.
+
+\subsection{Containers} 
+\label{edm}
+
+\subsection{Docker compose} 
 \label{edm}

### 2020-01-21 draft.tex
PNGs to Vector Pdfs

@@
 for instance, students developing a new algorithm variant or experimenting with
-a published work, a container-based platform can be more practical.
+a published work, a container-based platform can be more practical.Container
+technology enables researchers the local deployment of all the infrastructure
+needed to run the experiments. Later, if needed, the same deployment could be
+deployed to the cloud.
+
+In this section, we propose a design of a reactive container-based application
+for executing multi-population based optimization experiments.  For the design,
+we followed the design patterns highlighted in the previous section, and again
+we can go back to Figure \ref{fig:kafkEO} and use it as a guide for explaining
+the main components and their interactions. But first, we give a brief
+description of the technology we used.
 
 \subsection{Containers} 
-\label{edm}
+\label{containers}
+% TO DO
 
@@
 \label{edm}
+% TO DO
 

### 2020-01-21 draft.tex
Reference added, closes #3

@@
 
-\subsection{Docker compose} 
-\label{edm}
+\subsection{Docker Compose} 
+\label{compose}
 % TO DO
 
+\subsection{Redis Container} 
+\label{redis}
+
+\subsection{Controller Container} 
+\label{controller}
+
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=\textwidth]{worker}
+    \caption{Serverless function implementation details. Showing the worker algorithm in each container}
+    \label{fig:worker}
+\end{figure}
+
+\subsection{Worker Containers} 
+\label{workers}
+
+Figure \ref{fig:worker}.
+
 \subsection{Experimental Process} 
-\label{edm}
+\label{experiment_flow}
 
@@
     \caption{EvoSwarm Workflow}
+    \label{fig:experiment_flow}
 \end{sidewaysfigure}
 
-\begin{figure}[ht]
-    \centering
-    \includegraphics[width=\textwidth]{worker}
-    \caption{Serverless function implementation details. Showing the worker algorithm in each container}
-\end{figure}
+Figure \ref{fig:experiment_flow} shows the workflow users can follow to run
+experiments in EvoSwarm. The first step is to develop the search strategy
+algorithms that are going to be used by the multi-population algorithm.
+Algorithm Developers can upload the container definition or images directly to
+DockerHub. Another option is to upload the image definition (Dockerfile) to
+GitHub and define a trigger to re-upload the image to DockerHub.  Docker Compose
+configuration files can reference these containers. When a user needs to deploy
+the application, he or she needs to edit the Docker compose file to specify the
+type and number of containers needed for the application.  The Docker compose
+application is responsible for starting all the containers or downloading new
+versions if it is necessary. After starting the application, users can push
+Experiment Definition files and start monitoring the execution in the terminal.
+After the experiment is finished, the user must execute another script using the
+experiment id to generate a collection of files containing all the data
+generated by the experiment. The script generates these files in a JSON format.
+Researchers can use these files to analyze and plot experiment data.  Python
+scripts are included in the repository to plot the running times for the
+experiments, and all other plots used in this paper.  There is also a script to
+generate the files needed by the COCO framework, which can generate standard
+comparisons against other methods.
 

### 2020-01-22 draft.tex
Images in pdf, changed ignore file
package algorithm
marble controller diagram

@@
 \usepackage{rotating}
+\usepackage{algpseudocode,algorithm}
+
 \graphicspath{ {./images/} }
@@
 \label{controller}
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=\textwidth]{marble_controller}
+    \caption{Marble diagram for the Reactive Python implementation of the Controller}
+    \label{fig:marble_controller}
+\end{figure}
 
@@
 
+\begin{algorithm}
+    \caption{<your caption for this algorithm>}
+    \label{<your label for references later in your document>}
+    \begin{algorithmic}[1]
+            \Procedure{Euclid}{$a,b$}\Comment{The g.c.d. of a and b}
+            \State $r\gets a\bmod b$
+            \While{$r\not=0$}\Comment{We have the answer if r is 0}
+            \State $a\gets b$
+            \State $b\gets r$
+            \State $r\gets a\bmod b$
+            \EndWhile\label{euclidendwhile}
+            \State \textbf{return} $b$\Comment{The gcd is b}
+            \EndProcedure
+
+    \end{algorithmic}
+\end{algorithm}
+
 \subsection{Experimental Process} 
@@
 Figure \ref{fig:experiment_flow} shows the workflow users can follow to run
-experiments in EvoSwarm. The first step is to develop the search strategy
+experiments in Evo\-Swarm. The first step is to develop the search strategy
 algorithms that are going to be used by the multi-population algorithm.

### 2020-01-22 draft.tex
Controller section

@@
 \usepackage{algpseudocode,algorithm}
+\usepackage{hyperref}
 
@@
 \label{controller}
+
+The controller is an essential component of the architecture (Figure \ref{fig:kafkEO})
+because it is responsible for maintaining the evolutionary loop. It takes newly
+evolved populations from the Output Queue, mixes them,   and produces new
+messages from the result. At the same time, it must keep track of two conditions
+for ending the loop: the number of messages it has received reaches a maximum
+value, or the error of the best solution reaches a specific size.  Finally, it
+must filter out remnant messages from other experients. Messages from other
+experiments can remain in the queues because of the asynchronous nature of the
+system.  We chose to implement the controller in Python but using an API
+specialized in asynchronous event-driven programming over data streams.  The
+open-source library is called Reactive Extensions for Python (RxPY)
+\url{https://github.com/ReactiveX/RxPY} based on the Observer pattern
+\cite{gamma1995design} and functional programming.
+
 \begin{figure}[ht]
@@
 \begin{algorithm}
-    \caption{<your caption for this algorithm>}
-    \label{<your label for references later in your document>}
+    \caption{Migration}
+    \label{read}
     \begin{algorithmic}[1]
-            \Procedure{Euclid}{$a,b$}\Comment{The g.c.d. of a and b}
-            \State $r\gets a\bmod b$
-            \While{$r\not=0$}\Comment{We have the answer if r is 0}
-            \State $a\gets b$
-            \State $b\gets r$
-            \State $r\gets a\bmod b$
-            \EndWhile\label{euclidendwhile}
-            \State \textbf{return} $b$\Comment{The gcd is b}
+        \Procedure{cxBestFromEach}{$pop_1,pop_2$}
+            \State $pop_1.sort()$
+            \State $pop_2.sort()$
+            \State $size\gets min(len(pop_1), len(pop_2))$
+            \State $cxpoint\gets (size-1)/2$
+            \State $pop_1[cxpoint:]\gets pop_2[:cxpoint+2]$
+            \State \textbf{return} $pop_1$
         \EndProcedure

### 2020-01-23 draft.tex
General revision. Check out comments

@@
 
-The organization of the paper is as follows. First, in Section \ref{multi}, we present a
+The organization of the paper is as follows: First, in Section \ref{multi}, we present a
 background of the fundamental issues of integrating nature-inspired optimization
@@
 offer the conclusions of this paper and suggestions on future work in Section \ref{conclusions}.
-
-% A state of the art should be included here too - JJ
 
@@
 Section \ref{multi} and the general design principles of cloud-native
-applications we have mention earlier. As a first step, we describe the general
-architecture first, and then we describe each component in detail.
+applications we have mentioned earlier. As a first step, we describe the general
+architecture first, and then  every component in detail.
 
-The primary data element used throughout the system is the subpopulation, and at a higher
+The most elemental data structure used throughout the system is the subpopulation; and at a higher
 level, we can view the system as a continuous stream of subpopulations
@@
 % "A multi-population has many subpopulations or should we call them just populations?"
-% I will add a comment about this in the Multi-population section - Mario
+% I will add a comment about this in the Multi-population section -
+% Mario
+% use the usual "multi-deme" terminology. A population has different
+% sub-populations nad uses a multi-deme implementation - JJ
 from one process to the other. By "continuous stream," we mean that ideally,
@@
                                 % simply sent - JJ 
-                                % Meaning it does not wait for a respone, it continues doing other stuff - Mario 
+                                % Meaning it does not wait for a
+                                % respone, it continues doing other
+                                % stuff - Mario
+% That's rather "fire and forget". Or simply say it's sent without
+% waiting for a response - JJ
 to the queuing system; this
@@
 message queue by defining an event handler method that is going to be triggered when a
-message is pushed to the message queue. The two components at each end of the queue are
+message is pushed to it. The two components at each end of the queue are
 Producers that push messages to the queue and Consumers that pull them. This
-pattern is an essential component of a reactive architecture because it is
-highly scalable; one of the reasons for this are stateless functions. We can
-define stateless functions as methods that do not have side effects, meaning
-they do not need to read, keep, or modify data outside of the method. If
-Consumers are implemented using stateless functions, then there is no difference
+pattern is an essential component of a reactive architecture and makes
+it 
+highly scalable; one of the reasons for this is the use of functions
+that have no secondary effects or access global storage, that is,
+stateless functions. These functions do not need to read, keep, or modify data outside of the method. If
+Consumers are implemented using stateless functions, and thus, there is no difference
 between having one or many copies of the same function pulling work from the queue all at the
-same time, because no side effects are going to result from the
-concurrency.
+same time, because no side effects, including unwanted resource
+locking problems,  are going to result from this concurrency.
 % This is important and we should probably mention that in the
@@
     \centering
-    \includegraphics[width=\textwidth]{kafkEOsmall}
+    \includegraphics[width=\textwidth]{KafkEOsmall}
     \caption{The proposed general architecture, 
      showing each process in a swimlane, message-dataflow,
-     message queues, and high-level tasks for each process.} % This caption? - JJ
+     message queues, and high-level tasks for each process.} 
     \label{fig:kafkEO}
@@
 of the algorithm. We give more details about the configuration data structure
-in Section []. Once the configuration is read, we can follow the path of 
+in Section []. % Missing
+Once the configuration is read, we can follow the path of 
 messagess as follows:
@@
 for their services, and many offer academic discounts to students. But once we
-pass to the pay-as-you-go tier, computing resources need to be managed.  First,
-in many cases, there is a credit-card need to guarantee the payment of the
+pass to the pay-as-you-go tier, computing resources need to be
+managed, and, where possible, its cost minimized.  First,
+in many cases,  a credit-card is needed to guarantee the payment of the
 resources consumed and someone responsible for assigning quotas and, more
@@
 for instance, students developing a new algorithm variant or experimenting with
-a published work, a container-based platform can be more practical.Container
+a published work, a container-based platform can be more practical. Container
 technology enables researchers the local deployment of all the infrastructure
@@
 deployed to the cloud.
+% I don't see this line of argument. You can do exactly the same with
+% virtual machines instead of containers. Also, you can test a virtual
+% infrastructure locally and then only when it's needed deploy it to
+% the cloud. If you want to give good reasons for using containers
+% instead of virtual machines, there are quite a lot non-economic ones
+% - JJ
 
@@
 
+% Do we need a description of every container? Wouldn't a general
+% description of the containers used be enough? - JJ
+
 \subsection{Controller Container} 
@@
 for ending the loop: the number of messages it has received reaches a maximum
-value, or the error of the best solution reaches a specific size.  Finally, it
+value, or the error of the best solution goes below a specific threshold.  Finally, it
 must filter out remnant messages from other experients. Messages from other

### 2020-01-23 draft.tex
A few changes

@@
 % Again, you have to justify the need for a reactive architecture and define it before this paragraph - JJ
-
-% There is a big gap here where you go from generic considerations
-% about cloud computing to the presentation of this method. In the
-% middle, there should be motivation for it, what are you trying to
-% address precisely, and how you intend to address it. "Approaching
-% multi-population algorithms in cloud computing" is not nearly
-% enough. There are several points here, and every one of them needs
-% to be addressed and justified
-% * Using containers as opposed to a serverless architecture or a
-% virtual-machine based one.
-% * Combining several metaheuristics instead of using a single one.
-% * Reducing the amount of parameters.
-% There needs to be a motivation for every one, relate them to the
-% objectives of the paper (which need to be clarified) and eventually
-% prove the contribution of every one of them to the final result - JJ
 
@@
 
-% Based on a reactive architecture \cite{guervos2018introducing} , each
-% subpopulation is represented as an event message that is pushed asynchronously
-% into a message queue. Incoming messages trigger stateless functions that receive
-% the subpopulation and proceed to run an algorithm, using the parameters and
-% population included in the message. After the specified number of iterations,
-% each stateless function returns the evolved subpopulation by again pushing a
-% message to another queue, used for receiving the resulting subpopulations.
-% Subpopulations are received from the queue by a controller that is responsible
-% for mixing the individuals from different subpopulations and producing new
-% subpopulations. These new subpopulations are pushed again by the controller into
-% the message queue, creating a loop. The cycle stops when the controller receives
-% a subpopulation containing a candidate solution that satisfies a particular
-% condition, or a maximum number of messages were received.
-
 In this section, we present the design
@@
 % sub-populations nad uses a multi-deme implementation - JJ
+% Let's talk about it, I am using the Multi-population terminology 
+% maybe multi-dem is more common in the EA community but not 
+% much in the Swarm Intelligence side?  
+% Any how, we decide and explain the difference so readers of both 
+% camps are Ok with it. 
+% I prefer the use of population and something like the global population - Mario
+
 from one process to the other. By "continuous stream," we mean that ideally,
@@
 to communicate a subpopulation to another, it must pack the sub-population into
-a message, and then send the message asynchronously % I don't know if
+a message, and then send the message 
+% I don't know if
                                 % asynchronously applies here. It's
@@
 % waiting for a response - JJ
-to the queuing system; this
+% corrected - Mario
+to the queuing system without waiting for a response; this
 means that after the process pushes the message, it continues its execution
@@
 of the algorithm. We give more details about the configuration data structure
-in Section []. % Missing
+in Section \ref{experiment_flow}. % Missing  % added 
 Once the configuration is read, we can follow the path of 
@@
 % - JJ
+% OK, I will add an issue - Mario
 
@@
 
-\subsection{Redis Container} 
-\label{redis}
-
 % Do we need a description of every container? Wouldn't a general
 % description of the containers used be enough? - JJ 
+% Ok, I think only workers and the controller need to be described in detail - Mario
 
@@
 open-source library is called Reactive Extensions for Python (RxPY)
-\url{https://github.com/ReactiveX/RxPY} based on the Observer pattern
+\url{https://github.com/ReactiveX/RxPY}an it is based on the Observer pattern
 \cite{gamma1995design} and functional programming.

### 2020-01-24 draft.tex
Marble diagram notation explained

@@
 system.  We chose to implement the controller in Python but using an API
-specialized in asynchronous event-driven programming over data streams.  The
-open-source library is called Reactive Extensions for Python (RxPY)
-\url{https://github.com/ReactiveX/RxPY}an it is based on the Observer pattern
-\cite{gamma1995design} and functional programming.
+specialized in asynchronous event-driven programming over data streams.  
+The open-source library is called Reactive Extensions (ReactiveX) for Python 
+\url{https://github.com/ReactiveX/RxPY} an it is based on the Observer 
+pattern \cite{gamma1995design} and functional programming.
+
+In ReactiveX, an observer object subscribes to an Observable. Observables emit a
+sequence of items, and all the subscribed observers react to each emission. The
+ReactiveX library includes several reactive operators that we can use to
+transform and combine sequences of items. These operators provide reactive
+extensions that allow us to compose asynchronous sequences together in a
+declarative manner. In Figure /ref{marble}, we use a so-called marble diagram to
+represent the composition of Observables and reactive extensions operators. We
+represent the timeline of an Observable as a horizontal arrow, which indicates
+that time flows from left to right. In the diagram, items are represented as
+marbles. The position of each marble indicates the point in time when they were
+emitted by the Observable. Reactive operators are represented as text boxes,
+showing the transformation to be applied. Normally, a transformation results in
+another Observable, again emitting the results.
 

### 2020-01-24 draft.tex
Controller explained

@@
 
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=\textwidth]{marble_controller}
+    \caption{Marble diagram for the Reactive Python implementation of the Controller}
+    \label{fig:marble_controller}
+\end{figure}
+
 In ReactiveX, an observer object subscribes to an Observable. Observables emit a
@@
 showing the transformation to be applied. Normally, a transformation results in
-another Observable, again emitting the results.
+another Observable, again emitting new results.  For asynchronous programming, 
+this approach is more elegant than nested callbacks, that are more difficult 
+to code and debug.
 
-\begin{figure}[ht]
-    \centering
-    \includegraphics[width=\textwidth]{marble_controller}
-    \caption{Marble diagram for the Reactive Python implementation of the Controller}
-    \label{fig:marble_controller}
-\end{figure}
+Now we proceed to explain the reactive implementation of the controller.
+In Figure /ref{marble}, we  have the following composition of Observables:
+
+\begin{enumerate}
+
+\item The controller continually pulls new messages from the Output 
+message queue. These messages are instantly emitted by the consumed\_messages 
+Observable. In this particular period, we can see that the stream receives 
+two populations from another experiment, and are shown as two red marbles.  
+The filter operator removes all messages that belong to a different experiment; 
+in this case, we are only interested in blue marbles.
+
+\item The max\_iterations parameter indicates the number of populations that
+are going to be accepted. When this number of messages is reached, we must
+end the loop. In this example, max\_iterations = 6,  the take operator 
+assures that only 6 messages are received. After the Observable emits 
+the sixth message, it triggers the completed event.
+
+\item Many observers can be subscribed to the valid\_messages Observable because it
+emits all the valid items. At the moment, there are to additional methods
+subscribed that we are not showing, one for logging and the other for monitoring
+the search.
+
+\item The buffer\_with\_count(3) operator, waits until it receives 
+three valid messages to emit a single message that contains a list with
+the previous three. The migration method requires that list to mix them.
+
+\item The population\_mixer method pushes new populations to the messages Observable,
+a publish(population) observer is responsible for pushing 
+the newly generated populations back to the Input Message Queue.
+
+\end{enumerate}
 

### 2020-01-24 draft.tex
Update draft.tex

@@
 extensions that allow us to compose asynchronous sequences together in a
-declarative manner. In Figure /ref{marble}, we use a so-called marble diagram to
+declarative manner. In Figure \ref{fig:marble_controller}, we use a so-called marble diagram to
 represent the composition of Observables and reactive extensions operators. We
@@
 Now we proceed to explain the reactive implementation of the controller.
-In Figure /ref{marble}, we  have the following composition of Observables:
+In Figure \ref{fig:marble_controller}, we  have the following composition of Observables:
 
@@
 \item The controller continually pulls new messages from the Output 
-message queue. These messages are instantly emitted by the consumed\_messages 
+message queue. These messages are instantly emitted by the  \texttt{consumed\_messages}
 Observable. In this particular period, we can see that the stream receives 
@@
 
-\item The max\_iterations parameter indicates the number of populations that
+\item The \texttt{max\_iterations} parameter indicates the number of populations that
 are going to be accepted. When this number of messages is reached, we must
-end the loop. In this example, max\_iterations = 6,  the take operator 
+end the loop. In this example, \texttt{max\_iterations = 6},  the take operator 
 assures that only 6 messages are received. After the Observable emits 
@@
 
-\item Many observers can be subscribed to the valid\_messages Observable because it
+\item Many observers can be subscribed to the \texttt{valid\_messages}  Observable because it
 emits all the valid items. At the moment, there are to additional methods
@@
 
-\item The buffer\_with\_count(3) operator, waits until it receives 
+\item The   \texttt{buffer\_with\_count(3)} operator, waits until it receives 
 three valid messages to emit a single message that contains a list with
 the previous three. The migration method requires that list to mix them. 
+In our previous work, we needed a local buffer for storing a certain 
+number of populations to mix them with others. This design has the advantage
+of not needing extra memory, and it integrates better with the reactive paradigm.
+A possible disadvantage can be that it only mixes contiguous populations,
+but this can be mitigated with a larger buffer.
 
-\item The population\_mixer method pushes new populations to the messages Observable,
-a publish(population) observer is responsible for pushing 
+\item The  \texttt{population\_mixer} method pushes new populations to the messages Observable,
+a  \texttt{publish(population)} observer is responsible for pushing 
 the newly generated populations back to the Input Message Queue.
-
 \end{enumerate}

### 2020-01-26 draft.tex
image added
Caption added

@@
 experiments, and all other plots used in this paper.  There is also a script to
-generate the files needed by the COCO framework, which can generate standard
+generate the files needed by the COCO framework \cite{hansen2016coco}, which can generate standard
 comparisons against other methods.
@@
 
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=\textwidth]{sp_size_worker}
+    \caption{ Time required to complete experiments with different number of workers and populations sizes }
+    \label{fig:worker}
+\end{figure}
+
 \section{Conclusions} 

### 2020-01-26 draft.tex
Minor test
comment

@@
 
-Taking these factors into account, in this work, we present EcoSwarm, % maybe a less generic name? - JJ
-% agree - Mario
+Taking these factors into account, in this work, we present EcoSwarm, 
 a multi-container Docker application that reactively processes isolated and
@@
     \caption{ Time required to complete experiments with different number of workers and populations sizes }
-    \label{fig:worker}
+    \label{fig:spworker}
 \end{figure}

### 2020-01-26 draft.tex
A few minor corrections

@@
 The algorithm starts with the Setup process pulling a configuration message from
-the experiements queue, this is not shown in the diagram. The configuration messsage 
+the experiments queue, this is not shown in the diagram. The configuration message 
 includes all the parameters needed to execute the algorithm, among other things,
@@
 \begin{enumerate}
-\item In this step, the specified number of subpopulations is created acording to the 
+\item In this step, the specified number of subpopulations is created according to the 
 parameters found in the configuration structure. The population at this moment is just
@@
 \item Once they finish the execution, the population state
-is packed again along with aditional metadata about the execution of the algorithm.
+is packed again along with additional metadata about the execution of the algorithm.
 The resulting populations are now pushed to the Output queue. Once finished, another 
@@
 \item 
-Another advantege is that algorithm designers have many options for implementing this simple architecture. 
-The same basic components can be implmented as a single multithreaded program, or 
+Another advantage is that algorithm designers have many options for implementing this simple architecture. 
+The same basic components can be implmented as a single multi-threaded program, or 
 as a highly scalable serverless cloud application. Most modern languages include 
@@
 It is more costly to move entire subpopulations as messages than only passing certain individuals 
-from a process to another. Designers must conmsider this cost when working with large
+from a process to another. Designers must consider this cost when working with large
 individuals, and if possible send sub-populations using several messages, use compression or 
 adjust the size of subpopulations. Pool-based algorithms also suffer from this drawback, but 
-web based immplementations have been working with continous optimization problems of 1000 dimensions.
+web based implementations have been working with continuous optimization problems of 1000 dimensions.
 But, in other cases this is not a viable approach.
@@
 main components of a message-driven architecture. We can see that the algorithm
-and the subpopulations are separated, and the Population Queue keeps the state
+and subpopulations are separated, and the Population Queue keeps the state
 of subpopulations. While algorithms only need the subpopulations as parameters
@@
 state of the subpopulation is lost. But having the algorithm decoupled opens the
-possibility of having a serverless-function implementation.
+possibility of a serverless-function implementation.
 
@@
 value, or the error of the best solution goes below a specific threshold.  Finally, it
-must filter out remnant messages from other experients. Messages from other
+must filter out remnant messages from other experiments. Messages from other
 experiments can remain in the queues because of the asynchronous nature of the

### 2020-01-27 draft.tex
Experimental Setup

@@
 Figure \ref{fig:kafkEO}. Again, we can explain the model using the analogy
-of Producers and Consumers of messages. First, we can see that
-there are two queues, one labeled Input, and the other one Output. In the diagram, 
+of producers and consumers of messages. First, we can see that
+there are two queues, one labeled \texttt{Input}, and the other one \texttt{Output}. In the diagram, 
 push operations on a queue are represented by solid arrows connecting to the left side
-of the box, and pull or pop operations as solid arrows leaving from the right side.
+of the queue box, and pull or pop operations as solid arrows leaving from the right side.
 Also, the architecture has at least four processes indicated in the diagram as
-swimlanes: First, the Setup process, responsible for the reading a configuration
-file, and creating the initial subpopulations. Second, a Controller process,
+swimlanes: First, the \texttt{Setup} process, responsible for the reading a configuration
+file, and creating the initial subpopulations. Second, the \texttt{Controller} process,
 responsible for the migration between subpopulations, and keeping track of the
-iterations of the algorithm. The Message Queue process runs the Input and Output
-queues. Finally, there is at least one Stateless Function process responsible
-for running the isolated algorithms. In the example, two processes PSO and GA,
+iterations of the algorithm. The \texttt{Message Queue} process runs the \texttt{Input} and \texttt{Output}
+queues. Finally, there is at least one \texttt{Stateless Function} process responsible
+for running the isolated algorithms. In the example, two processes \texttt{PSO} and \texttt{GA},
 are shown.
 
-The algorithm starts with the Setup process pulling a configuration message from
-the experiments queue, this is not shown in the diagram. The configuration message 
+The algorithm starts with the \texttt{Setup} process pulling a configuration message from
+the \texttt{Experiments} queue, this is not shown in the diagram. The configuration message 
 includes all the parameters needed to execute the algorithm, among other things,
@@
 \begin{enumerate}
-\item In this step, the specified number of subpopulations is created according to the 
+\item In this step, the specified number of populations are created according to the 
 parameters found in the configuration structure. The population at this moment is just
@@
 
-\item Each subpopulation is then pushed to the Input queue, so they can be consumed 
+\item Each subpopulation is then pushed to the \texttt{Input} queue, so they can be consumed 
 by Stateless functions responsible for the execution of the search.  % Why the caps in Stateless? - JJ  
@@
 
-\item One or more Stateless functions are constantly pulling subpopulation messages
-from the Input queue. Taking the population data and metadata as parameters, these 
+\item One or more \texttt{Stateless functions} are constantly pulling subpopulation messages
+from the \texttt{Input} queue. Taking the population data and metadata as parameters, these 
 functions are responsible for the actual execution of the optimization algorithm. 
@@
 
-\item The Controller process is responsible for keeping track of the progress of 
-the search. It pulls current subpopulations from the Output queue, inspects the metadata 
+\item The \texttt{Controller} process is responsible for keeping track of the progress of 
+the search. It pulls current populations from the \texttt{Output} queue, inspects the metadata 
 and if an optimal solution has been found or the maximum number of iteration has been
@@
 are generated from this migration, and they are again pushed to the Input queue to continue 
-in a loop. The Controller is also responsible of logging the metadata recived with the
+in a loop. The \texttt{Controller} is also responsible of logging the metadata recived with the
 messages.  
@@
 On the other hand, there are several caveats as well:
-It is more costly to move entire subpopulations as messages than only passing certain individuals 
+It is more costly to move entire populations as messages than only passing certain individuals 
 from a process to another. Designers must consider this cost when working with large
-individuals, and if possible send sub-populations using several messages, use compression or 
-adjust the size of subpopulations. Pool-based algorithms also suffer from this drawback, but 
+individuals, and if possible send populations using several messages, use compression or 
+adjust the size of populations. Pool-based algorithms also suffer from this drawback, but 
 web based implementations have been working with continuous optimization problems of 1000 dimensions.
@@
 comparison on the coupling and communication between the main components:
-Subpopulations, Processing Nodes, and Algorithms.  In Figure \ref{fig:population_message} we show the
+subpopulations, processing nodes, and algorithms.  In Figure \ref{fig:population_message} we show the
 main components of a message-driven architecture. We can see that the algorithm
-and subpopulations are separated, and the Population Queue keeps the state
-of subpopulations. While algorithms only need the subpopulations as parameters
+and subpopulations are separated, and the \texttt{Population queue} keeps the state
+of subpopulations. While algorithms only need the populations as parameters
 for their execution.
@@
 we can go back to Figure \ref{fig:kafkEO} and use it as a guide for explaining
-the main components and their interactions. But first, we give a brief
-description of the technology we used.
+the main components and their interactions.
 
-\subsection{Containers} 
-\label{containers}
+% But first, we give a brief description of the technology we used.
+
+% Add if needed
+%\subsection{Containers} 
+%\label{containers}
 % TO DO
 
-\subsection{Docker Compose} 
-\label{compose}
+%\subsection{Docker Compose} 
+%\label{compose}
 % TO DO
@@
 % description  of the containers used be enough? - JJ 
-% Ok, I think only workers and the controller need to be described in detail - Mario
+% Ok, I thin only workers and the controller need to be described in detail - Mario 
+\subsection{Message Queue Container} 
+\label{message_container}
+
+In this work, we implemented all message
+queues in the Redis memory store. Each queue is a Redis List object, and we use
+the \texttt{LPOP} (left pop), and \texttt{RPOP} (right pop) commands for a queue like behavior.
+In those cases where we needed a blocking behavior, i.e., when a process needs
+to wait until a message is available, we used the blocking versions \texttt{BLPOP} and
+\texttt{BRPOP}. Redis in-memory operation are very fast, and all of the obove 
+operations have a time complexity of $\mathcal{O}(1)$. We use the official 
+\texttt{redis:alpine} container image from DockerHub.
 
@@
 
-The controller is an essential component of the architecture (Figure \ref{fig:kafkEO})
+The controller is an essential component of the architecture (see Figure \ref{fig:kafkEO})
 because it is responsible for maintaining the evolutionary loop. It takes newly
-evolved populations from the Output Queue, mixes them,   and produces new
+evolved populations from the \texttt{Output Queue}, mixes them,   and produces new
 messages from the result. At the same time, it must keep track of two conditions
@@
 
-In ReactiveX, an observer object subscribes to an Observable. Observables emit a
-sequence of items, and all the subscribed observers react to each emission. The
+In ReactiveX, an \texttt{observer} object subscribes to an \texttt{Observable} instance. 
+Observables emit a sequence of items, and all the subscribed observers react to each emission. The
 ReactiveX library includes several reactive operators that we can use to
@@
 extensions that allow us to compose asynchronous sequences together in a
-declarative manner. In Figure \ref{fig:marble_controller}, we use a so-called marble diagram to
-represent the composition of Observables and reactive extensions operators. We
-represent the timeline of an Observable as a horizontal arrow, which indicates
+declarative manner. In Figure \ref{fig:marble_controller}, we use a marble diagram to
+represent the composition of \texttt{Observables} and reactive extensions operators. We
+represent the timeline of an \texttt{Observable} as a horizontal arrow, which indicates
 that time flows from left to right. In the diagram, items are represented as
 marbles. The position of each marble indicates the point in time when they were
-emitted by the Observable. Reactive operators are represented as text boxes,
+emitted by the \texttt{Observable}. Reactive operators are represented as text boxes,
 showing the transformation to be applied. Normally, a transformation results in
-another Observable, again emitting new results.  For asynchronous programming, 
+another \texttt{Observable}, again emitting new results.  For asynchronous programming, 
 this approach is more elegant than nested callbacks, that are more difficult 
@@
 \begin{enumerate}
-
-\item The controller continually pulls new messages from the Output 
+\item The controller continually pulls new messages from the \texttt{Output} 
 message queue. These messages are instantly emitted by the  \texttt{consumed\_messages}
-Observable. In this particular period, we can see that the stream receives 
+\texttt{Observable}. In this particular period, we can see that the stream receives 
 two populations from another experiment, and are shown as two red marbles.  
@@
 end the loop. In this example, \texttt{max\_iterations = 6},  the take operator 
-assures that only 6 messages are received. After the Observable emits 
+assures that only 6 messages are received. After the \texttt{Observable} emits 
 the sixth message, it triggers the completed event.
 
-\item Many observers can be subscribed to the \texttt{valid\_messages}  Observable because it
+\item Many observers can be subscribed to the \texttt{valid\_messages} \texttt{Observable} because it
 emits all the valid items. At the moment, there are to additional methods
@@
 three valid messages to emit a single message that contains a list with
-the previous three. The migration method requires that list to mix them. 
+the previous three. The \texttt{population\_mixer}  method requires that list to mix them. 
 In our previous work, we needed a local buffer for storing a certain 
@@
 
-\item The  \texttt{population\_mixer} method pushes new populations to the messages Observable,
-a  \texttt{publish(population)} observer is responsible for pushing 
-the newly generated populations back to the Input Message Queue.
+\item The \texttt{population\_mixer} receives a list of three populations,
+let us say [A, B, C], and calls the migration method shown in
+Algorithm \ref{alg:migration}  for [A, B], [B, C] and [A, C]. 
+The migration algorithm sorts both populations and generates
+a new population containing the best half from each.
+This migration method is similar to those used successfully in previous 
+works on pool-based algorithms \ref{fig:pool_island}.  
+Finally, the method pushes the new populations to the 
+\texttt{messages Observable}. From there, a
+\texttt{publish(population)} observer is responsible for pushing
+the newly generated populations back to the \texttt{Input} message queue.
 \end{enumerate}
+
+\begin{algorithm}
+    \caption{Migration}
+    \label{alg:migration}
+    \begin{algorithmic}[1]
+        \Procedure{cxBestFromEach}{$pop_1,pop_2$}
+            \State $pop_1.sort()$
+            \State $pop_2.sort()$
+            \State $size\gets min(len(pop_1), len(pop_2))$
+            \State $cxpoint\gets (size-1)/2$
+            \State $pop_1[cxpoint:]\gets pop_2[:cxpoint+2]$
+            \State \textbf{return} $pop_1$
+        \EndProcedure 
+    \end{algorithmic}
+\end{algorithm}
+
+In the next section, we follow the algorithm flow and describe the 
+implementation details of the \texttt{worker} containers.
 
@@
 
-Figure \ref{fig:worker}.
+Worker containers include a Python script called \texttt{main.py} running an
+infinite loop, continually trying to pull a new population to work on it. Once
+\texttt{main.py} receives a message containing configuration data, it creates a
+worker object responsible for initializing and running the specified stateless
+function (i.e., GA or PSO). Once the function returns a new population, the
+worker pushes the results to the \texttt{Output} queue, finishes, and passes
+the execution to the main loop. This process is shown in Figure
+\ref{fig:worker}.
 
-\begin{algorithm}
-    \caption{Migration}
-    \label{read}
-    \begin{algorithmic}[1]
-        \Procedure{cxBestFromEach}{$pop_1,pop_2$}
-            \State $pop_1.sort()$
-            \State $pop_2.sort()$
-            \State $size\gets min(len(pop_1), len(pop_2))$
-            \State $cxpoint\gets (size-1)/2$
-            \State $pop_1[cxpoint:]\gets pop_2[:cxpoint+2]$
-            \State \textbf{return} $pop_1$
-        \EndProcedure
-
-    \end{algorithmic}
-\end{algorithm}
-
-\subsection{Experimental Process} 
+\subsection{Experimental Workflow in EvoSwarm} 
 \label{experiment_flow}
@@
 
+We aimed to learn if the proposed solution can be an efficient solution to
+improve the scalability and performance of population-based optimization
+algorithms. Hence, in this section, we verify the following:
+
+\begin{itemize}
+ \item If we can improve the execution time of the algorithm by adding populations
+ and serverless functions to the system. In particular, what is the effect of
+ these changes on the scalability of the system?
+
+\item Having a multi-population enabled platform, with heterogeneous populations and
+the support for mixing search strategies, can increase the performance of the
+search by needing fewer function evaluations than a homogeneous setting?
+
+\end{itemize}
+
+To validate these questions, we used benchmark functions from the Continuous
+Noiseless BBOB testbed, which is part of the Comparing Continuous Optimizers
+(COCO) framework \cite{hansen2016coco}. The testbed includes 24 real-parameter,
+single-objective benchmark functions, and the capability to provide additional
+instances of each function. Each instance of a function has a different optimal
+value. The standard benchmark of the testbed utilizes 15 instances per function
+over 2, 3, 5, 10, 20, and 40 dimensions. The maximum number of Function
+Evaluations (FEs) changes according to the dimension (D), and it is determined
+by the expression $10^5 \cdot D$. As an example, if we have $D = 2$, the
+maximum number of FEs is $200,000$
+
+The COCO framework offers several tools to compare the performance of
+algorithms, generating data sets, tables, and reports for an experiment. There
+is a repository of more than 200 results for the noiseless BOBB testbed,
+collected from BBOB workshops and special sessions between the years 2009 and
+2019 \url{https://coco.gforge.inria.fr/doku.php?id=algorithms-bbob}. The
+EvoSwarm application includes an adapted version of the noiseless BOBB testbed,
+compatible with the scripts of the framework, to compare with other algorithms
+in the repository.
+
+To validate the heterogeneous multi-population setting, we compare between
+homogeneous and an ensemble of multi-populations, using Genetic Algorithms
+(GAs) and Particle Swarm Optimization (PSO). For the GA implementation, we used
+the DEAP library \cite{fortin2012deap} and for the PSO, the EvoloPy library
+\cite{faris2016evolopy}. Both Python libraries are open-source. In EvoSwarm 
+these two algorithms are implemented as stateless functions.
+
+We show the parameters for each algorithm, GA in Table \ref{tab:GAparams} and
+for PSO in Table \ref{tab:PSOparams}. We obtained these parameters in a
+previous work \cite{}. To obtain the parameters, we tested first on the
+Rastrigin separable function with five dimensions. After about fifteen
+experiments, the most challenging targets were achieved for this particular
+function. We tested again with functions one to three, and after obtaining
+favorable results, the PSO and GA algorithm parameters were set.In the GA, we
+randomly set the mutation and crossover probabilities to have more
+heterogeneous workers; in these parameters, the range of values is specified.
+
+\begin{table}
+  \small
+  \caption{ DEAP GA EvoWorker Parameters }
+  \label{tab:GAparams} 
+  \centering
+  \small
+  \begin{tabular}{|l|c|}
+    \hline
+    Selection & Tournament size=12\\ \hline
+    Mutation & Gaussian $\mu=0.0$, $\sigma=0.5$, indbp=0.05  \\ \hline
+    Mutation Probability & [0.1,0.3]  \\ \hline
+    Crossover & Two Point  \\ \hline
+    Crossover Probability& [0.2,0.6]  \\ \hline
+  \end{tabular}
+\end{table}
+
+\begin{table}
+  \small
+  \caption{ EvoloPy PSO Parameters }
+  \label{tab:PSOparams} 
+  \centering
+  \small
+  \begin{tabular}{|l|c|}
+    \hline
+    $V_{max}$ & 6 \\ \hline
+    $W_{max}$ & $0.9$ \\ \hline
+    $W_{min}$ & $0.2$ \\ \hline
+    $C_1$ & 2 \\ \hline
+    $C_2$ & 2 \\ \hline
+  \end{tabular}
+\end{table}
+
+\begin{table}
+    \small
+    \caption{10 Populations experiment parameters
+    }
+    \label{tab:params:workers}
+    \centering
+    \small
+    \begin{tabular}{|l|c|c|c|c|c|c|}
+      \hline
+      Dimension        & 2  & 3  & 5  & 10 & 20  & 40  \\ \hline
+      Generations      & 40 & 25 & 28 & 50 & 66  & 80  \\ \hline
+      Population Size  & 50 & 60 & 60 & 70 & 100 & 125 \\ \hline
+      Populations      & 10 & 10 & 10 & 10 & 10  & 10  \\ \hline
+      Iterations       & 10 & 20 & 30 & 30 & 30  & 40  \\ \hline  
+    \end{tabular}
+\end{table}
+
+\begin{table}
+    \small
+    \caption{5 Populations experiment parameters
+    }
+    \label{tab:params:workers}
+    \centering
+    \small
+    \begin{tabular}{|l|c|c|c|c|c|c|}
+      \hline
+      Dimension        & 2  & 3  & 5  & 10 & 20  & 40  \\ \hline
+      Generations      & 40 & 25 & 28 & 50 & 66  & 80  \\ \hline
+      Population Size  & 100 & 120 & 120 & 140 & 200 & 250 \\ \hline
+      Populations      & 5 & 5 & 5 & 5 & 5  & 5  \\ \hline
+      Iterations       & 10 & 20 & 30 & 30 & 30  & 40  \\ \hline  
+    \end{tabular}
+\end{table}
+
 \section{Experimental Results} 

### 2020-01-27 draft.tex
Starting Experimental Setup

@@
 \usepackage{hyperref}
+\usepackage{listings}
 
@@
 \begin{document}
-
+\lstset{language=Python} 
 \begin{frontmatter}
@@
 algorithms, generating data sets, tables, and reports for an experiment. There
-is a repository of more than 200 results for the noiseless BOBB testbed,
-collected from BBOB workshops and special sessions between the years 2009 and
-2019 \url{https://coco.gforge.inria.fr/doku.php?id=algorithms-bbob}. The
+is a repository\footnote{\url{https://coco.gforge.inria.fr/doku.php?id=algorithms-bbob}} 
+of more than 200 results for the noiseless BOBB testbed, collected from 
+BBOB workshops and special sessions between the years 2009 and2019. The
 EvoSwarm application includes an adapted version of the noiseless BOBB testbed,
@@
 
-To validate the heterogeneous multi-population setting, we compare between
-homogeneous and an ensemble of multi-populations, using Genetic Algorithms
-(GAs) and Particle Swarm Optimization (PSO). For the GA implementation, we used
-the DEAP library \cite{fortin2012deap} and for the PSO, the EvoloPy library
-\cite{faris2016evolopy}. Both Python libraries are open-source. In EvoSwarm 
-these two algorithms are implemented as stateless functions.
+To test the heterogeneous multi-population capabilities,  we compare the
+performance between a homogeneous and an ensemble of multi-populations, using
+Genetic Algorithms (GAs) and Particle Swarm Optimization (PSO). For the GA
+implementation, we used the DEAP library \cite{fortin2012deap} and for the  PSO,
+the EvoloPy library \cite{faris2016evolopy}. Both Python libraries are
+open-source. In EvoSwarm these two algorithms are implemented as stateless
+functions.
 
-We show the parameters for each algorithm, GA in Table \ref{tab:GAparams} and
-for PSO in Table \ref{tab:PSOparams}. We obtained these parameters in a
-previous work \cite{}. To obtain the parameters, we tested first on the
+Next, we show the parameters for each algorithm, in Table \ref{tab:GAparams} 
+for the GA and in Table \ref{tab:PSOparams} for PSO. We obtained these parameters 
+following the same method as in a previous work \cite{garcia2017benchmarking}. 
+To obtain the parameters, we tested first on the
 Rastrigin separable function with five dimensions. After about fifteen
@@
     \hline
-    Selection & Tournament size=12\\ \hline
+    Selection & Tournament size=12                            \\ \hline
     Mutation & Gaussian $\mu=0.0$, $\sigma=0.5$, indbp=0.05   \\ \hline
@@
     Crossover & Two Point                                     \\ \hline
-    Crossover Probability& [0.2,0.6]  \\ \hline
+    Crossover Probability  & [0.2,0.6]                          \\ \hline
   \end{tabular}
@@
 
+% Experiment configuration
+
+\begin{table}
+    \small
+    \caption{ Experiment configuration example 
+    }
+    \label{tab:params}
+    \centering
+    \small
+    \begin{tabular}{|l|l|l|l|l|}
+      \hline
+      \multicolumn{3}{|l|}{Parameter}                   & Type             & Example         \\ \hline
+      \multicolumn{3}{|l|}{Worker Containers (Docker Compose)}        & \texttt{int}     & \texttt{8} \\ \hline
+      \multicolumn{3}{|l|}{GA-PSO Ratio}                 & \texttt{decimal} & \texttt{0.5}    \\  \hline
+      \multicolumn{3}{|l|}{Benchmark Functions}          & \texttt{list}    & \texttt{[1, 2, 3, 4, 5]} \\ \hline
+      \multicolumn{3}{|l|}{Dimensions}                   & \texttt{list}    & \texttt{[10, 20, 40]}        \\ \hline
+      \multicolumn{3}{|l|}{Crossover Probability Range}  & \texttt{list}    & \texttt{[0.2, 0.6]}      \\ \hline
+      \multicolumn{3}{|l|}{Mutation  Probability Rrange} & \texttt{list}    & \texttt{[0.1, 0.3]}      \\ \hline
+      \multicolumn{5}{|c|}{Dimensions}                                                      \\ \hline  
+      Dimension               & Generations & Population Size & Populations  &     Iterations    \\ \hline
+              10              & 50      & 140                 &      5                 & 30                \\ \hline
+              20              & 66      & 200                 &      5                 & 30               \\ \hline
+              40              & 80      & 250                 &      5                 & 40                \\ \hline
+    \end{tabular}
+\end{table}
+
 \begin{table}
@@
     }
-    \label{tab:params:workers}
+    \label{tab:params:10}
     \centering
@@
     }
-    \label{tab:params:workers}
+    \label{tab:params:5}
     \centering

### 2020-01-27 draft.tex
Updates bibliography on containers and evolutionary algorithms
Refs #2

@@
 while another algorithm, more suitable for a local search, finds the global
-optimum. This approach has been followed extensively in recent years, with
+optimum. This approach has been followed extensively in recent years with
 success. % References - JJ
@@
 % not that much to do - JJ
-Hadoop 
-    Comparison \cite{ferrucci2018using}
-    Hadoop\cite{fazenda2012} 
-    FlexGP \cite{FlexGP}, 
-    elephant56 \cite{salza2016elephant56},
+% Hadoop 
+%     Comparison \cite{ferrucci2018using}
+%     Hadoop\cite{fazenda2012} 
+%     FlexGP \cite{FlexGP}, 
+%     elephant56 \cite{salza2016elephant56},
 
-Pool
-    SofEA \cite{merelo2012sofea,sofea:cec2012} , 
-    EvoSpace \cite{garcia2015evospace},
-Volunteer
-    evospace-js \cite{garcia2017evospace}
-Queue
-    Containers \cite{valenzuela2015implementing} \cite{salza2016approach} 
-    Seamless local and Cloud \cite{leclerc2016seamless}
-    Serverless KafKEO \cite{guervos2018introducing}, 
-    Docker Orchestation \cite{dziurzanski2020scalable,zhao2019cloud}
-PSO 
-    PSO AWS \cite{li2015parallel}
+% Pool
+%     SofEA \cite{merelo2012sofea,sofea:cec2012} , 
+%     EvoSpace \cite{garcia2015evospace},
+% Volunteer
+%     evospace-js \cite{garcia2017evospace}
+% Queue
+%     Containers \cite{valenzuela2015implementing} \cite{salza2016approach} 
+%     Seamless local and Cloud \cite{leclerc2016seamless}
+%     Serverless KafKEO \cite{guervos2018introducing}, 
+%     Docker Orchestation \cite{dziurzanski2020scalable,zhao2019cloud}
+% PSO 
+% PSO AWS \cite{li2015parallel}
 

### 2020-01-27 draft.tex
Updates biblio and writes state of the art refs #2

@@
 \label{soa}
+
+Although on first sight containers look like simply a method that
+allows easy shipping of applications, they are much more than
+that. One of the first features that make them interesting is their
+ephemerality: fast startup and teardown time make them usable for
+machine-independent and cloud deployment of simple, ephemeral and
+maybe stateless functions. On the other hand, containerization also
+allows, through the use of orchestration engines like Docker Swarm or
+Kubernetes, or through simple, non-self-scaling specifications using
+Docker Compose, a definition of all the relationships between
+different parts of the application, or services. The full realization
+of Docker-based architectures has led to the creation of a series of
+patterns that, together, conform cloud-native application development
+\cite{kratzke2017understanding}.
+
+However, in the field of evolutionary algorithms, this has been a slow
+realization, from the origins where Salza and Ferrucci
+\cite{salza2016approach,salza2016develop} proposed an architecture
+that carried evolutionary algorithms to the cloud, developed further
+in \cite{de2017parallel} which starts to introduce cloud native
+aspects such as the use of messaging queues and CoreOS as an operating
+system designed from the ground up for container utilization. The way
+these containers are organized, however, is rather classical in
+nature: a master-worker approach, communicated using RabbitMQ, where
+replicated workers perform the tasks in parallel. They called their
+approach AMQPGA, inspired by the protocol, AMQP, used by RabbitMQ, the
+messaging system. Other EAs topologies have also been used in the
+cloud: \cite{dziurzanski2020scalable} adapts an island model to a
+container architecture, in a system that is functionally equivalent to
+classical models, but with the added advantage of being easy to deploy
+and run.
+
+Later approaches use microservices \cite{khalloof2018generic}, or were
+systematized \cite{salza2019speed} achieving speed-up through the
+creation of an abstraction layer over the evolutionary algorithm
+services, and also automatization of the whole workflow. This research
+bumped into some problems, since speedup is limited, mainly due to the
+non-automatic scaling of services and also the presence of a single
+master which is the one that runs the evolutionary
+algorithms. However, this master-worker architecture can be turned
+around with other kind of evolutionary architectures, such
+as pool-based systems
+\cite{valenzuela2015implementing,merelo2012sofea,sofea:cec2012} where
+a better match to the inherently heterogeneous nature of cloud-based
+architectures. Pool-based systems are closer to serverless systems
+\cite{malawski2017serverless}, and they pull from the pool and run
+whole algorithms, instead of just evaluation. The pool from where they
+pull is the common store of results, and can be accessed in an
+asynchronous way, which suits better the nature of cloud systems.
+
 % a lot of this has been done in the previous method... Maybe there's

### 2020-01-27 draft.tex
Some aclarations and references to literature

@@
 state of the subpopulation is lost. But having the algorithm decoupled opens the
-possibility of a serverless-function implementation.
+possibility of implementing the system using serverless functions.
 
@@
 In this work, we implemented all message
-queues in the Redis memory store. Each queue is a Redis List object, and we use
+queues in the Redis memory store. Early cloud architectures
+\cite{de2017parallel} used queues for communication, although in that
+case RabbitMQ was the chosen tool. Redis is faster, and also can act
+as a data store from where we can control the state of the application.
+Each queue is a Redis List object, and we use
 the \texttt{LPOP} (left pop), and \texttt{RPOP} (right pop) commands for a queue like behavior.

### 2020-01-27 draft.tex
Spell check

@@
 
-In the last decades, nature-inspired optimization algorithms have been succesfully
+In the last decades, nature-inspired optimization algorithms have been successfully
 applied to solve complex real-world problems \cite{yang2014nature}. Algorithms
@@
 
-Taking these factors into account, in this work, we present EcoSwarm, 
+Taking these factors into account, in this work, we present EvoSwarm, 
 a multi-container Docker application that reactively processes isolated and
@@
 additional components. Nevertheless, the application is compatible with other
-orchestration and deployment technologies if more scalabilty is needed. 
+orchestration and deployment technologies if more scalability is needed. 
 With additional configuration settings, applications can be deployed 
@@
 creation of an abstraction layer over the evolutionary algorithm
-services, and also automatization of the whole workflow. This research
+services, and also automation of the whole workflow. This research
 bumped into some problems, since speedup is limited, mainly due to the
@@
 Once the configuration is read, we can follow the path of 
-messagess as follows:
+messages as follows:
 
@@
 are generated from this migration, and they are again pushed to the Input queue to continue 
-in a loop. The \texttt{Controller} is also responsible of logging the metadata recived with the
+in a loop. The \texttt{Controller} is also responsible of logging the metadata received with the
 messages.  
@@
 the population-based algorithm. It is common that in a classic island-based algorithm
-each island is executed in a separate processing node, i.e. in a virtual machine, cpu,
+each island is executed in a separate processing node, i.e. in a virtual machine, CPU,
 or thread. In this case we can have just one processing node, or many nodes, each 
@@
 Another advantage is that algorithm designers have many options for implementing this simple architecture. 
-The same basic components can be implmented as a single multi-threaded program, or 
+The same basic components can be implemented as a single multi-threaded program, or 
 as a highly scalable serverless cloud application. Most modern languages include 
@@
 to wait until a message is available, we used the blocking versions \texttt{BLPOP} and
-\texttt{BRPOP}. Redis in-memory operation are very fast, and all of the obove 
+\texttt{BRPOP}. Redis in-memory operation are very fast, and all of the above 
 operations have a time complexity of $\mathcal{O}(1)$. We use the official 
@@
 is a repository\footnote{\url{https://coco.gforge.inria.fr/doku.php?id=algorithms-bbob}} 
-of more than 200 results for the noiseless BOBB testbed, collected from 
+of more than 200 results for the noiseless BBOB testbed, collected from 
 BBOB workshops and special sessions between the years 2009 and2019. The
-EvoSwarm application includes an adapted version of the noiseless BOBB testbed,
+EvoSwarm application includes an adapted version of the noiseless BBOB testbed,
 compatible with the scripts of the framework, to compare with other algorithms
@@
       \multicolumn{3}{|l|}{Crossover Probability Range}  & \texttt{list}    & \texttt{[0.2, 0.6]}      \\ \hline
-      \multicolumn{3}{|l|}{Mutation  Probability Rrange} & \texttt{list}    & \texttt{[0.1, 0.3]}      \\ \hline
+      \multicolumn{3}{|l|}{Mutation  Probability Range} & \texttt{list}    & \texttt{[0.1, 0.3]}      \\ \hline
       \multicolumn{5}{|c|}{Dimensions}                                                      \\ \hline  

### 2020-01-27 draft.tex
Minor corrections

@@
 algorithms that are going to be used by the multi-population algorithm.
-Algorithm Developers can upload the container definition or images directly to
+Algorithm developers can upload the container definition or images directly to
 DockerHub. Another option is to upload the image definition (Dockerfile) to
@@
 configuration files can reference these containers. When a user needs to deploy
-the application, he or she needs to edit the Docker compose file to specify the
-type and number of containers needed for the application.  The Docker compose
+the application, he or she needs to edit the {\tt docker-compose.yml} file to specify the
+type and number of containers needed for the application.  The {\tt docker-compose}
 application is responsible for starting all the containers or downloading new
@@
 open-source. In EvoSwarm these two algorithms are implemented as stateless
-functions.
+functions. % Did you release this also with a free license? Point to
+           % the URL - JJ
 
@@
 function. We tested again with functions one to three, and after obtaining
-favorable results, the PSO and GA algorithm parameters were set.In the GA, we
+favorable results, the PSO and GA algorithm parameters were set. In the GA, we
 randomly set the mutation and crossover probabilities to have more

### 2020-01-27 draft.tex
github url added

@@
 of more than 200 results for the noiseless BBOB testbed, collected from 
-BBOB workshops and special sessions between the years 2009 and2019. The
+BBOB workshops and special sessions between the years 2009 and 2019. The
 EvoSwarm application includes an adapted version of the noiseless BBOB testbed,
@@
 open-source. In EvoSwarm these two algorithms are implemented as stateless
-functions. % Did you release this also with a free license? Point to
+functions. EvoSwarm is also released with a free licence in \url{https://github.com/mariosky/EvoSwarm}.
+
+% Did you release this also with a free license? Point to
            % the URL - JJ

### 2020-01-28 draft.tex
Update draft.tex
Experimental config

@@
 configuration files can reference these containers. When a user needs to deploy
-the application, he or she needs to edit the {\tt docker-compose.yml} file to specify the
+the application, he or she needs to edit the Fi file to specify the
 type and number of containers needed for the application.  The {\tt docker-compose}
@@
 
-Next, we show the parameters for each algorithm, in Table \ref{tab:GAparams} 
-for the GA and in Table \ref{tab:PSOparams} for PSO. We obtained these parameters 
+Next, we show the parameters for each algorithm, in Table \ref{tab:GAparams} for
+the GA and in Table \ref{tab:PSOparams} for PSO. We obtained these parameters
 following the same method as in a previous work \cite{garcia2017benchmarking}.
-To obtain the parameters, we tested first on the
-Rastrigin separable function with five dimensions. After about fifteen
-experiments, the most challenging targets were achieved for this particular
-function. We tested again with functions one to three, and after obtaining
-favorable results, the PSO and GA algorithm parameters were set. In the GA, we
-randomly set the mutation and crossover probabilities to have more
-heterogeneous workers; in these parameters, the range of values is specified.
+To obtain the parameters, we tested first on the Rastrigin separable function
+with five dimensions. After about fifteen experiments, the most challenging
+targets were achieved for this particular function. We tested again with
+functions one to three, and after obtaining favorable results, the PSO and GA
+algorithm parameters were set. In the GA, we randomly set the mutation and
+crossover probabilities to have more heterogeneous workers; in these parameters,
+the range of values is specified. We did not change these parameters during the
+experiments, and only the population size and number of generations were
+provided as parameters.
 
@@
 
-% Experiment configuration
+% Experiment conf
+
+The EvoSwarm application uses configuration files to run noiseless BOBB testbed
+experiments. As a first step, the docker application must be deployed, using a
+{\tt docker-compose.yml} file, where the number of worker containers is
+specified. Then a JSON file containing the configuration parameters of the
+experiment has to be provided. Table \ref{tab:params} shows an example of these
+parameters. The {\em GA-PSO Ratio} parameter indicates the proportion of
+populations that will use the GA algorithm for the search. With a value of
+$0.50$ there will be about the same proportion of GA and PSO populations. If we
+specify a value of $0$ this will give us a PSO only algorithm, and finally, a
+value of $1$ means that all populations will run the GA algorithm. Next, we
+specify a list indicating which of the 24 benchmark functions will be tested.
+In the example, the experiment will use the first five functions. The
+{\em Instances} parameter, indicates how many instances of each function will be
+executed, this parameter has 15 as the default value. This value is used
+because is the standard for the BBOB benchmark \cite{hansen2016coco}. Then, in
+the {\em Dimensions} parameter we define a list of the dimensions that will be
+tested, we must select additional parameters for each dimension. According to
+the maximum number of Function Evaluations (FEs) we mentioned earlier, we
+define for each dimension: the number of populations that will be generated in
+the setup, and for each population, the number of generations, and population
+size. Finally, the {\em Iterations} parameter indicates how many complete loops
+will be performed. All these parameters give us the maximum number of FEs that
+will be performed. For instance, for $D = 2$, the maximum number of FEs is
+$200,000$ which is the same as $40*50*10*10$.
 
@@
       \multicolumn{3}{|l|}{Benchmark Functions}          & \texttt{list}    & \texttt{[1, 2, 3, 4, 5]} \\ \hline
+      \multicolumn{3}{|l|}{Instances}                    & \texttt{integer}    & \texttt{15} \\ \hline
       \multicolumn{3}{|l|}{Dimensions}                   & \texttt{list}    & \texttt{[10, 20, 40]}        \\ \hline

### 2020-01-28 draft.tex
Minor changes and reflow

@@
 \address[itt]{Department of Graduate Studies, Instituto Tecnol\'ogico de Tijuana, Tijuana BC, Mexico}
-\address[granada]{Universidad de Granada, Granada, Spain}
+\address[granada]{Department of Computer Architecture and Technology, Universidad de Granada, Granada, Spain}
 
 \begin{abstract} 
-Splitting a population into multiple instances is a technique that has been used extensively in
-recent years to improve the performance of nature-inspired optimization
-algorithms. Work on several populations can be done in parallel, and
-also asynchronously, which can be leveraged to create scalable
-implementations based on, among other methods, distributed, multi-threaded, parallel, and cloud-based computing. 
-% I have rephrased all this. Please check it out - JJ
-The design of these cloud-native, distributed, multi-population
-algorithms is not a trivial task. Coming from monolithic
-(single-instance) solutions, adaptations at several levels, from the
-algorithmic to the functional, must be made to leverage
-the scalability, elasticity, fault-tolerance, reproducibility, and
-cost-effectiveness of cloud systems while, at the same time,
-conserving the intended functionality. Instead of an evolutive
-approach, in this paper we propose to create from scratch a
+Splitting a population into multiple instances is a technique that has
+been used extensively in recent years in order to help improve the
+performance of nature-inspired optimization algorithms. Work on
+several populations can be done in parallel, and also asynchronously,
+a fact that can be leveraged to create scalable implementations based
+on, among other methods, distributed, multi-threaded, parallel, and
+cloud-based computing.  However, the design of these cloud-native,
+distributed, multi-population algorithms is not a trivial task. Coming
+from monolithic (single-instance) solutions, adaptations at several
+levels, from the algorithmic to the functional, must be made to
+leverage the scalability, elasticity, fault-tolerance,
+reproducibility, and cost-effectiveness of cloud systems while, at the
+same time, conserving the intended functionality. Instead of an
+evolutive approach, in this paper we propose to create from scratch a
 cloud-native optimization framework that is able to include multiple
@@
 performance and scalability of this solution, together with the effect
-other design parameters had on the 
-performance, in particular, the number and the size of populations
-with respect to problem size. The architecture and the implemented platform is an excellent
-alternative for running locally or in the cloud, thus proving that cloud-native
-bioinspired algorithms perform better in their "natural" environment than other
+other design parameters had on it; in particular, the number and the
+size of populations with respect to problem size. The architecture and
+the implemented platform is an excellent alternative for running
+locally or in the cloud, thus proving that cloud-native bioinspired
+algorithms perform better in their "natural" environment than other
 kinds of algorithms, and setting a new baseline for scaling and
 performance in the cloud.
-% Still a big longish. We need another iteration to reduce it - JJ
 \end{abstract}

### 2020-01-28 draft.tex
Some fixes and questions in experimental workflow

@@
 built on a single stack of services that communicate by layers, to microservices
-\cite{microservices} and serverless architectures \cite{varghese2018next,
-Varghese2018849} that favor the parallel and asynchronous communication of
+\cite{microservices} and serverless architectures \cite{varghese2018next,Varghese2018849} that favor the parallel and asynchronous communication of
 heterogeneous resources; in the process, a new process for designing {\em cloud
@@
 Algorithm developers can upload the container definition or images directly to
-DockerHub. Another option is to upload the image definition (Dockerfile) to
-GitHub and define a trigger to re-upload the image to DockerHub.  Docker Compose
+DockerHub; another option is to upload the container image definition (Dockerfile) to
+GitHub and define a trigger to re-upload the image to DockerHub; this
+can be defined in DockerHub (or any other registry) too.  Docker Compose
 configuration files can reference these containers. When a user needs to deploy
-the application, he or she needs to edit the Fi file to specify the
+the application, he or she needs to edit the Fi file
+% Fi file? - JJ+
+to specify the
 type and number of containers needed for the application.  The {\tt docker-compose}
-application is responsible for starting all the containers or downloading new
+application is responsible for starting all the containers, downloading new
 versions if it is necessary. After starting the application, users can push
 Experiment Definition files and start monitoring the execution in the terminal.
-After the experiment is finished, the user must execute another script using the
-experiment id to generate a collection of files containing all the data
-generated by the experiment. The script generates these files in a JSON format.
+After the experiment is finished, the user must execute another script
+that takes as argument the
+experiment id to generate a collection of files containing all the
+data generated by the experiment in a JSON format.
 Researchers can use these files to analyze and plot experiment data.  Python
 scripts are included in the repository to plot the running times for the
-experiments, and all other plots used in this paper.  There is also a script to
+experiments, and all other plots used in this paper;  there is also a script to
 generate the files needed by the COCO framework \cite{hansen2016coco}, which can generate standard

### 2020-01-28 draft.tex
Some more small changes

@@
 
-We aimed to learn if the proposed solution can be an efficient solution to
+We aimed to learn if the proposed solution can efficiently
 improve the scalability and performance of population-based optimization
@@
 \begin{itemize}
- \item If we can improve the execution time of the algorithm by adding populations
+ \item Whether we can improve the execution time of the algorithm by adding populations
  and serverless functions to the system. In particular, what is the effect of
@@
 
-\item Having a multi-population enabled platform, with heterogeneous populations and
-the support for mixing search strategies, can increase the performance of the
+\item Can having a multi-population enabled platform, with heterogeneous populations and
+the support for mixing search strategies,  increase the performance of the
 search by needing fewer function evaluations than a homogeneous setting?
@@
 
-% Did you release this also with a free license? Point to
-           % the URL - JJ
-
-Next, we show the parameters for each algorithm, in Table \ref{tab:GAparams} for
-the GA and in Table \ref{tab:PSOparams} for PSO. We obtained these parameters
+% Maybe this should be mentioned earlier, in the workflow? - JJ
+
+Next, we show the parameters for each algorithm, with Table \ref{tab:GAparams} for
+the GA and Table \ref{tab:PSOparams} for PSO. We obtained these parameters
 following the same method as in a previous work \cite{garcia2017benchmarking}.

### 2020-01-28 draft.tex
Update draft.tex

@@
 The EvoSwarm application uses configuration files to run noiseless BOBB testbed
-experiments. As a first step, the docker application must be deployed, using a
+experiments.  As a first step, we must deploy the docker application using a
 {\tt docker-compose.yml} file, where the number of worker containers is
@@
 experiment has to be provided.  Table \ref{tab:params} shows an example of these
-parameters. The {\em GA-PSO Ratio} parameter indicates the proportion of
-populations that will use the GA algorithm for the search. With a value of
+parameters. The \em{GA-PSO Ratio} parameter indicates the proportion of
+populations that will use the GA algorithm.  In the example, with a value of
 $0.50$ there will be about the same proportion of GA and PSO populations. If we
-specify a value of $0$ this will give us a PSO only algorithm, and finally, a
-value of $1$ means that all populations will run the GA algorithm. Next, we
-specify a list indicating which of the 24 benchmark functions will be tested.
-In the example, the experiment will use the first five functions. The
-{\em Instances} parameter, indicates how many instances of each function will be
-executed, this parameter has 15 as the default value. This value is used
-because is the standard for the BBOB benchmark \cite{hansen2016coco}. Then, in
-the {\em Dimensions} parameter we define a list of the dimensions that will be
-tested, we must select additional parameters for each dimension. According to
-the maximum number of Function Evaluations (FEs) we mentioned earlier, we
-define for each dimension: the number of populations that will be generated in
-the setup, and for each population, the number of generations, and population
-size. Finally, the {\em Iterations} parameter indicates how many complete loops
-will be performed. All these parameters give us the maximum number of FEs that
-will be performed. For instance, for $D = 2$, the maximum number of FEs is
-$200,000$ which is the same as $40*50*10*10$.
+specify a value of $0$, this will give us an algorithm with only PSO
+populations, and finally, a value of  $1$ means that all populations will run
+the GA algorithm. Next, we specify a list indicating which of the 24 benchmark
+functions will be tested. In the example, the experiment will use the first five
+functions.  The \em{Instances} parameter indicates how many instances of each
+function will be tested. \em{Instances}  has a defualt value of 15. We used this
+value because it is the standard for the BBOB benchmark \cite{hansen2016coco}.
+In the \em{Dimensions} parameter, we define a list of the dimensions that will
+be tested, and we must select additional parameters for each dimension.
+According to the maximum number of  Function Evaluations (FEs) we mentioned
+earlier, we define for each dimension: the number of populations that will be
+generated in the setup, and for each population,  the number of generations, and
+population size. Finally, the \em{Iterations} parameter indicates how many
+complete loops will be performed. All these parameters give us the maximum
+number of FEs that will be performed. For instance, for $D = 2$, the maximum
+number of FEs is $200,000$, which is the same as $40*50*10*10$.
 

### 2020-01-28 draft.tex
First Experiment Intro

@@
 configuration files can reference these containers. When a user needs to deploy
-the application, he or she needs to edit the Fi file
+the application, he or she needs to edit the {\tt docker-compose.yml} file
 % Fi file? - JJ+
-to specify the
-type and number of containers needed for the application.  The {\tt docker-compose}
+% fixed - Mario
+to specify the type and number of containers needed for the application.  The {\tt docker-compose}
 application is responsible for starting all the containers, downloading new
@@
  \item Whether we can improve the execution time of the algorithm by adding populations
- and serverless functions to the system. In particular, what is the effect of
+ and serverless functions to the system, in particular, what is the effect of
  these changes on the scalability of the system?
@@
 % Experiment conf
-
 The EvoSwarm application uses configuration files to run noiseless BOBB testbed
@@
 
+As we mentioned before, reactive systems can scale by adding additional copies
+of serverless functions. In our case, we can start additional worker containers
+to have the same effect. Regarding GAs using a similar architecture, the work of
+Salza et al., suggest that worker nodes, need to have at least a certain level
+of complexity, in terms of execution time,  to effectively scale on multiple
+nodes, tending to linear scalability. In a population-based algorithm like
+EvoSwarm, several parameters can increase or decrease the workload of workers.:
+
+\begin{itemize}
+    \item The number of populations: if there are more workers than populations, workers
+    must wait for work to arrive. If there are too many populations, they could be
+    standing in the queue for more time. 
+    \item The size of populations and the number of generations: 
+    These parameters naturally increase the execution time.
+    \item The complexity of the algorithm: For instance, in our case, 
+    the PSO implementation has a lower execution time than the GA.
+\end{itemize}
+
+To evaluate the effect of these parameters on the scalability of the system, we
+propose the following experiment:
+
 \begin{table}

### 2020-01-28 draft.tex
Exp 1 description intro

@@
 
-% Experiment conf
 The EvoSwarm application uses configuration files to run noiseless BOBB testbed
@@
 
+\subsection{Experiment 1}
+\label{sec:exp1}
+% Experiment 1
+
 As we mentioned before, reactive systems can scale by adding additional copies
@@
 propose the following experiment:
+
+For the experiment, we selected $f_4$ (Skew Rastrigin-Bueche separable) from the
+BBOB testbed. We selected this function because it is computationally demanding,
+and in higher dimensions has been difficult for PSO \cite{el2009black} and GAs
+\cite{nicolau2009application} to solve with the required FEs. As we are
+comparing just in terms of execution time, the results from a single function
+can be better understand, because there are fewer factors involved.
 

### 2020-01-28 draft.tex
Update draft.tex

@@
 experiment has to be provided.  Table \ref{tab:params} shows an example of these
-parameters. The \em{GA-PSO Ratio} parameter indicates the proportion of
+parameters. The { \em GA-PSO Ratio} parameter indicates the proportion of
 populations that will use the GA algorithm.  In the example, with a value of
@@
 functions will be tested. In the example, the experiment will use the first five
-functions.  The \em{Instances} parameter indicates how many instances of each
-function will be tested. \em{Instances}  has a defualt value of 15. We used this
+functions.  The {\em Instances} parameter indicates how many instances of each
+function will be tested. {\em Instances}  has a defualt value of 15. We used this
 value because it is the standard for the BBOB benchmark \cite{hansen2016coco}.
-In the \em{Dimensions} parameter, we define a list of the dimensions that will
+In the {\em Dimensions} parameter, we define a list of the dimensions that will
 be tested, and we must select additional parameters for each dimension.
@@
 generated in the setup, and for each population,  the number of generations, and
-population size. Finally, the \em{Iterations} parameter indicates how many
+population size. Finally, the {\em Iterations} parameter indicates how many
 complete loops will be performed. All these parameters give us the maximum
@@
 
+We executed two sets of experiments with five and ten initial populations,
+repeating each experiment using  1, 2, 4, and 8 workers.  We kept the same
+maximum number of FEs, changing the relevant parameters for this. 
+See Tables \ref{tab:params:10} and \ref{tab:params:5} 
+for the complete list.
+
 \begin{table}

### 2020-01-28 draft.tex
Experiment 2

@@
 
-\item Can having a multi-population enabled platform, with heterogeneous populations and
-the support for mixing search strategies,  increase the performance of the
+\item Having a multi-population enabled platform, with heterogeneous populations and
+the support for mixing search strategies,  increases the performance of the
 search by needing fewer function evaluations than a homogeneous setting?
@@
 
+\subsection{Experiment 2}
+\label{sec:exp2}
+% Experiment 1
+
+Having a multi-population based algorithm can decrease the execution time
+because of the parallel execution of the evolution. But, having heterogeneous
+populations increases the performance of the search by needing fewer function
+evaluations than a homogeneous setting.?
+
+In this experiment, we compare the performance of both heterogeneous and
+homogeneous settings. We compare a multi-population with only  GA and PSO
+populations, versus an ensemble of GA and PSO algorithms on the first five
+functions of the BOBB testbed.  We used ten populations and 8 workers for the
+experiment and the same parameters as before.
+
 \section{Experimental Results} 

### 2020-01-28 draft.tex
Some comments to the first experiment

@@
 
-The EvoSwarm application uses configuration files to run noiseless BOBB testbed
+The EvoSwarm application uses configuration files to run the noiseless BOBB testbed
 experiments.  As a first step, we must deploy the docker application using a
@@
 
+% What is the main objective of this experiment? Measure scalability?
+% - JJ
 As we mentioned before, reactive systems can scale by adding additional copies
 of serverless functions. In our case, we can start additional worker containers
-to have the same effect. Regarding GAs using a similar architecture, the work of
-Salza et al., suggest that worker nodes, need to have at least a certain level
+to have the same effect. Other authors, like Salza et al., % add
+                                % reference - JJ
+have used a
+similar architecture, which implies that worker nodes need to have at least a certain level
 of complexity, in terms of execution time,  to effectively scale on multiple
 nodes, tending to linear scalability. In a population-based algorithm like
-EvoSwarm, several parameters can increase or decrease the workload of workers.:
+EvoSwarm, several parameters can increase or decrease the workload of workers:
 
@@
 To evaluate the effect of these parameters on the scalability of the system, we
-propose the following experiment:
-
-For the experiment, we selected $f_4$ (Skew Rastrigin-Bueche separable) from the
-BBOB testbed. We selected this function because it is computationally demanding,
+propose  an experiment in which we selected $f_4$ (Skew Rastrigin-Bueche separable) from the
+BBOB testbed. This function because it is computationally demanding,
 and in higher dimensions has been difficult for PSO \cite{el2009black} and GAs
-\cite{nicolau2009application} to solve with the required FEs. As we are
-comparing just in terms of execution time, the results from a single function
-can be better understand, because there are fewer factors involved.
+\cite{nicolau2009application} to solve with the required FEs. As we
+are only comparing in terms of execution time, the results from a single function
+can be better understood, since there are fewer factors involved.
 

### 2020-01-28 draft.tex
Revision experiment 2

@@
 because of the parallel execution of the evolution. But, having heterogeneous
-populations increases the performance of the search by needing fewer function
-evaluations than a homogeneous setting.?
-
-In this experiment, we compare the performance of both heterogeneous and
+populations might enhance evolutionary search and need less
+evaluations than homogeneous systems; heterogeneous settings, if
+done right, will increase the diversity of the whole population % cita
+                                % multikulti o similar - JJ
+
+% Explain that heterogeneity comes from using different algorithms,
+% not from using different types of nodes - JJ
+But this is a rule of thumb, and it will depend on the degree of
+heterogeneity, as well as on the algorithm itself. In this experiment, we compare the performance of both heterogeneous and
 homogeneous settings. We compare a multi-population with only  GA and PSO

### 2020-01-28 draft.tex
Write BBOB correctly

@@
 populations, versus an ensemble of GA and PSO algorithms on the first five
-functions of the BOBB testbed.  We used ten populations and 8 workers for the
+functions of the BBOB testbed.  We used ten populations and 8 workers for the
 experiment and the same parameters as before.

### 2020-01-28 draft.tex
Resolving comments

@@
 
+%
+
 \subsection{Experiment 2}
@@
 because of the parallel execution of the evolution. But, having heterogeneous 
-populations might enhance evolutionary search and need less
+populations might enhance evolutionary search and needless
 evaluations than homogeneous systems; heterogeneous settings, if
-done right, will increase the diversity of the whole population % cita
+done right, increases the diversity of the whole population \cite{araujo2008multikulti}. % cita
                                 % multikulti o similar - JJ
@@
 % not from using different types of nodes - JJ 
-But this is a rule of thumb, and it will depend on the degree of
-heterogeneity, as well as on the algorithm itself. In this experiment, we compare the performance of both heterogeneous and
-homogeneous settings. We compare a multi-population with only  GA and PSO
-populations, versus an ensemble of GA and PSO algorithms on the first five
-functions of the BBOB testbed.  We used ten populations and 8 workers for the
-experiment and the same parameters as before.
+% done - Mario 
+But this is a rule of thumb, and it will depend on the degree of heterogeneity,
+as well as on the algorithm itself. Some level of heterogeneity can be
+implemented by just changing the configuration parameters of each population,
+but in this case, we are interested in a heterogeneous search strategy. 
+Therefore, in this experiment we compare a multi-population with only GA and PSO populations,
+versus an ensemble of GA and PSO algorithms. We tested on the first five functions of the
+BOBB testbed. We used ten populations and 8 workers for the experiment and the
+same parameters as before.
+
+\subsection{Hardware}
+\label{sec:hardware}
 
@@
 \label{results}
+
+We present the results of the experiments in this section. The effect of a
+different number of populations on execution time is reported in  Section
+\ref{sec:exp1results}. The results of the performance comparison between homogeneous and
+heterogeneous search strategies are presented in   Section \ref{sec:exp2results}.
+
+\subsection{Experiment 1}
+\label{sec:exp1results}
 
@@
 
+\subsection{Experiment 2}
+\label{sec:exp2results}
+
 \section{Conclusions} 

### 2020-01-28 draft.tex
Some corrections here and there

@@
     \includegraphics[width=\textwidth]{pool_island}
-    \caption{Population State and dataflow between processing nodes of a Pool-based multi-population algorithm.}
+    \caption{Population state and flow of data between processing nodes of a Pool-based multi-population algorithm.}
     \label{fig:pool_island}
@@
 evaluations than homogeneous systems; heterogeneous settings, if
-done right, increases the diversity of the whole population \cite{araujo2008multikulti}. % cita
-                                % multikulti o similar - JJ
-
-% Explain that heterogeneity comes from using different algorithms,
-% not from using different types of nodes - JJ 
-% done - Mario 
+done right, increases the diversity of the whole population
+\cite{araujo2008multikulti}.
+
 But this is a rule of thumb, and it will depend on the degree of heterogeneity,
@@
 versus an ensemble of GA and PSO algorithms. We tested on the first five functions of the
-BOBB testbed. We used ten populations and 8 workers for the experiment and the
+BOBB testbed. We used ten populations and eight workers for the experiment and the
 same parameters as before.
 
+% Shouldn't this be a subsection of experimental setup? - JJ
 \subsection{Hardware}
@@
 
-\begin{figure}[ht]
+\begin{figure}[h!tb]
+  % Is the axis correct? Seconds per instance or total time? Please
+  % change that - JJ
     \centering
     \includegraphics[width=\textwidth]{sp_size_worker}
-    \caption{ Time required to complete experiments with different number of workers and populations sizes }
+    \caption{Time required to complete experiments with different number of workers and populations sizes }
     \label{fig:spworker}

### 2020-01-29 draft.tex
BBOB again
Update draft.tex

@@
 versus an ensemble of GA and PSO algorithms. We tested on the first five functions of the
-BOBB testbed. We used ten populations and eight workers for the experiment and the
+BBOB testbed. We used ten populations and eight workers for the experiment and the
 same parameters as before.
@@
 
+The
+
 \begin{figure}[h!tb]
@@
   % change that - JJ
+  % Yes, they are instances - Mario
     \centering
     \includegraphics[width=\textwidth]{sp_size_worker}
-    \caption{Time required to complete experiments with different number of workers and populations sizes }
+    \caption{Time required, in seconds, to complete an instance of the function, with different number of workers and populations sizes }
     \label{fig:spworker}

### 2020-01-29 draft.tex
Some small corrections and adds acks

@@
 be tested, and we must select additional parameters for each dimension.
-According to the maximum number of  Function Evaluations (FEs) we mentioned
+According to the maximum number of function evaluations (FEs) we mentioned
 earlier, we define for each dimension: the number of populations that will be
@@
 propose  an experiment in which we selected $f_4$ (Skew Rastrigin-Bueche separable) from the
-BBOB testbed. This function because it is computationally demanding,
+BBOB testbed. This function has been used because it is computationally demanding,
 and in higher dimensions has been difficult for PSO \cite{el2009black} and GAs
@@
     \small
-    \caption{10 Populations experiment parameters
+    \caption{Parameters used in the experiments with ten populations
     }
     \label{tab:params:10}
+    \vspace{0.25cm}
     \centering
@@
     \small
-    \caption{5 Populations experiment parameters
+    \caption{Parameters used in the experiments that used five populations
     }
     \label{tab:params:5}
+    \vspace{0.25cm}
     \centering
@@
 
+\section{Acknowledgements}
+
+This paper has been supported in part by projects DeepBio
+(TIN2017-85727-C4-2-P).
+
 \bibliographystyle{elsarticle-num}

### 2020-01-29 draft.tex
Exp1 results

--- 2020-01-29 draft.tex
@@
 
-\subsection{Experiment 1}
-\label{sec:exp1results}
-
-The
-
-\begin{figure}[h!tb]
-  % Is the axis correct? Seconds per instance or total time? Please
-  % change that - JJ
-  % Yes, they are instances - Mario
+\begin{figure}[ht]
     \centering
@@
 
+\subsection{Experiment 1}
+\label{sec:exp1results}
+
+We summarize the results of the experiment in Figure \ref{fig:spworker}.  The
+experiment consisted of executing the same benchmark on EvoSwarm twice but
+changing the number of populations used in each run. The left column has the
+results when using five populations, and the right side shows the results for
+ten.  Graphics show the execution time achieved by a certain number of workers.
+Each box-plot indicates the time required for the execution of 15  function
+instances, in seconds.  Each row has a different scale on the y-axis, to
+compensate for the dimension increments. It is important to notice that the
+execution time can also be affected by the algorithm finding a solution with an
+error smaller than the threshold, before completing all the FEs.
+
+As expected, in most cases, increasing the number of workers reduces the time
+required for the execution of an instance. Mainly when using the same parameters
+(number of populations and number of workers ). But, when we compare the two
+options, we find some subtle differences:
+
+When using one or two workers, using 5 populations is better in most cases. A
+reason for this could be that the migration of individuals takes more time to
+reach the worker. For instance, when a single worker has 10 populations, only
+after processing all, it will start receiving mixed populations.  Although, in
+all cases, and even a single worker, populations are heterogeneous, this
+advantage takes more time to propagate. With four workers, almost the same
+results can be achieved regardless of our selection between five or ten
+populations. Also, there is no much difference between 4 and 8 workers, when
+using five populations. In general, we can observe that keeping a ratio close to
+1:1 between the number of workers and populations exploits better the computing
+resources available.
+
 \subsection{Experiment 2}

### 2020-01-29 draft.tex
subpop to pop

@@
 multi-population based methods when generally referring to techniques using
-subpopulations as part of their optimization strategy. % Some comment about asynchrony would be in order - JJ
+many populations as part of their optimization strategy. % Some comment about asynchrony would be in order - JJ
 % Comments below - Mario
@@
 When designing efficient multi-population algorithms, researchers need to
-consider additional issues \cite{Ma2019}. These include the number and size of subpopulations,
-the interaction between them, the search area of subpopulations, and the search
-strategy and parametrization of each subpopulation. We must take into account
+consider additional issues \cite{Ma2019}. These include the number and size of populations,
+the interaction between them, the search area of each population, and the search
+strategy and parametrization of each population. We must take into account
 these high-level issues when designing a parallel architecture. Researchers 
@@
     to a continuous stream of data. One of the options is to consider
-    sub-populations as messages to be modified asynchronously by functions.
+    populations as messages to be modified asynchronously by functions.
 
@@
 a multi-container Docker application that reactively processes isolated and
-heterogeneous subpopulations. The platform uses the Docker Compose tool for
+heterogeneous populations. The platform uses the Docker Compose tool for
 defining and running experiments as multi-container applications. 
@@
 Docker Container images host the stateless functions responsible for running 
-a search strategy on subpopulations, these functions could also be ported
+a search strategy on several populations, these functions could also be ported
 directly to a Function as a Service (FaaS) \cite{Roberts2016} implementation.
@@
 Queue implementation and logging and an Experiment Controller implemented in
-Python. Developers of sub-population processing functions have the freedom to
+Python. Developers of population processing functions have the freedom to
 implement new population-based algorithms or operators in the language they
@@
 Multi-population based methods divide the original population into
-smaller subpopulations or islands, with every subpopulation carrying out the
+smaller populations or islands, with every population carrying out the
 algorithm independently, with synchronous or asynchronous communication with the
@@
 This relative isolation helps in maintaining an overall
-diversity since each subpopulation will search in a particular area, at least
+diversity since each population will search in a particular area, at least
 between communications. The recombination mentioned above (mixing) or migration
-between subpopulations is needed to avoid a premature convergence of candidate
+between populations is needed to avoid a premature convergence of candidate
 solutions since smaller populations are known to perform better for a given
@@
 
-However, in most cases, algorithms applied to each subpopulation are
+However, in most cases, algorithms applied to each population are
 homogeneous, or at any rate, the same variant of the algorithm. As long as this
@@
 tip the balance between exploration and exploitation of the search space. On the
-other hand, current studies show that by having a high number of subpopulations
+other hand, current studies show that by having a high number of populations
 interacting in parallel, the effect of the individual parameters of each
-subpopulation is compensated by those selected in other
-subpopulations. % reference - JJ
-In this
-work, we will use random settings within a specific range as results have shown
+population is compensated by those selected in other
+populations . % reference - JJ
+
+In this work, we will use random settings within a specific range as results have shown
 this is a valid solution to this problem. % This should have been
@@
 kept fixed in order to control more easily the execution of the algorithm. For
-instance, by having the size of subpopulations fixed, it is easier to control
+instance, by having the size of populations fixed, it is easier to control
 the number of evaluations and the communication costs, when the algorithm is in
@@
 
-The most elemental data structure used throughout the system is the subpopulation; and at a higher
-level, we can view the system as a continuous stream of subpopulations
-flowing
-% We talk about "multi-population" elsewhere, including the
-% abstract. We should unify terminology - JJ
-% "multi-population" is the whole thing, a subpopulation is a member of that.
-% "A multi-population has many subpopulations or should we call them just populations?"
-% I will add a comment about this in the Multi-population section -
-% Mario
-% use the usual "multi-deme" terminology. A population has different
-% sub-populations nad uses a multi-deme implementation - JJ
-% Let's talk about it, I am using the Multi-population terminology 
-% maybe multi-dem is more common in the EA community but not 
-% much in the Swarm Intelligence side?  
-% Any how, we decide and explain the difference so readers of both 
-% camps are Ok with it. 
-% I prefer the use of population and something like the global population - Mario
-
-from one process to the other. By "continuous stream," we mean that ideally,
+The most elemental data structure used throughout the system is the population; and at a higher
+level, we can view the system as a continuous stream of populations
+flowing from one process to the other. By "continuous stream," we mean that ideally,
 everything must happen asynchronously without components needing to wait for
@@
 implemented by using a message queue system. For instance, if one process needs
-to communicate a subpopulation to another, it must pack the sub-population into
+to communicate a population to another, it must pack the population into
 a message, and then send the message 
-% I don't know if
-                                % asynchronously applies here. It's
-                                % simply sent - JJ 
-                                % Meaning it does not wait for a
-                                % respone, it continues doing other
-                                % stuff - Mario
-% That's rather "fire and forget". Or simply say it's sent without
-% waiting for a response - JJ
-% corrected - Mario
 to the queuing system without waiting for a response; this
@@
 swimlanes: First, the \texttt{Setup} process, responsible for the reading a configuration
-file, and creating the initial subpopulations. Second, the \texttt{Controller} process,
-responsible for the migration between subpopulations, and keeping track of the
+file, and creating the initial populations. Second, the \texttt{Controller} process,
+responsible for the migration between populations, and keeping track of the
 iterations of the algorithm. The \texttt{Message Queue} process runs the \texttt{Input} and \texttt{Output}
@@
 includes all the parameters needed to execute the algorithm, among other things,
-the number of subpopulations, the number of individuals, and the number of iterations
+the number of populations, the number of individuals, and the number of iterations
 of the algorithm. We give more details about the configuration data structure
@@
 parameters found in the configuration structure. The population at this moment is just
-static data, including each individual inside. Each subpopulation includes a metadata
+static data, including each individual inside. Each population includes a metadata
 section where its algorithm and execution parameters are specified. For instance, 
@@
 
-\item Each subpopulation is then pushed to the \texttt{Input} queue, so they can be consumed 
+\item Each population is then pushed to the \texttt{Input} queue, so they can be consumed 
 by Stateless functions responsible for the execution of the search.  % Why the caps in Stateless? - JJ  
@@
 
-\item One or more \texttt{Stateless functions} are constantly pulling subpopulation messages
+\item One or more \texttt{Stateless functions} are constantly pulling population messages
 from the \texttt{Input} queue. Taking the population data and metadata as parameters, these 
@@
 The resulting populations are now pushed to the Output queue. Once finished, another 
-subpopulation is pulled from the queue.
+population is pulled from the queue.
 
@@
 \item Otherwise, it passes the stream of messages 
-to a migration process, where subpopulations are mixed with each other. New populations 
+to a migration process, where populations are mixed with each other. New populations 
 are generated from this migration, and they are again pushed to the Input queue to continue 
@@
 comparison on the coupling and communication between the main components:
-subpopulations, processing nodes, and algorithms.  In Figure \ref{fig:population_message} we show the
+populations, processing nodes, and algorithms.  In Figure \ref{fig:population_message} we show the
 main components of a message-driven architecture. We can see that the algorithm
-and subpopulations are separated, and the \texttt{Population queue} keeps the state
-of subpopulations. While algorithms only need the populations as parameters
+and populations are separated, and the \texttt{Population queue} keeps the state
+of populations. While algorithms only need the populations as parameters
 for their execution.
@@
 In contrast, in the classic Island model shown in Figure
-\ref{fig:classicisland}, algorithms are coupled with a subpopulation, and the same processing
-node keeps both the algorithm process and the state of the subpopulation.  Other
+\ref{fig:classicisland}, algorithms are coupled with a population, and the same processing
+node keeps both the algorithm process and the state of the population.  Other
 processing nodes follow the same configuration and form of execution, and they
@@
 nodes.  In Figure \ref{fig:pool_island} we show the main components of a pool-based
-multi-population algorithm. Although algorithms and subpopulation state remains
+multi-population algorithm. Although algorithms and population state remains
 coupled, now, processing nodes do not communicate directly with each other.
@@
 external processing nodes, the system needs additional communication and
-processing for keeping track of each subpopulation.
+processing for keeping track of each population.
 
@@
 isolated algorithms take random samples of the global population and use this
-temporal subpopulation as parameters. An advantage of this design is that the
-sampling provides a type of migration between isolated subpopulations. A problem
-found is that when a processing node returns a subpopulation to the pool, the
-state of the subpopulation is lost. But having the algorithm decoupled opens the
+temporal population as parameters. An advantage of this design is that the
+sampling provides a type of migration between isolated populations. A problem
+found is that when a processing node returns a population to the pool, the
+state of the population is lost. But having the algorithm decoupled opens the
 possibility of implementing the system using serverless functions.

### 2020-01-29 draft.tex
biblio added

@@
 algorithm independently, with synchronous or asynchronous communication with the
-rest of the islands. % exemplify how this is done in different
+rest of the islands \cite{Ma2019}.% exemplify how this is done in different
                      % metaheuristics, especially those we will be
@@
 solutions since smaller populations are known to perform better for a given
-problem than bigger populations. % Smaller populations converge
+problem than bigger populations \cite{li2016multi,wu2016differential}. % Smaller populations converge
                                 % prematurely or perform better? - JJ
@@
 population is compensated by those selected in other
-populations . % reference - JJ
+populations \cite{li2016multi,tanabe2013evaluation}. % reference - JJ
 
 In this work, we will use random settings within a specific range as results have shown
-this is a valid solution to this problem. % This should have been
+this is a valid solution to this problem \cite{garcia2014randomized}. % This should have been
                                 % mentioned in the introduction - JJ
@@
 optimum. This approach has been followed extensively in recent years with
-success. % References - JJ
+success \cite{li2015multi,godio2016multi,biswas2014co}. % References - JJ
 Moreover, there is a need for frameworks, architecture, and

### 2020-01-29 draft.tex
Comments results

@@
 
+% Section headers should be headlines. Like "Measuring scalability in
+% heterogenous systems", or something like that. Experiment 1 says
+% nothing about what's inside (this also applies below) - JJ
 \subsection{Experiment 1}
@@
 experiment consisted of executing the same benchmark on EvoSwarm twice but
-changing the number of populations used in each run. The left column has the
+changing the number of populations used in each run. %What did you
+                                %want to prove? - JJ
+The left column has the
 results when using five populations, and the right side shows the results for
-ten.  Graphics show the execution time achieved by a certain number of workers.
+ten. % Why only 5 and 10? - JJ
+Charts show the execution time achieved by a certain number of workers.
 Each box-plot indicates the time required for the execution of 15 function
@@
 (number of populations and number of workers ). But, when we compare the two
-options, we find some subtle differences:
-
-When using one or two workers, using 5 populations is better in most cases. A
+options, we find some subtle differences: When using one or two
+workers, using 5 populations is better in most cases. A 
 reason for this could be that the migration of individuals takes more time to
 reach the worker. For instance, when a single worker has 10 populations, only
-after processing all, it will start receiving mixed populations.  Although, in
+after processing every one of them, it will start receiving mixed populations.  Although, in
 all cases, and even a single worker, populations are heterogeneous, this
@@
 results can be achieved regardless of our selection between five or ten
-populations. Also, there is no much difference between 4 and 8 workers, when
+populations. Also, there is not much difference between 4 and 8 workers, when
 using five populations. In general, we can observe that keeping a ratio close to
-1:1 between the number of workers and populations exploits better the computing
+1:1 between the number of workers and populations  better exploits the computing
 resources available.
+% I don't really understand this too well... JJ
+% Also, it might just be that populations are waiting on the queue to
+% be processed if there are more sub-populations than workers. The
+% fact that 10 subpopulations take more with 1w than 5 subpopulations
+% might be due to late startup as you say above, but you need to
+% explain it better.
+% Actually, what I see is no significant difference, or diminishing
+% difference, between 5 and 10 sub-populations as the number of
+% workers increases, and as long as the number of workers is less than
+% the number of subpopulations.
 

### 2020-01-29 draft.tex
Justifying containers

@@
 \item Each population is then pushed to the \texttt{Input} queue, so they can be consumed 
-by Stateless functions responsible for the execution of the search.  % Why the caps in Stateless? - JJ  
-% It refers to the swimlines in the diagram, maybe I should point it out  - Mario
-% Added a comment above. - Mario
+by Stateless functions responsible for the execution of the search.
 
@@
 In contrast, in the classic Island model shown in Figure
-\ref{fig:classicisland}, algorithms are coupled with a population, and the same processing
+\ref{fig:classicisland}, algorithms are  population methods, and the same processing
 node keeps both the algorithm process and the state of the population.  Other
@@
 processing nodes, on-runtime, can be more difficult, because they need to have
-an addressing mechanism. % JJ, help I don't know if this is clearly understood, a mean like a DNS - Mario
+an addressing mechanism or simply be aware of these new nodes. This is
+solved in part in Peer to Peer (P2P) evolutionary algorithms
+\cite{juanlu:europar}, but coupling still exists, meaning that every
+node must run the whole evolutionary algorithm and hold full copies of
+its subpopulation. 
+% JJ, help I don't know if this is clearly understood, a mean like a
+% DNS - Mario
+% Added something - JJ
 
@@
 technology enables researchers the local deployment of all the infrastructure
-needed to run the experiments. Later, if needed, the same deployment could be
+needed to run the experiments. Later, if needed, the same system could be
 deployed to the cloud.
+
+The main advantage of container deployments over virtual machine
+instances is that they add much less overhead to communication between
+them, and they are much more economical to deploy. Scaling and
+monitoring can be done much more easily through cloud orchestration
+systems such as Kubernetes; these are the reasons why we chose it for
+this work.
 % I don't see this line of argument. You can do exactly the same with

### 2020-01-30 draft.tex
Exp 2 images

--- 2020-01-29 draft.tex
@@
 multi-population based methods when generally referring to techniques using
-many populations as part of their optimization strategy. % Some comment about asynchrony would be in order - JJ
+subpopulations as part of their optimization strategy. % Some comment about asynchrony would be in order - JJ
 % Comments below - Mario
@@
 When designing efficient multi-population algorithms, researchers need to
-consider additional issues \cite{Ma2019}. These include the number and size of populations,
-the interaction between them, the search area of each population, and the search
-strategy and parametrization of each population. We must take into account
+consider additional issues \cite{Ma2019}. These include the number and size of subpopulations,
+the interaction between them, the search area of subpopulations, and the search
+strategy and parametrization of each subpopulation. We must take into account
 these high-level issues when designing a parallel architecture. Researchers 
@@
     to a continuous stream of data. One of the options is to consider
-    populations as messages to be modified asynchronously by functions.
+    sub-populations as messages to be modified asynchronously by functions.
 
@@
 a multi-container Docker application that reactively processes isolated and
-heterogeneous populations. The platform uses the Docker Compose tool for
+heterogeneous subpopulations. The platform uses the Docker Compose tool for
 defining and running experiments as multi-container applications. 
@@
 Docker Container images host the stateless functions responsible for running 
-a search strategy on several populations, these functions could also be ported
+a search strategy on subpopulations, these functions could also be ported
 directly to a Function as a Service (FaaS) \cite{Roberts2016} implementation.
@@
 Queue implementation and logging and an Experiment Controller implemented in
-Python. Developers of population processing functions have the freedom to
+Python. Developers of sub-population processing functions have the freedom to
 implement new population-based algorithms or operators in the language they
@@
 Multi-population based methods divide the original population into
-smaller populations or islands, with every population carrying out the
+smaller subpopulations or islands, with every subpopulation carrying out the
 algorithm independently, with synchronous or asynchronous communication with the
-rest of the islands \cite{Ma2019}.% exemplify how this is done in different
+rest of the islands. % exemplify how this is done in different
                      % metaheuristics, especially those we will be
@@
 This relative isolation helps in maintaining an overall
-diversity since each population will search in a particular area, at least
+diversity since each subpopulation will search in a particular area, at least
 between communications. The recombination mentioned above (mixing) or migration
-between populations is needed to avoid a premature convergence of candidate
+between subpopulations is needed to avoid a premature convergence of candidate
 solutions since smaller populations are known to perform better for a given
-problem than bigger populations \cite{li2016multi,wu2016differential}. % Smaller populations converge
+problem than bigger populations. % Smaller populations converge
                                 % prematurely or perform better? - JJ
@@
 
-However, in most cases, algorithms applied to each population are
+However, in most cases, algorithms applied to each subpopulation are
 homogeneous, or at any rate, the same variant of the algorithm. As long as this
@@
 tip the balance between exploration and exploitation of the search space. On the
-other hand, current studies show that by having a high number of populations
+other hand, current studies show that by having a high number of subpopulations
 interacting in parallel, the effect of the individual parameters of each
-population is compensated by those selected in other
-populations \cite{li2016multi,tanabe2013evaluation}. % reference - JJ
-
-In this work, we will use random settings within a specific range as results have shown
-this is a valid solution to this problem \cite{garcia2014randomized}. % This should have been
+subpopulation is compensated by those selected in other
+subpopulations. % reference - JJ
+In this
+work, we will use random settings within a specific range as results have shown
+this is a valid solution to this problem. % This should have been
                                 % mentioned in the introduction - JJ
@@
 kept fixed in order to control more easily the execution of the algorithm. For
-instance, by having the size of populations fixed, it is easier to control
+instance, by having the size of subpopulations fixed, it is easier to control
 the number of evaluations and the communication costs, when the algorithm is in
@@
 optimum. This approach has been followed extensively in recent years with
-success \cite{li2015multi,godio2016multi,biswas2014co}. % References - JJ
+success. % References - JJ
 Moreover, there is a need for frameworks, architecture, and
@@
 
-The most elemental data structure used throughout the system is the population; and at a higher
-level, we can view the system as a continuous stream of populations
-flowing from one process to the other. By "continuous stream," we mean that ideally,
+The most elemental data structure used throughout the system is the subpopulation; and at a higher
+level, we can view the system as a continuous stream of subpopulations
+flowing
+% We talk about "multi-population" elsewhere, including the
+% abstract. We should unify terminology - JJ
+% "multi-population" is the whole thing, a subpopulation is a member of that.
+% "A multi-population has many subpopulations or should we call them just populations?"
+% I will add a comment about this in the Multi-population section -
+% Mario
+% use the usual "multi-deme" terminology. A population has different
+% sub-populations nad uses a multi-deme implementation - JJ
+% Let's talk about it, I am using the Multi-population terminology 
+% maybe multi-dem is more common in the EA community but not 
+% much in the Swarm Intelligence side?  
+% Any how, we decide and explain the difference so readers of both 
+% camps are Ok with it. 
+% I prefer the use of population and something like the global population - Mario
+
+from one process to the other. By "continuous stream," we mean that ideally,
 everything must happen asynchronously without components needing to wait for
@@
 implemented by using a message queue system. For instance, if one process needs
-to communicate a population to another, it must pack the population into
+to communicate a subpopulation to another, it must pack the sub-population into
 a message, and then send the message 
+% I don't know if
+                                % asynchronously applies here. It's
+                                % simply sent - JJ 
+                                % Meaning it does not wait for a
+                                % respone, it continues doing other
+                                % stuff - Mario
+% That's rather "fire and forget". Or simply say it's sent without
+% waiting for a response - JJ
+% corrected - Mario
 to the queuing system without waiting for a response; this
@@
 swimlanes: First, the \texttt{Setup} process, responsible for the reading a configuration
-file, and creating the initial populations. Second, the \texttt{Controller} process,
-responsible for the migration between populations, and keeping track of the
+file, and creating the initial subpopulations. Second, the \texttt{Controller} process,
+responsible for the migration between subpopulations, and keeping track of the
 iterations of the algorithm. The \texttt{Message Queue} process runs the \texttt{Input} and \texttt{Output}
@@
 includes all the parameters needed to execute the algorithm, among other things,
-the number of populations, the number of individuals, and the number of iterations
+the number of subpopulations, the number of individuals, and the number of iterations
 of the algorithm. We give more details about the configuration data structure
@@
 parameters found in the configuration structure. The population at this moment is just
-static data, including each individual inside. Each population includes a metadata
+static data, including each individual inside. Each subpopulation includes a metadata
 section where its algorithm and execution parameters are specified. For instance, 
@@
 
-\item Each population is then pushed to the \texttt{Input} queue, so they can be consumed 
+\item Each subpopulation is then pushed to the \texttt{Input} queue, so they can be consumed 
 by Stateless functions responsible for the execution of the search.  % Why the caps in Stateless? - JJ  
@@
 
-\item One or more \texttt{Stateless functions} are constantly pulling population messages
+\item One or more \texttt{Stateless functions} are constantly pulling subpopulation messages
 from the \texttt{Input} queue. Taking the population data and metadata as parameters, these 
@@
 The resulting populations are now pushed to the Output queue. Once finished, another 
-population is pulled from the queue.
+subpopulation is pulled from the queue.
 
@@
 \item Otherwise, it passes the stream of messages 
-to a migration process, where populations are mixed with each other. New populations 
+to a migration process, where subpopulations are mixed with each other. New populations 
 are generated from this migration, and they are again pushed to the Input queue to continue 
@@
 comparison on the coupling and communication between the main components:
-populations, processing nodes, and algorithms.  In Figure \ref{fig:population_message} we show the
+subpopulations, processing nodes, and algorithms.  In Figure \ref{fig:population_message} we show the
 main components of a message-driven architecture. We can see that the algorithm
-and populations are separated, and the \texttt{Population queue} keeps the state
-of populations. While algorithms only need the populations as parameters
+and subpopulations are separated, and the \texttt{Population queue} keeps the state
+of subpopulations. While algorithms only need the populations as parameters
 for their execution.
@@
 In contrast, in the classic Island model shown in Figure
-\ref{fig:classicisland}, algorithms are coupled with a population, and the same processing
-node keeps both the algorithm process and the state of the population.  Other
+\ref{fig:classicisland}, algorithms are coupled with a subpopulation, and the same processing
+node keeps both the algorithm process and the state of the subpopulation.  Other
 processing nodes follow the same configuration and form of execution, and they
@@
 nodes.  In Figure \ref{fig:pool_island} we show the main components of a pool-based
-multi-population algorithm. Although algorithms and population state remains
+multi-population algorithm. Although algorithms and subpopulation state remains
 coupled, now, processing nodes do not communicate directly with each other.
@@
 external processing nodes, the system needs additional communication and
-processing for keeping track of each population.
+processing for keeping track of each subpopulation.
 
@@
 isolated algorithms take random samples of the global population and use this
-temporal population as parameters. An advantage of this design is that the
-sampling provides a type of migration between isolated populations. A problem
-found is that when a processing node returns a population to the pool, the
-state of the population is lost. But having the algorithm decoupled opens the
+temporal subpopulation as parameters. An advantage of this design is that the
+sampling provides a type of migration between isolated subpopulations. A problem
+found is that when a processing node returns a subpopulation to the pool, the
+state of the subpopulation is lost. But having the algorithm decoupled opens the
 possibility of implementing the system using serverless functions.
@@
 be tested, and we must select additional parameters for each dimension.
-According to the maximum number of function evaluations (FEs) we mentioned
+According to the maximum number of  Function Evaluations (FEs) we mentioned
 earlier, we define for each dimension: the number of populations that will be
@@
 propose  an experiment in which we selected $f_4$ (Skew Rastrigin-Bueche separable) from the
-BBOB testbed. This function has been used because it is computationally demanding,
+BBOB testbed. This function because it is computationally demanding,
 and in higher dimensions has been difficult for PSO \cite{el2009black} and GAs
@@
     \small
-    \caption{Parameters used in the experiments with ten populations
+    \caption{10 Populations experiment parameters
     }
     \label{tab:params:10}
-    \vspace{0.25cm}
     \centering
@@
     \small
-    \caption{Parameters used in the experiments that used five populations
+    \caption{5 Populations experiment parameters
     }
     \label{tab:params:5}
-    \vspace{0.25cm}
     \centering
@@
 
-\begin{figure}[ht]
+\subsection{Experiment 1}
+\label{sec:exp1results}
+
+\begin{figure}[h!tb]
+  % Is the axis correct? Seconds per instance or total time? Please
+  % change that - JJ
     \centering
     \includegraphics[width=\textwidth]{sp_size_worker}
-    \caption{Time required, in seconds, to complete an instance of the function, with different number of workers and populations sizes }
+    \caption{Time required to complete experiments with different number of workers and populations sizes }
     \label{fig:spworker}
@@
 
-\subsection{Experiment 1}
-\label{sec:exp1results}
-
-We summarize the results of the experiment in Figure \ref{fig:spworker}.  The
-experiment consisted of executing the same benchmark on EvoSwarm twice but
-changing the number of populations used in each run. The left column has the
-results when using five populations, and the right side shows the results for
-ten.  Graphics show the execution time achieved by a certain number of workers.
-Each box-plot indicates the time required for the execution of 15  function
-instances, in seconds.  Each row has a different scale on the y-axis, to
-compensate for the dimension increments. It is important to notice that the
-execution time can also be affected by the algorithm finding a solution with an
-error smaller than the threshold, before completing all the FEs.
-
-As expected, in most cases, increasing the number of workers reduces the time
-required for the execution of an instance. Mainly when using the same parameters
-(number of populations and number of workers ). But, when we compare the two
-options, we find some subtle differences:
-
-When using one or two workers, using 5 populations is better in most cases. A
-reason for this could be that the migration of individuals takes more time to
-reach the worker. For instance, when a single worker has 10 populations, only
-after processing all, it will start receiving mixed populations.  Although, in
-all cases, and even a single worker, populations are heterogeneous, this
-advantage takes more time to propagate. With four workers, almost the same
-results can be achieved regardless of our selection between five or ten
-populations. Also, there is no much difference between 4 and 8 workers, when
-using five populations. In general, we can observe that keeping a ratio close to
-1:1 between the number of workers and populations exploits better the computing
-resources available.
-
 \subsection{Experiment 2}
@@
 
+\begin{figure*}[h!tb]
+    \begin{tabular}
+        {c@{\hspace*{-0.00001\textwidth}}
+         c@{\hspace*{-0.00001\textwidth}}
+         c@{\hspace*{-0.00001\textwidth}}
+        }
+    GA  &  PSO & GA \& PSO\\   
+    \includegraphics[width=0.33\textwidth]{GAOnly_f001}&
+    \includegraphics[width=0.33\textwidth]{PSOOnly_f001}&
+    \includegraphics[width=0.33\textwidth]{GAPSO_f001}\\
+
+    \includegraphics[width=0.33\textwidth]{GAOnly_f002}&
+    \includegraphics[width=0.33\textwidth]{PSOOnly_f002}&
+    \includegraphics[width=0.33\textwidth]{GAPSO_f002}\\
+
+    \includegraphics[width=0.33\textwidth]{GAOnly_f003}&
+    \includegraphics[width=0.33\textwidth]{PSOOnly_f003}&
+    \includegraphics[width=0.33\textwidth]{GAPSO_f003}\\
+
+    \includegraphics[width=0.33\textwidth]{GAOnly_f004}&
+    \includegraphics[width=0.33\textwidth]{PSOOnly_f004}&
+    \includegraphics[width=0.33\textwidth]{GAPSO_f004}\\
+
+    \includegraphics[width=0.33\textwidth]{GAOnly_f005}&
+    \includegraphics[width=0.33\textwidth]{PSOOnly_f005}&
+    \includegraphics[width=0.33\textwidth]{GAPSO_f005}\\
+    \end{tabular}
+    \vspace{-3ex}
+     \caption{test}
+    \end{figure*}
+
+    \begin{figure}[h!tb]
+        \begin{tabular}
+            {c@{\hspace*{-0.00001\textwidth}}
+             c@{\hspace*{-0.00001\textwidth}}
+            }
+
+        \includegraphics[width=0.45\textwidth]{ppfigs_f001}&
+        \includegraphics[width=0.45\textwidth]{ppfigs_f002}\\
+
+        \includegraphics[width=0.45\textwidth]{ppfigs_f003}&
+        \includegraphics[width=0.45\textwidth]{ppfigs_f004}\\
+
+        \includegraphics[width=0.45\textwidth]{ppfigs_f005}&
+        \\
+        \end{tabular}
+        \vspace{-3ex}
+         \caption{test}
+    \end{figure}
+
 \section{Conclusions} 
@@
 
-\section{Acknowledgements}
-
-This paper has been supported in part by projects DeepBio
-(TIN2017-85727-C4-2-P).
-
 \bibliographystyle{elsarticle-num}

### 2020-01-30 draft.tex
Captions added

--- 2020-01-29 draft.tex
@@
 \usepackage{listings}
+\usepackage{xcolor}
 
@@
 
+\begin{figure*}[h!tb]
+    \begin{tabular}
+        {c@{\hspace*{-0.00001\textwidth}}
+         c@{\hspace*{-0.00001\textwidth}}
+         c@{\hspace*{-0.00001\textwidth}}
+        }
+    GA  &  PSO & GA \& PSO\\   
+    \includegraphics[width=0.33\textwidth]{GAOnly_f001}&
+    \includegraphics[width=0.33\textwidth]{PSOOnly_f001}&
+    \includegraphics[width=0.33\textwidth]{GAPSO_f001}\\
+
+    \includegraphics[width=0.33\textwidth]{GAOnly_f002}&
+    \includegraphics[width=0.33\textwidth]{PSOOnly_f002}&
+    \includegraphics[width=0.33\textwidth]{GAPSO_f002}\\
+
+    \includegraphics[width=0.33\textwidth]{GAOnly_f003}&
+    \includegraphics[width=0.33\textwidth]{PSOOnly_f003}&
+    \includegraphics[width=0.33\textwidth]{GAPSO_f003}\\
+
+    \includegraphics[width=0.33\textwidth]{GAOnly_f004}&
+    \includegraphics[width=0.33\textwidth]{PSOOnly_f004}&
+    \includegraphics[width=0.33\textwidth]{GAPSO_f004}\\
+
+    \includegraphics[width=0.33\textwidth]{GAOnly_f005}&
+    \includegraphics[width=0.33\textwidth]{PSOOnly_f005}&
+    \includegraphics[width=0.33\textwidth]{GAPSO_f005}\\
+    \end{tabular}
+    \vspace{-3ex}
+     \caption{
+ Scaling of runtime with dimension to reach certain target values $\Delta f$. Lines:
+ average runtime (aRT); Cross ($+$): median runtime of successful runs to reach
+ the most difficult target that was reached at least once (but not always);
+ Cross (\textcolor{red}{$\times$}): maximum number of f-evaluations in any trial. Notched boxes:
+ interquartile range with median of simulated runs; All values are divided by
+ dimension and plotted as $log_{10}$ values versus dimension. 
+ %Shown is the aRT for
+ %fixed values of $\Delta f = 10k$ with $k$ given in the legend. 
+ Numbers above aRT-symbols
+ (if appearing) indicate the number of trials reaching the respective target.
+ %The light thick line with diamonds indicates the best algorithm from BBOB 2009
+ %for the most difficult target. Horizontal lines mean linear scaling, slanted
+ %grid lines depict quadratic scaling.
+ }
+    \end{figure*}
+
+    \begin{figure}[h!tb]
+        \begin{tabular}
+            {c@{\hspace*{-0.00001\textwidth}}
+             c@{\hspace*{-0.00001\textwidth}}
+            }
+
+        \includegraphics[width=0.48\textwidth]{ppfigs_f001}&
+        \includegraphics[width=0.48\textwidth]{ppfigs_f002}\\
+
+        \includegraphics[width=0.48\textwidth]{ppfigs_f003}&
+        \includegraphics[width=0.48\textwidth]{ppfigs_f004}\\
+
+        \includegraphics[width=0.48\textwidth]{ppfigs_f005}&
+        \\
+        \end{tabular}
+        \vspace{-3ex}
+         \caption{Average running time (aRT in number of f-evaluations as $log_{10}$ value),
+          divided by dimension for target function value $10^{-8}$ versus dimension. 
+          Slanted grid lines indicate quadratic scaling with the dimension. 
+          Different symbols correspond to different algorithms given in the 
+          legend of $f_1$. Light symbols give the maximum number of 
+          function evaluations from the longest trial divided by dimension. 
+          Black stars indicate a statistically better result compared to 
+          all other algorithms with $p < 0.01$ and Bonferroni 
+          correction number of dimensions (six).}
+    \end{figure}
+
 \section{Conclusions} 
@@
 
-This paper has been supported in part by projects DeepBio
+This paper has been supported in part by projects TecNM-5654.19-P and DeepBio
 (TIN2017-85727-C4-2-P).

### 2020-01-30 draft.tex
Comments on comments
Comment out experiment 2

--- 2020-01-29 draft.tex
@@
 
-\subsection{Experiment 2}
-\label{sec:exp2}
-% Experiment 1
-
-Having a multi-population based algorithm can decrease the execution time
-because of the parallel execution of the evolution. But, having heterogeneous 
-populations might enhance evolutionary search and needless
-evaluations than homogeneous systems; heterogeneous settings, if
-done right, increases the diversity of the whole population
-\cite{araujo2008multikulti}.
-
-But this is a rule of thumb, and it will depend on the degree of heterogeneity,
-as well as on the algorithm itself. Some level of heterogeneity can be
-implemented by just changing the configuration parameters of each population,
-but in this case, we are interested in a heterogeneous search strategy. 
-Therefore, in this experiment we compare a multi-population with only GA and PSO populations,
-versus an ensemble of GA and PSO algorithms. We tested on the first five functions of the
-BBOB testbed. We used ten populations and eight workers for the experiment and the
-same parameters as before.
+% \subsection{Experiment 2}
+% \label{sec:exp2}
+% % Experiment 1
+
+% Having a multi-population based algorithm can decrease the execution time
+% because of the parallel execution of the evolution. But, having heterogeneous 
+% populations might enhance evolutionary search and needless
+% evaluations than homogeneous systems; heterogeneous settings, if
+% done right, increases the diversity of the whole population
+% \cite{araujo2008multikulti}.
+
+% But this is a rule of thumb, and it will depend on the degree of heterogeneity,
+% as well as on the algorithm itself. Some level of heterogeneity can be
+% implemented by just changing the configuration parameters of each population,
+% but in this case, we are interested in a heterogeneous search strategy. 
+% Therefore, in this experiment we compare a multi-population with only GA and PSO populations,
+% versus an ensemble of GA and PSO algorithms. We tested on the first five functions of the
+% BBOB testbed. We used ten populations and eight workers for the experiment and the
+% same parameters as before.
 
@@
 ten. % Why only 5 and 10? - JJ
+     % mmm I really do not have an answer for that, I was thinking about 
+     % testing from 1 to 15, and also more workers - Mario
+
 Charts show the execution time achieved by a certain number of workers.
@@
 % explain it better.
+
+% I was thinking also in terms of the heterogeneous populations.
+% If population 1 to 6 are exploiting, and 7 to 9 are exploring
+% they will take longer to catchup. 
+% While with more workers this happens almost at the start
+% If it makes sense, I will try to explain it better tomorrow with a 
+% clear head - Mario
+
 % Actually, what I see is no significant difference, or diminishing
@@
 
-\subsection{Experiment 2}
-\label{sec:exp2results}
+% I think that was the point, but not too much less for instance 1w and 10pop
+% Also I see a very big jump between 1w and 2w - Mario
+
+% \subsection{Experiment 2}
+% \label{sec:exp2results}
+
+% \begin{figure*}[h!tb]
+%     \begin{tabular}
+%         {c@{\hspace*{-0.00001\textwidth}}
+%          c@{\hspace*{-0.00001\textwidth}}
+%          c@{\hspace*{-0.00001\textwidth}}
+%         }
+%     GA  &  PSO & GA \& PSO\\   
+%     \includegraphics[width=0.33\textwidth]{GAOnly_f001}&
+%     \includegraphics[width=0.33\textwidth]{PSOOnly_f001}&
+%     \includegraphics[width=0.33\textwidth]{GAPSO_f001}\\
+
+%     \includegraphics[width=0.33\textwidth]{GAOnly_f002}&
+%     \includegraphics[width=0.33\textwidth]{PSOOnly_f002}&
+%     \includegraphics[width=0.33\textwidth]{GAPSO_f002}\\
+
+%     \includegraphics[width=0.33\textwidth]{GAOnly_f003}&
+%     \includegraphics[width=0.33\textwidth]{PSOOnly_f003}&
+%     \includegraphics[width=0.33\textwidth]{GAPSO_f003}\\
+
+%     \includegraphics[width=0.33\textwidth]{GAOnly_f004}&
+%     \includegraphics[width=0.33\textwidth]{PSOOnly_f004}&
+%     \includegraphics[width=0.33\textwidth]{GAPSO_f004}\\
+
+%     \includegraphics[width=0.33\textwidth]{GAOnly_f005}&
+%     \includegraphics[width=0.33\textwidth]{PSOOnly_f005}&
+%     \includegraphics[width=0.33\textwidth]{GAPSO_f005}\\
+%     \end{tabular}
+%     \vspace{-3ex}
+%      \caption{
+%  Scaling of runtime with dimension to reach certain target values $\Delta f$. Lines:
+%  average runtime (aRT); Cross ($+$): median runtime of successful runs to reach
+%  the most difficult target that was reached at least once (but not always);
+%  Cross (\textcolor{red}{$\times$}): maximum number of f-evaluations in any trial. Notched boxes:
+%  interquartile range with median of simulated runs; All values are divided by
+%  dimension and plotted as $log_{10}$ values versus dimension. 
+%  %Shown is the aRT for
+%  %fixed values of $\Delta f = 10k$ with $k$ given in the legend. 
+%  Numbers above aRT-symbols
+%  (if appearing) indicate the number of trials reaching the respective target.
+%  %The light thick line with diamonds indicates the best algorithm from BBOB 2009
+%  %for the most difficult target. Horizontal lines mean linear scaling, slanted
+%  %grid lines depict quadratic scaling.
+%  }
+%     \end{figure*}
+
+%     \begin{figure}[h!tb]
+%         \begin{tabular}
+%             {c@{\hspace*{-0.00001\textwidth}}
+%              c@{\hspace*{-0.00001\textwidth}}
+%             }
+
+%         \includegraphics[width=0.48\textwidth]{ppfigs_f001}&
+%         \includegraphics[width=0.48\textwidth]{ppfigs_f002}\\
+
+%         \includegraphics[width=0.48\textwidth]{ppfigs_f003}&
+%         \includegraphics[width=0.48\textwidth]{ppfigs_f004}\\
+
+%         \includegraphics[width=0.48\textwidth]{ppfigs_f005}&
+%         \\
+%         \end{tabular}
+%         \vspace{-3ex}
+%          \caption{Average running time (aRT in number of f-evaluations as $log_{10}$ value),
+%           divided by dimension for target function value $10^{-8}$ versus dimension. 
+%           Slanted grid lines indicate quadratic scaling with the dimension. 
+%           Different symbols correspond to different algorithms given in the 
+%           legend of $f_1$. Light symbols give the maximum number of 
+%           function evaluations from the longest trial divided by dimension. 
+%           Black stars indicate a statistically better result compared to 
+%           all other algorithms with $p < 0.01$ and Bonferroni 
+%           correction number of dimensions (six).}
+%     \end{figure}
 

### 2020-01-30 draft.tex
Small corrections and merging experiment 1 with the whole section

@@
 % nothing about what's inside (this also applies below) - JJ
-\subsection{Experiment 1}
-\label{sec:exp1}
+%\subsection{Experiment 1}
+%\label{sec:exp1}
 % Experiment 1
@@
     \small
-    \caption{Parameters used in the experiments that used five populations
+    \caption{Parameters used in the experiments with five populations
     }
@@
 % Shouldn't this be a subsection of experimental setup? - JJ
-\subsection{Hardware}
-\label{sec:hardware}
+% \subsection{Hardware}
+% \label{sec:hardware}
 
@@
 different number of populations on execution time is reported in  Section
-\ref{sec:exp1results}. The results of the performance comparison between homogeneous and
-heterogeneous search strategies are presented in Section \ref{sec:exp2results}.
+\ref{sec:exp1results}.
+% The results of the performance comparison between homogeneous and
+% heterogeneous search strategies are presented in Section \ref{sec:exp2results}.
 
@@
 
-\subsection{Experiment 1}
-\label{sec:exp1results}
+% \subsection{Experiment 1}
+% \label{sec:exp1results}
 
@@
 required for the execution of an instance. Mainly when using the same parameters
-(number of populations and number of workers ). But, when we compare the two
+(number of populations and number of workers). But, when we compare the two
 options, we find some subtle differences: When using one or two

### 2020-01-30 draft.tex
Explaining away results

@@
 
-We present the results of the experiments in this section. The effect of a
-different number of populations on execution time is reported in  Section
-\ref{sec:exp1results}.
+We present the results of the experiments in this section.
+% The effect of a
+% different number of populations on execution time is reported in  Section
+% \ref{sec:exp1results}.
 % The results of the performance comparison between homogeneous and
 % heterogeneous search strategies are presented in Section \ref{sec:exp2results}.     
-
+%
 \begin{figure}[ht]
@@
 \end{figure}
-
+%
+%
 % \subsection{Experiment 1}
 % \label{sec:exp1results}
-
-We summarize the results of the experiment in Figure \ref{fig:spworker}.  The
+%
+with results summarized in Figure \ref{fig:spworker}.  The
 experiment consisted of executing the same benchmark on EvoSwarm twice but
-changing the number of populations used in each run. %What did you
-                                %want to prove? - JJ
-The left column has the
+changing the number of populations used in each run. This is mainly
+intended as a showcase of the flexibility of the tool presented in
+this paper, but we also wanted to test how the number of workers
+affected the time needed to reach a solution. Adding more populations
+might add more overhead, so this might have a noticeable effect.
+The left column in that figure has the
 results when using five populations, and the right side shows the results for
-ten. % Why only 5 and 10? - JJ
-     % mmm I really do not have an answer for that, I was thinking about 
-     % testing from 1 to 15, and also more workers - Mario
-
-Charts show the execution time achieved by a certain number of workers.
+ten. In principle the main issue is how to find the right number of
+workers for a particular problem, but in general we wanted to check
+the communication overhead when the size of the problem increases and
+how the presented framework reacted to it. % Added some explanation
+% here - JJ
+
+Charts show the execution time achieved by a certain number of
+workers, for every problem dimension, which, besides, changes the
+population size.
 Each box-plot indicates the time required for the execution of 15 function
@@
 
-As expected, in most cases, increasing the number of workers reduces the time
-required for the execution of an instance. Mainly when using the same parameters
-(number of populations and number of workers). But, when we compare the two
+As expected, and in most cases, increasing the number of workers reduces the time
+required for the execution of an instance, mainly when using the same parameters
+(number of populations and number of workers), that is, performance
+scales with the number of workers for a fixed problem and number of
+sub-populations used. However, when we compare the two
 options, we find some subtle differences: When using one or two
-workers, using 5 populations is better in most cases. A 
+workers, using 5 populations takes less time in most cases. A 
 reason for this could be that the migration of individuals takes more time to
@@
 populations. Also, there is not much difference between 4 and 8 workers, when
-using five populations. In general, we can observe that keeping a ratio close to
-1:1 between the number of workers and populations  better exploits the computing
-resources available.
-% I don't really understand this too well... JJ
-% Also, it might just be that populations are waiting on the queue to
-% be processed if there are more sub-populations than workers. The
-% fact that 10 subpopulations take more with 1w than 5 subpopulations
-% might be due to late startup as you say above, but you need to
-% explain it better.
-
-% I was thinking also in terms of the heterogeneous populations.
-% If population 1 to 6 are exploiting, and 7 to 9 are exploring
-% they will take longer to catchup. 
-% While with more workers this happens almost at the start
-% If it makes sense, I will try to explain it better tomorrow with a 
-% clear head - Mario
-
-% Actually, what I see is no significant difference, or diminishing
-% difference, between 5 and 10 sub-populations as the number of
-% workers increases, and as long as the number of workers is less than
-% the number of subpopulations.
-
-% I think that was the point, but not too much less for instance 1w and 10pop
-% Also I see a very big jump between 1w and 2w - Mario
+using five populations, which might indicate that some workers are
+idling while populations are being processed by other workers. In
+general, we can observe that keeping a ratio close to 1:1, but a bit
+better than that, between the number of workers and populations, better exploits the computing
+resources available. This also decreases the relation between the
+overhead needed to migrate and process incoming and outgoing
+population and the total computing time: more populations will need
+more time to do that.
+
+Another additional issue is related to the algorithm itself. Many more
+workers than populations will have a wider range of states of
+evolution in the experiment; populations that will be mixed will have
+very different fitness states, which means that a part of them will
+probably have been not so useful, following the intermediate
+disturbance hypothesis. When the number of workers is increased, the
+difference between the state of evolution of populations will be
+intermediate, leading to better results; eventually, when there are as
+many workers as populations, the state will be quite similar, which is
+shown with a lower gap in the time won adding more workers.
 

### 2020-01-30 draft.tex
Moves figure closer to text
Moving some stuff around
Added conclusion
This clarifies the main selling points of the paper, closes #1
Also changed title, and checked bibliography, not much more to add,
closes #2

@@
 
-\title{An Event-based Architecture for Multi-population Optimization Algorithms}
+\title{EvoSwarm: An Event-based Architecture for Multi-population Optimization Algorithms}
 
@@
 heterogeneous populations. The platform uses the Docker Compose tool for
-defining and running experiments as multi-container applications. 
-% Docker compose is not really for production, more for development environments, 
-% you have to justify this. Also, it does not handle scaling so need to justify it too - JJ
-% justification: - Mario
+defining and running experiments as multi-container applications. EvoSwarm is also released with a free licence in \url{https://github.com/mariosky/EvoSwarm}.
 We preferred the use of a container-based application for this prototype because
@@
 open-source. In EvoSwarm these two algorithms are implemented as stateless
-functions. EvoSwarm is also released with a free licence in \url{https://github.com/mariosky/EvoSwarm}.
-
-% Maybe this should be mentioned earlier, in the workflow? - JJ
+functions.
 
@@
 %
-\begin{figure}[ht]
+\begin{figure}[h!p]
     \centering
@@
 
+In this paper we have presented the EvoSwarm framework that implements
+an event-based, cloud native architecture suitable for hosting any
+kind of multi-population optimization algorithm. The framework is
+based in industry-standard development tools such as Docker and Docker
+Compose, making easy to develop and deploy multi-algorithm distributed
+experiments.
+
+The set up of this kind of algorithms will have a big influence on the
+results, since diversity is important for this kind of algorithms and
+its creation and destruction play a big part of the algorithm's
+performance. One of the main levers to control it in EvoSwarm is via
+the number of workers that will be allocated from the beginning, which
+is why we have performed an experiment mainly intended as a proof of
+concept, but also devoted to show how two parameters in the framework,
+number of workers and number of populations used to initially seed the
+system, interact with each other. The conclusion of this experiment is
+that there is an interesting interplay, with the most benefit being
+obtained with a number of populations equal to the number of workers
+or slightly inferior or superior, which brings algorithmic benefit
+(experiment taking less time because it needs less evaluations) and
+also performance benefits (experiment taking less time because
+evaluations are being split between different workers.
+
+As a conclusion, EvoSwarm is a framework designed and tested according
+to industry standars which is able to bring many benefits to the
+bioinspired algorithm community, by showing the possibility of
+creating new, high performance, mixed algorithms, which can also be
+easily deployed to the cloud, or locally to systems that simply need
+to have Docker and Docker Compose installed.
+
+It also opens new possibilities. From the dessign point of view, it
+would be interesting to try and use automatic scaling, instead of
+setting the number of workers by hand. This would need a certain
+amount of redesign, the most important of which would be using
+container orchestration systems such as Kubernetes or Docker
+Swarm. Instead of setting a fixed number of workers, only the maximum
+amount would need to be established, which could be done directly or
+by setting an evaluation budget.
+
+From the algorithmic point of view, there are many possible lines of
+research. Mixing population-based algorithms with other algorithms
+would be a possibility, but also using different instances of
+population based algorithms, such as checking how Estimation of
+Distribution Algorithms would work together with evolutionary
+algorithms. EvoloPy also includes other population based algorithms,
+which could be tested, trying to find out which sets of algorithms are
+a better match to each other. All these will be explored as future
+lines of work.
+
 \section{Acknowledgements}

### 2020-01-30 draft.tex
Spell check. I think it's ready to go

@@
 Moreover, parallel implementations can run in two ways: synchronous and
-asynchronous. When operations run in parallel but must maintain synchronism with
+asynchronous. When operations run in parallel but must maintain synchrony with
 the other operations,  all operations need to finish to continue to the next
@@
 
-The EvoSwarm application uses configuration files to run the noiseless BOBB testbed
+The EvoSwarm application uses configuration files to run the noiseless BBOB testbed
 experiments.  As a first step, we must deploy the docker application using a
@@
 functions.  The {\em Instances} parameter indicates how many instances of each
-function will be tested. {\em Instances}  has a defualt value of 15. We used this
+function will be tested. {\em Instances}  has a default value of 15. We used this
 value because it is the standard for the BBOB benchmark \cite{hansen2016coco}.
@@
 As a conclusion, EvoSwarm is a framework designed and tested according
-to industry standars which is able to bring many benefits to the
+to industry standards which is able to bring many benefits to the
 bioinspired algorithm community, by showing the possibility of
@@
 
-It also opens new possibilities. From the dessign point of view, it
+It also opens new possibilities. From the design point of view, it
 would be interesting to try and use automatic scaling, instead of

### 2020-01-30 draft.tex
Changed highlight of the paper, there is no space for the algorithmic comparison

@@
     applicability of our application,  measuring the execution time, 
-    speedup, and a comparison between homogeneous and an ensemble of 
-    multi-populations, using Genetic Algorithms (GAs) and Particle 
-    Swarm Optimization (PSO) in a benchmark for the optimization of 
-    continuous functions. % "Validate the applicability" falls short
+    speedup.
+    %, and a comparison between homogeneous and an ensemble of 
+    %multi-populations, using Genetic Algorithms (GAs) and Particle 
+    %Swarm Optimization (PSO) in a benchmark for the optimization of 
+    %continuous functions. % "Validate the applicability" falls short
                           % of being also a contribution. It must

### 2020-01-30 draft.tex
Quotes corrected

@@
 locally or in the cloud, thus proving that cloud-native bioinspired
-algorithms perform better in their "natural" environment than other
+algorithms perform better in their ``natural'' environment than other
 kinds of algorithms, and setting a new baseline for scaling and
@@
 level, we can view the system as a continuous stream of populations
-flowing from one process to the other. By "continuous stream," we mean that ideally,
+flowing from one process to the other. By ``continuous stream,'' we mean that ideally,
 everything must happen asynchronously without components needing to wait for

### 2020-04-30 draft.tex
images moved
Rewriting paragraph to avoid, closes #12

@@
 \cite{salza2016approach,salza2016develop} proposed an architecture
-that carried evolutionary algorithms to the cloud, developed further
-in \cite{de2017parallel} which starts to introduce cloud native
-aspects such as the use of messaging queues and CoreOS as an operating
-system designed from the ground up for container utilization. The way
-these containers are organized, however, is rather classical in
-nature: a master-worker approach, communicated using RabbitMQ, where
-replicated workers perform the tasks in parallel. They called their
-approach AMQPGA, inspired by the protocol, AMQP, used by RabbitMQ, the
-messaging system. Other EAs topologies have also been used in the
-cloud: \cite{dziurzanski2020scalable} adapts an island model to a
-container architecture, in a system that is functionally equivalent to
-classical models, but with the added advantage of being easy to deploy
-and run.
+that carried evolutionary algorithms to the cloud, and which were
+developed further in the next paper \cite{de2017parallel}. These
+papers started to introduce cloud native aspects such as the use of
+messaging queues and CoreOS as an operating system designed from the
+ground up for container utilization in the area of evolutionary
+algorithms. The way these containers are managed, however, is rather
+classical in nature from the point of view of distributed EC: a
+master-worker approach, communicated using RabbitMQ (a messaging
+queue), where replicated workers perform the tasks in parallel. They
+called their approach AMQPGA, inspired by the protocol, AMQP, used by
+RabbitMQ, the messaging system. This classical approach has been also
+followed by other authors by using other EA ``classical'' topologies
+in the cloud: \cite{dziurzanski2020scalable} adapts an island model to
+a container architecture, in a system that is functionally equivalent
+to classical models, but with the added advantage of being easy to
+deploy and run.
 
@@
 
-% \subsection{Experiment 2}
-% \label{sec:exp2results}
-
-% \begin{figure*}[h!tb]
-%     \begin{tabular}
-%         {c@{\hspace*{-0.00001\textwidth}}
-%          c@{\hspace*{-0.00001\textwidth}}
-%          c@{\hspace*{-0.00001\textwidth}}
-%         }
-%     GA  &  PSO & GA \& PSO\\   
-%     \includegraphics[width=0.33\textwidth]{GAOnly_f001}&
-%     \includegraphics[width=0.33\textwidth]{PSOOnly_f001}&
-%     \includegraphics[width=0.33\textwidth]{GAPSO_f001}\\
-
-%     \includegraphics[width=0.33\textwidth]{GAOnly_f002}&
-%     \includegraphics[width=0.33\textwidth]{PSOOnly_f002}&
-%     \includegraphics[width=0.33\textwidth]{GAPSO_f002}\\
-
-%     \includegraphics[width=0.33\textwidth]{GAOnly_f003}&
-%     \includegraphics[width=0.33\textwidth]{PSOOnly_f003}&
-%     \includegraphics[width=0.33\textwidth]{GAPSO_f003}\\
-
-%     \includegraphics[width=0.33\textwidth]{GAOnly_f004}&
-%     \includegraphics[width=0.33\textwidth]{PSOOnly_f004}&
-%     \includegraphics[width=0.33\textwidth]{GAPSO_f004}\\
-
-%     \includegraphics[width=0.33\textwidth]{GAOnly_f005}&
-%     \includegraphics[width=0.33\textwidth]{PSOOnly_f005}&
-%     \includegraphics[width=0.33\textwidth]{GAPSO_f005}\\
-%     \end{tabular}
-%     \vspace{-3ex}
-%      \caption{
-%  Scaling of runtime with dimension to reach certain target values $\Delta f$. Lines:
-%  average runtime (aRT); Cross ($+$): median runtime of successful runs to reach
-%  the most difficult target that was reached at least once (but not always);
-%  Cross (\textcolor{red}{$\times$}): maximum number of f-evaluations in any trial. Notched boxes:
-%  interquartile range with median of simulated runs; All values are divided by
-%  dimension and plotted as $log_{10}$ values versus dimension. 
-%  %Shown is the aRT for
-%  %fixed values of $\Delta f = 10k$ with $k$ given in the legend. 
-%  Numbers above aRT-symbols
-%  (if appearing) indicate the number of trials reaching the respective target.
-%  %The light thick line with diamonds indicates the best algorithm from BBOB 2009
-%  %for the most difficult target. Horizontal lines mean linear scaling, slanted
-%  %grid lines depict quadratic scaling.
-%  }
-%     \end{figure*}
-
-%     \begin{figure}[h!tb]
-%         \begin{tabular}
-%             {c@{\hspace*{-0.00001\textwidth}}
-%              c@{\hspace*{-0.00001\textwidth}}
-%             }
-
-%         \includegraphics[width=0.48\textwidth]{ppfigs_f001}&
-%         \includegraphics[width=0.48\textwidth]{ppfigs_f002}\\
-
-%         \includegraphics[width=0.48\textwidth]{ppfigs_f003}&
-%         \includegraphics[width=0.48\textwidth]{ppfigs_f004}\\
-
-%         \includegraphics[width=0.48\textwidth]{ppfigs_f005}&
-%         \\
-%         \end{tabular}
-%         \vspace{-3ex}
-%          \caption{Average running time (aRT in number of f-evaluations as $log_{10}$ value),
-%           divided by dimension for target function value $10^{-8}$ versus dimension. 
-%           Slanted grid lines indicate quadratic scaling with the dimension. 
-%           Different symbols correspond to different algorithms given in the 
-%           legend of $f_1$. Light symbols give the maximum number of 
-%           function evaluations from the longest trial divided by dimension. 
-%           Black stars indicate a statistically better result compared to 
-%           all other algorithms with $p < 0.01$ and Bonferroni 
-%           correction number of dimensions (six).}
-%     \end{figure}
-
 \section{Conclusions} 

### 2020-04-30 draft.tex
Adds clarification on why this is new refs #4

@@
 that carried evolutionary algorithms to the cloud, and which were
-developed further in the next paper \cite{de2017parallel}. These
+developed further in their next paper \cite{de2017parallel}. These
 papers started to introduce cloud native aspects such as the use of
@@
 Section \ref{multi} and the general design principles of cloud-native
-applications we have mentioned earlier. As a first step, we describe the general
+applications we have mentioned earlier; our intention is to steer
+clear of a direct translation of ``classical'' GA architecturs to the
+cloud and design from the ground up a cloud-native method. As a first step, we describe the general
 architecture first, and then  every component in detail.

### 2020-04-30 draft.tex
Add comment about #5
Adds references
Some hints on #4

@@
 been used extensively in recent years in order to help improve the
-performance of nature-inspired optimization algorithms. Work on
-several populations can be done in parallel, and also asynchronously,
+performance of nature-inspired optimization algorithms. Work on those
+populations can be done in parallel, and they can interact asynchronously,
 a fact that can be leveraged to create scalable implementations based
 on, among other methods, distributed, multi-threaded, parallel, and
-cloud-based computing.  However, the design of these cloud-native,
+cloud-native computing.  However, the design of these cloud-native,
 distributed, multi-population algorithms is not a trivial task. Coming
@@
 levels, from the algorithmic to the functional, must be made to
-leverage the scalability, elasticity, fault-tolerance,
+leverage the scalability, elasticity, (limited) fault-tolerance,
 reproducibility, and cost-effectiveness of cloud systems while, at the
@@
 algorithms perform better in their ``natural'' environment than other
-kinds of algorithms, and setting a new baseline for scaling and
+kinds of algorithms, % Did we compare it to other kind of algorithms?
+                     % If we didn't, this should go - JJ
+and setting a new baseline for scaling and
 performance in the cloud.
@@
 In the last decades, nature-inspired optimization algorithms have been successfully
-applied to solve complex real-world problems \cite{yang2014nature}. Algorithms
-inspired by natural processes include evolutionary algorithms (EAs)
-\cite{back1996evolutionary} and swarm intelligence (SI) \cite{kennedy2006swarm},
-among others. These population-based algorithms share the common characteristic
-of using an initial set of random candidate solutions that are later used to
-generate a new set of candidates, using a nature-inspired heuristic. Popular EAs
+applied to solve many complex real-world problems
+\cite{yang2014nature}; these algorithms can be broadly grouped into evolutionary algorithms (EAs)
+\cite{back1996evolutionary} and swarm intelligence (SI)
+\cite{kennedy2006swarm}, as well as other catetories; popular EAs
 are Genetic Algorithms (GAs) \cite{holland1992adaptation,eiben2003genetic}, 
@@
 
+Besides being all population-based, they also share the common characteristic
+of creating an initial set of random candidate solutions that are later used to
+generate a new set of candidates, using a nature-inspired heuristic.
 Population-based algorithms are intrinsically parallel, and can be implemented to run asynchronously.
-The fitness of each individual can be evaluated independently of 
-others, and similarly, each population could evolve in isolation. Since 
+The fitness of each individual can be (in general) evaluated independently of 
+others, and similarly, each population could evolve in isolation. Researchers use the term
+multi-population based methods when generally referring to techniques using
+many populations as part of their optimization strategy. 
+Since 
 earlier works, researchers have been proposing some form of parallelization
-\cite{muhlenbein1988evolution} to increase the scalability of these algorithms.
+\cite{muhlenbein1988evolution} to increase the scalability of these
+algorithms that adds to this simple parallel execution some form of
+interaction between the independently running populations.
 The island model was one of the first techniques proposed for parallelization,
-which lead to an increased performance \cite{gorges1990explicit,grosso1985computer}. 
+which led to an early proof of speedup   \cite{gorges1990explicit,grosso1985computer}. 
 The concept was to divide a large population into communicated subpopulations. 
@@
 besides the execution speed; these include avoiding a premature convergence and
-maintaining the diversity of the global population \cite{li2015multi}. Researchers use the term
-multi-population based methods when generally referring to techniques using
-many populations as part of their optimization strategy. % Some comment about asynchrony would be in order - JJ
-% Comments below - Mario
+maintaining the diversity of the global population \cite{li2015multi}.
+
 Moreover, parallel implementations can run in two ways: synchronous and
@@
 the other operations,  all operations need to finish to continue to the next
-step. For instance, in a  parent/worker model, the parent needs to wait for all
+step. For instance, in a  master/worker model, the master needs to wait for all
 workers to finish their operations before moving to the next iteration.  In
 contrast, in an asynchronous execution,  operations are not synchronized with
-each other. Following the previous example, now the parent operation can
+each other, so there's no time lost waiting for other populations to
+arribe to the synchronization point. Following the previous example, now the master operation can
 continue to the next iteration, even if a worker has not finished its work.  In
@@
 map-reduce \cite{fazenda2012},  grid \cite{munawar2010design,Gonzalez09},
-voluntary computing \cite{MilkyWay}, and cloud computing
-\cite{GValdez2015,salza2019speed,valenzuela2015implementing,FlexGP}. % many references here too - JJ 
+voluntary computing \cite{MilkyWay}, and, more importantly, cloud computing
+\cite{GValdez2015,salza2019speed,valenzuela2015implementing,FlexGP}.
+
 Cloud computing has become the standard way of running
 enterprise applications, not only because of the convenience of the
-pay-as-you-go model or the non-existent cost of administration, but also because
+pay-as-you-go model or the non-existent cost of physical facilities, but also because
 it offers a way of describing the infrastructure as part of the code, so that it
@@
 computing.  However,  cloud computing has also been evolving, going from simply
-putting in the cloud old-style monolithic applications, that is, applications
+putting using virtual machines elsewhere than on premises old-style monolithic applications, that is, applications
 built on a single stack of services that communicate by layers, to microservices
@@
 stateless functions} \cite{malawski2017serverless}, that react to events and
-only come into existence while they are doing some processing. % many references more here - JJ 
+only come into existence while they are doing some processing.
 Systems developed with the patterns outlined above have the
@@
 Taking these factors into account, in this work, we present EvoSwarm, 
-a multi-container Docker application that reactively processes isolated and
+a multi-container Docker-based application that reactively processes
+isolated and
+% Maybe container-based? You can create containers with podman - JJ
 heterogeneous populations. The platform uses the Docker Compose tool for
@@
 
-Docker Container images host the stateless functions responsible for running 
+Docker container images host the stateless functions responsible for running 
 a search strategy on several populations, these functions could also be ported
@@
 
+% The reviewer says that we should clarify the contributions, and the
+% contribution should answer a research question. Maybe we should
+% clarify these contributions to answer to that. Contribution should
+% be something like "We have proved that cloud-native applications are
+% able to easily accommodate multi-paradigm bioinspired algorithms
+% with little efforts of parametrization and
+% that a combination of algorithms has better speedup and performance
+% than any single one taken independently.
+% These items above could be preceded by "We have proved this in the
+% following steps" - JJ
 \begin{itemize}
@@
 % to better understand the state of the art.
+% I don't know. Maybe it should be directly merged with the state of
+% the art, if there's some essential to our paper it should go to the
+% introduction - JJ
 
@@
 
+%%%%% ----------------- This gets lost in this section -----------------------------
 In this work, we will use random settings within a specific range as results have shown
@@
 
+%%%%%% -------------------- This too. Should go up, to intro, or down,
+%%%%%% to method - JJ
 Some parameters, specially the population size, are
@@
 
+%%%% ---------------------- Also to the introduction
 Combining multiple algorithms, with different parameters, interacting with each
@@
 optimum. This approach has been followed extensively in recent years with
-success \cite{li2015multi,godio2016multi,biswas2014co}. % References - JJ
+success \cite{li2015multi,godio2016multi,biswas2014co}. 
 Moreover, there is a need for frameworks, architecture, and
@@
 
-Although on first sight containers look like simply a method that
+Although at first sight containers look like simply a method that
 allows easy shipping of applications, they are much more than
@@
 cloud and design from the ground up a cloud-native method. As a first step, we describe the general
-architecture first, and then  every component in detail.
+architecture, and then  every component in detail.
 
@@
 
-In contrast, in the classic Island model shown in Figure
+In contrast, in the classic island model shown in Figure
 \ref{fig:classicisland}, algorithms are  population methods, and the same processing
@@
 % OK, I will add an issue - Mario
+% Did you finally? - JJ
 
@@
 
+% Maybe you should say here how everything wors: launch with
+% Docker-compose, whatever. That's the issue that one of the reviewers
+% say - JJ
 \subsection{Worker Containers} 
@@
 
+\section*{References}
+
 \bibliographystyle{elsarticle-num}

### 2020-04-30 draft.tex
This expansion could be useful for #5 and #6

@@
 algorithm parameters were set. In the GA, we randomly set the mutation and
-crossover probabilities to have more heterogeneous workers; in these parameters,
+crossover probabilities to have more heterogeneous workers; % highlight this and add citation - JJ
+in these parameters,
 the range of values is specified. We did not change these parameters during the
@@
 
-The EvoSwarm application uses configuration files to run the noiseless BBOB testbed
-experiments.  As a first step, we must deploy the docker application using a
+The experiments can be easily reproduced by using the configuration
+files the the EvoSwarm application uses to run the noiseless BBOB testbed
+experiments. % Specify this a bit more (with commands such as
+             % docker-compose up and so on and use an itemize
+             % environment to highlight it and answer to the reviewer
+             % - JJ
+As a first step, we must deploy the docker application using a
 {\tt docker-compose.yml} file, where the number of worker containers is
@@
 experiment has to be provided.  Table \ref{tab:params} shows an example of these
-parameters. The { \em GA-PSO Ratio} parameter indicates the proportion of
+parameters. The {\em GA-PSO Ratio} parameter indicates the proportion of
 populations that will use the GA algorithm.  In the example, with a value of
@@
 populations, and finally, a value of  $1$ means that all populations will run
-the GA algorithm. Next, we specify a list indicating which of the 24 benchmark
+the GA algorithm. % Opening up the door to ask for testing with
+                  % different proportion. You have to justify this
+                  % value - JJ 
+Next, we specify a list indicating which of the 24 benchmark
 functions will be tested. In the example, the experiment will use the first five
@@
 
+% this is not a conclusion. The conclusion is an answer to the
+% research question. This works, and works precisely for the reasons
+% we say it works. - JJ
 As a conclusion, EvoSwarm is a framework designed and tested according

### 2020-04-30 draft.tex
Addressing first comments

@@
                      % If we didn't, this should go - JJ 
+                     % Compared against GA and PSO and others from BBOB - Mario
+                     % on the algorithm, not speedup 
 and setting a new baseline for scaling and
@@
 Taking these factors into account, in this work, we present EvoSwarm, 
-a multi-container Docker-based application that reactively processes
+a container-based application that reactively processes
 isolated and
 % Maybe container-based? You can create containers with podman - JJ
+% yes, and is shorter - M
 heterogeneous populations. The platform uses the Docker Compose tool for
@@
 follows:
-
+We have proved that cloud-native applications are
+able to easily accommodate multi-paradigm bioinspired algorithms
+with little efforts of parametrization and
+that a combination of algorithms has better speedup and performance
+than any single one taken independently.
+
+We have proved this in the following steps:
 % The reviewer says that we should clarify the contributions, and the
@@
 % following steps" - JJ
+% I like it will put it just like that. 
+% A reviewer also put enphasis on the workflow and reproductivility 
+% and extracting a Model? we can put something on that too.
+
 \begin{itemize}
@@
                                 % the art - JJ
-    \item Lastly, we present an empirical study to validate the 
-    applicability of our application,  measuring the execution time, 
+    \item Fourth, we present an empirical study to validate the 
+    applicability of our application,  measuring the execution time and  
     speedup.
-    %, and a comparison between homogeneous and an ensemble of 
-    %multi-populations, using Genetic Algorithms (GAs) and Particle 
-    %Swarm Optimization (PSO) in a benchmark for the optimization of 
-    %continuous functions. % "Validate the applicability" falls short
+
+    \item Lastly, we present a comparison between homogeneous and an ensemble of 
+    multi-populations, using Genetic Algorithms (GAs) and Particle 
+    Swarm Optimization (PSO) in a benchmark for the optimization of 
+    continuous functions. % "Validate the applicability" falls short
                           % of being also a contribution. It must
@@
 classical in nature from the point of view of distributed EC: a
-master-worker approach, communicated using RabbitMQ (a messaging
+master-worker approach, comunicated using RabbitMQ (a messaging
 queue), where replicated workers perform the tasks in parallel. They
@@
 followed by other authors by using other EA ``classical'' topologies
-in the cloud: \cite{dziurzanski2020scalable} adapts an island model to
+in the cloud: Dziurzanski et al. \cite{dziurzanski2020scalable} adapts an island model to
 a container architecture, in a system that is functionally equivalent

### 2020-04-30 draft.tex
Proposal for #5
Starting new section

@@
 % Did you finally? - JJ
-
+% I did not, will add and assign it - M 
 In this section, we propose a design of a reactive container-based application
@@
 % say - JJ
+
+% This issue is also commanted in the experimental section
+% I will explain it and then we can see where it goes. - M
+
 \subsection{Worker Containers} 
@@
 
-\subsection{Experimental Workflow in EvoSwarm} 
+\section{Experimental Workflow in EvoSwarm} % We could upgrade this to a section and explain in detail 
+% taking care of #5 - M
 \label{experiment_flow}
+
+Figure \ref{fig:experiment_flow} shows the general workflow to run experiments 
+in Evo\-Swarm, and three roles users can play. The first step is to develop the 
+search strategy algorithms that are going to be used by the multi-population algorithm.
+
+Figure \ref{fig:experiment_flow} shows the general workflow to run experiments
+in Evo\-Swarm, and three roles users can play. The first step is to develop the
+search strategy algorithms that are going to be used by the multi-population
+algorithm; algorithm developers do this. Then administrators are responsible for
+the installation and configuration of the application. Once the application is
+running, users can start one or several experiments, monitor the execution, and
+analyze the results. Depending on the type of experiment executed, users could
+have additional needs. As a use case, there are additional scripts for
+processing the results of experiments to be used with the COCO framework. In the
+next sections, we give more details about these roles and their workflow.
 
@@
 
-Figure \ref{fig:experiment_flow} shows the workflow users can follow to run
-experiments in Evo\-Swarm. The first step is to develop the search strategy
+\subsection{Developer}
+
+\subsection{EvoSwarm Configuration}
+
+\subsection{Experiment Execution}
+
+\subsection{Using the COCO Framework}
+% BBOB Cases
+
+The first step is to develop the search strategy
 algorithms that are going to be used by the multi-population algorithm.
@@
 the application, he or she needs to edit the {\tt docker-compose.yml} file
-% Fi file? - JJ+
-% fixed - Mario
 to specify the type and number of containers needed for the application.  The {\tt docker-compose}

### 2020-05-01 draft.tex
Justifying containers vs vms, closes #13

@@
 
-The main advantage of container deployments over virtual machine
-instances is that they add much less overhead to communication between
-them, and they are much more economical to deploy. Scaling and
-monitoring can be done much more easily through cloud orchestration
-systems such as Kubernetes; these are the reasons why we chose it for
-this work.
+% The main advantage of container deployments over virtual machine
+% instances is that they add much less overhead to communication between
+% them, and they are much more economical to deploy. Scaling and
+% monitoring can be done much more easily through cloud orchestration
+% systems such as Kubernetes; these are the reasons why we chose it for
+% this work.
 % I don't see this line of argument. You can do exactly the same with
@@
 % OK, I will add an issue - Mario
-% Did you finally? - JJ
+% Did you finally? - JJrehe
 % I did not, will add and assign it - M
+There are many advantages of container-based architectures over one
+based on virtual machines. The main ones are architectural: it makes
+the design of an application gravitate towards a set of decoupled
+tasks, that use reactive, event based programming, to interact with
+the rest of the application; containers hold micro-services that are
+much easier to test, integrate and deploy. It also simplifies
+the interaction APIs. Of course, that makes them more economical in a
+pay-per-use cloud provider, but that is not the main reason why we
+have chosen them in this paper; reproductibility of experiments and a
+clear and idempotent description of the architecture through docker
+compose are the main reasons.
+
 In this section, we propose a design of a reactive container-based application

### 2020-05-01 draft.tex
Spell check and fix mostly my mistakes

@@
 \cite{back1996evolutionary} and swarm intelligence (SI)
-\cite{kennedy2006swarm}, as well as other catetories; popular EAs
+\cite{kennedy2006swarm}, as well as other categories; popular EAs
 are Genetic Algorithms (GAs) \cite{holland1992adaptation,eiben2003genetic}, 
@@
 each other, so there's no time lost waiting for other populations to
-arribe to the synchronization point. Following the previous example, now the master operation can
+arrive to the synchronization point. Following the previous example, now the master operation can
 continue to the next iteration, even if a worker has not finished its work.  In
@@
 classical in nature from the point of view of distributed EC: a
-master-worker approach, comunicated using RabbitMQ (a messaging
+master-worker approach, communicated using RabbitMQ (a messaging
 queue), where replicated workers perform the tasks in parallel. They
@@
 applications we have mentioned earlier; our intention is to steer
-clear of a direct translation of ``classical'' GA architecturs to the
+clear of a direct translation of ``classical'' GA architectures to the
 cloud and design from the ground up a cloud-native method. As a first step, we describe the general
@@
 pay-per-use cloud provider, but that is not the main reason why we
-have chosen them in this paper; reproductibility of experiments and a
+have chosen them in this paper; reproducibility of experiments and a
 clear and idempotent description of the architecture through docker

### 2020-05-04 draft.tex
a minor change
Starting cloud deplyment
@jj please check the starting point of a proposed short subsection on cloud implementation. Please add question remaing and things that are not needed.

@@
     \centering
-    \includegraphics[width=\textwidth]{worker}
+    \includegraphics[width=\textwidth]{worker_new}
     \caption{Serverless function implementation details. Showing the worker algorithm in each container}
@@
 
+\subsection{Deploying to a Cloud Provider}
+
+% This is only an idea of the section
+% 
+Several options:
+    A virtual machine with a container-optimized operating system, for example,
+     Google Compute Engine VMs, or Amazon EC2.  We still need to provision the
+      virtual machine and run docker-compose commands as we would do in a local 
+      workstation.
+
+    A Kubernetes cluster, this provides auto-scaling and management of containerized
+    application, but it needs additional configuration for provisioning. All major
+    cloud-providers have Kubernetes services.
+
+    A container orchestration service, these services are compatible with docker-compose
+    files for configuration and provide transparent provisioning. Examples of these
+    services are Amazon Elastic Container Service,  Red Hat Openshift, and Marantis Docker Enterprise.
+
+As a proof-of-concept, we deployed the EvoSwarm application in Amazon Elastic
+Container Service (ECS).  Because it needs fewer changes in configuration, 
+and it is has a lower cost than a Kubernetes alternative.
+
+We can define an ECS cluster directly on AWS Elastic Cloud VMs or through AWS Fargate.
+Fargate is a serverless compute engine for containers that provides automatic provision 
+and management of AWS ECS resources. With this service, calculating the cost of execution
+is simplified because users only specify the computing units and memory for all the application,
+instead of every component.
+
+Using the ECS CLI, we can provision a cluster with a configuration file
+specifying the computing resources needed. Multicontainer applications are
+deployed to ECS as tasks, and a single cluster can run multiple tasks in
+parallel. When using the Fargate service as a backend, tasks are limited to a
+maximum of 10 containers.
+
+We deployed an AWS Fargate cluster and defined a Task configuration file with
+8GBs of memory and 4096 computing units, equivalent to 4 CPUs. Because of the
+ten container limit, we only used seven workers and tested the same function as
+in the previous experiment, but save resources only for dimensions 10 and 20.
+
+The results ..
+
 \section{Conclusions} 

### 2020-05-04 draft.tex
Update draft.tex

@@
 worker object responsible for initializing and running the specified stateless
-function (i.e., GA or PSO). Once the function returns a new population, the
-worker pushes the results to the \texttt{Output} queue, finishes, and passes
-the execution to the main loop. This process is shown in Figure
+function (i.e., GA or PSO) with the population and configuration in the message.
+Once the function returns the evolved population, the
+the main script pushes the results to the \texttt{Output} queue, and continues
+the execution to the infinite loop. This process is shown in Figure
 \ref{fig:worker}.
@@
 Figure \ref{fig:experiment_flow} shows the general workflow to run experiments
-in Evo\-Swarm, and three roles users can play. The first step is to develop the 
-search strategy algorithms that are going to be used by the multi-population algorithm.
-
-Figure \ref{fig:experiment_flow} shows the general workflow to run experiments
-in Evo\-Swarm, and three roles users can play. The first step is to develop the
+in Evo\-Swarm, and the three roles users can play. The first step is to develop the
 search strategy algorithms that are going to be used by the multi-population
@@
 analyze the results. Depending on the type of experiment executed, users could
-have additional needs. As a use case, there are additional scripts for
+have additional needs, in this case there are additional scripts for
 processing the results of experiments to be used with the COCO framework. In the
-next sections, we give more details about these roles and their workflow.
+next sections, we give more details about these roles and their workflows.
 
@@
 can be defined in DockerHub (or any other registry) too.  Docker Compose
-configuration files can reference these containers. When a user needs to deploy
+configuration files can reference these images. When a user needs to deploy
 the application, he or she needs to edit the {\tt docker-compose.yml} file

### 2020-05-04 draft.tex
Depeloper point of view
@jj please check this section regarding #5

@@
 
+Developers can add new algorithm stateless functions in
+several ways and with two levels of integration. In the first level of
+integration, the algorithm execution is not dependent on an initial
+configuration created in the setup, and it only needs to take messages from the
+input queue and return the modified population to the output queue. On the
+second level, the algorithm receives a configuration from the initial setup,
+specifying initial algorithm parameters. We show an example of the configuration
+file received by workers in Figure x. % Adding to Annex or is not even needed? - M
+
+In the case of a decoupled worker, developers can implement a new stateless
+docker container that executes a demon that continuously pulls messages from the
+input queue on the Redis host specified as an environmental variable \tt{REDIS\_HOST}.
+If the function is written in python, developers can base the implementation on
+the provided worker container image, adding a new type of function.
+
+If developers use an existing library for population-based algorithms, they need
+to change the standard model of execution, that first creates a random
+population because, in this case, we need to start with the population provided
+as a parameter. Depending on the implementation, it required to extract the
+population from the JSON message and create the population object needed by the
+library. Again, the state of the population must be encoded as a JSON message to
+be returned as a result. Depending on the type of experiment, additional data
+can be returned in the message. For instance, the message can include the id of
+the container, or even records giving information for each iteration of the
+algorithm. The format of the JSON message is depicted in Figure y. % Adding to Annex or is not even needed? - M
+
+For the next level of integration, developers need to add code to the setup
+container to specify additional configuration options that can be related to
+other workers.
+
+Migration between populations it is independent of the type of algorithm, as the
+controller treats messages as the same type of objects. But, developers could
+also change the controller to specify new rules of operation, for instance, to
+change the type of migration depending on the algorithm.
+
 \subsection{EvoSwarm Configuration}

### 2020-05-04 draft.tex
Update draft.tex

@@
 
-\subsection{EvoSwarm Configuration}
-
-\subsection{Experiment Execution}
-
-\subsection{Using the COCO Framework}
-% BBOB Cases
-
-The first step is to develop the search strategy
-algorithms that are going to be used by the multi-population algorithm.
 Algorithm developers can upload the container definition or images directly to
@@
 GitHub and define a trigger to re-upload the image to DockerHub; this
-can be defined in DockerHub (or any other registry) too.  Docker Compose
-configuration files can reference these images. When a user needs to deploy
-the application, he or she needs to edit the {\tt docker-compose.yml} file
-to specify the type and number of containers needed for the application.  The {\tt docker-compose}
-application is responsible for starting all the containers, downloading new
-versions if it is necessary. After starting the application, users can push
+can be defined in DockerHub (or any other registry) too.
+
+\subsection{EvoSwarm Configuration}
+
+The administrator role is responsible for
+editing the Docker Compose configuration files that reference the images created
+by developers. To deploy the application, he or she needs to edit the {\tt
+docker-compose.yml} file to specify the type and number of containers required
+for the experiments. The {\tt docker-compose} application is responsible for
+starting all the containers, and downloading new versions if it is necessary.
+
+\subsection{Experiment Execution}
+
+After starting the application, users can push
 Experiment Definition files and start monitoring the execution in the terminal.
 After the experiment is finished, the user must execute another script
-that takes as argument the
-experiment id to generate a collection of files containing all the
+that takes as argument the experiment id to generate a collection of files containing all the
 data generated by the experiment in a JSON format.
+
+\subsection{Using the COCO Framework}
+
 Researchers can use these files to analyze and plot experiment data.  Python

### 2020-05-04 draft.tex
Update draft.tex

@@
 file received by workers in Figure x. % Adding to Annex or is not even needed? - M
-
 In the case of a decoupled worker, developers can implement a new stateless
@@
 the provided worker container image, adding a new type of function.  
-
 If developers use an existing library for population-based algorithms, they need
@@
 container to specify additional configuration options that can be related to
-other workers.
+other workers. As with all optimization libraries, developers also need 
+to write the fitness function that is going to be optimized.
 

### 2020-05-04 draft.tex
Reviews #5

@@
 
-\subsection{Developer}
-
-Developers can add new algorithm stateless functions in
+\subsection{EvoSwarm from the developer point of view}
+
+Developers can add new algorithm as stateless functions in
 several ways and with two levels of integration. In the first level of
@@
 file received by workers in Figure x. % Adding to Annex or is not even needed? - M
+
+% Clarify this is the first level.
 In the case of a decoupled worker, developers can implement a new stateless
-docker container that executes a demon that continuously pulls messages from the
-input queue on the Redis host specified as an environmental variable \tt{REDIS\_HOST}.
-If the function is written in python, developers can base the implementation on
+Docker container that executes a daemon that continuously pulls messages from the
+input queue on the Redis host specified as an environmental variable {\tt REDIS\_HOST}.
+If the function is written in Python, developers can base the implementation on
 the provided worker container image, adding a new type of function.  
@@
 algorithm. The format of the JSON message is depicted in Figure y. % Adding to Annex or is not even needed? - M
+% it's probably OK if you add it, but instead of Figure it could be a listing - JJ
 

### 2020-05-04 draft.tex
Many comments re #5 for @mariosky

@@
 If the function is written in Python, developers can base the implementation on
-the provided worker container image, adding a new type of function.  
+the provided worker container image, adding a new type of function.  % say which function, what's its API, how can you test it - JJ
 If developers use an existing library for population-based algorithms, they need
@@
 population because, in this case, we need to start with the population provided
-as a parameter. Depending on the implementation, it required to extract the
+as a parameter. Depending on the implementation, % Depending on which implementation? - JJ
+it is required to extract the
 population from the JSON message and create the population object needed by the
@@
 be returned as a result. Depending on the type of experiment, additional data
-can be returned in the message. For instance, the message can include the id of
+can be returned in the message. % how do you know the type of experiment? how do you add that? Which JSON key? - JJ
+For instance, the message can include the id of
 the container, or even records giving information for each iteration of the
-algorithm. The format of the JSON message is depicted in Figure y. % Adding to Annex or is not even needed? - M
+algorithm. % Is this mandatory? Optional? How is the algorithm going to use it? - JJ
+The format of the JSON message is depicted in Figure y. % Adding to Annex or is not even needed? - M
 % it's probably OK if you add it, but instead of Figure it could be a listing - JJ
@@
 For the next level of integration, developers need to add code to the setup
-container to specify additional configuration options that can be related to
+container to specify additional configuration options that can be related to % related to? What do you mean here?
 other workers. As with all optimization libraries, developers also need 
-to write the fitness function that is going to be optimized.
-
-Migration between populations it is independent of the type of algorithm, as the
-controller treats messages as the same type of objects. But, developers could
+to write the fitness function that is going to be optimized. % Not clear how you integrate this with the rest. An example would probably help. - JJ
+
+Migration between populations is independent of the type of algorithm, as the
+controller treats messages as the same type of objects. But developers could
 also change the controller to specify new rules of operation, for instance, to
-change the type of migration depending on the algorithm.
-
+change the type of migration depending on the algorithm. % Not clear how you would do this. Maybe a simple example would help - JJ
+
+% This is the second level? - JJ
 Algorithm developers can upload the container definition or images directly to
@@
 GitHub and define a trigger to re-upload the image to DockerHub; this
-can be defined in DockerHub (or any other registry) too.
+can be defined in DockerHub (or any other registry) too.u 
+% This is probably accessory. What difference would it make to upload it to GitHub?
+% You also need to say in this section how it helps your overall objective.
+Does it contribute to reproducibility?
+We've talked about that elsewhere.
+Does it help diversity by making easy to add new algorithms?
+- JJ
 

### 2020-05-04 draft.tex
revise cloud deployment, refs #6

@@
 
-\subsection{Deploying to a Cloud Provider}
+\subsection{Deploying to a cloud provider}
 
@@
 %
+% First you have to say that cloud is a way of describing deployment to any kind of infrastructure in a reproducible way, and it needn't be in the cloud proper, it can be deployed to anything, that's precisely the thing.
 Several options:
@@
       virtual machine and run docker-compose commands as we would do in a local 
-      workstation.
+      workstation. % Remember that Salza et al. used CoreOS, which is optimized for containers - JJ
 
@@
 
+    % This is the second option. The first option was already tested, right? - JJ
 As a proof-of-concept, we deployed the EvoSwarm application in Amazon Elastic
-Container Service (ECS).  Because it needs fewer changes in configuration, 
+Container Service (ECS),  since it needs fewer changes in configuration (from deploying it locally using docker-compose)
 and it is has a lower cost than a Kubernetes alternative.
@@
 specifying the computing resources needed. Multicontainer applications are
+% Maybe an example - JJ
 deployed to ECS as tasks, and a single cluster can run multiple tasks in

### 2020-05-05 draft.tex

--- 2020-05-04 draft.tex
@@
 
-\subsection{EvoSwarm from the developer point of view}
-
-Developers can add new algorithm as stateless functions in
+\subsection{Developer}
+
+Developers can add new algorithm stateless functions in
 several ways and with two levels of integration. In the first level of
@@
 file received by workers in Figure x. % Adding to Annex or is not even needed? - M
-
-% Clarify this is the first level.
 In the case of a decoupled worker, developers can implement a new stateless
-Docker container that executes a daemon that continuously pulls messages from the
+docker container that executes a demon that continuously pulls messages from the
 input queue on the Redis host specified as an environmental variable {\tt REDIS\_HOST}.
-If the function is written in Python, developers can base the implementation on
+If the function is written in python, developers can base the implementation on
 the provided worker container image, adding a new type of function.  
@@
 algorithm. The format of the JSON message is depicted in Figure y. % Adding to Annex or is not even needed? - M
-% it's probably OK if you add it, but instead of Figure it could be a listing - JJ
 

### 2020-05-06 draft.tex
Tables for experiment 1

--- 2020-05-04 draft.tex
@@
 %
+
+\begin{table}[]
+    \caption{Speedup by worker and dimension, taking one worker as the baseline.}
+    \label{tab:speedup-table}
+    \vspace{0.25cm}
+    \centering
+
+    \begin{tabular}{c|r|r|r|r|r|r|}
+    \cline{2-7}
+    \multicolumn{1}{l|}{}           & \multicolumn{3}{c|}{5 Populations}                                          & \multicolumn{3}{c|}{10 Populations}                                         \\ \hline
+    \multicolumn{1}{|l|}{Dimension} & \multicolumn{1}{c|}{2W} & \multicolumn{1}{c|}{4W} & \multicolumn{1}{c|}{8W} & \multicolumn{1}{c|}{2W} & \multicolumn{1}{c|}{4W} & \multicolumn{1}{c|}{8W} \\ \hline
+    \multicolumn{1}{|c|}{2}         & 1.94                    & 3.12                    & 3.86                    & 1.84                    & 3.38                    & 5.98                    \\ \hline
+    \multicolumn{1}{|c|}{3}         & 1.93                    & 2.77                    & 3.15                    & 1.90                    & 3.26                    & 5.96                    \\ \hline
+    \multicolumn{1}{|c|}{5}         & 1.91                    & 2.90                    & 3.09                    & 1.91                    & 3.46                    & 6.14                    \\ \hline
+    \multicolumn{1}{|c|}{10}        & 1.92                    & 3.03                    & 3.30                    & 1.90                    & 3.61                    & 6.41                    \\ \hline
+    \multicolumn{1}{|c|}{20}        & 1.91                    & 3.14                    & 3.27                    & 1.89                    & 3.73                    & 6.71                    \\ \hline
+    \multicolumn{1}{|c|}{40}        & 2.00                    & 2.92                    & 3.12                    & 1.95                    & 3.80                    & 7.52                    \\ \hline
+    \end{tabular}
+\end{table}
+
+\begin{table}[]
+    \small
+    \caption{ Median of evaluation rate per second, for each number of workers and dimension}
+    \label{tab:medians}
+    \vspace{0.25cm}
+    \centering
+
+    \begin{tabular}{ccrr}
+    \hline
+               &         & 5 Populations    & 10 Populations \\
+     Dimension & Workers & Median of rate & Median of rate  \\
+    \hline
+    2         & 1w      & 10,815.95      & 10,771.30      \\
+
+              & 2w      & 20,704.11      & 19,197.58      \\
+              & 4w      & 32,563.61      & 35,248.83      \\
+              & 8w      & 40,106.09      & 59,744.88      \\
+    \hline
+    3         & 1w      & 9,599.62       & 9,389.69       \\
+              & 2w      & 18,634.43      & 18,064.24      \\
+              & 4w      & 27,607.12      & 30,419.51      \\
+              & 8w      & 30,189.27      & 56,102.75      \\
+    \hline
+    5         & 1w      & 9,473.63       & 8,671.92       \\
+              & 2w      & 18,181.18      & 16,753.68      \\
+              & 4w      & 27,351.11      & 31,145.09      \\
+              & 8w      & 29,257.72      & 52,985.83      \\
+    \hline
+    10        & 1w      & 8,353.71       & 8,145.61       \\
+              & 2w      & 15,957.87      & 15,577.79      \\
+              & 4w      & 25,696.70      & 29,176.94      \\
+              & 8w      & 28,571.52      & 54,348.60      \\
+    \hline
+    20        & 1w      & 6,920.57       & 6,723.48       \\
+              & 2w      & 13,297.58      & 13,136.11      \\
+              & 4w      & 22,307.73      & 25,271.93      \\
+              & 8w      & 24,179.41      & 46,988.12      \\
+    \hline
+    40        & 1w      & 5,086.88       & 4,976.05       \\
+              & 2w      & 9,896.65       & 9,736.12       \\
+              & 4w      & 16,326.54      & 18,695.84      \\
+              & 8w      & 19,183.37      & 37,302.60      \\
+    \hline  
+    \end{tabular}
+\end{table}
+
+\begin{table}[]
+    \begin{tabular}{lllllll}
+              &               &               & \multicolumn{2}{l}{5 Populations} & \multicolumn{2}{l}{10 Populations} \\
+    Dimension & \multicolumn{2}{l}{Increment} & Speedup         & p-value         & Speedup          & p-value         \\
+    2         & 1w            & 2w            & 1.94            & 6E-16           & 1.84             & 2E-16           \\
+              & 2w            & 4w            & 1.61            & 2E-15           & 1.83             & 1E-15           \\
+              & 4w            & 8w            & 1.24            & 1E-06           & 1.77             & 2E-16           \\
+    3         & 1w            & 2w            & 1.93            & 2E-16           & 1.90             & 2E-16           \\
+              & 2w            & 4w            & 1.44            & 8E-15           & 1.71             & 2E-16           \\
+              & 4w            & 8w            & 1.14            & 3E-05           & 1.83             & 2E-16           \\
+    5         & 1w            & 2w            & 1.91            & 9E-16           & 1.91             & 2E-16           \\
+              & 2w            & 4w            & 1.52            & 3E-11           & 1.82             & 2E-16           \\
+              & 4w            & 8w            & 1.06            & 5E-03           & 1.77             & 2E-16           \\
+    10        & 1w            & 2w            & 1.92            & 3E-16           & 1.90             & 2E-16           \\
+              & 2w            & 4w            & 1.58            & 3E-16           & 1.89             & 2E-16           \\
+              & 4w            & 8w            & 1.09            & 1E-03           & 1.78             & 2E-16           \\
+    20        & 1w            & 2w            & 1.91            & 2E-16           & 1.89             & 2E-16           \\
+              & 2w            & 4w            & 1.64            & 2E-15           & 1.97             & 2E-16           \\
+              & 4w            & 8w            & 1.04            & 1E-02           & 1.80             & 2E-16           \\
+    40        & 1w            & 2w            & 2.00            & 2E-16           & 1.95             & 2E-16           \\
+              & 2w            & 4w            & 1.46            & 1E-09           & 1.95             & 2E-16           \\
+              & 4w            & 8w            & 1.07            & 3E-02           & 1.98             & 2E-16          
+    \end{tabular}
+    \end{table}
+
 \begin{figure}[h!p]

### 2020-05-06 draft.tex
Added Figures and a table caption

@@
 \begin{table}[]
+    \small
+    \caption{ Speedup of incrementing the number of workers for each dimension showing the p-rate for the Wilcoxon test}
+    \label{tab:medians}
+    \vspace{0.25cm}
+    \centering
     \begin{tabular}{lllllll}
@@
     \centering
-    \includegraphics[width=\textwidth]{sp_size_worker}
-    \caption{Time required, in seconds, to complete an instance of the function, with different number of workers and populations sizes }
+    \includegraphics[width=\textwidth]{evalrate}
+    \caption{Rate of number of evaluations per second, for 30 instances of the function, with different number of workers and populations sizes }
     \label{fig:spworker}
 \end{figure}
+
+\begin{figure}[h!p]
+    \centering
+    \includegraphics[width=\textwidth]{evalsworker}
+    \caption{Total number of evaluations for 30 instances of the function, with different number of workers and populations sizes }
+    \label{fig:spworker}
+\end{figure}
+
 %

### 2020-05-06 draft.tex
listings

@@
 
+\lstset{
+basicstyle=\small\ttfamily,
+}
 \graphicspath{ {./images/} }
@@
 
+\begin{lstlisting}[language=Python, caption=Python example]
+{"problem":{
+  "name":"BBOB","instance": 1,"error":1e-8,"function":4,
+  "dim":3,"search_space":[-5,5],
+  "problem_id":"1515-4-1-3",
+  "max_iterations": 20,"population":[],
+  "population_size": 100,
+  "algorithm": "GA",  
+  "params":{
+      "GA":{"crossover": { "type": "cxTwoPoint", 
+            "CXPB_RND": [0.2, 0.6], "CXPB": 0.26},
+             "mutation": { "MUTPB_RND": [0.1, 0.3], 
+             "indpb": 0.05, "sigma": 0.5, 
+             "type":"mutGaussian","mu":0,"MUTPB":0.12},
+             "selection": { 
+               "type": "tools.selTournament", 
+               "tournsize": 2 },
+               "iterations": 50},
+       "PSO": {"Vmax":5,"wMax":0.9,"wMin":0.2,
+               "c1":2,"c2":2,"iterations":50 } 
+    },
+  "experiment":{"type":"benchmark","experiment_id": 1515}
+  }
+}
+\end{lstlisting}
+
 % Clarify this is the first level.
@@
 
+\begin{lstlisting}[language=Python, caption=Python example]
+{
+"time_stamp": 1588455316.698875,
+"evals": [
+  {
+    "gen_num": 0, "best_fitness": -259.96925945939404,
+    "best_solution": [ -4.9446848321611405, 
+                        2.593007989886873, 
+                        2.0011025031167673],
+    "num_of_evals": 17
+  },
+  {
+    "gen_num": 49, "best_fitness": -419.47555780783347,
+    "best_solution": [ -0.9165294450108963, 
+                        2.291966676472422, 
+                        1.8527006568795485],
+    "num_of_evals": 28
+  }
+],
+"worker_id": "4f5eb12e-8278-40de-b0f3-4ac18171566d",
+"message_counter": 1,
+"message_id": "cbcd2e82-e232-4818-8aec-2cd072c7cfe9",
+"best_score": -419.47555780783347
+}
+\end{lstlisting}
+
 For the next level of integration, developers need to add code to the setup
@@
 - JJ
+\begin{lstlisting}[language=Python, caption=Python example]
+{
+ "DIM_CONFIGURATION" : 
+ {
+ "2":{"GEN":40,"POP_SIZE": 50,"MAX_ITER":10,"MESSAGES":10},
+ "3":{"GEN":25,"POP_SIZE": 60,"MAX_ITER":20,"MESSAGES":10},
+ "5":{"GEN":28,"POP_SIZE": 60,"MAX_ITER":30,"MESSAGES":10},
+"10":{"GEN":50,"POP_SIZE": 70,"MAX_ITER":30,"MESSAGES":10},
+"20":{"GEN":66,"POP_SIZE":100,"MAX_ITER":30,"MESSAGES":10},
+"40":{"GEN":80,"POP_SIZE":125,"MAX_ITER":40,"MESSAGES":10}
+},
+"GA_WORKER_RATIO" : 0.5,
+"FUNCTIONS": [4],
+"DIMENSIONS" : [2, 3, 5, 10, 20, 40],
+"CXPB_RND": [0.2, 0.6],
+"MUTPB_RND": [0.1, 0.3]
+}
+\end{lstlisting}
 

### 2020-05-06 draft.tex
listing labels

--- 2020-05-06 draft.tex
@@
 specifying initial algorithm parameters. We show an example of the configuration
-file received by workers in Figure x. % Adding to Annex or is not even needed? - M
+file received by workers in  Listing \ref{code:config}. % Adding to Annex or is not even needed? - M
+
+\begin{lstlisting}[language=Python, caption = Configuration message example, label=code:config]
+{"problem":{
+  "name":"BBOB","instance": 1,"error":1e-8,"function":4,
+  "dim":3,"search_space":[-5,5],
+  "problem_id":"1515-4-1-3",
+  "max_iterations": 20,"population":[],
+  "population_size": 100,
+  "algorithm": "GA",  
+  "params":{
+      "GA":{"crossover": { "type": "cxTwoPoint", 
+            "CXPB_RND": [0.2, 0.6], "CXPB": 0.26},
+             "mutation": { "MUTPB_RND": [0.1, 0.3], 
+             "indpb": 0.05, "sigma": 0.5, 
+             "type":"mutGaussian","mu":0,"MUTPB":0.12},
+             "selection": { 
+               "type": "tools.selTournament", 
+               "tournsize": 2 },
+               "iterations": 50},
+       "PSO": {"Vmax":5,"wMax":0.9,"wMin":0.2,
+               "c1":2,"c2":2,"iterations":50 } 
+    },
+  "experiment":{"type":"benchmark","experiment_id": 1515}
+  }
+}
+\end{lstlisting}
 
@@
 algorithm. % Is this mandatory? Optional? How is the algorithm going to use it? - JJ
-The format of the JSON message is depicted in Figure y. % Adding to Annex or is not even needed? - M
+The format of the JSON message is depicted in  Listing \ref{code:result}. % Adding to Annex or is not even needed? - M
 % it's probably OK if you add it, but instead of Figure it could be a listing - JJ
+
+\begin{lstlisting}[language=Python, caption=Fragment of output message, label=code:result ]
+{
+"time_stamp": 1588455316.698875,
+"evals": [
+  {
+    "gen_num": 0, "best_fitness": -259.96925945939404,
+    "best_solution": [ -4.9446848321611405, 
+                        2.593007989886873, 
+                        2.0011025031167673],
+    "num_of_evals": 17
+  },
+  {
+    "gen_num": 49, "best_fitness": -419.47555780783347,
+    "best_solution": [ -0.9165294450108963, 
+                        2.291966676472422, 
+                        1.8527006568795485],
+    "num_of_evals": 28
+  }
+],
+"worker_id": "4f5eb12e-8278-40de-b0f3-4ac18171566d",
+"message_counter": 1,
+"message_id": "cbcd2e82-e232-4818-8aec-2cd072c7cfe9",
+"best_score": -419.47555780783347
+}
+\end{lstlisting}
 
@@
 can be defined in DockerHub (or any other registry) too.u
+
 % This is probably accessory. What difference would it make to upload it to GitHub?
@@
 that takes as argument the experiment id to generate a collection of files containing all the
-data generated by the experiment in a JSON format.
+data generated by the experiment in a JSON format as shown in Listing \ref{code:exp}.
+
+\begin{lstlisting}[language=Python, caption=Experiment configuration example, label=code:exp]
+{
+"DIM_CONFIGURATION" : 
+ {
+ "2":{"GEN":40,"POP_SIZE": 50,"MAX_ITER":10,"MESSAGES":10},
+ "3":{"GEN":25,"POP_SIZE": 60,"MAX_ITER":20,"MESSAGES":10},
+ "5":{"GEN":28,"POP_SIZE": 60,"MAX_ITER":30,"MESSAGES":10},
+ "10":{"GEN":50,"POP_SIZE": 70,"MAX_ITER":30,"MESSAGES":10},
+ "20":{"GEN":66,"POP_SIZE":100,"MAX_ITER":30,"MESSAGES":10},
+ "40":{"GEN":80,"POP_SIZE":125,"MAX_ITER":40,"MESSAGES":10}
+ },
+"GA_WORKER_RATIO" : 0.5,
+"FUNCTIONS": [4],
+"DIMENSIONS" : [2, 3, 5, 10, 20, 40],
+"CXPB_RND": [0.2, 0.6],
+"MUTPB_RND": [0.1, 0.3]
+}
+\end{lstlisting}
 

### 2020-05-06 draft.tex
Some comments here and there. Please @mariosky check them out

@@
 
+% Maybe these three sections should be merged; they are all experimental set up and the two next ones start with "experimental"
+% (and he next one too) - JJ
 \section{Container-based application} 
@@
 
-We present the results of the experiments in this section.
+We present the results of the experiments in this section. % that's pretty obvious. Say what you intend to prove with the experiments and why - JJ
 % The effect of a
@@
 
-\begin{table}[]
+\begin{table}[h!tbp]
     \small
-    \caption{ Median of evaluation rate per second, for each number of workers and dimension}
+    \caption{Median evaluation rate per second, for each number of workers and dimension}
+        % mark significant differences in some way - JJ
+
     \label{tab:medians}
@@
     \small
-    \caption{ Speedup of incrementing the number of workers for each dimension showing the p-rate for the Wilcoxon test}
+    \caption{Speedup achieved with incrementing the number of workers for each dimension showing the p-rate for the Wilcoxon test}
+    % mark significant differences in some way - JJ
     \label{tab:medians}
@@
     \includegraphics[width=\textwidth]{evalrate}
-    \caption{Rate of number of evaluations per second, for 30 instances of the function, with different number of workers and populations sizes }
+    \caption{Boxplot of the number of evaluations per second, for 30 instances of the function, with different number of workers and populations sizes }
     \label{fig:spworker}
@@
     \includegraphics[width=\textwidth]{evalsworker}
-    \caption{Total number of evaluations for 30 instances of the function, with different number of workers and populations sizes }
+    \caption{Boxplot of the total number of evaluations for 30 instances of the function, with different number of workers and populations sizes }
+    % Which function? - JJ
     \label{fig:spworker}
@@
 % First you have to say that cloud is a way of describing deployment to any kind of infrastructure in a reproducible way, and it needn't be in the cloud proper, it can be deployed to anything, that's precisely the thing.
+But make clear that when you use a cloud provider, there are more options to do it...
+- JJ
 Several options:

### 2020-05-06 draft.tex
Upadate Developer point of view

@@
 second level, the algorithm receives a configuration from the initial setup,
-specifying initial algorithm parameters. We show an example of the configuration
-file received by workers in  Listing \ref{code:config}. % Adding to Annex or is not even needed? - M
+specifying the initial algorithm parameters. We show an example of the configuration
+file received by workers in Listing \ref{code:config}  in which we abbreviated
+the population key. The message includes the current state
+of the population, together with data about the problem, algorithm,
+and its parameters.
 
@@
   "problem_id":"1515-4-1-3",
-  "max_iterations": 20,"population":[],
+  "max_iterations": 20,
+  "population":[
+          [ -3.323452321611405, 
+             2.593007989886873, 
+             3.001112341167673],
+             ...
+          [ -2.22e343232411405, 
+             2.593007989886873, 
+             2.0011025031167673] 
+          ],
   "population_size": 100,
@@
 % Clarify this is the first level.
-In the case of a decoupled worker, developers can implement a new stateless
-Docker container that executes a daemon that continuously pulls messages from the
-input queue on the Redis host specified as an environmental variable {\tt REDIS\_HOST}.
-If the function is written in Python, developers can base the implementation on
-the provided worker container image, adding a new type of function.  % say which function, what's its API, how can you test it - JJ
+In the case of the first level of integration, a decoupled worker, developers
+can implement a new stateless docker container following the description of a
+worker container described in the previous section (\ref{workers}). The worker
+container executes a demon that continuously pulls messages from the input queue
+on the Redis host specified as an environmental variable REDIS_HOST. If the new
+population-based algorithm is written in python, developers can base the new
+implementation on the provided worker container image, adding a new algorithm
+function and worker and modifying the \texttt{main.py} file.
+% say which function, what's its API, how can you test it - JJ
+% this is not an API the need to write their own - Mario
+% We can write an API later - M
 If developers use an existing library for population-based algorithms, they need
@@
 population because, in this case, we need to start with the population provided
-as a parameter. Depending on the implementation, % Depending on which implementation? - JJ
-it is required to extract the
-population from the JSON message and create the population object needed by the
-library. Again, the state of the population must be encoded as a JSON message to
-be returned as a result. Depending on the type of experiment, additional data
-can be returned in the message. % how do you know the type of experiment? how do you add that? Which JSON key? - JJ
-For instance, the message can include the id of
-the container, or even records giving information for each iteration of the
-algorithm. % Is this mandatory? Optional? How is the algorithm going to use it? - JJ
-The format of the JSON message is depicted in  Listing \ref{code:result}. % Adding to Annex or is not even needed? - M
-% it's probably OK if you add it, but instead of Figure it could be a listing - JJ
+as a parameter. Depending on the library, % Depending on which implementation? - JJ
+it may be required to extract the population from the JSON message and replace
+the population object created by the library. Again, the state of 
+the population must be encoded as a JSON message to
+be returned as a result. Additional data
+must be returned in the JSON message, % how do you know the type of experiment? how do you add that? Which JSON key? - JJ
+% I was wrong, it does not depend on the problem. - M
+as depicted in  Listing \ref{code:result}. 
+The output message contains the id of the worker container together
+with records giving information about each iteration of the algorithm, 
+including the best solution found. The controller uses this information 
+to track the execution of the algorithm and keeping a log in the Redis 
+container for posterior analyses. . % Is this mandatory? Optional? How is the algorithm going to use it? - JJ
+% Mandatory and used for everything - M
 
@@
 
-For the next level of integration, developers need to add code to the setup
-container to specify additional configuration options that can be related to % related to? What do you mean here?
-other workers. As with all optimization libraries, developers also need 
-to write the fitness function that is going to be optimized. % Not clear how you integrate this with the rest. An example would probably help. - JJ
-
+For the secondary level of integration, developers need to add code to the setup
+container to specify additional configuration options. For instance,  % related to? What do you mean here?
+in the current configuration message (see Listing \ref{code:config})
+workers executing a GA, initialize the mutation and crossover 
+probabilities randomly from a specific range of values. Users specify
+this range on the experiment configuration.  % Not clear how you integrate this with the rest. An example would probably help. - JJ
+% done - M
 Migration between populations is independent of the type of algorithm, as the
@@
 change the type of migration depending on the algorithm. % Not clear how you would do this. Maybe a simple example would help - JJ
+For example, in the case of swarm-based algorithms,  migration between  % Added - M
+populations (swarms) can follow a topology, indicating valid connections 
+between them (i.e., ring or hypercube). The controller
+can apply these rules selectively only to swarm-based algorithms.
 
 % This is the second level? - JJ
-Algorithm developers can upload the container definition or images directly to
+In either level, algorithm developers can upload the container definition or images directly to
 DockerHub; another option is to upload the container image definition (Dockerfile) to
 GitHub and define a trigger to re-upload the image to DockerHub; this
-can be defined in DockerHub (or any other registry) too.u
+can be defined in DockerHub (or any other registry) too.
 
@@
 % You also need to say in this section how it helps your overall objective. 
-Does it contribute to reproducibility?
-We've talked about that elsewhere.
-Does it help diversity by making easy to add new algorithms?
-- JJ
+% Does it contribute to reproducibility? We've talked about that elsewhere. 
+% Does it help diversity by making easy to add new algorithms? - JJ
 

### 2020-05-06 draft.tex
More comments addressed

@@
 % This is the second level? - JJ
-In either level, algorithm developers can upload the container definition or images directly to
-DockerHub; another option is to upload the container image definition (Dockerfile) to
-GitHub and define a trigger to re-upload the image to DockerHub; this
-can be defined in DockerHub (or any other registry) too.
+
+On either level of integration, algorithm developers can upload the container
+definition or images directly to a public repository like DockerHub. Another
+option is to upload the container image definition (Dockerfile) to GitHub and
+define a trigger to re-upload the image to the registry after an update. Having
+images publicly available can contribute to the reproducibility of experiments.
+Moreover, having more algorithms available can benefit researchers trying to
+compare against other algorithms or increase the diversity of their algorithms
+by adding more search methods. Finally, in some cloud services, having hosted 
+images is a requirement, as it is the case of Amazon ECS.
 
@@
 % Does it help diversity by making easy to add new algorithms? - JJ
+% Explain it a bit more - M
 
@@
 
-After starting the application, users can push
-Experiment Definition files and start monitoring the execution in the terminal.
-After the experiment is finished, the user must execute another script
+After starting the application, users can push several Experiment Definition 
+files at a time and start monitoring the current execution in the terminal.
+After an experiment is finished, the user must execute another script
 that takes as argument the experiment id to generate a collection of files containing all the
-data generated by the experiment in a JSON format as shown in Listing \ref{code:exp}.
-
-\begin{lstlisting}[language=Python, caption=Experiment configuration example, label=code:exp]
+data generated by the experiment, again in a JSON format. An example of an 
+experiment definition is shown in Listing \ref{code:exp}. We give more details about the 
+data contained in the file in the next section.
+
+\begin{lstlisting}[language=Python, caption=Experiment definition example, label=code:exp]
 {

### 2020-05-06 draft.tex
Moved sections for the three experiments

@@
 container executes a demon that continuously pulls messages from the input queue
-on the Redis host specified as an environmental variable REDIS_HOST. If the new
+on the Redis host specified as an environmental variable {\texttt REDIS\_HOST}. If the new
 population-based algorithm is written in python, developers can base the new
@@
 to track the execution of the algorithm and keeping a log in the Redis 
-container for posterior analyses. . % Is this mandatory? Optional? How is the algorithm going to use it? - JJ
+container for posterior analyses.  % Is this mandatory? Optional? How is the algorithm going to use it? - JJ
 % Mandatory and used for everything - M
@@
 % This is the second level? - JJ
-
+% either - M
 On either level of integration, algorithm developers can upload the container
@@
 % Does it help diversity by making easy to add new algorithms? - JJ
-% Explain it a bit more - M
+% I explain it a bit more - M
 
@@
 \subsection{Experiment Execution}
+\label{sec:exeperimentalDefinition}
 
@@
 
-\section{Experimental Setup} 
+\section{General Experimental Setup} 
 \label{setup}
@@
 improve the scalability and performance of population-based optimization
-algorithms. Hence, in this section, we verify the following:
+algorithms. Hence, in the following sections, we verify the following:
 
@@
 search by needing fewer function evaluations than a homogeneous setting?
-
 \end{itemize}
@@
 crossover probabilities to have more heterogeneous workers; % highlight this and add citation - JJ 
-in these parameters,
-the range of values is specified. We did not change these parameters during the
+% done below - M
+in these parameters, the range of values is specified. We did not change these parameters during the 
 experiments, and only the population size and number of generations were
-provided as parameters.
+provided as parameters. This strategy has been applied successfully for increasing 
+the performance of multi-population algorithms without requiring the setting 
+of initial parameters \cite{garcia2014randomized}.
 
@@
 files the the EvoSwarm application uses to run the noiseless BBOB testbed
-experiments. % Specify this a bit more (with commands such as
+experiments as seen in Section~\ref{sec:exeperimentalDefinition}. % Specify this a bit more (with commands such as
              % docker-compose up and so on and use an itemize
@@
                   % value - JJ  
+                  % We only tested 0, 1 and .5, there is not really a justification - M 
 Next, we specify a list indicating which of the 24 benchmark
@@
 % nothing about what's inside (this also applies below) - JJ
-%\subsection{Experiment 1}
-%\label{sec:exp1}
+\section{Measuring Scalability}
+\label{sec:exp1}
 % Experiment 1
-
-% What is the main objective of this experiment? Measure scalability?
-% - JJ
 As we mentioned before, reactive systems can scale by adding additional copies
 of serverless functions. In our case, we can start additional worker containers
-to have the same effect. Other authors, like Salza et al., % add
+to have the same effect. Other authors, like Salza et al. \cite{salza2019speed},   % add
                                 % reference - JJ
@@
 
+% What is the main objective of this experiment? Measure scalability?
+% - JJ 
+% Yes, see below - M
+
 To evaluate the effect of these parameters on the scalability of the system, we
@@
 
+\subsection{Setup}
+
 We executed two sets of experiments with five and ten initial populations,
-repeating each experiment using  1, 2, 4, and 8 workers.  We kept the same
+repeating each experiment 30 times using  1, 2, 4, and 8 workers.  We kept the same 
 maximum number of FEs, changing the relevant parameters for this. 
@@
 for the complete list.
+
+with results summarized in Figure \ref{fig:spworker}.  The
+experiment consisted of executing the same benchmark on EvoSwarm twice but
+changing the number of populations used in each run. This is mainly
+intended as a showcase of the flexibility of the tool presented in
+this paper, but we also wanted to test how the number of workers
+affected the time needed to reach a solution. Adding more populations
+might add more overhead, so this might have a noticeable effect.
+The left column in that figure has the
+results when using five populations, and the right side shows the results for
+ten. In principle the main issue is how to find the right number of
+workers for a particular problem, but in general we wanted to check
+the communication overhead when the size of the problem increases and
+how the presented framework reacted to it. % Added some explanation
+% here - JJ
+
+Charts show the execution time achieved by a certain number of
+workers, for every problem dimension, which, besides, changes the
+population size.
+Each box-plot indicates the time required for the execution of 15 function
+instances, in seconds.  Each row has a different scale on the y-axis, to
+compensate for the dimension increments. It is important to notice that the
+execution time can also be affected by the algorithm finding a solution with an
+error smaller than the threshold, before completing all the FEs.
+
+As expected, and in most cases, increasing the number of workers reduces the time
+required for the execution of an instance, mainly when using the same parameters
+(number of populations and number of workers), that is, performance
+scales with the number of workers for a fixed problem and number of
+sub-populations used. However, when we compare the two
+options, we find some subtle differences: When using one or two
+workers, using 5 populations takes less time in most cases. A 
+reason for this could be that the migration of individuals takes more time to
+reach the worker. For instance, when a single worker has 10 populations, only
+after processing every one of them, it will start receiving mixed populations.  Although, in
+all cases, and even a single worker, populations are heterogeneous, this
+advantage takes more time to propagate. With four workers, almost the same
+results can be achieved regardless of our selection between five or ten
+populations. Also, there is not much difference between 4 and 8 workers, when
+using five populations, which might indicate that some workers are
+idling while populations are being processed by other workers. In
+general, we can observe that keeping a ratio close to 1:1, but a bit
+better than that, between the number of workers and populations, better exploits the computing
+resources available. This also decreases the relation between the
+overhead needed to migrate and process incoming and outgoing
+population and the total computing time: more populations will need
+more time to do that.
+
+Another additional issue is related to the algorithm itself. Many more
+workers than populations will have a wider range of states of
+evolution in the experiment; populations that will be mixed will have
+very different fitness states, which means that a part of them will
+probably have been not so useful, following the intermediate
+disturbance hypothesis. When the number of workers is increased, the
+difference between the state of evolution of populations will be
+intermediate, leading to better results; eventually, when there are as
+many workers as populations, the state will be quite similar, which is
+shown with a lower gap in the time won adding more workers.
 
@@
 
-%
-
-% \subsection{Experiment 2}
-% \label{sec:exp2}
-% % Experiment 1
-
-% Having a multi-population based algorithm can decrease the execution time
-% because of the parallel execution of the evolution. But, having heterogeneous 
-% populations might enhance evolutionary search and needless
-% evaluations than homogeneous systems; heterogeneous settings, if
-% done right, increases the diversity of the whole population
-% \cite{araujo2008multikulti}.
-
-% But this is a rule of thumb, and it will depend on the degree of heterogeneity,
-% as well as on the algorithm itself. Some level of heterogeneity can be
-% implemented by just changing the configuration parameters of each population,
-% but in this case, we are interested in a heterogeneous search strategy. 
-% Therefore, in this experiment we compare a multi-population with only GA and PSO populations,
-% versus an ensemble of GA and PSO algorithms. We tested on the first five functions of the
-% BBOB testbed. We used ten populations and eight workers for the experiment and the
-% same parameters as before.
-
-% Shouldn't this be a subsection of experimental setup? - JJ
-% \subsection{Hardware}
-% \label{sec:hardware}
-
-\section{Experimental Results} 
+\subsection{Experimental Results} 
 \label{results}
-
-We present the results of the experiments in this section. % that's pretty obvious. Say what you intend to prove with the experiments and why - JJ
+% that's pretty obvious. Say what you intend to prove with the experiments and why - JJ 
+% Removed - M
 % The effect of a
@@
 
-    \label{tab:medians}
+    \label{tab:medians:rate}
     \vspace{0.25cm}
@@
     % mark significant differences in some way - JJ
-    \label{tab:medians}
+    \label{tab:speedup:test}
     \vspace{0.25cm}
@@
     \includegraphics[width=\textwidth]{evalrate}
-    \caption{Boxplot of the number of evaluations per second, for 30 instances of the function, with different number of workers and populations sizes }
-    \label{fig:spworker}
+    \caption{Boxplot of the number of evaluations per second, for 30 instances of function $f_4$, with different number of workers and populations sizes }
+    \label{fig:spworker:ratio}
 \end{figure}
@@
     \includegraphics[width=\textwidth]{evalsworker}
-    \caption{Boxplot of the total number of evaluations for 30 instances of the function, with different number of workers and populations sizes }
+    \caption{Boxplot of the total number of evaluations for 30 instances of function $f_4$ , with different number of workers and populations sizes }
     % Which function? - JJ
@@
 
-%
-%
-% \subsection{Experiment 1}
-% \label{sec:exp1results}
-%
-with results summarized in Figure \ref{fig:spworker}.  The
-experiment consisted of executing the same benchmark on EvoSwarm twice but
-changing the number of populations used in each run. This is mainly
-intended as a showcase of the flexibility of the tool presented in
-this paper, but we also wanted to test how the number of workers
-affected the time needed to reach a solution. Adding more populations
-might add more overhead, so this might have a noticeable effect.
-The left column in that figure has the
-results when using five populations, and the right side shows the results for
-ten. In principle the main issue is how to find the right number of
-workers for a particular problem, but in general we wanted to check
-the communication overhead when the size of the problem increases and
-how the presented framework reacted to it. % Added some explanation
-% here - JJ
-
-Charts show the execution time achieved by a certain number of
-workers, for every problem dimension, which, besides, changes the
-population size.
-Each box-plot indicates the time required for the execution of 15 function
-instances, in seconds.  Each row has a different scale on the y-axis, to
-compensate for the dimension increments. It is important to notice that the
-execution time can also be affected by the algorithm finding a solution with an
-error smaller than the threshold, before completing all the FEs.
-
-As expected, and in most cases, increasing the number of workers reduces the time
-required for the execution of an instance, mainly when using the same parameters
-(number of populations and number of workers), that is, performance
-scales with the number of workers for a fixed problem and number of
-sub-populations used. However, when we compare the two
-options, we find some subtle differences: When using one or two
-workers, using 5 populations takes less time in most cases. A 
-reason for this could be that the migration of individuals takes more time to
-reach the worker. For instance, when a single worker has 10 populations, only
-after processing every one of them, it will start receiving mixed populations.  Although, in
-all cases, and even a single worker, populations are heterogeneous, this
-advantage takes more time to propagate. With four workers, almost the same
-results can be achieved regardless of our selection between five or ten
-populations. Also, there is not much difference between 4 and 8 workers, when
-using five populations, which might indicate that some workers are
-idling while populations are being processed by other workers. In
-general, we can observe that keeping a ratio close to 1:1, but a bit
-better than that, between the number of workers and populations, better exploits the computing
-resources available. This also decreases the relation between the
-overhead needed to migrate and process incoming and outgoing
-population and the total computing time: more populations will need
-more time to do that.
-
-Another additional issue is related to the algorithm itself. Many more
-workers than populations will have a wider range of states of
-evolution in the experiment; populations that will be mixed will have
-very different fitness states, which means that a part of them will
-probably have been not so useful, following the intermediate
-disturbance hypothesis. When the number of workers is increased, the
-difference between the state of evolution of populations will be
-intermediate, leading to better results; eventually, when there are as
-many workers as populations, the state will be quite similar, which is
-shown with a lower gap in the time won adding more workers.
-
-\subsection{Deploying to a cloud provider}
+\section{Performance of Heterogeneous Populations in the BBOB Benchmark}
+\label{sec:exp2}
+
+\subsection{Setup}
+
+% % Experiment 1
+
+% Having a multi-population based algorithm can decrease the execution time
+% because of the parallel execution of the evolution. But, having heterogeneous 
+% populations might enhance evolutionary search and needless
+% evaluations than homogeneous systems; heterogeneous settings, if
+% done right, increases the diversity of the whole population
+% \cite{araujo2008multikulti}.
+
+% But this is a rule of thumb, and it will depend on the degree of heterogeneity,
+% as well as on the algorithm itself. Some level of heterogeneity can be
+% implemented by just changing the configuration parameters of each population,
+% but in this case, we are interested in a heterogeneous search strategy. 
+% Therefore, in this experiment we compare a multi-population with only GA and PSO populations,
+% versus an ensemble of GA and PSO algorithms. We tested on the first five functions of the
+% BBOB testbed. We used ten populations and eight workers for the experiment and the
+% same parameters as before.
+
+% Shouldn't this be a subsection of experimental setup? - JJ
+% \subsection{Hardware}
+% \label{sec:hardware}
+
+\section{Deploying to a cloud provider}
 

### 2020-05-06 draft.tex
Hardware uised

@@
 
+For local experiments we have deployed the EvoSwarm application in a
+Desktop PC with AMD Ryzen 9 3900x 12-core processor with 24 threads 
+and 16 GB RAM. We used Docker version 19.03.3, build a872fc2f86, and docker-compose version 1.21.0, 
+in Ubuntu Linux 18.04, and Python 3.7.5 code. Container images and Docker compose file are available at
+(https://hub.docker.com/u/mariosky), and (https://github.com/mariosky/EvoSwarm) respectively. 
+The BOBB experiments were performed with COCO \cite{hansen2016coco} version bbob.v15.03 in python, 
+the plots were produced with version 2.3.2.
+
 % Section headers should be headlines. Like "Measuring scalability in

### 2020-05-07 draft.tex
Changing Images

@@
                   % value - JJ  
-                  % We only tested 0, 1 and .5, there is not really a justification - M 
+                  % We mention earlier that we replicated the params of another paper - M 
 Next, we specify a list indicating which of the 24 benchmark
@@
 % Yes, see below - M
+\subsection{Setup}
 
@@
 can be better understood, since there are fewer factors involved.
-
-\subsection{Setup}
-
-We executed two sets of experiments with five and ten initial populations,
+Following the procedure described in the previous Section, we executed 
+two sets of experiments with five and ten initial populations,
 repeating each experiment 30 times using  1, 2, 4, and 8 workers.  We kept the same 
@@
 
-with results summarized in Figure \ref{fig:spworker}.  The
+\begin{figure}[h!p]
+  \centering
+  \includegraphics[width=\textwidth]{time_wrkr}
+  \caption{time required for the execution of 15 function
+  instances, in seconds }
+
+  % Which function? - JJ
+  \label{fig:spworker}
+\end{figure}
+
+\begin{figure}[h!p]
+  \centering
+  \includegraphics[width=\textwidth]{evalsworker}
+  \caption{Boxplot of the number of evaluations per second, for 30 instances of function $f_4$, with different number of workers and populations sizes }
+
+  % Which function? - JJ
+  \label{fig:spworker}
+\end{figure}
+
+\subsection{Experimental Results} 
+\label{results}
+
+With results summarized in Figure \ref{fig:spworker}.  The
 experiment consisted of executing the same benchmark on EvoSwarm twice but
@@
 
-\subsection{Experimental Results} 
-\label{results}
 % that's pretty obvious. Say what you intend to prove with the experiments and why - JJ 
@@
     \small
-    \caption{Median evaluation rate per second, for each number of workers and dimension}
+    \caption{Boxplot of the total number of evaluations for 30 instances of function $f_4$ , with different number of workers and populations sizes }
       % mark significant differences in some way - JJ
@@
 
-\begin{figure}[h!p]
-    \centering
-    \includegraphics[width=\textwidth]{evalsworker}
-    \caption{Boxplot of the total number of evaluations for 30 instances of function $f_4$ , with different number of workers and populations sizes }
-    % Which function? - JJ
-    \label{fig:spworker}
-\end{figure}
-
 \section{Performance of Heterogeneous Populations in the BBOB Benchmark}

### 2020-05-07 draft.tex
Introduction to the General Exp Setup
@jj  please check
Addresses @mariosky query about experimental setup

@@
 improve the scalability and performance of population-based optimization
-algorithms. Hence, in the following sections, we verify the following:
+algorithms. % If you want to prove it *improves*, you need to provide a baseline. What will be your baseline? You need to specify anyway in which way it really improves - JJ
+Hence, in the following sections, we verify the following:
+
+% JJ these are the three experiments, please check them. The last two addressing
+% issues from a reviewer - M
 
 \begin{itemize}
-\item Whether we can improve the execution time of the algorithm by adding populations
- and serverless functions to the system, in particular, what is the effect of
- these changes on the scalability of the system?
-
-\item Having a multi-population enabled platform, with heterogeneous populations and
+\item Can we improve the execution time of the algorithm by adding populations
+ and serverless functions to the system and, in particular, what is its effect on the scalability of the system?
+
+\item Does having a multi-population enabled platform, with heterogeneous populations and
 the support for mixing search strategies,  increases the performance of the
-search by needing fewer function evaluations than a homogeneous setting?
+search by needing fewer function evaluations than a homogeneous setting? % But in this case the answer is no, right? - JJ
+
+\item Can EvoSwarm as a cloud-native, container-based application,  be deployed
+locally or in the cloud by just specifying the configuration of container images
+and resources to an orchestration service?
 \end{itemize}
@@
 by the expression $10^5 \cdot D$. As an example, if we have $D = 2$, the
-maximum number of FEs is $200,000$
+maximum number of FEs is $200,000$.
+% How do you manage to stop exactly at that number of evaluations in a distributed system? - JJ
 
@@
 - JJ
+% You should start writing this out soon... Don't leave it this way - JJ
 Several options:

### 2020-05-07 draft.tex
REformatting a bit

@@
 of initial parameters \cite{garcia2014randomized}.
-
-\begin{table}
+%
+\begin{table}[h!tbp]
   \small
-  \caption{ DEAP GA EvoWorker Parameters }
+  \caption{DEAP GA EvoWorker Parameters }
   \label{tab:GAparams} 
@@
 \end{table}
-
-\begin{table}
+%
+\begin{table}[h!tbp]
   \small
@@
 number of FEs is $200,000$, which is the same as $40*50*10*10$.
-
-\begin{table}
+%
+\begin{table}[h!tbp]
     \small
@@
 
-% What is the main objective of this experiment? Measure scalability?
-% - JJ 
-% Yes, see below - M
 \subsection{Setup}
@@
 for the complete list. 
-
-\begin{figure}[h!p]
+%
+\begin{figure}[h!tbp]
   \centering
   \includegraphics[width=\textwidth]{time_wrkr}
-  \caption{time required for the execution of 15 function
+  \caption{Time required for the execution of 15 function
   instances, in seconds }
-
   % Which function? - JJ
@@
 \end{figure}
-
-\begin{figure}[h!p]
+%
+%
+%
+\begin{figure}[h!tbp]
   \centering
@@
 many workers as populations, the state will be quite similar, which is
-shown with a lower gap in the time won adding more workers.
-
-\begin{table}
+shown with a smaller gap in the time won adding more workers.
+%
+%
+\begin{table}[h!tbp]
     \small
@@
 \end{table}
-
-\begin{table}
+%
+\begin{table}[h!tbp]
     \small
@@
 \end{table}
-
-% that's pretty obvious. Say what you intend to prove with the experiments and why - JJ 
-% Removed - M
+%
+%
 % The effect of a
@@
 %
-
-\begin{table}[]
+%
+\begin{table}[h!tbp]
     \caption{Speedup by worker and dimension, taking one worker as the baseline.}
@@
 \end{table}
-
+%
+%
 \begin{table}[h!tbp]
@@
 \end{table}
-
-\begin{table}[]
+%
+%
+\begin{table}[h!tbp]
     \small
@@
     \end{table}
-
-\begin{figure}[h!p]
+%
+%
+\begin{figure}[h!tbp]
     \centering

### 2020-05-07 draft.tex
Details checked and revised, closes #7 (when you add to response letter)

@@
 For local experiments we have deployed the EvoSwarm application in a
-Desktop PC with AMD Ryzen 9 3900x 12-core processor with 24 threads 
+Desktop PC with an AMD Ryzen 9 3900x 12-core processor with 24 threads 
 and 16 GB RAM. We used Docker version 19.03.3, build a872fc2f86, and docker-compose version 1.21.0, 
-in Ubuntu Linux 18.04, and Python 3.7.5 code. Container images and Docker compose file are available at
-(https://hub.docker.com/u/mariosky), and (https://github.com/mariosky/EvoSwarm) respectively. 
-The BOBB experiments were performed with COCO \cite{hansen2016coco} version bbob.v15.03 in python, 
-the plots were produced with version 2.3.2.
+in Ubuntu Linux 18.04; coce is run using Python 3.7.5. Container images and the docker-compose file are available at
+\url{https://hub.docker.com/u/mariosky}, and \url{https://github.com/mariosky/EvoSwarm} respectively. 
+The BOBB experiments were performed with COCO \cite{hansen2016coco} version bbob.v15.03 in Python, 
+the plots were produced with version 2.3.2. % Are we using BBOB plots? - JJ
 

### 2020-05-07 draft.tex
comments
Revises #5, still a few issues to check. @mariosky check comments

@@
 file received by workers in Listing \ref{code:config}  in which we abbreviated
-the population key. The message includes the current state
+the population key (the ellipsis is used instead). The message includes the current state
 of the population, together with data about the problem, algorithm,
@@
 
-% Clarify this is the first level.
-In the case of the first level of integration, a decoupled worker, developers
-can implement a new stateless docker container following the description of a
-worker container described in the previous section (\ref{workers}). The worker
-container executes a demon that continuously pulls messages from the input queue
-on the Redis host specified as an environmental variable {\texttt REDIS\_HOST}. If the new
-population-based algorithm is written in python, developers can base the new
+In the case of the first level of integration, a decoupled worker, % a decoupler worker what? - JJ
+developers
+can implement a new stateless Docker container following the description in the previous section (\ref{workers}). The worker
+container executes a daemon that continuously pulls messages from the input queue
+on the Redis host specified as an environment variable {\texttt REDIS\_HOST}. If the new
+population-based algorithm is written in Python, developers can base the new
 implementation on the provided worker container image, adding a new algorithm
@@
 % We can write an API later - M
+% Anyway you need to say what is the signature of the function. Also, if they can use the existing implementation, you need to specify the license - JJ
 If developers use an existing library for population-based algorithms, they need
 to change the standard model of execution, that first creates a random
-population because, in this case, we need to start with the population provided
-as a parameter. Depending on the library, % Depending on which implementation? - JJ
+population because, in this case, the algorithm need to start with the population provided
+as a parameter. Depending on the library, % Depending on which library? - JJ
 it may be required to extract the population from the JSON message and replace
@@
 be returned as a result. Additional data
-must be returned in the JSON message, % how do you know the type of experiment? how do you add that? Which JSON key? - JJ
-% I was wrong, it does not depend on the problem. - M
+must be returned in the JSON message, 
 as depicted in  Listing \ref{code:result}. 
@@
 including the best solution found. The controller uses this information 
-to track the execution of the algorithm and keeping a log in the Redis 
-container for posterior analyses.  % Is this mandatory? Optional? How is the algorithm going to use it? - JJ
-% Mandatory and used for everything - M
-
-\begin{lstlisting}[language=Python, caption=Fragment of output message, label=code:result ]
+to track the execution of the algorithm and keep a log in the Redis 
+container for posterior analyses.
+
+\begin{lstlisting}[language=JSON, caption=Fragment of output message, label=code:result ]
 {
@@
 probabilities randomly from a specific range of values. Users specify
-this range on the experiment configuration.  % Not clear how you integrate this with the rest. An example would probably help. - JJ
-% done - M
+this range on the experiment configuration. 
 Migration between populations is independent of the type of algorithm, as the
@@
 also change the controller to specify new rules of operation, for instance, to
-change the type of migration depending on the algorithm. % Not clear how you would do this. Maybe a simple example would help - JJ
-For example, in the case of swarm-based algorithms,  migration between  % Added - M
+change the type of migration depending on the algorithm. 
+For example, in the case of swarm-based algorithms,  migration between  
 populations (swarms) can follow a topology, indicating valid connections 
@@
 
-% This is the second level? - JJ
-% either - M
 On either level of integration, algorithm developers can upload the container
@@
 images is a requirement, as it is the case of Amazon ECS.
-
-% This is probably accessory. What difference would it make to upload it to GitHub?
-% You also need to say in this section how it helps your overall objective. 
-% Does it contribute to reproducibility? We've talked about that elsewhere. 
-% Does it help diversity by making easy to add new algorithms? - JJ
-% I explain it a bit more - M
 
@@
 improve the scalability and performance of population-based optimization
-algorithms. % If you want to prove it *improves*, you need to provide a baseline. What will be your baseline? You need to specify anyway in which way it really improves - JJ
+algorithms. % If you want to prove it *improves*, you need to provide a baseline. 
+            %What will be your baseline? You need to specify anyway 
+            %in which way it really improves - JJ
+            % The baseline is a single population GA or PSO? - M
 Hence, in the following sections, we verify the following:
@@
 search by needing fewer function evaluations than a homogeneous setting? % But in this case the answer is no, right? - JJ
-
+                                                                         % Is yes, but is in the second experiment - M
 \item Can EvoSwarm as a cloud-native, container-based application,  be deployed
@@
 % How do you manage to stop exactly at that number of evaluations in a distributed system? - JJ
+% In the controller we check the number of messages received, we assume that each message 
+% needs a X amount of #FEs we stop at the maximum number of messages parameter.
+% If the solution is found we stop sending messages and ignore messages from that
+% experiment ID - M
 

### 2020-05-08 draft.tex
Updated Developer
addressing @jj comments

@@
 
-In the case of the first level of integration, a decoupled worker, % a decoupler worker what? - JJ
-developers
-can implement a new stateless Docker container following the description in the previous section (\ref{workers}). The worker
+In the case of the first level of integration, developers % a decoupler worker what? - JJ
+% I think it was redundant, I was saying that in this level the worker container
+% is not tied to the rest of the containers. - M 
+can implement a new stateless Docker container following
+the description in the previous section (\ref{workers}). The worker
 container executes a daemon that continuously pulls messages from the input queue
@@
 function and worker and modifying the \texttt{main.py} file. 
+The worker must include a constructor with the configuration dictionary 
+(see Listing \ref{code:config}) as the only parameter, and call a the new function 
+with the required parameters, unpacking them from the configuration dictionary if necessary.
+All the code of the EvoSwarm Application is published with the open source MIT license.
 % say which function, what's its API, how can you test it - JJ
@@
 % We can write an API later - M
-% Anyway you need to say what is the signature of the function. Also, if they can use the existing implementation, you need to specify the license - JJ
+% Anyway you need to say what is the signature of the function. 
+%Also, if they can use the existing implementation, you need to specify the license - JJ
+
 If developers use an existing library for population-based algorithms, they need
-to change the standard model of execution, that first creates a random
-population because, in this case, the algorithm need to start with the population provided
-as a parameter. Depending on the library, % Depending on which library? - JJ
-it may be required to extract the population from the JSON message and replace
-the population object created by the library. Again, the state of 
-the population must be encoded as a JSON message to
-be returned as a result. Additional data
-must be returned in the JSON message, 
-as depicted in  Listing \ref{code:result}. 
-The output message contains the id of the worker container together
-with records giving information about each iteration of the algorithm, 
-including the best solution found. The controller uses this information 
-to track the execution of the algorithm and keep a log in the Redis 
+to change the standard model of execution, that creates first a random
+population because, in this case, the algorithm needs to start with the
+population provided as a parameter. Depending on the library used, % Depending on which library? - JJ 
+it may be required to extract the population from the
+JSON message and replace the population object created by the library. For
+example, the DEAP library we used for the GA algorithm, creates the type of the
+population at runtime (i.e., bit, float), so we have to replace the population
+once DEAP created it. In the case of EvoloPy, the PSO method directly uses a
+list of floats for storing the swarm, so we just added a parameter to receive
+the population as a list of floats. Both libraries store information about each
+iteration, so we just read it and put it in the format required.
+Again, the state of the population must be encoded as a JSON message to be
+returned as a result, with additional data as
+depicted in Listing \ref{code:result}. The output message contains the id of the
+worker container together with records giving information about each iteration
+of the algorithm, including the best solution found. The controller uses this
+information to track the execution of the algorithm and keep a log in the Redis
 container for posterior analyses.
 
-\begin{lstlisting}[language=JSON, caption=Fragment of output message, label=code:result ]
+\begin{lstlisting}[language=Python, caption=Fragment of output message, label=code:result ]
 {

### 2020-05-08 draft.tex
Changed table and Figure positions

@@
 \subsection{EvoSwarm Configuration} 
-
+\label{sec:evoswarm:config}
 The administrator role is responsible for
@@
 for the experiments. The {\tt docker-compose} application is responsible for
-starting all the containers, and downloading new versions if it is necessary.
+starting all the containers, and downloading new versions if it is necessary,
+by executing the command {\tt docker-compose up}  inside the EvoSwarm directory
 
@@
 
+%
+
 \section{General Experimental Setup} 
@@
 
+\begin{table}[h!tbp]
+  \small
+  \caption{DEAP GA EvoWorker Parameters }
+  \label{tab:GAparams} 
+  \centering
+  \small
+  \begin{tabular}{|l|c|}
+    \hline
+    Selection & Tournament size=12                            \\ \hline
+    Mutation & Gaussian $\mu=0.0$, $\sigma=0.5$, indbp=0.05   \\ \hline
+    Mutation Probability & [0.1,0.3]                          \\ \hline
+    Crossover & Two Point                                     \\ \hline
+    Crossover Probability  & [0.2,0.6]                          \\ \hline
+  \end{tabular}
+\end{table}
+%
+\begin{table}[h!tbp]
+  \small
+  \caption{ EvoloPy PSO Parameters }
+  \label{tab:PSOparams} 
+  \centering
+  \small
+  \begin{tabular}{|l|c|}
+    \hline
+    $V_{max}$ & 6 \\ \hline
+    $W_{max}$ & $0.9$ \\ \hline
+    $W_{min}$ & $0.2$ \\ \hline
+    $C_1$ & 2 \\ \hline
+    $C_2$ & 2 \\ \hline
+  \end{tabular}
+\end{table}
+
 Next, we show the parameters for each algorithm, with Table \ref{tab:GAparams} for
@@
 of initial parameters \cite{garcia2014randomized}.
-%
-\begin{table}[h!tbp]
-  \small
-  \caption{DEAP GA EvoWorker Parameters }
-  \label{tab:GAparams} 
-  \centering
-  \small
-  \begin{tabular}{|l|c|}
-    \hline
-    Selection & Tournament size=12                            \\ \hline
-    Mutation & Gaussian $\mu=0.0$, $\sigma=0.5$, indbp=0.05   \\ \hline
-    Mutation Probability & [0.1,0.3]                          \\ \hline
-    Crossover & Two Point                                     \\ \hline
-    Crossover Probability  & [0.2,0.6]                          \\ \hline
-  \end{tabular}
-\end{table}
-%
-\begin{table}[h!tbp]
-  \small
-  \caption{ EvoloPy PSO Parameters }
-  \label{tab:PSOparams} 
-  \centering
-  \small
-  \begin{tabular}{|l|c|}
-    \hline
-    $V_{max}$ & 6 \\ \hline
-    $W_{max}$ & $0.9$ \\ \hline
-    $W_{min}$ & $0.2$ \\ \hline
-    $C_1$ & 2 \\ \hline
-    $C_2$ & 2 \\ \hline
-  \end{tabular}
-\end{table}
 
 The experiments can be easily reproduced by using the configuration
-files the the EvoSwarm application uses to run the noiseless BBOB testbed
+files the EvoSwarm application uses to run the noiseless BBOB testbed
 experiments as seen in Section~\ref{sec:exeperimentalDefinition}. % Specify this a bit more (with commands such as
@@
              % - JJ
+             % Done, a few lines down - M
 As a first step, we must deploy the docker application using a
-{\tt docker-compose.yml} file, where the number of worker containers is
-specified. Then a JSON file containing the configuration parameters of the
+{\tt docker-compose.yml} file, as described in Section~\ref{sec:evoswarm:config}. 
+Then a JSON file containing the configuration parameters of the
 experiment has to be provided.  Table \ref{tab:params} shows an example of these
@@
                   % We mention earlier that we replicated the params of another paper - M
-Next, we specify a list indicating which of the 24 benchmark
-functions will be tested. In the example, the experiment will use the first five
-functions.  The {\em Instances} parameter indicates how many instances of each
-function will be tested. {\em Instances}  has a default value of 15. We used this
-value because it is the standard for the BBOB benchmark \cite{hansen2016coco}.
-In the {\em Dimensions} parameter, we define a list of the dimensions that will
-be tested, and we must select additional parameters for each dimension.
-According to the maximum number of function evaluations (FEs) we mentioned
-earlier, we define for each dimension: the number of populations that will be
-generated in the setup, and for each population,  the number of generations, and
-population size. Finally, the {\em Iterations} parameter indicates how many
-complete loops will be performed. All these parameters give us the maximum
-number of FEs that will be performed. For instance, for $D = 2$, the maximum
-number of FEs is $200,000$, which is the same as $40*50*10*10$.
-%
+
 \begin{table}[h!tbp]
@@
 
+Next, we specify a list indicating which of the 24 benchmark
+functions will be tested. In the example, the experiment will use the first five
+functions.  The {\em Instances} parameter indicates how many instances of each
+function will be tested. {\em Instances}  has a default value of 15. We used this
+value because it is the standard for the BBOB benchmark \cite{hansen2016coco}.
+In the {\em Dimensions} parameter, we define a list of the dimensions that will
+be tested, and we must select additional parameters for each dimension.
+According to the maximum number of function evaluations (FEs) we mentioned
+earlier, we define for each dimension: the number of populations that will be
+generated in the setup, and for each population,  the number of generations, and
+population size. Finally, the {\em Iterations} parameter indicates how many
+complete loops will be performed. All these parameters give us the maximum
+number of FEs that will be performed. For instance, for $D = 2$, the maximum
+number of FEs is $200,000$, which is the same as $40*50*10*10$.
+%
+
 For local experiments we have deployed the EvoSwarm application in a
@@
                                 % reference - JJ
+                                % done - Mario
 have used a
@@
 
-\subsection{Setup}
-
-To evaluate the effect of these parameters on the scalability of the system, we
-propose  an experiment in which we selected $f_4$ (Skew Rastrigin-Bueche separable) from the
-BBOB testbed. This function has been used because it is computationally demanding,
-and in higher dimensions has been difficult for PSO \cite{el2009black} and GAs
-\cite{nicolau2009application} to solve with the required FEs. As we
-are only comparing in terms of execution time, the results from a single function
-can be better understood, since there are fewer factors involved.
-Following the procedure described in the previous Section, we executed 
-two sets of experiments with five and ten initial populations,
-repeating each experiment 30 times using  1, 2, 4, and 8 workers.  We kept the same 
-maximum number of FEs, changing the relevant parameters for this. 
-See Tables \ref{tab:params:10} and \ref{tab:params:5} 
-for the complete list. 
-%
-\begin{figure}[h!tbp]
-  \centering
-  \includegraphics[width=\textwidth]{time_wrkr}
-  \caption{Time required for the execution of 15 function
-  instances, in seconds }
-  % Which function? - JJ
-  \label{fig:spworker}
-\end{figure}
-%
-%
-%
-\begin{figure}[h!tbp]
-  \centering
-  \includegraphics[width=\textwidth]{evalsworker}
-  \caption{Boxplot of the number of evaluations per second, for 30 instances of function $f_4$, with different number of workers and populations sizes }
-
-  % Which function? - JJ
-  \label{fig:spworker}
-\end{figure}
-
-\subsection{Experimental Results} 
-\label{results}
-
-With results summarized in Figure \ref{fig:spworker}.  The
-experiment consisted of executing the same benchmark on EvoSwarm twice but
-changing the number of populations used in each run. This is mainly
-intended as a showcase of the flexibility of the tool presented in
-this paper, but we also wanted to test how the number of workers
-affected the time needed to reach a solution. Adding more populations
-might add more overhead, so this might have a noticeable effect.
-The left column in that figure has the
-results when using five populations, and the right side shows the results for
-ten. In principle the main issue is how to find the right number of
-workers for a particular problem, but in general we wanted to check
-the communication overhead when the size of the problem increases and
-how the presented framework reacted to it. % Added some explanation
-% here - JJ
-
-Charts show the execution time achieved by a certain number of
-workers, for every problem dimension, which, besides, changes the
-population size.
-Each box-plot indicates the time required for the execution of 15 function
-instances, in seconds.  Each row has a different scale on the y-axis, to
-compensate for the dimension increments. It is important to notice that the
-execution time can also be affected by the algorithm finding a solution with an
-error smaller than the threshold, before completing all the FEs.
-
-As expected, and in most cases, increasing the number of workers reduces the time
-required for the execution of an instance, mainly when using the same parameters
-(number of populations and number of workers), that is, performance
-scales with the number of workers for a fixed problem and number of
-sub-populations used. However, when we compare the two
-options, we find some subtle differences: When using one or two
-workers, using 5 populations takes less time in most cases. A 
-reason for this could be that the migration of individuals takes more time to
-reach the worker. For instance, when a single worker has 10 populations, only
-after processing every one of them, it will start receiving mixed populations.  Although, in
-all cases, and even a single worker, populations are heterogeneous, this
-advantage takes more time to propagate. With four workers, almost the same
-results can be achieved regardless of our selection between five or ten
-populations. Also, there is not much difference between 4 and 8 workers, when
-using five populations, which might indicate that some workers are
-idling while populations are being processed by other workers. In
-general, we can observe that keeping a ratio close to 1:1, but a bit
-better than that, between the number of workers and populations, better exploits the computing
-resources available. This also decreases the relation between the
-overhead needed to migrate and process incoming and outgoing
-population and the total computing time: more populations will need
-more time to do that.
-
-Another additional issue is related to the algorithm itself. Many more
-workers than populations will have a wider range of states of
-evolution in the experiment; populations that will be mixed will have
-very different fitness states, which means that a part of them will
-probably have been not so useful, following the intermediate
-disturbance hypothesis. When the number of workers is increased, the
-difference between the state of evolution of populations will be
-intermediate, leading to better results; eventually, when there are as
-many workers as populations, the state will be quite similar, which is
-shown with a smaller gap in the time won adding more workers.
-%
-%
 \begin{table}[h!tbp]
@@
 \end{table}
+
+\subsection{Setup}
+
+To evaluate the effect of these parameters on the scalability of the system, we
+propose  an experiment in which we selected $f_4$ (Skew Rastrigin-Bueche separable) from the
+BBOB testbed. This function has been used because it is computationally demanding,
+and in higher dimensions has been difficult for PSO \cite{el2009black} and GAs
+\cite{nicolau2009application} to solve with the required FEs. As we
+are only comparing in terms of execution time, the results from a single function
+can be better understood, since there are fewer factors involved.
+Following the procedure described in the previous Section, we executed 
+two sets of experiments with five and ten initial populations,
+repeating each experiment 30 times using  1, 2, 4, and 8 workers.  We kept the same 
+maximum number of FEs, changing the relevant parameters for this. 
+See Tables \ref{tab:params:10} and \ref{tab:params:5} 
+for the complete list. 
+%
+\begin{figure}[h!tbp]
+  \centering
+  \includegraphics[width=\textwidth]{time_wrkr}
+  \caption{Boxplot of the median time required for the execution of 30 
+  instances of function $f_4$, in seconds, different number of workers and populations sizes.  }
+  % Which function? - JJ
+  \label{fig:spworker}
+\end{figure}
+%
+\begin{table}[]
+  \caption{The median of the time required to finish an instance of the function $f_4$, 30 runs
+  with different number of workers and populations}
+  \label{tab:time}
+  \vspace{0.25cm}
+    \centering
+    \begin{tabular}{ccrr}
+    \hline
+              &         & 5 Populations    & 10 Populations \\
+     Dimension& Workers & Median of rate & Median of rate  \\
+    \hline
+          2   & 1w      & 2.96   & 3.15   \\
+              & 2w      & 1.84   & 1.72   \\
+              & 4w      & 1.22   & 0.96   \\
+              & 8w      & 1.05   & 0.59   \\
+    \hline
+          3   & 1w      & 6.53   & 7.76   \\
+              & 2w      & 3.64   & 3.83   \\
+              & 4w      & 2.33   & 2.26   \\
+              & 8w      & 2.17   & 1.36   \\
+    \hline
+          5   & 1w      & 12.33  & 14.19  \\
+              & 2w      & 6.63   & 7.41   \\
+              & 4w      & 4.25   & 4.34   \\
+              & 8w      & 3.67   & 2.41   \\
+    \hline
+          10  & 1w      & 31.50  & 38.30  \\
+              & 2w      & 16.07  & 19.31  \\
+              & 4w      & 9.99   & 10.35  \\
+              & 8w      & 8.97   & 5.92   \\
+    \hline
+          20  & 1w      & 116.35 & 124.51 \\
+              & 2w      & 51.18  & 63.68  \\
+              & 4w      & 32.26  & 33.07  \\
+              & 8w      & 28.64  & 18.25  \\
+    \hline
+          40  & 1w      & 369.89 & 448.89 \\
+              & 2w      & 198.43 & 234.50 \\
+              & 4w      & 122.39 & 120.62 \\
+              & 8w      & 122.31 & 63.35 \\
+    \hline
+  \end{tabular}
+\end{table}
+
 %
 %
-% The effect of a
-% different number of populations on execution time is reported in  Section
-% \ref{sec:exp1results}.
-% The results of the performance comparison between homogeneous and
-% heterogeneous search strategies are presented in Section \ref{sec:exp2results}.     
-%
-%
+\begin{figure}[h!tbp]
+  \centering
+  \includegraphics[width=\textwidth]{evalspersecond}
+  \caption{Boxplot of the number of evaluations per second, for 30 instances of function $f_4$, with different number of workers and populations sizes }
+
+  % Which function? - JJ
+  \label{fig:spworker}
+\end{figure}
+
 \begin{table}[h!tbp]
@@
 %
-%
-\begin{table}[h!tbp]
-    \small
-    \caption{Boxplot of the total number of evaluations for 30 instances of function $f_4$ , with different number of workers and populations sizes }
-      % mark significant differences in some way - JJ
-
-    \label{tab:medians:rate}
-    \vspace{0.25cm}
-    \centering
-
-    \begin{tabular}{ccrr}
-    \hline
-               &         & 5 Populations    & 10 Populations \\
-     Dimension & Workers & Median of rate & Median of rate  \\
-    \hline
-    2         & 1w      & 10,815.95      & 10,771.30      \\
-
-              & 2w      & 20,704.11      & 19,197.58      \\
-              & 4w      & 32,563.61      & 35,248.83      \\
-              & 8w      & 40,106.09      & 59,744.88      \\
-    \hline
-    3         & 1w      & 9,599.62       & 9,389.69       \\
-              & 2w      & 18,634.43      & 18,064.24      \\
-              & 4w      & 27,607.12      & 30,419.51      \\
-              & 8w      & 30,189.27      & 56,102.75      \\
-    \hline
-    5         & 1w      & 9,473.63       & 8,671.92       \\
-              & 2w      & 18,181.18      & 16,753.68      \\
-              & 4w      & 27,351.11      & 31,145.09      \\
-              & 8w      & 29,257.72      & 52,985.83      \\
-    \hline
-    10        & 1w      & 8,353.71       & 8,145.61       \\
-              & 2w      & 15,957.87      & 15,577.79      \\
-              & 4w      & 25,696.70      & 29,176.94      \\
-              & 8w      & 28,571.52      & 54,348.60      \\
-    \hline
-    20        & 1w      & 6,920.57       & 6,723.48       \\
-              & 2w      & 13,297.58      & 13,136.11      \\
-              & 4w      & 22,307.73      & 25,271.93      \\
-              & 8w      & 24,179.41      & 46,988.12      \\
-    \hline
-    40        & 1w      & 5,086.88       & 4,976.05       \\
-              & 2w      & 9,896.65       & 9,736.12       \\
-              & 4w      & 16,326.54      & 18,695.84      \\
-              & 8w      & 19,183.37      & 37,302.60      \\
-    \hline  
-    \end{tabular}
-\end{table}
-%
-%
+
 \begin{table}[h!tbp]
@@
   \end{table}
+
+\subsection{Experimental Results} 
+\label{results}
+
+With results summarized in Figure \ref{fig:spworker}.  The
+experiment consisted of executing the same benchmark on EvoSwarm twice but
+changing the number of populations used in each run. This is mainly
+intended as a showcase of the flexibility of the tool presented in
+this paper, but we also wanted to test how the number of workers
+affected the time needed to reach a solution. Adding more populations
+might add more overhead, so this might have a noticeable effect.
+The left column in that figure has the
+results when using five populations, and the right side shows the results for
+ten. In principle the main issue is how to find the right number of
+workers for a particular problem, but in general we wanted to check
+the communication overhead when the size of the problem increases and
+how the presented framework reacted to it. % Added some explanation
+% here - JJ
+
+Charts show the execution time achieved by a certain number of
+workers, for every problem dimension, which, besides, changes the
+population size.
+Each box-plot indicates the time required for the execution of 15 function
+instances, in seconds.  Each row has a different scale on the y-axis, to
+compensate for the dimension increments. It is important to notice that the
+execution time can also be affected by the algorithm finding a solution with an
+error smaller than the threshold, before completing all the FEs.
+
+As expected, and in most cases, increasing the number of workers reduces the time
+required for the execution of an instance, mainly when using the same parameters
+(number of populations and number of workers), that is, performance
+scales with the number of workers for a fixed problem and number of
+sub-populations used. However, when we compare the two
+options, we find some subtle differences: When using one or two
+workers, using 5 populations takes less time in most cases. A 
+reason for this could be that the migration of individuals takes more time to
+reach the worker. For instance, when a single worker has 10 populations, only
+after processing every one of them, it will start receiving mixed populations.  Although, in
+all cases, and even a single worker, populations are heterogeneous, this
+advantage takes more time to propagate. With four workers, almost the same
+results can be achieved regardless of our selection between five or ten
+populations. Also, there is not much difference between 4 and 8 workers, when
+using five populations, which might indicate that some workers are
+idling while populations are being processed by other workers. In
+general, we can observe that keeping a ratio close to 1:1, but a bit
+better than that, between the number of workers and populations, better exploits the computing
+resources available. This also decreases the relation between the
+overhead needed to migrate and process incoming and outgoing
+population and the total computing time: more populations will need
+more time to do that.
+
+Another additional issue is related to the algorithm itself. Many more
+workers than populations will have a wider range of states of
+evolution in the experiment; populations that will be mixed will have
+very different fitness states, which means that a part of them will
+probably have been not so useful, following the intermediate
+disturbance hypothesis. When the number of workers is increased, the
+difference between the state of evolution of populations will be
+intermediate, leading to better results; eventually, when there are as
+many workers as populations, the state will be quite similar, which is
+shown with a smaller gap in the time won adding more workers.
+%
+% Boxplot of the total number of evaluations for 30 instances of function f4 , with
+% different number of workers and populations sizes
+
 %
 %
-\begin{figure}[h!tbp]
-    \centering
-    \includegraphics[width=\textwidth]{evalrate}
-    \caption{Boxplot of the number of evaluations per second, for 30 instances of function $f_4$, with different number of workers and populations sizes }
-    \label{fig:spworker:ratio}
-\end{figure}
+% The effect of a
+% different number of populations on execution time is reported in  Section
+% \ref{sec:exp1results}.
+% The results of the performance comparison between homogeneous and
+% heterogeneous search strategies are presented in Section \ref{sec:exp2results}.     
+%
+%
+
+%
+%
+%\begin{figure}[h!tbp]
+%%    \centering
+%    \includegraphics[width=\textwidth]{evalrate}
+%%    \caption{Boxplot of the number of evaluations per second, for 30 instances of function $f_4$, with different number of workers and populations sizes }
+%    \label{fig:spworker:ratio}
+%\end{figure}
 

### 2020-05-10 draft.tex
Revising section for developers, @mariosky another round is going to be needed

@@
 
-In the case of the first level of integration, developers % a decoupler worker what? - JJ
-% I think it was redundant, I was saying that in this level the worker container
-% is not tied to the rest of the containers. - M 
+In the case of the first level of integration, developers
 can implement a new stateless Docker container following
@@
 % Anyway you need to say what is the signature of the function. 
-%Also, if they can use the existing implementation, you need to specify the license - JJ
+% Also, if they can use the existing implementation, you need to specify the license - JJ u
+% Any reaction to this? - JJ
 
@@
 population because, in this case, the algorithm needs to start with the
-population provided as a parameter. Depending on the library used, % Depending on which library? - JJ 
+population provided as a parameter. Depending on the evolutionary (or other) algorithm library used,
 it may be required to extract the population from the
-JSON message and replace the population object created by the library. For
+JSON message and replace the population object created by the default library implementation. For
 example, the DEAP library we used for the GA algorithm, creates the type of the
@@
 For the secondary level of integration, developers need to add code to the setup
-container to specify additional configuration options. For instance,  % related to? What do you mean here?
+container %maybe you should specify more, at the beginning of the section, what is the current setup, what containers are there, what libraries are being used, and so no. - JJ 
+to specify additional configuration options. For instance,  % related to? What do you mean here?
 in the current configuration message (see Listing \ref{code:config})
@@
 between them (i.e., ring or hypercube). The controller
-can apply these rules selectively only to swarm-based algorithms.
+can apply these rules selectively only to swarm-based algorithms. In general, these configuration options will be algorithm and library specific.
 
 On either level of integration, algorithm developers can upload the container
-definition or images directly to a public repository like DockerHub. Another
+definition or images directly to a public repository like DockerHub. % Clarify what you need to change in those images. Will you need to just rebuild the image with new code, or something else? - JJ
+Another
 option is to upload the container image definition (Dockerfile) to GitHub and
@@
 images is a requirement, as it is the case of Amazon ECS.
+
+% Besides addressing the concerns of a reviewer, you should mention how this is related to the overall objective of the paper, saying this contributes to reproducibility or to be able to do it on the cloud  - Jj
 

### 2020-05-11 draft.tex
Update draft.tex
Update developer  point of view

@@
 % (and he next one too) - JJ
+% done - Mario 
 \section{Container-based application} 
@@
 % description  of the containers used be enough? - JJ 
-% Ok, I thin only workers and the controller need to be described in detail - Mario 
+% Ok, I think only workers and the controller need to be described in detail - Mario 
 \subsection{Message Queue Container} 
@@
 % Any reaction to this? - JJ
+% I forgot to put done, but I added in the above paragraphs. It was correct? - Mario
 
@@
 For the secondary level of integration, developers need to add code to the setup
-container %maybe you should specify more, at the beginning of the section, what is the current setup, what containers are there, what libraries are being used, and so no. - JJ 
-to specify additional configuration options. For instance,  % related to? What do you mean here?
+
+container %maybe you should specify more, at the beginning of the section, 
+% what is the current setup, what containers are there, what libraries are being used, and so on. - JJ
+% The previous section talks about, can you have look? Maybe I should move it to this one. - Mario
+% If not, I will add a remainder, something like: following from the setup described in the previous section... - M  
+to specify additional configuration options. For instance,  % related to? What do you mean here? - done Mario 
 in the current configuration message (see Listing \ref{code:config})
@@
 between them (i.e., ring or hypercube). The controller
-can apply these rules selectively only to swarm-based algorithms. In general, these configuration options will be algorithm and library specific.
-
-On either level of integration, algorithm developers can upload the container
-definition or images directly to a public repository like DockerHub. % Clarify what you need to change in those images. Will you need to just rebuild the image with new code, or something else? - JJ
-Another
-option is to upload the container image definition (Dockerfile) to GitHub and
+can apply these rules selectively only to swarm-based algorithms. 
+In general, these configuration options will be algorithm and library specific.
+
+On either level of integration and after making the changes outlined before,
+algorithm developers can upload the container 
+definition or images directly to a public repository like DockerHub. 
+% Clarify what you need to change in those images. 
+% Added a brief comment at the beginning  - M
+% Will you need to 
+% just rebuild the image with new code, or something else? - JJ
+% That and document your work, talk about this at the end - M
+% We need to document our own images :) - M
+
+Another option is to upload the container image definition (Dockerfile) to GitHub and
 define a trigger to re-upload the image to the registry after an update. Having
@@
 
-% Besides addressing the concerns of a reviewer, you should mention how this is related to the overall objective of the paper, saying this contributes to reproducibility or to be able to do it on the cloud  - Jj
+% Besides addressing the concerns of a reviewer, you should mention how this is 
+% related to the overall objective of the paper, saying this contributes to
+% reproducibility or to be able to do it on the cloud  - Jj
+% Added the lines bellow - Mario
+
+Moreover, public repositories have the functionality for documenting, sharing,
+and tracking different versions of both images and image definitions by using
+tags. This infrastructure supports the objective of the paper, which is to
+contribute to the reproducibility of experiments and platform independence.
 
@@
 %
-% First you have to say that cloud is a way of describing deployment to any kind of infrastructure in a reproducible way, and it needn't be in the cloud proper, it can be deployed to anything, that's precisely the thing.
-But make clear that when you use a cloud provider, there are more options to do it...
-- JJ
+
+% First you have to say that cloud is a way of describing deployment to any kind
+% of infrastructure in a reproducible way, and it needn't be in the cloud proper,
+% it can be deployed to anything, that's precisely the thing. But make clear that
+% when you use a cloud provider, there are more options to do it... - JJ
+
 % You should start writing this out soon... Don't leave it this way - JJ

### 2020-05-12 draft.tex
Cloud experiment setup

@@
 }
+
 \graphicspath{ {./images/} }
@@
 number of FEs is $200,000$, which is the same as $40*50*10*10$.
-%
 
@@
 
-% Section headers should be headlines. Like "Measuring scalability in
-% heterogenous systems", or something like that. Experiment 1 says
-% nothing about what's inside (this also applies below) - JJ
 \section{Measuring Scalability}
@@
 shown with a smaller gap in the time won adding more workers.
-%
-% Boxplot of the total number of evaluations for 30 instances of function f4 , with
-% different number of workers and populations sizes
-
-%
-%
-% The effect of a
-% different number of populations on execution time is reported in  Section
-% \ref{sec:exp1results}.
-% The results of the performance comparison between homogeneous and
-% heterogeneous search strategies are presented in Section \ref{sec:exp2results}.     
-%
-%
-
-%
-%
-%\begin{figure}[h!tbp]
-%%    \centering
-%    \includegraphics[width=\textwidth]{evalrate}
-%%    \caption{Boxplot of the number of evaluations per second, for 30 instances of function $f_4$, with different number of workers and populations sizes }
-%    \label{fig:spworker:ratio}
-%\end{figure}
 
@@
 
-% You should start writing this out soon... Don't leave it this way - JJ
-Several options:
-    A virtual machine with a container-optimized operating system, for example,
+Container-based applications describe the deployment of isolated processes in a
+consistent and portable way, going from development, testing, and production
+independently on the kind of infrastructure. The term cloud-native describes
+these container-based environments because they are deployed on an
+infrastructure that abstracts the underlying compute, storage, and networking
+primitives. This infrastructure does need to be a cloud provider; it could be a
+developer's laptop, a single server, or a production environment. When we deploy
+a cloud-native application to a cloud provider, we have more options to choose:
+
+\begin{description}
+
+  \item[A virtual machine] with a container-optimized operating system, for example,
   Google Compute Engine VMs, or Amazon EC2.  We still need to provision the
   virtual machine and run docker-compose commands as we would do in a local 
-      workstation. % Remember that Salza et al. used CoreOS, which is optimized for containers - JJ
-
-    A Kubernetes cluster, this provides auto-scaling and management of containerized
-    application, but it needs additional configuration for provisioning. All major
+  workstation. Salza et al. \cite{salza2019speed} tested this option by using
+  CoreOs cluster. 
+  % Remember that Salza et al. used CoreOS, which is optimized for containers - JJ
+
+  \item[A Kubernetes cluster] provides auto-scaling and management of containerized
+  applications, but it needs additional configuration for provisioning. All major
   cloud-providers have Kubernetes services.
 
-    A container orchestration service, these services are compatible with docker-compose
+  \item[A container orchestration service] these services are compatible with docker-compose
     files for configuration and provide transparent provisioning. Examples of these
@@
 
+\end{description}
+
 % This is the second option. The first option was already tested, right? - JJ
-As a proof-of-concept, we deployed the EvoSwarm application in Amazon Elastic
-Container Service (ECS),  since it needs fewer changes in configuration (from deploying it locally using docker-compose)
-and it is has a lower cost than a Kubernetes alternative.
+% This is the third option.  - M
+
+As a proof-of-concept, we chose the third option and deployed the EvoSwarm application in Amazon Elastic
+Container Service (ECS),  since it needs fewer changes in configuration 
+(from deploying it locally using docker-compose) and it is has a lower cost than a Kubernetes alternative.
 
@@
 
-Using the ECS CLI, we can provision a cluster with a configuration file
-specifying the computing resources needed. Multicontainer applications are
+Multicontainer applications are 
 % Maybe an example - JJ
+% This is shown next - Mario
 deployed to ECS as tasks, and a single cluster can run multiple tasks in
@@
 
-We deployed an AWS Fargate cluster and defined a Task configuration file with
+Using the ECS CLI, we can provision a cluster with a configuration file
+specifying the computing resources needed. We can specify the configuration a Task 
+by using a task definition file (see Listing~\ref{code:ecs}) in YAML format.
+The main attributes are the {\texttt task\_size} indicating the memory and CPU units
+assigned, and network configuration. The complete configuration of the task is generated
+by the {\texttt compose command} that uses a {\texttt docker-compose.yml} to generate 
+the configuration on-the-fly. For example to deployed the controller, setup and redis 
+containers we used the file in Listing ~\ref{code:compose}.
+
+We deployed two configurations, one that included all containers in a single Task,
+and another one, where we assigned each worker to a single Task and the controller, 
+redis, and setup in another independent Task.
+
+For the first option, we deployed an AWS Fargate cluster and defined a Task configuration file with
 8GBs of memory and 4096 computing units, equivalent to 4 CPUs. Because of the
-ten container limit, we only used seven workers and tested the same function as
-in the previous experiment, but save resources only for dimensions 10 and 20.
-
-The results ..
+ten container limit, we only used seven workers and tested function $f_4$ as
+in the previous experiment, and to save resources we only tested for dimensions 10 and 20.
+
+For the second option we assigned each worker to a single task with 4GBs of memory and 2048 computing units,
+and scale the number of worker Tasks to eight.
+
+\begin{lstlisting}[caption=ecs-params.yml , label=code:ecs]
+version: 1
+task_definition:
+  task_execution_role: ecsTaskExecutionRole
+  ecs_network_mode: awsvpc
+  task_size:
+    mem_limit: 8GB
+    cpu_limit: 4096
+run_params:
+  network_configuration:
+    awsvpc_configuration:
+      subnets:
+        - "subnet-0ca2a542bbd74239e"
+        - "subnet-0dfd367b5ca531b07"
+      security_groups:
+        - "sg-0681c18469874b391"
+      assign_public_ip: ENABLED
+\end{lstlisting}
+
+\begin{lstlisting}[caption={docker-compose.yml for setup, controller and redis containers}, label=code:compose]
+  version: '3'
+  services:
+    redis:
+      image: "redis:alpine"
+      ports:
+        - "6379:6379"
+      logging:
+        driver: awslogs
+        options:
+          awslogs-region: us-west-2
+          awslogs-group: docker-logs
+          awslogs-stream-prefix: redis
+
+    setup:
+      image: mariosky/setup
+      command: python experiment.py
+      volumes:
+        - /code
+      environment:
+        PYTHONUNBUFFERED: 1
+        REDIS_HOST: localhost
+      logging:
+        driver: awslogs
+        options:
+          awslogs-region: us-west-2
+          awslogs-group: docker-logs
+          awslogs-stream-prefix: setup
+
+    controller:
+      image: mariosky/controller
+      volumes:
+          - /code
+      command: python controller.py
+      environment:
+          PYTHONUNBUFFERED: 1
+          REDIS_HOST: localhost
+      logging:
+        driver: awslogs
+        options:
+          awslogs-region: us-west-2
+          awslogs-group: docker-logs
+          awslogs-stream-prefix: controller
+  \end{lstlisting}
+
+The results are shown in Table X.
 

### 2020-05-12 draft.tex
OK, revised, this is done

@@
 % done - Mario
+% But also change title? - JJ
 \section{Container-based application} 
@@
 All the code of the EvoSwarm Application is published with the open source MIT license.
-% say which function, what's its API, how can you test it - JJ
-% this is not an API the need to write their own - Mario
-% We can write an API later - M
-% Anyway you need to say what is the signature of the function. 
-% Also, if they can use the existing implementation, you need to specify the license - JJ u
-% Any reaction to this? - JJ
-% I forgot to put done, but I added in the above paragraphs. It was correct? - Mario
 
@@
 
-container %maybe you should specify more, at the beginning of the section, 
-% what is the current setup, what containers are there, what libraries are being used, and so on. - JJ
-% The previous section talks about, can you have look? Maybe I should move it to this one. - Mario
-% If not, I will add a remainder, something like: following from the setup described in the previous section... - M  
-to specify additional configuration options. For instance,  % related to? What do you mean here? - done Mario 
+container
 in the current configuration message (see Listing \ref{code:config})
@@
 definition or images directly to a public repository like DockerHub.
-% Clarify what you need to change in those images. 
-% Added a brief comment at the beginning  - M
-% Will you need to 
-% just rebuild the image with new code, or something else? - JJ
-% That and document your work, talk about this at the end - M
-% We need to document our own images :) - M
 
@@
 images is a requirement, as it is the case of Amazon ECS.
-
-% Besides addressing the concerns of a reviewer, you should mention how this is 
-% related to the overall objective of the paper, saying this contributes to
-% reproducibility or to be able to do it on the cloud  - Jj
-% Added the lines bellow - Mario
 

### 2020-05-12 draft.tex
REvised cloud deployment section

@@
 
-Container-based applications describe the deployment of isolated processes in a
-consistent and portable way, going from development, testing, and production
+Container-based applications describe the deployment of loosely coupled processes in a
+consistent and portable way, in every phase of the life cycle of an application, from development through testing to production
 independently on the kind of infrastructure. The term cloud-native describes
@@
 developer's laptop, a single server, or a production environment. When we deploy
-a cloud-native application to a cloud provider, we have more options to choose:
+a cloud-native application to a cloud provider, we have some more options to choose:
 
@@
 Container Service (ECS),  since it needs fewer changes in configuration 
-(from deploying it locally using docker-compose) and it is has a lower cost than a Kubernetes alternative.
-
-We can define an ECS cluster directly on AWS Elastic Cloud VMs or through AWS Fargate.
+(from deploying it locally using docker-compose) and it is has a lower cost than the Kubernetes alternative. % Remember to mention cost at the end - JJ
+
+We can define an ECS cluster directly on AWS Elastic Cloud VMs or through the AWS Fargate service.
 Fargate is a serverless compute engine for containers that provides automatic provision 
@@
 
-Multicontainer applications are 
-% Maybe an example - JJ
-% This is shown next - Mario
-deployed to ECS as tasks, and a single cluster can run multiple tasks in
+Multicontainer applications are deployed to ECS as tasks, and a single cluster can run multiple tasks in
 parallel. When using the Fargate service as a backend, tasks are limited to a
@@
 Using the ECS CLI, we can provision a cluster with a configuration file
-specifying the computing resources needed. We can specify the configuration a Task 
+specifying the computing resources needed. We can specify the configuration of a Task 
 by using a task definition file (see Listing~\ref{code:ecs}) in YAML format.
@@
 assigned, and network configuration. The complete configuration of the task is generated
-by the {\texttt compose command} that uses a {\texttt docker-compose.yml} to generate 
+by the {\texttt compose command} that uses a {\texttt docker-compose.yml} file to generate 
 the configuration on-the-fly. For example to deployed the controller, setup and redis 
@@
 For the second option we assigned each worker to a single task with 4GBs of memory and 2048 computing units,
-and scale the number of worker Tasks to eight.
+and scale the number of worker tasks to eight.  % Please use caps or not for Tasks in a consistent way - JJ
 
@@
 
+% Please clarify that those images are published in Docker hub by yourself - JJ
 The results are shown in Table X.

### 2020-05-13 draft.tex
Cloud and BBOB added
Experiments 2 and 3 added. Refs #6 and  #9.

@@
 \usepackage{xcolor}
+\usepackage[normalem]{ulem}
+\useunder{\uline}{\ul}{}
 
@@
 For local experiments we have deployed the EvoSwarm application in a
-Desktop PC with an AMD Ryzen 9 3900x 12-core processor with 24 threads 
+Desktop PC with an AMD Ryzen 9 3900x 12-core processor with 24 threads with a 3.8 GHz base clock, 
 and 16 GB RAM. We used Docker version 19.03.3, build a872fc2f86, and docker-compose version 1.21.0, 
@@
 
+\begin{figure*}[h!tb]
+  \begin{tabular}
+      {c@{\hspace*{-0.00001\textwidth}}
+       c@{\hspace*{-0.00001\textwidth}}
+       c@{\hspace*{-0.00001\textwidth}}
+      }
+  GA  &  PSO & GA \& PSO\\   
+  \includegraphics[width=0.30\textwidth]{GAOnly_f001}&
+  \includegraphics[width=0.30\textwidth]{PSOOnly_f001}&
+  \includegraphics[width=0.30\textwidth]{GAPSO_f001}\\
+
+  \includegraphics[width=0.30\textwidth]{GAOnly_f002}&
+  \includegraphics[width=0.30\textwidth]{PSOOnly_f002}&
+  \includegraphics[width=0.30\textwidth]{GAPSO_f002}\\
+
+  \includegraphics[width=0.30\textwidth]{GAOnly_f003}&
+  \includegraphics[width=0.30\textwidth]{PSOOnly_f003}&
+  \includegraphics[width=0.30\textwidth]{GAPSO_f003}\\
+
+  \includegraphics[width=0.30\textwidth]{GAOnly_f004}&
+  \includegraphics[width=0.30\textwidth]{PSOOnly_f004}&
+  \includegraphics[width=0.30\textwidth]{GAPSO_f004}\\
+
+  \includegraphics[width=0.30\textwidth]{GAOnly_f005}&
+  \includegraphics[width=0.30\textwidth]{PSOOnly_f005}&
+  \includegraphics[width=0.30\textwidth]{GAPSO_f005}\\
+  \end{tabular}
+  \vspace{-3ex}
+   \caption{
+Scaling of running time with problem dimension to reach certain target values $\Delta f$. Lines:
+average runtime (aRT); Cross ($+$): median runtime of successful runs to reach
+the most difficult target that was reached at least once (but not always);
+Cross (\textcolor{red}{$\times$}): maximum number of f-evaluations in any trial. Notched boxes:
+interquartile range with median of simulated runs; all values are divided by
+dimension and plotted as $log_{10}$ values versus dimension. 
+}
+\label{fig:bbob}
+\end{figure*}
+
 \subsection{Setup}
 
-% % Experiment 1
-
-% Having a multi-population based algorithm can decrease the execution time
-% because of the parallel execution of the evolution. But, having heterogeneous 
-% populations might enhance evolutionary search and needless
-% evaluations than homogeneous systems; heterogeneous settings, if
-% done right, increases the diversity of the whole population
-% \cite{araujo2008multikulti}.
-
-% But this is a rule of thumb, and it will depend on the degree of heterogeneity,
-% as well as on the algorithm itself. Some level of heterogeneity can be
-% implemented by just changing the configuration parameters of each population,
-% but in this case, we are interested in a heterogeneous search strategy. 
-% Therefore, in this experiment we compare a multi-population with only GA and PSO populations,
-% versus an ensemble of GA and PSO algorithms. We tested on the first five functions of the
-% BBOB testbed. We used ten populations and eight workers for the experiment and the
-% same parameters as before.
-
-% Shouldn't this be a subsection of experimental setup? - JJ
-% \subsection{Hardware}
-% \label{sec:hardware}
+Having a multi-population based algorithm can decrease the execution time
+because of the parallel execution of the evolution. But, having heterogeneous 
+populations might enhance evolutionary search and needless
+evaluations than homogeneous systems; heterogeneous settings, if
+done right, increases the diversity of the whole population\cite{araujo2008multikulti}.
+
+But this is a rule of thumb, and it will depend on the degree of heterogeneity,
+as well as on the algorithm itself. Some level of heterogeneity can be
+implemented by just changing the configuration parameters of each population,
+as in the previous experiment, but in this case, we are interested in a heterogeneous search strategy.
+
+Therefore, in this experiment we compare a multi-population with only GA and PSO populations,
+versus an ensemble of GA and PSO algorithms. We tested on the first five functions of the
+BBOB testbed. We used ten populations and eight workers for the experiment and the
+same parameters as before.
+
+\subsection{Experimental Results}
+
+The results obtained by the experiments show (see Figure \ref{fig:bbob}) how
+the runtime scales with dimension to reach certain target values $\Delta f$;
+The color of the lines indicates the average runtime, for each target reached
+at least one time. A red circle without a number on top indicates the algorithm
+has reached the $10^{-8}$ target on all instances for that dimension. If there
+is a number on the top, it indicates how many times the algorithm has reached
+the target. Fixed values of targets $\Delta f = 10^{k}$ with $k$ colors are
+given in the legend of $f_1$, $k$ values are $[-8,-5,-3,-2,-1,0,1]$. Results
+from each configuration are presented in each column, GA, PSO, and GA\&PSO
+populations.
+
+We can see that the GA algorithm (on the left column) has a hard time reaching even the
+$10^{-5}$ target in functions ($f_1-f_4$) for all dimensions, and has the worst
+performance in 20 and 40 dimensions. Finally, it performs better on $f_5$
+reaching the $10^{-8}$ on all dimensions but with a higher runtime on lower
+dimensions.
+
+On the other hand, the multi-population PSO (taking the column on the middle) 
+has better performance; its results take a smaller range,
+reaching all targets for functions $f_1$ and $f_2$ but with a slightly higher
+runtime than the multi-strategy configuration. It is not able to reach the
+$10^{-8}$ target on the higher dimensions of $f_3$ and $f_4$. Moreover, the PSO
+multi-population has the best runtime for the lower dimensions of $f_5$.
+
+Finally, the proposed PSO\&GA mixed algorithm has reached the $10^{-8}$ target
+value on all functions ($f_1-f_5$) and scales well to higher dimensions, even on
+$f_3$ and $f_4$ (charts on the lower rows), which are usually considered
+difficult functions. This experiment also exemplifies the type of analyses that
+we can perform with the EvoSwarm application by reusing the COCO framework.
 
@@
 
-Container-based applications describe the deployment of loosely coupled processes in a
-consistent and portable way, in every phase of the life cycle of an application, from development through testing to production
-independently on the kind of infrastructure. The term cloud-native describes
-these container-based environments because they are deployed on an
-infrastructure that abstracts the underlying compute, storage, and networking
-primitives. This infrastructure does need to be a cloud provider; it could be a
-developer's laptop, a single server, or a production environment. When we deploy
-a cloud-native application to a cloud provider, we have some more options to choose:
+Container-based applications describe the deployment of loosely coupled
+processes in a consistent and portable way, in every phase of the life cycle of
+an application, from development through testing to production independently on
+the kind of infrastructure. The term cloud-native describes these
+container-based environments because they are deployed on an infrastructure that
+abstracts the underlying compute, storage, and networking primitives. This
+infrastructure does need to be a cloud provider; it could be a developer's
+laptop, a single server, or a production environment. When we deploy a
+cloud-native application to a cloud provider, we have some more options to
+choose:
 
@@
 Using the ECS CLI, we can provision a cluster with a configuration file
-specifying the computing resources needed. We can specify the configuration of a Task 
+specifying the computing resources needed. We can specify the configuration of a task 
 by using a task definition file (see Listing~\ref{code:ecs}) in YAML format.
 The main attributes are the {\texttt task\_size} indicating the memory and CPU units
-assigned, and network configuration. The complete configuration of the task is generated
-by the {\texttt compose command} that uses a {\texttt docker-compose.yml} file to generate 
-the configuration on-the-fly. For example to deployed the controller, setup and redis 
-containers we used the file in Listing ~\ref{code:compose}.
-
-We deployed two configurations, one that included all containers in a single Task,
-and another one, where we assigned each worker to a single Task and the controller, 
-redis, and setup in another independent Task.
-
-For the first option, we deployed an AWS Fargate cluster and defined a Task configuration file with
+assigned, and network configuration. The complete configuration is generated
+by the {\texttt ecs compose command} that uses a {\texttt docker-compose.yml} file to generate 
+the configuration on-the-fly. For example to deploy the controller, setup and redis 
+containers we used the file in Listing ~\ref{code:compose}. We have 
+provided the container images and are published in Docker hub \url{https://hub.docker.com/u/mariosky}.
+
+We deployed two configurations, one that included all containers in a single task,
+and another one, where we assigned each worker to a different task and the controller, 
+redis, and setup in another, for a total of nine concurrent Fargate tasks.
+
+For the first option, we deployed an AWS Fargate cluster and defined a task configuration file with
 8GBs of memory and 4096 computing units, equivalent to 4 CPUs. Because of the
 ten container limit, we only used seven workers and tested function $f_4$ as
-in the previous experiment, and to save resources we only tested for dimensions 10 and 20.
-
-For the second option we assigned each worker to a single task with 4GBs of memory and 2048 computing units,
-and scale the number of worker tasks to eight.  % Please use caps or not for Tasks in a consistent way - JJ
+in the previous experiment, and to save resources we only tested for dimensions 10 and 20 for 15 instances.
+
+For the second option, we used the same cluster and additionally
+assigned each worker to a task with 4GBs of memory and 2048 computing units. Then scale the 
+number of worker tasks to eight. We can scale the number of tasks for up to 100 instances in 
+a Fargate configuration. 
+                                  % Please use caps or not for Tasks in a consistent way - JJ
+                                  % Done -Mario
+                                  % Done above - M
 
@@
 
+\subsection{Experimental Results}
+
 % Please clarify that those images are published in Docker hub by yourself - JJ
-The results are shown in Table X.
+The results are shown in Figure~\ref{fig:aws-results} and Table~\ref{tab:aws}.
+
+As we can see, the local environment (see section \ref{setup}) achieved a higher
+ratio of evaluations per second (in bold). Still, we must consider that there is a minimum
+network overhead in a local workstation and a clock speed of 3.8 GHz. The first
+option (labeled Single) is straightforward to deploy, needing only a single
+docker-compose file, and one command to execute. But with the limit of 4CPUs and
+ten containers, it is the slower option. The second option (labeled Multi), is
+faster, and not too far from the local execution, especially when solving the
+$f_4$ in 20 dimensions (underline in table). With the capacity for increasing the number of worker
+containers, it is reasonable to think that we can achieve a higher number of
+evaluations with an appropriate configuration. Nevertheless, the goal of this
+experiment was not to increase the number of evaluations but to test the
+flexibility of a cloud-native application.
+
+\begin{figure}[h!tbp]
+  \centering
+  \includegraphics[width=\textwidth]{aws}
+  \caption{Boxplot of the number of evaluations per second ratio, for 15 instances of function $f_4$,
+   with three execution environments Local, Single task and Multi-task.
+   }
+  \label{fig:aws-results}
+\end{figure}
+
+\begin{table}[]
+  \small
+  \caption{ Median number of the evaluations per second ratio, for 15 instances of function $f_4$,
+  with three execution environments Local, Single task and Multi task}
+  \label{tab:aws}
+  \vspace{0.25cm}
+  \centering
+
+  \begin{tabular}{lll}
+  \hline
+  Dim & Env         & Median of ratio \\
+  \hline
+  10  & Local       & \textbf{47,166.29}       \\
+      & Multi task  & 40,466.45    \\
+      & Single task & 18,193.05      \\
+  \hline
+  20  & Local       & \textbf{36,101.86}   \\
+      & Multi task  & {\ul 33,916.02}       \\
+      & Single task & 16,546.95  \\
+  \hline
+  \end{tabular}
+\end{table}
 

### 2020-05-13 draft.tex
Time of execution with statistical test

@@
 of complexity, in terms of execution time,  to effectively scale on multiple
-nodes, tending to linear scalability. In a population-based algorithm like
-EvoSwarm, several parameters can increase or decrease the workload of workers:
+nodes, tending to linear scalability. In that paper, Salza et al. tested scalability
+by simulating the work of nodes by using a sleep function and messages of different sizes.
+In a population-based algorithm like EvoSwarm, several parameters can increase or 
+decrease the workload of workers:
 
@@
 can be better understood, since there are fewer factors involved.
-Following the procedure described in the previous Section, we executed 
+Following the procedure described in the previous section, we executed 
 two sets of experiments with five and ten initial populations,
@@
   \centering
-  \includegraphics[width=\textwidth]{time_wrkr}
+  \includegraphics[width=\textwidth]{time_diff}
   \caption{Boxplot of the median time required for the execution of 30 
@@
               & 2w      & 1.84   & 1.72   \\
-              & 4w      & 1.22   & 0.96   \\
-              & 8w      & 1.05   & 0.59   \\
+              & 4w      & 1.22   & \textbf{0.96}   \\
+              & 8w      & 1.05   & \textbf{0.59}   \\
     \hline
-          3   & 1w      & 6.53   & 7.76   \\
+          3   & 1w      & 6.53   & \textbf{7.76}   \\
               & 2w      & 3.64   & 3.83   \\
               & 4w      & 2.33   & 2.26   \\
-              & 8w      & 2.17   & 1.36   \\
+              & 8w      & 2.17   &  \textbf{1.36}   \\
     \hline
-          5   & 1w      & 12.33  & 14.19  \\
-              & 2w      & 6.63   & 7.41   \\
+          5   & 1w      & \textbf{12.33}  & 14.19  \\
+              & 2w      & 6.63   & \textbf{7.41}   \\
               & 4w      & 4.25   & 4.34   \\
-              & 8w      & 3.67   & 2.41   \\
+              & 8w      & 3.67   &  \textbf{2.41}   \\
     \hline
-          10  & 1w      & 31.50  & 38.30  \\
-              & 2w      & 16.07  & 19.31  \\
+          10  & 1w      & \textbf{31.50}  & 38.30  \\
+              & 2w      & \textbf{16.07}  & 19.31  \\
               & 4w      &  9.99   & 10.35  \\
-              & 8w      & 8.97   & 5.92   \\
+              & 8w      & 8.97   &  \textbf{5.92}  \\
     \hline
           20  & 1w      &  116.35 & 124.51 \\
-              & 2w      & 51.18  & 63.68  \\
+              & 2w      & \textbf{51.18}  & 63.68  \\
               & 4w      & 32.26  & 33.07  \\
-              & 8w      & 28.64  & 18.25  \\
+              & 8w      & 28.64  & \textbf{18.25}  \\
     \hline
-          40  & 1w      & 369.89 & 448.89 \\
-              & 2w      & 198.43 & 234.50 \\
+          40  & 1w      & \textbf{369.89} & 448.89 \\
+              & 2w      & \textbf{198.43} & 234.50 \\
               & 4w      & 122.39 &  120.62 \\
-              & 8w      & 122.31 & 63.35 \\
+              & 8w      & 122.31 & \textbf{63.35} \\
     \hline
@@
 
-With results summarized in Figure \ref{fig:spworker}.  The
-experiment consisted of executing the same benchmark on EvoSwarm twice but
-changing the number of populations used in each run. This is mainly
-intended as a showcase of the flexibility of the tool presented in
-this paper, but we also wanted to test how the number of workers
-affected the time needed to reach a solution. Adding more populations
-might add more overhead, so this might have a noticeable effect.
-The left column in that figure has the
-results when using five populations, and the right side shows the results for
-ten. In principle the main issue is how to find the right number of
-workers for a particular problem, but in general we wanted to check
-the communication overhead when the size of the problem increases and
-how the presented framework reacted to it. % Added some explanation
-% here - JJ
-
-Charts show the execution time achieved by a certain number of
+In this experiment, we wanted to test how the number of workers affected the
+time needed to reach a solution and the speedup that each configuration can
+produce. Adding more populations might add more overhead, so this might have a
+noticeable effect. In principle, the main issue is how to find the right number
+of workers for a particular problem, but in general, we wanted to check the
+communication overhead when the size of the problem increases and how the
+presented framework reacted to it.    % Added some explanation % here - JJ
+We report the comparison regarding the execution time in Section~\ref{sec:exec-time}
+and the summary of the speedup in section Section~ref{sec:speedup}.
+
+\subsubsection{Execution Time}
+\label{sec:exec-time}
+
+Figure \ref{fig:spworker} shows the execution time achieved by a certain number of
 workers, for every problem dimension, which, besides, changes the
 population size.
-Each box-plot indicates the time required for the execution of 15 function
+Each box-plot indicates the time required for the execution of 30 function
 instances, in seconds. Each row has a different scale on the y-axis, to
@@
 
-As expected, and in most cases, increasing the number of workers reduces the time
-required for the execution of an instance, mainly when using the same parameters
+In Table~\ref{tab:time} we can see that as expected, and in most cases, 
+increasing the number of workers reduces the time
+required for the execution of an instance, 
+mainly when using the same parameters
 (number of populations and number of workers), that is, performance
 scales with the number of workers for a fixed problem and number of
-sub-populations used. However, when we compare the two
+sub-populations used. See Table~\ref{tab:speedup:test} for a statistical test 
+on speedup, this explain further in the next section.
+
+However, when we compare the two
 options, we find some subtle differences: When using one or two
-workers, using 5 populations takes less time in most cases. A 
-reason for this could be that the migration of individuals takes more time to
+workers, using 5 populations takes less time in most cases. 
+In Table~\ref{tab:time} results with p-value of the two-tailed Wilcoxon
+signed-ranks test was grater than the level of significance of 0.05 
+(i.e., accepting the null hypothesis of inequality) are marked in boldface.
+A reason for this could be that the migration of individuals takes more time to
 reach the worker. For instance, when a single worker has 10 populations, only
@@
 
-Another additional issue is related to the algorithm itself. Many more
-workers than populations will have a wider range of states of
-evolution in the experiment; populations that will be mixed will have
-very different fitness states, which means that a part of them will
-probably have been not so useful, following the intermediate
-disturbance hypothesis. When the number of workers is increased, the
-difference between the state of evolution of populations will be
-intermediate, leading to better results; eventually, when there are as
-many workers as populations, the state will be quite similar, which is
-shown with a smaller gap in the time won adding more workers.
+% I think this is not the case? 
+% Another additional issue is related to the algorithm itself. Many more
+% workers than populations will have a wider range of states of
+% evolution in the experiment; populations that will be mixed will have
+% very different fitness states, which means that a part of them will
+% probably have been not so useful, following the intermediate
+% disturbance hypothesis. When the number of workers is increased, the
+% difference between the state of evolution of populations will be
+% intermediate, leading to better results; eventually, when there are as
+% many workers as populations, the state will be quite similar, which is
+% shown with a smaller gap in the time won adding more workers.
+
+\subsubsection{Speedup}
+\label{sec:speedup}
 

### 2020-05-13 draft.tex
Correct images

@@
 scales with the number of workers for a fixed problem and number of
-sub-populations used. See Table~\ref{tab:speedup:test} for a statistical test 
-on speedup, this explain further in the next section.
+sub-populations used. The exception is the case when using five populations;
+the time does not decrease when progressing from four to eight workers.  
+That is because most of the time, workers are waiting for a population 
+to be available in the message queue.
 

### 2020-05-13 draft.tex
Starting Results of Speed

--- 2020-05-13 draft.tex
@@
   \centering
-  \includegraphics[width=\textwidth]{time_diff}
+  \includegraphics[width=\textwidth]{time_wrkr}
   \caption{Boxplot of the median time required for the execution of 30 
@@
     \hline
-          2   & 1w      & 2.96   & 3.15   \\
-              & 2w      & 1.84   & 1.72   \\
+          2   & 1w      & \textbf{2.96}   & 3.15   \\
+              & 2w      & 1.84   & \textbf{1.72}   \\
               & 4w      & 1.22   & \textbf{0.96}   \\
@@
     \hline
-          3   & 1w      & 6.53   & \textbf{7.76}   \\
+          3   & 1w      & \textbf{6.53}   & 7.76   \\
               & 2w      & 3.64   & 3.83   \\
@@
           5   & 1w      & \textbf{12.33}  & 14.19  \\
-              & 2w      & 6.63   & \textbf{7.41}   \\
+              & 2w      & \textbf{6.63}   & 7.41   \\
               & 4w      & 4.25   & 4.34   \\
@@
               & 2w      & \textbf{16.07}  & 19.31  \\
-              & 4w      &  9.99   & 10.35  \\
+              & 4w      & \textbf{9.99}   & 10.35  \\
               & 8w      & 8.97   & \textbf{5.92}   \\
     \hline
-          20  & 1w      &  116.35 & 124.51 \\
+          20  & 1w      & \textbf{116.35} & 124.51 \\
               & 2w      & \textbf{51.18}  & 63.68  \\
@@
               & 2w      & \textbf{198.43} & 234.50 \\
-              & 4w      & 122.39 &  120.62 \\
+              & 4w      & 122.39 & \textbf{120.62} \\
               & 8w      & 122.31 & \textbf{63.35} \\
@@
 increasing the number of workers reduces the time
-required for the execution of an instance, 
-mainly when using the same parameters
+required for the execution of an instance (shown in boldface), mainly when using the same parameters
 (number of populations and number of workers), that is, performance
 scales with the number of workers for a fixed problem and number of
-sub-populations used. See Table~\ref{tab:speedup:test} for a statistical test 
-on speedup, this explain further in the next section.
+sub-populations used.
 
@@
 workers, using 5 populations takes less time in most cases. 
-In Table~\ref{tab:time} results with p-value of the two-tailed Wilcoxon
-signed-ranks test was grater than the level of significance of 0.05 
-(i.e., accepting the null hypothesis of inequality) are marked in boldface.
-A reason for this could be that the migration of individuals takes more time to
+Results 
+A 
+reason for this could be that the migration of individuals takes more time to
 reach the worker. For instance, when a single worker has 10 populations, only
@@
 
-% I think this is not the case? 
-% Another additional issue is related to the algorithm itself. Many more
-% workers than populations will have a wider range of states of
-% evolution in the experiment; populations that will be mixed will have
-% very different fitness states, which means that a part of them will
-% probably have been not so useful, following the intermediate
-% disturbance hypothesis. When the number of workers is increased, the
-% difference between the state of evolution of populations will be
-% intermediate, leading to better results; eventually, when there are as
-% many workers as populations, the state will be quite similar, which is
-% shown with a smaller gap in the time won adding more workers.
+Another additional issue is related to the algorithm itself. Many more
+workers than populations will have a wider range of states of
+evolution in the experiment; populations that will be mixed will have
+very different fitness states, which means that a part of them will
+probably have been not so useful, following the intermediate
+disturbance hypothesis. When the number of workers is increased, the
+difference between the state of evolution of populations will be
+intermediate, leading to better results; eventually, when there are as
+many workers as populations, the state will be quite similar, which is
+shown with a smaller gap in the time won adding more workers.
 

### 2020-05-14 draft.tex
Some revision requests here and there

--- 2020-05-13 draft.tex
@@
 dimension and plotted as $log_{10}$ values versus dimension.
+% More explanation for those not familiar with BBOB - JJ
 }
@@
 
-Having a multi-population based algorithm can decrease the execution time
-because of the parallel execution of the evolution. But, having heterogeneous 
-populations might enhance evolutionary search and needless
-evaluations than homogeneous systems; heterogeneous settings, if
+A multi-population based algorithm can decrease the execution time
+because of parallel execution. But, having heterogeneous 
+populations might enhance evolutionary search and avoid unneeded
+evaluations more than homogeneous systems systems do; heterogeneous settings, if
 done right, increases the diversity of the whole population\cite{araujo2008multikulti}.
@@
 
-Therefore, in this experiment we compare a multi-population with only GA and PSO populations,
-versus an ensemble of GA and PSO algorithms. We tested on the first five functions of the
-BBOB testbed. We used ten populations and eight workers for the experiment and the
+Therefore, in this experiment we compare a multi-population with  GA- and PSO- only populations,
+versus an ensemble of GA and PSO algorithms. We have tested on the first five functions of the
+BBOB testbed; we have used ten populations and eight workers for the experiment and the
 same parameters as before.
@@
 The color of the lines indicates the average runtime, for each target reached
-at least one time. A red circle without a number on top indicates the algorithm
+at least once. A red circle without a number on top indicates the algorithm
 has reached the $10^{-8}$ target on all instances for that dimension. If there
-is a number on the top, it indicates how many times the algorithm has reached
+is that number, it indicates how many times the algorithm has reached
 the target. Fixed values of targets $\Delta f = 10^{k}$ with $k$ colors are
@@
 from each configuration are presented in each column, GA, PSO, and GA\&PSO
-populations.
-
-We can see that the GA algorithm (on the left column) has a hard time reaching even the
+populations. % clarify which colors for every level - JJ
+
+We can see that the GA algorithm (on the leftmost column) has a hard time reaching even the
 $10^{-5}$ target in functions ($f_1-f_4$) for all dimensions, and has the worst
@@
 On the other hand, the multi-population PSO (taking the column on the middle) 
-has better performance; its results take a smaller range,
+has better performance; its results are spread over a smaller range,
 reaching all targets for functions $f_1$ and $f_2$ but with a slightly higher
@@
 difficult functions. This experiment also exemplifies the type of analyses that
-we can perform with the EvoSwarm application by reusing the COCO framework.
+we can perform with the results of the EvoSwarm application by reusing the COCO framework.
+% If this is what you want to prove, it's not a very strong point.
+Maybe it would be stronger if you proved that these mixed algorithms can be tested in the same conditions as single-algorithm, and also that these mixed algorithms work better than single-algorithm versions - JJ
 
@@
   CoreOs cluster.
-  % Remember that Salza et al. used CoreOS, which is optimized for containers - JJ
 
@@
 
-% This is the second option. The first option was already tested, right? - JJ
-% This is the third option.  - M
-
-As a proof-of-concept, we chose the third option and deployed the EvoSwarm application in Amazon Elastic
+As a proof-of-concept, we chose the third option, that is, a container orchestration service, and deployed the EvoSwarm application in Amazon Elastic
 Container Service (ECS),  since it needs fewer changes in configuration 
@@
 a Fargate configuration.
-                                  % Please use caps or not for Tasks in a consistent way - JJ
-                                  % Done -Mario
-                                  % Done above - M
 
@@
 \end{lstlisting}
+% Do you need the details about the subnets? As long as they are here, you'll have to explain them. Why two subnets? It's probably better to leave that network configuration out anyway - JJ
+
+% Please clarify that those images are published in Docker hub by yourself - JJ
 
@@
   \end{lstlisting}
+% You will probably need to explain this a bit - JJ
 
@@
 
-% Please clarify that those images are published in Docker hub by yourself - JJ
 The results are shown in Figure~\ref{fig:aws-results} and Table~\ref{tab:aws}.

### 2020-05-14 draft.tex
Continuing with the revision of intro

@@
 
-Moreover, parallel implementations can run in two ways: synchronous and
-asynchronous. When operations run in parallel but must maintain synchrony with
+% Probably it's not the time for this, but maybe it would be interesting to mention that asynchronous implementations in most cases need a change in the underlying algorithm, while synchronous might not need it - JJ
+Moreover, parallel implementations can run in two different ways: synchronous and
+asynchronous. When operations run in parallel but every node must maintain synchrony with
 the other operations,  all operations need to finish to continue to the next
@@
 each other, so there's no time lost waiting for other populations to
-arrive to the synchronization point. Following the previous example, now the master operation can
+arrive to the synchronization point; results arriving from other workers are incorporated as soon as they are received. Following the previous example, now the master operation can
 continue to the next iteration, even if a worker has not finished its work.  In
@@
 
-The trend concerning the parallel execution multi-population algorithms go from
+The trend concerning the parallel execution multi-population algorithms goes from
 earlier hardware-based implementations using transputers
-\cite{gorges1990explicit}, multi-threaded \cite{merelo2019scaling}, and
-multi-core systems \cite{Serrano2008,lai2019adaptive}. % many citations here - JJ 
+\cite{gorges1990explicit} through 
+multi-core systems \cite{Serrano2008,lai2019adaptive} to multi-threaded \cite{merelo2019scaling} ones.
 And recently the focus has been on exploiting a higher number of processing
 units by using GPUs \cite{tan2015survey,li2007efficient}, or distributed
-systems; techniques used to achieve this include web-based \cite{JSON},
-map-reduce \cite{fazenda2012},  grid \cite{munawar2010design,Gonzalez09},
-voluntary computing \cite{MilkyWay}, and, more importantly, cloud computing
-\cite{GValdez2015,salza2019speed,valenzuela2015implementing,FlexGP}.
-
-Cloud computing has become the standard way of running
+systems, including web-based \cite{JSON} systems,
+map-reduce \cite{fazenda2012} implementations,  grids \cite{munawar2010design,Gonzalez09},
+voluntary computing systems \cite{MilkyWay}, % Maybe add one of our own voluntary computing references? - JJ
+and, more importantly, cloud computing
+\cite{GValdez2015,salza2019speed,valenzuela2015implementing,FlexGP}. Cloud computing has become the standard way of running
 enterprise applications, not only because of the convenience of the
-pay-as-you-go model or the non-existent cost of physical facilities, but also because
+pay-as-you-go model or the non-existent sunk cost of physical facilities, but also because
 it offers a way of describing the infrastructure as part of the code, so that it
@@
 computing.  However,  cloud computing has also been evolving, going from simply
-putting using virtual machines elsewhere than on premises old-style monolithic applications, that is, applications
+putting on virtual machines elsewhere than on premises old-style monolithic applications, that is, applications
 built on a single stack of services that communicate by layers, to microservices
 \cite{microservices} and serverless architectures \cite{varghese2018next,Varghese2018849} that favor the parallel and asynchronous communication of
-heterogeneous resources; in the process, a new process for designing {\em cloud
-native applications} has been created, with brand new methodologies and
+heterogeneous resources; in the process, a new methodology for designing {\em cloud
+native applications} has been created, with brand new methods and
 techniques \cite{Baldini2016287}. In these architectures, services or even
-functions are seen as single processing points, departing from the monolithic
-and even distributed paradigms, become a loosely coupled collection of {\em
+functions are seen as independent processing nodes, departing from the monolithic
+and even distributed paradigms, and becoming a loosely coupled collection of {\em
 stateless functions} \cite{malawski2017serverless}, that react to events and
@@
 properties of a reactive system \cite{boner2014reactive} and they are commonly
-more flexible, loosely-coupled, and scalable. % Some properies and reference - Mario  
+more flexible, fault-tolerant, and scalable. % Some properies and reference - Mario  
 % Again, you have to justify the need for a reactive architecture and define it before this paragraph - JJ
 
-When designing efficient multi-population algorithms, researchers need to
-consider additional issues \cite{Ma2019}. These include the number and size of populations,
+% There's a gap here. You've finished the previous paragraph talking about the cloud, and now you talk about something else. I'll make some intro - JJ
+Beyond the choice of design methodology and implementation, researchers need to
+consider additional issues \cite{Ma2019} when designing efficient multi-population algorithms. These include the number and size of populations,
 the interaction between them, the search area of each population, and the search

### 2020-05-14 draft.tex
REvised introduction
Please @mariosky check it carefully; we should ensure that by adding new sections we have not drifted away the original narrative of the paper. We certainly have from the last paragraph of the introduction

@@
 
-% There's a gap here. You've finished the previous paragraph talking about the cloud, and now you talk about something else. I'll make some intro - JJ
-Beyond the choice of design methodology and implementation, researchers need to
-consider additional issues \cite{Ma2019} when designing efficient multi-population algorithms. These include the number and size of populations,
-the interaction between them, the search area of each population, and the search
-strategy and parametrization of each population. We must take into account
-these high-level issues when designing a parallel architecture. Researchers 
+Researchers 
 and algorithm designers willing to  adapt to the cloud their
@@
 
-    \item To scale the system, when it is possible, they must use stateless functions for
+    \item To scale the system, when it is possible, there are several options, including the  use of stateless functions for
     processing.
@@
 
+% I have moved these sentences here, since they seem to fit the best.
+Beyond the choice of design methodology and implementation, researchers need to
+consider additional issues \cite{Ma2019} when designing efficient multi-population algorithms. These include the number and size of populations,
+the interaction between them, the search area of each population, and the search
+strategy and parametrization of each population. We must take into account
+these high-level issues when designing a parallel architecture.
+
 Taking these factors into account, in this work, we present EvoSwarm, 
@@
 isolated and
-% Maybe container-based? You can create containers with podman - JJ
-% yes, and is shorter - M
 heterogeneous populations. The platform uses the Docker Compose tool for
@@
 Docker container images host the stateless functions responsible for running 
-a search strategy on several populations, these functions could also be ported
+a search strategy on several populations, % clarify "on several populations" Sequentially? At the same time? - JJ
+these functions could also be ported
 directly to a Function as a Service (FaaS) \cite{Roberts2016} implementation.
@@
 % and extracting a Model? we can put something on that too.
-
+% yep.
+
+% Maybe here you should make reference to the sections where you are actually doing it (after conveniently rephrasing it) - JJ
 \begin{itemize}
@@
     % achieve the contribution - JJ
+    % Maybe you should check this out again ... - JJ
     \item Second, we propose a method for the deployment and execution of 
@@
 our work. In Section \ref{method}, we present the proposed method and the container-based
-application in Section \ref{docker} .  Section \ref{setup} describes
+application in Section \ref{docker} . % No reference to section 6.
+Section \ref{setup} describes
 the design of the empirical evaluation we designed to assess the effectiveness
-of the method, and in Section \ref{results}, we report and discuss the results. Finally, we
+of the method, and in Section \ref{results}, % This is a subsection, fix it. 
+we report and discuss the results. % No references to sections 9 and 10
+Finally, we
 offer the conclusions of this paper and suggestions on future work in Section \ref{conclusions}.

### 2020-05-14 draft.tex
Comments on general structure of the paper, as well as some more comments

@@
   \title{EvoSwarm: An Event-based Architecture for Multi-population Optimization Algorithms}
+  % maybe introduce "reproducible" here, somewhere, as well as "cloud native"
 
@@
 % But also change title? - JJ
+% Experimental design of a container-based application or something like that - JJ
 \section{Container-based application} 
@@
 multi-population based algorithms, but there are in our experience several
-practical issues of this approach. In general, cloud platforms have free-tiers
+practical issues of this approach. % several issues that make it practical or useful instead? - JJ
+In general, cloud platforms have free-tiers
 for their services, and many offer academic discounts to students. But once we
@@
 clear and idempotent description of the architecture through docker
-compose are the main reasons.
+compose % Be consistent with docker compose. Either {\tt docker-compose} or {\sf docker-compose}, but not docker compose, for instance. - JJ
+are the main reasons.
 
@@
 we can go back to Figure \ref{fig:kafkEO} and use it as a guide for explaining
-the main components and their interactions.
+the main components and their interactions. The containers used in this design will be explained next.
 
@@
 
+% You can't start with this.
+Explain what this section is about, and link it with the previous one.
+Maybe it's better to create a big section with "Experimental setup", and make this a subsection, with subsections becoming subsubsections - JJ
 Figure \ref{fig:experiment_flow} shows the general workflow to run experiments
@@
 
+% You said "subsections will talk about roles", this is not a role. Change the title to that of a role. Would that be administrator? - JJ
+
 \subsection{EvoSwarm Configuration}
@@
 
+% Also, it's a bit short. Maybe merge it into the next? - JJ
+
+% Again, not a role - JJ
 \subsection{Experiment Execution}
@@
 
+% This is totally unexpected. Whose role is this? You didn't even introduce it. Also, extremely short - JJ
 \subsection{Using the COCO Framework}
@@
 
-%
+% You should clarify that, before, you were explaining how to set up an experiment.
+In this section, you are explaining the experiments that you are doing in this paper.
+But you should relate both sections to the general objectives of the paper.
+In the prior section, you want to prove that EvoSwarm can be used by the public to create reproductive research (and maybe you should add a small conclusion sub or sub-sub section), in the next you explain the experiment we are doing for this paper.
+It would be interesting to make references from the next section to the previous one, indicating how the general setup of EvoSwarm helps _us_ to create this experiment - JJ
 
@@
             % The baseline is a single population GA or PSO? - M
+% For instance. But I was actually looking for another word that does not imply a comparison - JJ
 Hence, in the following sections, we verify the following:
@@
  and serverless functions to the system and, in particular, what is its effect on the scalability of the system?
+% Improve with respect to what?
+Single population?
+Please explain carefully the issues and how scaling is related to the number of populations.
+- JJ
 
 \item Does having a multi-population enabled platform, with heterogeneous populations and
-the support for mixing search strategies,  increases the performance of the
+the support for mixing search strategies,  increase the performance of the
 search by needing fewer function evaluations than a homogeneous setting? % But in this case the answer is no, right? - JJ
@@
 and resources to an orchestration service?
+% This does not fit in here...
+It's related to reproducibility, and as such belongs in the previous section.
+The subsections that answer this should also go to the previous section - JJ
 \end{itemize}
+
+% Also, point forward to the section (or subsections) that will be devoted to proving each point. BTW, they should probably be subsections.
 
@@
 
+% Maybe subsection? Also say the question you're answering - JJ
 \section{Measuring Scalability}
@@
 % Experiment 1
-As we mentioned before, reactive systems can scale by adding additional copies
+As we mentioned before % say where - JJ
+, reactive systems can scale by adding additional copies
 of serverless functions. In our case, we can start additional worker containers
-to have the same effect. Other authors, like Salza et al. \cite{salza2019speed},   % add
-                                % reference - JJ
-                                % done - Mario
+to have the same effect. Other authors, like Salza et al. \cite{salza2019speed},
 have used a
@@
   \caption{The median of the time required to finish an instance of the function $f_4$, 30 runs
-  with different number of workers and populations}
+  with different number of workers and populations} % time or rate? - JJ
   \label{tab:time}
@@
 
-% I think this is not the case? 
-% Another additional issue is related to the algorithm itself. Many more
-% workers than populations will have a wider range of states of
-% evolution in the experiment; populations that will be mixed will have
-% very different fitness states, which means that a part of them will
-% probably have been not so useful, following the intermediate
-% disturbance hypothesis. When the number of workers is increased, the
-% difference between the state of evolution of populations will be
-% intermediate, leading to better results; eventually, when there are as
-% many workers as populations, the state will be quite similar, which is
-% shown with a smaller gap in the time won adding more workers.
-
 \subsubsection{Speedup}
@@
 
+% TBD? - JJ
+
+% No idea where this fits in. I think it's the second question in 7, so it should probably be a subsection - JJ
 \section{Performance of Heterogeneous Populations in the BBOB Benchmark}
@@
 \section{Deploying to a cloud provider}
+
+% This should probably be a subsection of section 5, where you describe what any user needs to do to get this running and you  talk about reproducibility - JJ
 
@@
   \includegraphics[width=\textwidth]{aws}
-  \caption{Boxplot of the number of evaluations per second ratio, for 15 instances of function $f_4$,
+  \caption{Boxplot of the number of evaluations per second, for 15 instances of function $f_4$,
    with three execution environments Local, Single task and Multi-task.
@@
 \end{figure}
-
-\begin{table}[]
+%
+\begin{table}[h!tbp]
   \small
@@
 % we say it works. - JJ
+% Again, revise search questions and explain answers - JJ
 As a conclusion, EvoSwarm is a framework designed and tested according

### 2020-05-14 draft.tex
More explanation of BBOB figure
@JJ please check it is super clear :angel

@@
    \caption{
-Scaling of running time with problem dimension to reach certain target values $\Delta f$. Lines:
-average runtime (aRT); Cross ($+$): median runtime of successful runs to reach
+Scaling of expected runtime (RT defined as the number of function
+evaluations to reach a target function value for the first time) for  
+each problem dimension. Target values $\Delta f$ = 10, 1, $10^{-1}$, $10^{-2}$,$10^{-3}$,$10^{-5}$,$10^{-8}$
+(the exponent and line colors are given in the legend of $f_1$ in the top row). 
+Lines are the average RT (aRT); Cross ($+$): median RT of successful runs to reach
 the most difficult target that was reached at least once (but not always);
-Cross (\textcolor{red}{$\times$}): maximum number of f-evaluations in any trial. Notched boxes:
-interquartile range with median of simulated runs; all values are divided by
-dimension and plotted as $log_{10}$ values versus dimension.
+Cross (\textcolor{red}{$\times$}): maximum number of f-evaluations in any trial. 
+All values are divided by dimension and plotted as $log_{10}$ values versus dimension.
 % More explanation for those not familiar with BBOB - JJ
+% I think is more clear? - M
 }

### 2020-05-14 draft.tex
Commenty aboout colors in legend
Addressing comment about BBOB experiment

@@
 the target. Fixed values of targets $\Delta f = 10^{k}$ with $k$ colors are
-given in the legend of $f_1$, $k$ values are $[-8,-5,-3,-2,-1,0,1]$. Results
+given in the legend of $f_1$ in the top row, $k$ values are $[-8,-5,-3,-2,-1,0,1]$. Results
 from each configuration are presented in each column, GA, PSO, and GA\&PSO
 populations. % clarify which colors for every level - JJ
+% colors are shown in the figure, I put a hint that is in the top row. 
+% But I can list them if that is better - M
 
@@
 we can perform with the results of the EvoSwarm application by reusing the COCO framework.
-% If this is what you want to prove, it's not a very strong point.
-Maybe it would be stronger if you proved that these mixed algorithms can be tested in the same conditions as single-algorithm, and also that these mixed algorithms work better than single-algorithm versions - JJ
+We also proved that multi-population algorithms can be compared under the same 
+conditions, including single and multi-algorithm versions, and results show 
+that the later have better performance.
+
+% If this is what you want to prove, it's not a very strong point. Maybe it
+% would be stronger if you proved that these mixed algorithms can be tested in the
+% same conditions as single-algorithm, and also that these mixed algorithms work
+% better than single-algorithm versions - JJ
+% Ok, stated like that - M
 
 \section{Deploying to a cloud provider}
-
-% This should probably be a subsection of section 5, where you describe what any user needs to do to get this running and you  talk about reproducibility - JJ
-
-% This is only an idea of the section
-%
-
-% First you have to say that cloud is a way of describing deployment to any kind
-% of infrastructure in a reproducible way, and it needn't be in the cloud proper,
-% it can be deployed to anything, that's precisely the thing. But make clear that
-% when you use a cloud provider, there are more options to do it... - JJ
 

### 2020-05-14 draft.tex
Removed network section

@@
     cpu_limit: 4096
-run_params:
-  network_configuration:
-    awsvpc_configuration:
-      subnets:
-        - "subnet-0ca2a542bbd74239e"
-        - "subnet-0dfd367b5ca531b07"
-      security_groups:
-        - "sg-0681c18469874b391"
-      assign_public_ip: ENABLED
 \end{lstlisting}
-% Do you need the details about the subnets? As long as they are here, you'll have to explain them. Why two subnets? It's probably better to leave that network configuration out anyway - JJ
-
+% Do you need the details about the subnets? As long as they are here,
+% you'll have to explain them. Why two subnets? It's probably better to
+% leave that network configuration out anyway - JJ
+% Ok, I removed that section. those parameters are required, but are used 
+% internally by the cluster. - M 
+% The security group was required to have external access to redis. - M
 % Please clarify that those images are published in Docker hub by yourself - JJ
+% I have done this a few commits back. (4 paragraphs up) - M
 

### 2020-05-14 draft.tex
Explained a bit the compose file

@@
 the configuration on-the-fly. For example to deploy the controller, setup and redis 
-containers we used the file in Listing ~\ref{code:compose}. We have 
-provided the container images and are published in Docker hub \url{https://hub.docker.com/u/mariosky}.
+containers we used the file in Listing ~\ref{code:compose}.
+Each container is defined as a  {\texttt service} , with the following attributes: {\texttt image},
+that reference the container images we have published in Docker hub
+\url{https://hub.docker.com/u/mariosky}, open {\texttt ports}, a {\texttt command} to be executed at
+startup, configuration for the AWS {\texttt logging} service, and finally {\texttt environment}
+variables used by code running inside containers.
 
@@
       command: python experiment.py
-      volumes:
-        - /code
       environment:
@@
       image: mariosky/controller
-      volumes:
-          - /code
       command: python controller.py
@@
 % You will probably need to explain this a bit - JJ 
-
+% done - M  
 \subsection{Experimental Results}

### 2020-05-14 draft.tex
Comment about asynchronous implementations
@jj  please change at will
sentence clarified

@@
 
-% Probably it's not the time for this, but maybe it would be interesting to mention that asynchronous implementations in most cases need a change in the underlying algorithm, while synchronous might not need it - JJ
+% Probably it's not the time for this, but maybe it would be interesting to 
+% mention that asynchronous implementations in most cases need a change in 
+% the underlying algorithm, while synchronous might not need it - JJ 
+% TO DO - Mario 
 Moreover, parallel implementations can run in two different ways: synchronous and
@@
 each other, so there's no time lost waiting for other populations to
-arrive to the synchronization point; results arriving from other workers are incorporated as soon as they are received. Following the previous example, now the master operation can
+arrive to the synchronization point; results arriving from other workers are 
+incorporated as soon as they are received. Following the previous example, now the master operation can
 continue to the next iteration, even if a worker has not finished its work.  In
@@
 
-% I have moved these sentences here, since they seem to fit the best.
+% I have moved these sentences here, since they seem to fit the best.(OK - M)
 Beyond the choice of design methodology and implementation, researchers need to
@@
 
-Docker container images host the stateless functions responsible for running 
-a search strategy on several populations, % clarify "on several populations" Sequentially? At the same time? - JJ
-these functions could also be ported
+Docker container images host the stateless functions responsible for 
+running a search strategy on populations they receive as messages. 
+A single container can pull many messages on its lifetime, processing one at a time. 
+% clarify "on several populations" Sequentially? At the same time? - JJ 
+% done - M
+These functions could also be ported
 directly to a Function as a Service (FaaS) \cite{Roberts2016} implementation.

### 2020-05-14 draft.tex
References to sections corrected.

@@
 % I like it will put it just like that. 
-% A reviewer also put enphasis on the workflow and reproductivility 
+% A reviewer also put enph   asis on the workflow and reproductivility 
 % and extracting a Model? we can put something on that too.
@@
     % Maybe you should check this out again ... - JJ
-    \item Second, we propose a method for the deployment and execution of 
+    \item Second, we propose a new method for the deployment and execution of 
     multiple experiments by specifying the infrastructure as part of an 
@@
 our work. In Section \ref{method}, we present the proposed method and the container-based
-application in Section \ref{docker} . % No reference to section 6.
+application in Section \ref{docker} and the workflow for experiment design in Section \ref{experiment_flow}. % No reference to section 6.
 Section \ref{setup} describes
 the design of the empirical evaluation we designed to assess the effectiveness
-of the method, and in Section \ref{results}, % This is a subsection, fix it. 
-we report and discuss the results. % No references to sections 9 and 10
+of the method, scalability is analysed in Section \ref{sec:exp1}, % This is a subsection, fix it. 
+the performance of heterogeneous populations in Second \label{sec:exp2} 
+and a cloud deployment example in Section \ref{cloud-aws}. % No references to sections 9 and 10
 Finally, we
@@
 \section{Deploying to a cloud provider}
-
+\label{cloud-aws}
 Container-based applications describe the deployment of loosely coupled

### 2020-05-14 draft.tex
Title proposal following @jj recommendation
consistent docker-compose name

@@
 
-  \title{EvoSwarm: An Event-based Architecture for Multi-population Optimization Algorithms}
+  \title{EvoSwarm: A Cloud-Native Architecture for the Reproducible Execution of Multi-population Optimization Algorithms}
   % maybe introduce "reproducible" here, somewhere, as well as "cloud native"
+  % I like cloud-native more than event-based. 
+  % How about that one?  - M
 
@@
 isolated and
-heterogeneous populations. The platform uses the Docker Compose tool for
+heterogeneous populations. The platform uses the {\tt docker-compose} tool for
 defining and running experiments as multi-container applications. EvoSwarm is also released with a free licence in \url{https://github.com/mariosky/EvoSwarm}.
@@
 Kubernetes, or through simple, non-self-scaling specifications using
-Docker Compose, a definition of all the relationships between
+{\tt docker-compose}, a definition of all the relationships between
 different parts of the application, or services. The full realization
@@
 % Experimental design of a container-based application or something like that - JJ 
-\section{Container-based application} 
+% Like it - M
+\section{Experimental design of a container-based application} 
 \label{docker}
@@
 We have mentioned several advantages of a cloud-native application for running
-multi-population based algorithms, but there are in our experience several
-practical issues of this approach. % several issues that make it practical or useful instead? - JJ
+multi-population based algorithms, but there are in our experience, several
+issues that make this approach more practical. % several issues that make it practical or useful instead? - JJ 
 In general, cloud platforms have free-tiers    
@@
 have chosen them in this paper; reproducibility of experiments and a
-clear and idempotent description of the architecture through docker
-compose % Be consistent with docker compose. Either {\tt docker-compose} or {\sf docker-compose}, but not docker compose, for instance. - JJ
+clear and idempotent description of the architecture through {\tt docker-compose} 
+% Be consistent with docker compose. Either {\tt docker-compose} or {\sf docker-compose}, but not docker compose, for instance. - JJ - ok
 are the main reasons.
@@
     \multicolumn{3}{|l|}{Parameter}                    & Type             & Example         \\ \hline
-    \multicolumn{3}{|l|}{Worker Containers (Docker Compose)}        & \texttt{int}     & \texttt{8} \\ \hline
+    \multicolumn{3}{|l|}{Worker Containers}        & \texttt{int}     & \texttt{8} \\ \hline
     \multicolumn{3}{|l|}{GA-PSO Ratio}                 & \texttt{decimal} & \texttt{0.5}    \\  \hline
@@
 Desktop PC with an AMD Ryzen 9 3900x 12-core processor with 24 threads with a 3.8 GHz base clock, 
-and 16 GB RAM. We used Docker version 19.03.3, build a872fc2f86, and docker-compose version 1.21.0, 
-in Ubuntu Linux 18.04; coce is run using Python 3.7.5. Container images and the docker-compose file are available at
+and 16 GB RAM. We used Docker version 19.03.3, build a872fc2f86, and {\tt docker-compose} version 1.21.0, 
+in Ubuntu Linux 18.04; coce is run using Python 3.7.5. Container images and the {\tt docker-compose} file are available at
 \url{https://hub.docker.com/u/mariosky}, and \url{https://github.com/mariosky/EvoSwarm} respectively. 
@@
   Google Compute Engine VMs, or Amazon EC2.  We still need to provision the
-  virtual machine and run docker-compose commands as we would do in a local 
+  virtual machine and run {\tt docker-compose} commands as we would do in a local 
   workstation. Salza et al. \cite{salza2019speed} tested this option by using
@@
 
-  \item[A container orchestration service] these services are compatible with docker-compose
+  \item[A container orchestration service] these services are compatible with {\tt docker-compose}
     files for configuration and provide transparent provisioning. Examples of these
@@
 Container Service (ECS),  since it needs fewer changes in configuration 
-(from deploying it locally using docker-compose) and it is has a lower cost than the Kubernetes alternative. % Remember to mention cost at the end - JJ
+(from deploying it locally using {\tt docker-compose}) and it is has a lower cost than the Kubernetes alternative. % Remember to mention cost at the end - JJ
 
@@
 option (labeled Single) is straightforward to deploy, needing only a single
-docker-compose file, and one command to execute. But with the limit of 4CPUs and
+{\tt docker-compose} file, and one command to execute. But with the limit of 4CPUs and
 ten containers, it is the slower option. The second option (labeled Multi), is
@@
 easily deployed to the cloud, or locally to systems that simply need
-to have Docker and Docker Compose installed.
+to have Docker and {\tt docker-compose} installed.
 

### 2020-05-14 draft.tex
Commit before the Big Move

@@
 map-reduce \cite{fazenda2012} implementations,  grids \cite{munawar2010design,Gonzalez09},
-voluntary computing systems \cite{MilkyWay}, % Maybe add one of our own voluntary computing references? - JJ
+voluntary computing systems \cite{MilkyWay,merelo2016nodio}, 
 and, more importantly, cloud computing
@@
 
-\section{Multi-population methods} % Is this part of the state of the
+\section{Multi-population Methods} % Is this part of the state of the
 \label{multi}
@@
 
-\subsection{Comparison with other cloud-based works} 
+\subsection{Comparison with Other Cloud-based Works} 
 \label{comparison}
@@
 % Like it - M
-\section{Experimental design of a container-based application} 
+\section{Experimental Design of a Container-based Application} 
 \label{docker}
@@
 
-\subsection{EvoSwarm from the developer point of view}
+\subsection{EvoSwarm from the Developer Point of View}
 
@@
 
-\section{Deploying to a cloud provider}
+\section{Deploying to a Cloud Provider}
 \label{cloud-aws}
@@
 \end{description}
+
+\subsection{Setup}
 

### 2020-05-14 draft.tex
Moving Sections

@@
 properties of a reactive system \cite{boner2014reactive} and they are commonly
-more flexible, fault-tolerant, and scalable. % Some properies and reference - Mario  
+more flexible, fault-tolerant, and scalable. % Some properties and reference - Mario  
 % Again, you have to justify the need for a reactive architecture and define it before this paragraph - JJ
@@
 % I like it will put it just like that. 
-% A reviewer also put enph   asis on the workflow and reproductivility 
+% A reviewer also put enphasis on the workflow and reproducivility 
 % and extracting a Model? we can put something on that too.
@@
 our work. In Section \ref{method}, we present the proposed method and the container-based
-application in Section \ref{docker} and the workflow for experiment design in Section \ref{experiment_flow}. % No reference to section 6.
-Section \ref{setup} describes
-the design of the empirical evaluation we designed to assess the effectiveness
-of the method, scalability is analysed in Section \ref{sec:exp1}, % This is a subsection, fix it. 
-the performance of heterogeneous populations in Second \label{sec:exp2} 
-and a cloud deployment example in Section \ref{cloud-aws}. % No references to sections 9 and 10
-Finally, we
-offer the conclusions of this paper and suggestions on future work in Section \ref{conclusions}.
+application in Section \ref{docker} and the workflow for reproducible experiment design in Section \ref{experiment_flow}. % No reference to section 6.
+Section \ref{setup} describes the design of the empirical evaluation to assess the effectiveness
+of the method,  with respect to scalability and % This is a subsection, fix it. 
+the performance of heterogeneous populations.  % No references to sections 9 and 10
+Finally, we offer the conclusions of this paper and suggestions on future work in Section \ref{conclusions}.
 
@@
 asynchronous way, which suits better the nature of cloud systems.
-
-% a lot of this has been done in the previous method... Maybe there's
-% not that much to do - JJ
-% Hadoop 
-%     Comparison \cite{ferrucci2018using}
-%     Hadoop\cite{fazenda2012} 
-%     FlexGP \cite{FlexGP}, 
-%     elephant56 \cite{salza2016elephant56},
-
-% Pool
-%     SofEA \cite{merelo2012sofea,sofea:cec2012} , 
-%     EvoSpace \cite{garcia2015evospace},
-% Volunteer
-%     evospace-js \cite{garcia2017evospace}
-% Queue
-%     Containers \cite{valenzuela2015implementing} \cite{salza2016approach} 
-%     Seamless local and Cloud \cite{leclerc2016seamless}
-%     Serverless KafKEO \cite{guervos2018introducing}, 
-%     Docker Orchestation \cite{dziurzanski2020scalable,zhao2019cloud}
-% PSO 
-% PSO AWS \cite{li2015parallel}
 
@@
 
-% Maybe you should say here how everything wors: launch with
+% Maybe you should say here how everything works: launch with
 % Docker-compose, whatever. That's the issue that one of the reviewers
@@
 % This is totally unexpected. Whose role is this? You didn't even introduce it. Also, extremely short - JJ
+% This is one of the applications described in the Workflow Figure. But this is better explained in other places.
+% We can delete it (even from the Figure)  
+% Will leave it until we decide - M 
 \subsection{Using the COCO Framework}
@@
 
+% TODO: connection to cloud subsection  - M
+
+\subsection{Deploying to a Cloud Provider}
+\label{cloud-aws}
+Container-based applications describe the deployment of loosely coupled
+processes in a consistent and portable way, in every phase of the life cycle of
+an application, from development through testing to production independently on
+the kind of infrastructure. The term cloud-native describes these
+container-based environments because they are deployed on an infrastructure that
+abstracts the underlying compute, storage, and networking primitives. This
+infrastructure does need to be a cloud provider; it could be a developer's
+laptop, a single server, or a production environment. When we deploy a
+cloud-native application to a cloud provider, we have some more options to
+choose:
+
+\begin{description}
+
+  \item[A virtual machine] with a container-optimized operating system, for example,
+  Google Compute Engine VMs, or Amazon EC2.  We still need to provision the
+  virtual machine and run {\tt docker-compose} commands as we would do in a local 
+  workstation. Salza et al. \cite{salza2019speed} tested this option by using
+  CoreOs cluster.
+
+  \item[A Kubernetes cluster] provides auto-scaling and management of containerized
+  applications, but it needs additional configuration for provisioning. All major
+  cloud-providers have Kubernetes services.
+
+  \item[A container orchestration service] these services are compatible with {\tt docker-compose}
+    files for configuration and provide transparent provisioning. Examples of these
+    services are Amazon Elastic Container Service,  Red Hat Openshift, and Marantis Docker Enterprise.
+
+\end{description}
+
+\subsubsection{Setup}
+
+As a proof-of-concept, we chose the third option, that is, a container orchestration service, and deployed the EvoSwarm application in Amazon Elastic
+Container Service (ECS),  since it needs fewer changes in configuration 
+(from deploying it locally using {\tt docker-compose}) and it is has a lower cost than the Kubernetes alternative. % Remember to mention cost at the end - JJ
+
+We can define an ECS cluster directly on AWS Elastic Cloud VMs or through the AWS Fargate service.
+Fargate is a serverless compute engine for containers that provides automatic provision 
+and management of AWS ECS resources. With this service, calculating the cost of execution
+is simplified because users only specify the computing units and memory for all the application,
+instead of every component.
+
+Multicontainer applications are deployed to ECS as tasks, and a single cluster can run multiple tasks in
+parallel. When using the Fargate service as a backend, tasks are limited to a
+maximum of 10 containers.
+
+Using the ECS CLI, we can provision a cluster with a configuration file
+specifying the computing resources needed. We can specify the configuration of a task 
+by using a task definition file (see Listing~\ref{code:ecs}) in YAML format.
+The main attributes are the {\texttt task\_size} indicating the memory and CPU units
+assigned, and network configuration. The complete configuration is generated
+by the {\texttt ecs compose command} that uses a {\texttt docker-compose.yml} file to generate 
+the configuration on-the-fly. For example to deploy the controller, setup and redis 
+containers we used the file in Listing ~\ref{code:compose}.
+Each container is defined as a  {\texttt service} , with the following attributes: {\texttt image},
+that reference the container images we have published in Docker hub
+\url{https://hub.docker.com/u/mariosky}, open {\texttt ports}, a {\texttt command} to be executed at
+startup, configuration for the AWS {\texttt logging} service, and finally {\texttt environment}
+variables used by code running inside containers.
+
+We deployed two configurations, one that included all containers in a single task,
+and another one, where we assigned each worker to a different task and the controller, 
+redis, and setup in another, for a total of nine concurrent Fargate tasks.
+
+For the first option, we deployed an AWS Fargate cluster and defined a task configuration file with
+8GBs of memory and 4096 computing units, equivalent to 4 CPUs. Because of the
+ten container limit, we only used seven workers and tested function $f_4$ as
+in the previous experiment, and to save resources we only tested for dimensions 10 and 20 for 15 instances.
+
+For the second option, we used the same cluster and additionally
+assigned each worker to a task with 4GBs of memory and 2048 computing units. Then scale the 
+number of worker tasks to eight. We can scale the number of tasks for up to 100 instances in 
+a Fargate configuration.
+
+\begin{lstlisting}[caption=ecs-params.yml , label=code:ecs]
+version: 1
+task_definition:
+  task_execution_role: ecsTaskExecutionRole
+  ecs_network_mode: awsvpc
+  task_size:
+    mem_limit: 8GB
+    cpu_limit: 4096
+\end{lstlisting}
+% Do you need the details about the subnets? As long as they are here,
+% you'll have to explain them. Why two subnets? It's probably better to
+% leave that network configuration out anyway - JJ
+% Ok, I removed that section. those parameters are required, but are used 
+% internally by the cluster. - M 
+% The security group was required to have external access to redis. - M
+% Please clarify that those images are published in Docker hub by yourself - JJ
+% I have done this a few commits back. (4 paragraphs up) - M
+
+\begin{lstlisting}[caption={docker-compose.yml for setup, controller and redis containers}, label=code:compose]
+  version: '3'
+  services:
+    redis:
+      image: "redis:alpine"
+      ports:
+        - "6379:6379"
+      logging:
+        driver: awslogs
+        options:
+          awslogs-region: us-west-2
+          awslogs-group: docker-logs
+          awslogs-stream-prefix: redis
+
+    setup:
+      image: mariosky/setup
+      command: python experiment.py
+      environment:
+        PYTHONUNBUFFERED: 1
+        REDIS_HOST: localhost
+      logging:
+        driver: awslogs
+        options:
+          awslogs-region: us-west-2
+          awslogs-group: docker-logs
+          awslogs-stream-prefix: setup
+
+    controller:
+      image: mariosky/controller
+      command: python controller.py
+      environment:
+          PYTHONUNBUFFERED: 1
+          REDIS_HOST: localhost
+      logging:
+        driver: awslogs
+        options:
+          awslogs-region: us-west-2
+          awslogs-group: docker-logs
+          awslogs-stream-prefix: controller
+  \end{lstlisting}
+% You will probably need to explain this a bit - JJ 
+% done - M  
+\subsubsection{Experimental Results}
+
+The results are shown in Figure~\ref{fig:aws-results} and Table~\ref{tab:aws}.
+
+As we can see, the local environment (see section \ref{setup}) achieved a higher
+ratio of evaluations per second (in bold). Still, we must consider that there is a minimum
+network overhead in a local workstation and a clock speed of 3.8 GHz. The first
+option (labeled Single) is straightforward to deploy, needing only a single
+{\tt docker-compose} file, and one command to execute. But with the limit of 4CPUs and
+ten containers, it is the slower option. The second option (labeled Multi), is
+faster, and not too far from the local execution, especially when solving the
+$f_4$ in 20 dimensions (underline in table). With the capacity for increasing the number of worker
+containers, it is reasonable to think that we can achieve a higher number of
+evaluations with an appropriate configuration. Nevertheless, the goal of this
+experiment was not to increase the number of evaluations but to test the
+flexibility of a cloud-native application.
+
+\begin{figure}[h!tbp]
+  \centering
+  \includegraphics[width=\textwidth]{aws}
+  \caption{Boxplot of the number of evaluations per second, for 15 instances of function $f_4$,
+   with three execution environments Local, Single task and Multi-task.
+   }
+  \label{fig:aws-results}
+\end{figure}
+%
+\begin{table}[h!tbp]
+  \small
+  \caption{ Median number of the evaluations per second ratio, for 15 instances of function $f_4$,
+  with three execution environments Local, Single task and Multi task}
+  \label{tab:aws}
+  \vspace{0.25cm}
+  \centering
+
+  \begin{tabular}{lll}
+  \hline
+  Dim & Env         & Median of ratio \\
+  \hline
+  10  & Local       & \textbf{47,166.29}       \\
+      & Multi task  & 40,466.45    \\
+      & Single task & 18,193.05      \\
+  \hline
+  20  & Local       & \textbf{36,101.86}   \\
+      & Multi task  & {\ul 33,916.02}       \\
+      & Single task & 16,546.95  \\
+  \hline
+  \end{tabular}
+\end{table}
+
 % You should clarify that, before, you were explaining how to set up an experiment.
@@
 
-\section{General Experimental Setup} 
-\label{setup}
+\section{Experimental Study} 
+\label{setup} % To be renamed
 
@@
 % Also, point forward to the section (or subsections) that will be devoted to proving each point. BTW, they should probably be subsections.
+
+\subsection{General Experimental Setup} 
+\label{sec:general-setup} % To be renamed
 
@@
 % Maybe subsection? Also say the question you're answering - JJ
-\section{Measuring Scalability}
+\subsection{Measuring Scalability}
 \label{sec:exp1}
@@
 
-\subsection{Setup}
+\subsubsection{Setup}
 
@@
 
-\subsection{Experimental Results} 
-\label{results}
+%\subsubsection{Experimental Results} 
+%\label{results}
 
@@
 % No idea where this fits in. I think it's the second question in 7, so it should probably be a subsection - JJ
-\section{Performance of Heterogeneous Populations in the BBOB Benchmark}
+\subsection{Performance of Heterogeneous Populations in the BBOB Benchmark}
 \label{sec:exp2}
@@
 
-\subsection{Experimental Results}
+\subsubsection{Experimental Results}
 
@@
 % Ok, stated like that - M
-
-\section{Deploying to a Cloud Provider}
-\label{cloud-aws}
-Container-based applications describe the deployment of loosely coupled
-processes in a consistent and portable way, in every phase of the life cycle of
-an application, from development through testing to production independently on
-the kind of infrastructure. The term cloud-native describes these
-container-based environments because they are deployed on an infrastructure that
-abstracts the underlying compute, storage, and networking primitives. This
-infrastructure does need to be a cloud provider; it could be a developer's
-laptop, a single server, or a production environment. When we deploy a
-cloud-native application to a cloud provider, we have some more options to
-choose:
-
-\begin{description}
-
-  \item[A virtual machine] with a container-optimized operating system, for example,
-  Google Compute Engine VMs, or Amazon EC2.  We still need to provision the
-  virtual machine and run {\tt docker-compose} commands as we would do in a local 
-  workstation. Salza et al. \cite{salza2019speed} tested this option by using
-  CoreOs cluster.
-
-  \item[A Kubernetes cluster] provides auto-scaling and management of containerized
-  applications, but it needs additional configuration for provisioning. All major
-  cloud-providers have Kubernetes services.
-
-  \item[A container orchestration service] these services are compatible with {\tt docker-compose}
-    files for configuration and provide transparent provisioning. Examples of these
-    services are Amazon Elastic Container Service,  Red Hat Openshift, and Marantis Docker Enterprise.
-
-\end{description}
-
-\subsection{Setup}
-
-As a proof-of-concept, we chose the third option, that is, a container orchestration service, and deployed the EvoSwarm application in Amazon Elastic
-Container Service (ECS),  since it needs fewer changes in configuration 
-(from deploying it locally using {\tt docker-compose}) and it is has a lower cost than the Kubernetes alternative. % Remember to mention cost at the end - JJ
-
-We can define an ECS cluster directly on AWS Elastic Cloud VMs or through the AWS Fargate service.
-Fargate is a serverless compute engine for containers that provides automatic provision 
-and management of AWS ECS resources. With this service, calculating the cost of execution
-is simplified because users only specify the computing units and memory for all the application,
-instead of every component.
-
-Multicontainer applications are deployed to ECS as tasks, and a single cluster can run multiple tasks in
-parallel. When using the Fargate service as a backend, tasks are limited to a
-maximum of 10 containers.
-
-Using the ECS CLI, we can provision a cluster with a configuration file
-specifying the computing resources needed. We can specify the configuration of a task 
-by using a task definition file (see Listing~\ref{code:ecs}) in YAML format.
-The main attributes are the {\texttt task\_size} indicating the memory and CPU units
-assigned, and network configuration. The complete configuration is generated
-by the {\texttt ecs compose command} that uses a {\texttt docker-compose.yml} file to generate 
-the configuration on-the-fly. For example to deploy the controller, setup and redis 
-containers we used the file in Listing ~\ref{code:compose}.
-Each container is defined as a  {\texttt service} , with the following attributes: {\texttt image},
-that reference the container images we have published in Docker hub
-\url{https://hub.docker.com/u/mariosky}, open {\texttt ports}, a {\texttt command} to be executed at
-startup, configuration for the AWS {\texttt logging} service, and finally {\texttt environment}
-variables used by code running inside containers.
-
-We deployed two configurations, one that included all containers in a single task,
-and another one, where we assigned each worker to a different task and the controller, 
-redis, and setup in another, for a total of nine concurrent Fargate tasks.
-
-For the first option, we deployed an AWS Fargate cluster and defined a task configuration file with
-8GBs of memory and 4096 computing units, equivalent to 4 CPUs. Because of the
-ten container limit, we only used seven workers and tested function $f_4$ as
-in the previous experiment, and to save resources we only tested for dimensions 10 and 20 for 15 instances.
-
-For the second option, we used the same cluster and additionally
-assigned each worker to a task with 4GBs of memory and 2048 computing units. Then scale the 
-number of worker tasks to eight. We can scale the number of tasks for up to 100 instances in 
-a Fargate configuration.
-
-\begin{lstlisting}[caption=ecs-params.yml , label=code:ecs]
-version: 1
-task_definition:
-  task_execution_role: ecsTaskExecutionRole
-  ecs_network_mode: awsvpc
-  task_size:
-    mem_limit: 8GB
-    cpu_limit: 4096
-\end{lstlisting}
-% Do you need the details about the subnets? As long as they are here,
-% you'll have to explain them. Why two subnets? It's probably better to
-% leave that network configuration out anyway - JJ
-% Ok, I removed that section. those parameters are required, but are used 
-% internally by the cluster. - M 
-% The security group was required to have external access to redis. - M
-% Please clarify that those images are published in Docker hub by yourself - JJ
-% I have done this a few commits back. (4 paragraphs up) - M
-
-\begin{lstlisting}[caption={docker-compose.yml for setup, controller and redis containers}, label=code:compose]
-  version: '3'
-  services:
-    redis:
-      image: "redis:alpine"
-      ports:
-        - "6379:6379"
-      logging:
-        driver: awslogs
-        options:
-          awslogs-region: us-west-2
-          awslogs-group: docker-logs
-          awslogs-stream-prefix: redis
-
-    setup:
-      image: mariosky/setup
-      command: python experiment.py
-      environment:
-        PYTHONUNBUFFERED: 1
-        REDIS_HOST: localhost
-      logging:
-        driver: awslogs
-        options:
-          awslogs-region: us-west-2
-          awslogs-group: docker-logs
-          awslogs-stream-prefix: setup
-
-    controller:
-      image: mariosky/controller
-      command: python controller.py
-      environment:
-          PYTHONUNBUFFERED: 1
-          REDIS_HOST: localhost
-      logging:
-        driver: awslogs
-        options:
-          awslogs-region: us-west-2
-          awslogs-group: docker-logs
-          awslogs-stream-prefix: controller
-  \end{lstlisting}
-% You will probably need to explain this a bit - JJ 
-% done - M  
-\subsection{Experimental Results}
-
-The results are shown in Figure~\ref{fig:aws-results} and Table~\ref{tab:aws}.
-
-As we can see, the local environment (see section \ref{setup}) achieved a higher
-ratio of evaluations per second (in bold). Still, we must consider that there is a minimum
-network overhead in a local workstation and a clock speed of 3.8 GHz. The first
-option (labeled Single) is straightforward to deploy, needing only a single
-{\tt docker-compose} file, and one command to execute. But with the limit of 4CPUs and
-ten containers, it is the slower option. The second option (labeled Multi), is
-faster, and not too far from the local execution, especially when solving the
-$f_4$ in 20 dimensions (underline in table). With the capacity for increasing the number of worker
-containers, it is reasonable to think that we can achieve a higher number of
-evaluations with an appropriate configuration. Nevertheless, the goal of this
-experiment was not to increase the number of evaluations but to test the
-flexibility of a cloud-native application.
-
-\begin{figure}[h!tbp]
-  \centering
-  \includegraphics[width=\textwidth]{aws}
-  \caption{Boxplot of the number of evaluations per second, for 15 instances of function $f_4$,
-   with three execution environments Local, Single task and Multi-task.
-   }
-  \label{fig:aws-results}
-\end{figure}
-%
-\begin{table}[h!tbp]
-  \small
-  \caption{ Median number of the evaluations per second ratio, for 15 instances of function $f_4$,
-  with three execution environments Local, Single task and Multi task}
-  \label{tab:aws}
-  \vspace{0.25cm}
-  \centering
-
-  \begin{tabular}{lll}
-  \hline
-  Dim & Env         & Median of ratio \\
-  \hline
-  10  & Local       & \textbf{47,166.29}       \\
-      & Multi task  & 40,466.45    \\
-      & Single task & 18,193.05      \\
-  \hline
-  20  & Local       & \textbf{36,101.86}   \\
-      & Multi task  & {\ul 33,916.02}       \\
-      & Single task & 16,546.95  \\
-  \hline
-  \end{tabular}
-\end{table}
 

### 2020-05-15 draft.tex
Some comments here and there, check them out
Deletes some addressed comments and remarking others

@@
   % How about that one?  - M
+  % better than the previous one.
 
@@
 
+% Write some introduction to this subsubsection. 
 The results are shown in Figure~\ref{fig:aws-results} and Table~\ref{tab:aws}.
@@
 {\tt docker-compose} file, and one command to execute. But with the limit of 4CPUs and
-ten containers, it is the slower option. The second option (labeled Multi), is
+ten containers, it is the slowest option. The second option (labeled Multi), is
 faster, and not too far from the local execution, especially when solving the
@@
 flexibility of a cloud-native application.
+% Mention costs - JJ
 
@@
   \small
-  \caption{ Median number of the evaluations per second ratio, for 15 instances of function $f_4$,
-  with three execution environments Local, Single task and Multi task}
+  \caption{ Median number of the evaluations per second, for 15 instances of function $f_4$,
+  with three execution environments Local, Single task and Multi task. The best result is shown in {\bf boldface}.}
   \label{tab:aws}
@@
   20  & Local       & \textbf{36,101.86}   \\
-      & Multi task  & {\ul 33,916.02}       \\
+      & Multi task  & {\ul 33,916.02}       \\ % why the underline?
       & Single task & 16,546.95  \\
@@
 
-% You should clarify that, before, you were explaining how to set up an experiment.
-In this section, you are explaining the experiments that you are doing in this paper.
-But you should relate both sections to the general objectives of the paper.
-In the prior section, you want to prove that EvoSwarm can be used by the public to create reproductive research (and maybe you should add a small conclusion sub or sub-sub section), in the next you explain the experiment we are doing for this paper.
-It would be interesting to make references from the next section to the previous one, indicating how the general setup of EvoSwarm helps _us_ to create this experiment - JJ
+% Emphasize these results for the purpose of the paper - JJ
+
+% You should clarify that, before, you were explaining how to set up
+% an experiment. In this section, you are explaining the experiments
+% that you are doing in this paper. But you should relate both
+% sections to the general objectives of the paper. In the prior
+% section, you want to prove that EvoSwarm can be used by the public
+% to create reproductive research (and maybe you should add a small
+% conclusion sub or sub-sub section), in the next you explain the
+% experiment we are doing for this paper. It would be interesting to
+% make references from the next section to the previous one,
+% indicating how the general setup of EvoSwarm helps _us_ to create
+% this experiment - JJ
 
@@
 locally or in the cloud by just specifying the configuration of container images
-and resources to an orchestration service?
-% This does not fit in here...
-It's related to reproducibility, and as such belongs in the previous section.
-The subsections that answer this should also go to the previous section - JJ
+and resources to an orchestration service? % This does not fit in
+                                % here... It's related to
+                                % reproducibility, and as such belongs
+                                % in the previous section. The
+                                % subsections that answer this should
+                                % also go to the previous section - JJ
+% Still need to check this out - JJ
 \end{itemize}
@@
 maximum number of FEs is $200,000$.
-% How do you manage to stop exactly at that number of evaluations in a distributed system? - JJ
-% In the controller we check the number of messages received, we assume that each message 
-% needs a X amount of #FEs we stop at the maximum number of messages parameter.
-% If the solution is found we stop sending messages and ignore messages from that
-% experiment ID - M
 
@@
 % I think is more clear? - M
+% Almost everything, but the cream-colored rhombs in the bottom...
 }

### 2020-05-15 draft.tex
Explanationa added.
If I add more text the figure moves to the last page. I had to remove some words in order to make the text fit,

@@
 Scaling of expected runtime (RT defined as the number of function 
-evaluations to reach a target function value for the first time) for  
-each problem dimension. Target values $\Delta f$ = 10, 1, $10^{-1}$, $10^{-2}$,$10^{-3}$,$10^{-5}$,$10^{-8}$
-(the exponent and line colors are given in the legend of $f_1$ in the top row). 
-Lines are the average RT (aRT); Cross ($+$): median RT of successful runs to reach
-the most difficult target that was reached at least once (but not always);
-Cross (\textcolor{red}{$\times$}): maximum number of f-evaluations in any trial. 
-All values are divided by dimension and plotted as $log_{10}$ values versus dimension.
+evaluations \#FE) to reach a target for  
+each dimension. Target values $\Delta f$ = 10, 1, $10^{-1}$, $10^{-2}$,$10^{-3}$,$10^{-5}$,$10^{-8}$
+(the exponent and colors are given in the legend of $f_1$). 
+Lines are the average RT; Cross ($+$): median RT of successful runs to reach
+the most difficult target that was reached at least once;
+Cross (\textcolor{red}{$\times$}): maximum \#FE in any trial. 
+Values are divided by dimension and plotted as $log_{10}$.
+The light thick line with diamonds indicates the theoretical best algorithm from BBOB 2009 for the 
+most difficult target.
 % More explanation for those not familiar with BBOB - JJ

### 2020-05-15 draft.tex
comment
Eliminates some comments, small changes

@@
   20  & Local       & \textbf{36,101.86}   \\
-      & Multi task  & {\ul 33,916.02}       \\ % why the underline?
+      & Multi task  & {\ul 33,916.02}       \\ % why the underline? [Best between cloud options :) - Mario] Then why not above too? - JJ
       & Single task & 16,546.95  \\
@@
   \vspace{-3ex}
-   \caption{
-Scaling of expected runtime (RT defined as the number of function 
+   \caption{Scaling of expected runtime (RT defined as the number of function 
 evaluations \#FE) to reach a target for  
@@
 most difficult target.
-% More explanation for those not familiar with BBOB - JJ
-% I think is more clear? - M
-% Almost everything, but the cream-colored rhombs in the bottom...
 }
@@
 A multi-population based algorithm can decrease the execution time
-because of parallel execution. But, having heterogeneous 
-populations might enhance evolutionary search and avoid unneeded
-evaluations more than homogeneous systems systems do; heterogeneous settings, if
-done right, increases the diversity of the whole population\cite{araujo2008multikulti}.
+thanks to running the algorithm on several populations at the same
+time. But, if the populations are heterogeneous 
+it might enhance evolutionary search and avoid unneeded
+evaluations to a greater extent than homogeneous systems systems do; heterogeneous settings, if
+done right, increase the diversity of the whole population\cite{araujo2008multikulti}.
 
@@
 from each configuration are presented in each column, GA, PSO, and GA\&PSO
-populations. % clarify which colors for every level - JJ
-% colors are shown in the figure, I put a hint that is in the top row. 
-% But I can list them if that is better - M
+populations.
 

### 2020-05-15 draft.tex
More comments addressed
Moving question about cloud to reproducible
@JJ changed the name of the section
Deleting resolved comments
Updating the now Reproducible section
I'm keeping just two roles. I will not mention installation.

@@
   \title{EvoSwarm: A Cloud-Native Architecture for the Reproducible Execution of Multi-population Optimization Algorithms}
-  % maybe introduce "reproducible" here, somewhere, as well as "cloud native"
-  % I like cloud-native more than event-based. 
-  % How about that one?  - M
-  % better than the previous one.
 
@@
                      % on the algorithm, not speedup 
+                     % I like the sentence, If we did not prove it, 
+                     % at least we build a tool that will allow
+                     % algorithms to run in their natural environment.
+                     % The focus of the Journal after all is in computer systems - M
+
 and setting a new baseline for scaling and
@@
 
-% But first, we give a brief description of the technology we used.
-
-% Add if needed
-%\subsection{Containers} 
-%\label{containers}
-% TO DO
-
-%\subsection{Docker Compose} 
-%\label{compose}
-% TO DO
-
-% Do we need a description of every container? Wouldn't a general 
-% description  of the containers used be enough? - JJ 
-% Ok, I think only workers and the controller need to be described in detail - Mario 
 \subsection{Message Queue Container} 
@@
 
-\section{Experimental Workflow in EvoSwarm} % We could upgrade this to a section and explain in detail 
-% taking care of #5 - M
+\section{Workflow for Reproducible Experiment Design in EvoSwarm} 
 \label{experiment_flow}
-
-% You can't start with this.
-Explain what this section is about, and link it with the previous one.
-Maybe it's better to create a big section with "Experimental setup", and make this a subsection, with subsections becoming subsubsections - JJ
-Figure \ref{fig:experiment_flow} shows the general workflow to run experiments
-in Evo\-Swarm, and the three roles users can play. The first step is to develop the
-search strategy algorithms that are going to be used by the multi-population
-algorithm; algorithm developers do this. Then administrators are responsible for
-the installation and configuration of the application. Once the application is
-running, users can start one or several experiments, monitor the execution, and
-analyze the results. Depending on the type of experiment executed, users could
-have additional needs, in this case there are additional scripts for
-processing the results of experiments to be used with the COCO framework. In the
-next sections, we give more details about these roles and their workflows.
+% Changed the section title to give more focus on reproducibility - M
+% You can't start with this. Explain what this section is about, and link it with 
+% the previous one. Maybe it's better to create a big section with "Experimental setup", 
+% and make this a subsection, with subsections becoming subsubsections - JJ 
+% done - M
+
+An essential goal of our EvoSwarm application is to allow researchers the
+description and orchestration of reproducible multi-population experiments.
+Following the application design described in the previous section, we now
+present the workflow of an experiment's entire life-cycle viewed from the
+perspective of the two roles users can play. On one side is the user role that
+designs and runs experiments and, on the other, the role of a developer of
+algorithms. The scenario presented in Figure~\ref{fig:experiment_flow} shows the
+workflow of an experiment in Evo\-Swarm the two roles users can play. The first
+step is to develop the search strategy algorithms that are going to be used by
+the multi-population algorithm; algorithm developers do this. Then users are
+responsible for the configuration and deployment of the resources needed for the
+experiment. Once the application is running, users can start one or several
+experiments, monitor the execution, and analyze the results. Depending on the
+type of experiment executed, users could have additional needs, in this case
+there are additional scripts for processing the results of experiments to be
+used with the COCO framework. In the next sections, we give more details about
+these roles and their workflows, and at the end we present a proof-of-concept example of a
+reproducible experiment in EvoSwarm.
+
+%\begin{itemize}
+  % Question move to here for the moment, I mention this in this intro. - Mario
+
+%  \item Can EvoSwarm as a cloud-native, container-based application,  be deployed
+%  locally or in the cloud by just specifying the configuration of container images
+%  and resources to an orchestration service? 
+%\end{itemize}
 
@@
 For the secondary level of integration, developers need to add code to the setup
-
-container
+container to specify additional configuration options. For instance,
 in the current configuration message (see Listing \ref{code:config})
@@
 
-% You said "subsections will talk about roles", this is not a role. Change the title to that of a role. Would that be administrator? - JJ
-
-\subsection{EvoSwarm Configuration}
+% You said "subsections will talk about roles", this is not a role. 
+% Change the title to that of a role. Would that be administrator? - JJ
+
+\subsection{EvoSwarm from the User Point of View}
+
+\subsubsection{EvoSwarm Configuration}
 \label{sec:evoswarm:config}
@@
 % Again, not a role - JJ
-\subsection{Experiment Execution}
+\subsubsection{Experiment Execution}
 \label{sec:exeperimentalDefinition}
@@
 % Will leave it until we decide - M 
-\subsection{Using the COCO Framework}
+\subsubsection{Using the COCO Framework}
 
@@
 \end{lstlisting}
-% Do you need the details about the subnets? As long as they are here,
-% you'll have to explain them. Why two subnets? It's probably better to
-% leave that network configuration out anyway - JJ
-% Ok, I removed that section. those parameters are required, but are used 
-% internally by the cluster. - M 
-% The security group was required to have external access to redis. - M
-% Please clarify that those images are published in Docker hub by yourself - JJ
-% I have done this a few commits back. (4 paragraphs up) - M
 
@@
   \end{lstlisting}
-% You will probably need to explain this a bit - JJ 
-% done - M  
+
 \subsubsection{Experimental Results}
@@
 
-% JJ these are the three experiments, please check them. The last two addressing
-% issues from a reviewer - M
-
 \begin{itemize}
@@
 the support for mixing search strategies,  increase the performance of the
-search by needing fewer function evaluations than a homogeneous setting? % But in this case the answer is no, right? - JJ
-                                                                         % Is yes, but is in the second experiment - M
-\item Can EvoSwarm as a cloud-native, container-based application,  be deployed
-locally or in the cloud by just specifying the configuration of container images
-and resources to an orchestration service? % This does not fit in
-                                % here... It's related to
-                                % reproducibility, and as such belongs
-                                % in the previous section. The
-                                % subsections that answer this should
-                                % also go to the previous section - JJ
-% Still need to check this out - JJ
+search by needing fewer function evaluations than a homogeneous setting?
+
+% Moved reproducibility question from here - M
 \end{itemize}
 
-% Also, point forward to the section (or subsections) that will be devoted to proving each point. BTW, they should probably be subsections.
+% Also, point forward to the section (or subsections) that will be devoted to proving each point.
+% BTW, they should probably be subsections.
 
@@
 files the EvoSwarm application uses to run the noiseless BBOB testbed
-experiments as seen in Section~\ref{sec:exeperimentalDefinition}. % Specify this a bit more (with commands such as
-             % docker-compose up and so on and use an itemize
-             % environment to highlight it and answer to the reviewer
-             % - JJ
-             % Done, a few lines down - M
+experiments as seen in Section~\ref{sec:exeperimentalDefinition}. 
 As a first step, we must deploy the docker application using a
@@
 populations, and finally, a value of  $1$ means that all populations will run
-the GA algorithm. % Opening up the door to ask for testing with
-                  % different proportion. You have to justify this
-                  % value - JJ  
-                  % We mention earlier that we replicated the params of another paper - M
+the GA algorithm.
 
@@
 The BOBB experiments were performed with COCO \cite{hansen2016coco} version bbob.v15.03 in Python, 
-the plots were produced with version 2.3.2. % Are we using BBOB plots? - JJ
+the plots were produced with version 2.3.2.
 
@@
   \caption{Boxplot of the number of evaluations per second, for 30 instances of function $f_4$, with different number of workers and populations sizes }
-
-  % Which function? - JJ
   \label{fig:spworker}
@@
   \small
-  \caption{Speedup achieved with incrementing the number of workers for each dimension showing the p-rate for the Wilcoxon test}
+  \caption{Speedup achieved with incrementing the number of workers for
+   each dimension showing the p-rate for the Wilcoxon test}
   % mark significant differences in some way - JJ
@@
 
-% No idea where this fits in. I think it's the second question in 7, so it should probably be a subsection - JJ
+% No idea where this fits in. I think it's the second question in 7,
+% so it should probably be a subsection - JJ
+
 \subsection{Performance of Heterogeneous Populations in the BBOB Benchmark}
@@
 
-% If this is what you want to prove, it's not a very strong point. Maybe it
-% would be stronger if you proved that these mixed algorithms can be tested in the
-% same conditions as single-algorithm, and also that these mixed algorithms work
-% better than single-algorithm versions - JJ
-% Ok, stated like that - M
-
 \section{Conclusions} 

### 2020-05-15 draft.tex
More updates

@@
 \subsection{EvoSwarm from the User Point of View} 
-
-\subsubsection{EvoSwarm Configuration}
+% All in a single section - Mario
+%\subsubsection{Experiment Configuration and Execution}
 \label{sec:evoswarm:config}
-The administrator role is responsible for
-editing the Docker Compose configuration files that reference the images created
+
+The user role is responsible for
+editing the {\tt docker-compose} configuration files that reference the images created
 by developers. To deploy the application, he or she needs to edit the {\tt
 docker-compose.yml} file to specify the type and number of containers required
-for the experiments. The {\tt docker-compose} application is responsible for
+for the experiments. Then, the {\tt docker-compose} application is responsible for
 starting all the containers, and downloading new versions if it is necessary,
-by executing the command {\tt docker-compose up}  inside the EvoSwarm directory
-
-% Also, it's a bit short. Maybe merge it into the next? - JJ
-
-% Again, not a role - JJ
-\subsubsection{Experiment Execution}
-\label{sec:exeperimentalDefinition}
+by executing the command {\tt docker-compose up}  inside the EvoSwarm directory.
 
@@
 experiment definition is shown in Listing \ref{code:exp}. We give more details about the 
-data contained in the file in the next section.
+data contained in the file in the next section. Users can use these files 
+to analyze and plot experiment data.  Python
+scripts are included in the repository to plot the running times for the
+experiments, and all other plots used in this paper;  there is also a script to
+generate the files needed by the COCO framework \cite{hansen2016coco}, which can generate standard
+comparisons against other methods.
 
@@
  {
- "2":{"GEN":40,"POP_SIZE": 50,"MAX_ITER":10,"MESSAGES":10},
- "3":{"GEN":25,"POP_SIZE": 60,"MAX_ITER":20,"MESSAGES":10},
- "5":{"GEN":28,"POP_SIZE": 60,"MAX_ITER":30,"MESSAGES":10},
  "10":{"GEN":50,"POP_SIZE": 70,"MAX_ITER":30,"MESSAGES":10},
  "20":{"GEN":66,"POP_SIZE":100,"MAX_ITER":30,"MESSAGES":10},
- "40":{"GEN":80,"POP_SIZE":125,"MAX_ITER":40,"MESSAGES":10}
  },
@@
 "FUNCTIONS": [4],
-"DIMENSIONS" : [2, 3, 5, 10, 20, 40],
+"DIMENSIONS" : [10, 20],
 "CXPB_RND": [0.2, 0.6],
@@
 
-% This is totally unexpected. Whose role is this? You didn't even introduce it. Also, extremely short - JJ
-% This is one of the applications described in the Workflow Figure. But this is better explained in other places.
-% We can delete it (even from the Figure)  
-% Will leave it until we decide - M 
-\subsubsection{Using the COCO Framework}
-
-Researchers can use these files to analyze and plot experiment data.  Python
-scripts are included in the repository to plot the running times for the
-experiments, and all other plots used in this paper;  there is also a script to
-generate the files needed by the COCO framework \cite{hansen2016coco}, which can generate standard
-comparisons against other methods.
-
 % TODO: connection to cloud subsection  - M
@@
 \label{cloud-aws}
-Container-based applications describe the deployment of loosely coupled
-processes in a consistent and portable way, in every phase of the life cycle of
-an application, from development through testing to production independently on
-the kind of infrastructure. The term cloud-native describes these
-container-based environments because they are deployed on an infrastructure that
-abstracts the underlying compute, storage, and networking primitives. This
-infrastructure does need to be a cloud provider; it could be a developer's
-laptop, a single server, or a production environment. When we deploy a
-cloud-native application to a cloud provider, we have some more options to
-choose:
+
+As a proof-of-concept, we present a brief experiment that is deployed in three
+distinct environments, taking advantage that as a container-based application,
+EvoSwarm describes the deployment of loosely coupled processes in a consistent
+and portable way.
+
+The term cloud-native describes these container-based
+environments because we can deploy them on an infrastructure that abstracts the
+underlying compute, storage, and networking primitives. This infrastructure does not
+need to be a cloud provider; it could be a developer's laptop, a single server,
+or a production environment. When we deploy a cloud-native application to a
+cloud provider, we have some more options to choose:
+
+%Container-based applications describe the deployment of loosely coupled
+%processes in a consistent and portable way, in every phase of the life cycle of
+%an application, from development through testing to production independently on
+%the kind of infrastructure. The term cloud-native describes these
+%container-based environments because they are deployed on an infrastructure that
+%abstracts the underlying compute, storage, and networking primitives. This
+%infrastructure does need to be a cloud provider; it could be a developer's
+%laptop, a single server, or a production environment. When we deploy a
+%cloud-native application to a cloud provider, we have some more options to
+%choose:
 
@@
 \subsubsection{Setup}
-
-As a proof-of-concept, we chose the third option, that is, a container orchestration service, and deployed the EvoSwarm application in Amazon Elastic
-Container Service (ECS),  since it needs fewer changes in configuration 
-(from deploying it locally using {\tt docker-compose}) and it is has a lower cost than the Kubernetes alternative. % Remember to mention cost at the end - JJ
+\label{setup:local-cloud}
+
+We chose the third option, that is, a container orchestration service, and
+deployed the EvoSwarm application in Amazon Elastic Container Service (ECS),
+since it needs fewer changes in configuration (from deploying it locally using
+{\tt docker-compose}) and it is has a lower cost than the Kubernetes
+alternative.
+
+% Remember to mention cost at the end - JJ
 
@@
 
-We deployed two configurations, one that included all containers in a single task,
+We deployed two ECS configurations, one that included all containers in a single task,
 and another one, where we assigned each worker to a different task and the controller, 
@@
 a Fargate configuration.
+
+For local experiments, we have deployed the EvoSwarm application in a
+Desktop PC with an AMD Ryzen 9 3900x 12-core processor with 24 threads with a 3.8 GHz base clock, 
+and 16 GB RAM. We used Docker version 19.03.3, build a872fc2f86, and {\tt docker-compose} version 1.21.0, 
+in Ubuntu Linux 18.04; code is run using Python 3.7.5. Container images and the {\tt docker-compose} file are available at
+\url{https://hub.docker.com/u/mariosky}, and \url{https://github.com/mariosky/EvoSwarm} respectively.
+
+We reproduced the experiment of Listing~\ref{code:exp} in each environment. The experiment consists of
+solving $f_4$ for 10 and 20 dimensions. We give more details about additional parameters of the configuration in Section~\ref{setup}.
 
@@
 % Write some introduction to this subsubsection.
+
+In this experiment, we wanted to test several deployment options for the 
+cloud-native application, and have an idea of the computing capacity of each option
+measured as the number of function evaluations per second.
 The results are shown in Figure~\ref{fig:aws-results} and Table~\ref{tab:aws}. 
-
 As we can see, the local environment (see section \ref{setup}) achieved a higher
@@
 experiment was not to increase the number of evaluations but to test the
-flexibility of a cloud-native application.
+flexibility of a cloud-native application. The total cost of testing, configuration, 
+and execution of the experiments in ECS was less than \$1.00 US Dollar.
+
 % Mention costs - JJ
@@
 
+The same workflow and configuration files used by this brief experiment, will be
+used on the next section, to carry a more extensive evaluation of the framework.
+Again, each configuration describe the deployment of containers in a consistent
+and portable way, an it was used in each phase of the life cycle, from
+development through testing to production independently on the kind of
+infrastructure.
+
 % Emphasize these results for the purpose of the paper - JJ
+% done - Mario
+
+\section{Experimental Study} 
+\label{setup} % To be renamed
 
@@
 % this experiment - JJ  
-
-\section{Experimental Study} 
-\label{setup} % To be renamed
-
-We aimed to learn if the proposed solution can efficiently
+% Connected below - M
+
+In the prior section, we wanted to prove that EvoSwarm can be used by 
+the public to create reproductive research, but in this section are
+using the same framework to learn if the proposed solution can efficiently
 improve the scalability and performance of population-based optimization
@@
 % For instance. But I was actually looking for another word that does not imply a comparison - JJ
-Hence, in the following sections, we verify the following:
+Hence, in the following sections, we verify the following questions:
 
@@
 \item Can we improve the execution time of the algorithm by adding populations
- and serverless functions to the system and, in particular, what is its effect on the scalability of the system?
-% Improve with respect to what?
-Single population?
-Please explain carefully the issues and how scaling is related to the number of populations.
-- JJ
+ and serverless functions (workers)  to the system and, in particular,
+ what is its effect on the scalability of the system? % Improve with respect to what? Single population? 
+                         %Please explain carefully the issues and how scaling is related to the number of populations. - JJ
+                         % The question could be that, How scaling is related to number of populations and workers in the system ? - M
 
@@
 search by needing fewer function evaluations than a homogeneous setting? 
-
 % Moved reproducibility question from here - M
+
 \end{itemize}
+
+Section~\label{sec:exp1} is devoted to answering the first question, and Section~\label{sec:exp2} to the second.
+Next, we describe the general experimental setup to both experiments.
 
@@
 % BTW, they should probably be subsections.
+% Done - M
 
 \subsection{General Experimental Setup} 
-\label{sec:general-setup} % To be renamed
+\label{sec:general-setup}
 
@@
 files the EvoSwarm application uses to run the noiseless BBOB testbed
-experiments as seen in Section~\ref{sec:exeperimentalDefinition}. 
+experiments as seen in Section~\ref{sec:evoswarm:config}. 
 As a first step, we must deploy the docker application using a
@@
 
-For local experiments we have deployed the EvoSwarm application in a
-Desktop PC with an AMD Ryzen 9 3900x 12-core processor with 24 threads with a 3.8 GHz base clock, 
-and 16 GB RAM. We used Docker version 19.03.3, build a872fc2f86, and {\tt docker-compose} version 1.21.0, 
-in Ubuntu Linux 18.04; coce is run using Python 3.7.5. Container images and the {\tt docker-compose} file are available at
-\url{https://hub.docker.com/u/mariosky}, and \url{https://github.com/mariosky/EvoSwarm} respectively. 
-The BOBB experiments were performed with COCO \cite{hansen2016coco} version bbob.v15.03 in Python, 
-the plots were produced with version 2.3.2.
+We have deployed the EvoSwarm application locally in the same hardware and
+software described in Section~\ref{setup:local-cloud}, additionally the BOBB
+experiments were performed with COCO \cite{hansen2016coco} version bbob.v15.03
+in Python, the plots were produced with version 2.3.2.
 
 % Maybe subsection? Also say the question you're answering - JJ
+% Done, I added the research question at the beginning of the setup question - M
 \subsection{Measuring Scalability}
@@
 % Experiment 1
-As we mentioned before % say where - JJ
-, reactive systems can scale by adding additional copies
+
+As we mentioned before in Section~\ref{method}, reactive systems can scale by adding additional copies
 of serverless functions. In our case, we can start additional worker containers
 to have the same effect. Other authors, like Salza et al. \cite{salza2019speed},
-have used a
-similar architecture, which implies that worker nodes need to have at least a certain level
+have used a similar architecture, which implies that worker nodes need to have at least a certain level
 of complexity, in terms of execution time,  to effectively scale on multiple
@@
 
-To evaluate the effect of these parameters on the scalability of the system, we
+To answer the first research question, we need to evaluate the effect of these parameters on the scalability of the system,   we
 propose  an experiment in which we selected $f_4$ (Skew Rastrigin-Bueche separable) from the
@@
   instances of function $f_4$, in seconds, different number of workers and populations sizes.  }
-  % Which function? - JJ
   \label{fig:spworker}
@@
 implemented by just changing the configuration parameters of each population,
-as in the previous experiment, but in this case, we are interested in a heterogeneous search strategy.
+as in the previous experiment. But in this case, we are interested in 
+answering the second research question, and we want to validate if by using 
+heterogeneous populations and mixing search strategies,  we can increase the performance of the
+search.
 

### 2020-05-16 draft.tex
Some more comments deleted
Speedup results

@@
 % Maybe here you should make reference to the sections where you are actually doing it (after conveniently rephrasing it) - JJ
+% These highlights go in the web page of the paper, they need to be short and to the point, 
+% I mention them briefly when presenting the paper's sections - M
 \begin{itemize}
@@
 
-% Also, multipopulation design of what?x
 Multi-population based methods divide the original population into
@@
 its subpopulation.
-% JJ, help I don't know if this is clearly understood, a mean like a
-% DNS - Mario
-% Added something - JJ
 
@@
 
-% The main advantage of container deployments over virtual machine
-% instances is that they add much less overhead to communication between
-% them, and they are much more economical to deploy. Scaling and
-% monitoring can be done much more easily through cloud orchestration
-% systems such as Kubernetes; these are the reasons why we chose it for
-% this work.
-% I don't see this line of argument. You can do exactly the same with
-% virtual machines instead of containers. Also, you can test a virtual
-% infrastructure locally and then only when it's needed deploy it to
-% the cloud. If you want to give good reasons for using containers
-% instead of virtual machines, there are quite a lot non-economic ones
-% - JJ
-% OK, I will add an issue - Mario
-% Did you finally? - JJrehe
-% I did not, will add and assign it - M
 There are many advantages of container-based architectures over one
@@
 \end{figure}
-
-% Maybe you should say here how everything works: launch with
-% Docker-compose, whatever. That's the issue that one of the reviewers
-% say - JJ
-
-% This issue is also commanted in the experimental section
-% I will explain it and then we can see where it goes. - M
 
@@
 
-% You said "subsections will talk about roles", this is not a role. 
-% Change the title to that of a role. Would that be administrator? - JJ
-
 \subsection{EvoSwarm from the User Point of View} 
-% All in a single section - Mario
-%\subsubsection{Experiment Configuration and Execution}
 \label{sec:evoswarm:config}
@@
 % TODO: connection to cloud subsection  - M
+% Done - M
 
@@
 cloud provider, we have some more options to choose:
-
-%Container-based applications describe the deployment of loosely coupled
-%processes in a consistent and portable way, in every phase of the life cycle of
-%an application, from development through testing to production independently on
-%the kind of infrastructure. The term cloud-native describes these
-%container-based environments because they are deployed on an infrastructure that
-%abstracts the underlying compute, storage, and networking primitives. This
-%infrastructure does need to be a cloud provider; it could be a developer's
-%laptop, a single server, or a production environment. When we deploy a
-%cloud-native application to a cloud provider, we have some more options to
-%choose:
 
@@
 % Write some introduction to this subsubsection.
-
+% Done - M
 In this experiment, we wanted to test several deployment options for the 
@@
 As we can see, the local environment (see section \ref{setup}) achieved a higher
-ratio of evaluations per second (in bold). Still, we must consider that there is a minimum
+ratio of evaluations per second (in bold). Still, we must consider that there is minimum
 network overhead in a local workstation and a clock speed of 3.8 GHz. The first
 option (labeled Single) is straightforward to deploy, needing only a single
-{\tt docker-compose} file, and one command to execute. But with the limit of 4CPUs and
+{\tt docker-compose} file, and one command to execute. But with the limit of 4 CPUs and
 ten containers, it is the slowest option. The second option (labeled Multi), is
 faster, and not too far from the local execution, especially when solving the
-$f_4$ in 20 dimensions (underline in table). With the capacity for increasing the number of worker
+$f_4$ in 20 dimensions (underlined in table). With the capacity for increasing the number of worker
 containers, it is reasonable to think that we can achieve a higher number of
@@
 % Mention costs - JJ
+% Done  - M
 
@@
   20  & Local       & \textbf{36,101.86}   \\
-      & Multi task  & {\ul 33,916.02}       \\ % why the underline? [Best between cloud options :) - Mario] Then why not above too? - JJ
+      & Multi task  & {\ul 33,916.02}       \\ 
       & Single task & 16,546.95  \\
@@
 (COCO) framework \cite{hansen2016coco}. The testbed includes 24 real-parameter,
-single-objective benchmark functions, and the capability to provide additional
+single-objective ben chmark functions, and the capability to provide additional
 instances of each function. Each instance of a function has a different optimal
@@
   instances of function $f_4$, in seconds, different number of workers and populations sizes.  }
-  \label{fig:spworker}
+  \label{fig:spworker:time}
 \end{figure}
@@
   \caption{Boxplot of the number of evaluations per second, for 30 instances of function $f_4$, with different number of workers and populations sizes }
-  \label{fig:spworker}
+  \label{fig:spworker:evalspersecond}
 \end{figure}
@@
   \multicolumn{1}{|l|}{Dimension} & \multicolumn{1}{c|}{2W} & \multicolumn{1}{c|}{4W} & \multicolumn{1}{c|}{8W} & \multicolumn{1}{c|}{2W} & \multicolumn{1}{c|}{4W} & \multicolumn{1}{c|}{8W} \\ \hline
-  \multicolumn{1}{|c|}{2}         & 1.94                    & 3.12                    & 3.86                    & 1.84                    & 3.38                    & 5.98                    \\ \hline
-  \multicolumn{1}{|c|}{3}         & 1.93                    & 2.77                    & 3.15                    & 1.90                    & 3.26                    & 5.96                    \\ \hline
-  \multicolumn{1}{|c|}{5}         & 1.91                    & 2.90                    & 3.09                    & 1.91                    & 3.46                    & 6.14                    \\ \hline
-  \multicolumn{1}{|c|}{10}        & 1.92                    & 3.03                    & 3.30                    & 1.90                    & 3.61                    & 6.41                    \\ \hline
-  \multicolumn{1}{|c|}{20}        & 1.91                    & 3.14                    & 3.27                    & 1.89                    & 3.73                    & 6.71                    \\ \hline
-  \multicolumn{1}{|c|}{40}        & 2.00                    & 2.92                    & 3.12                    & 1.95                    & 3.80                    & 7.52                    \\ \hline
+  \multicolumn{1}{|c|}{2}         & \textbf{1.94}           & 3.12                    & 3.86                    & 1.84                    & \textbf{3.38}                    & \textbf{5.98}                    \\ \hline
+  \multicolumn{1}{|c|}{3}         & \textbf{1.93}           & 2.77                    & 3.15                    & 1.90                    & \textbf{3.26}                   & \textbf{5.96}                    \\ \hline
+  \multicolumn{1}{|c|}{5}         & 1.91                    & 2.90                    & 3.09                    & 1.91                    & \textbf{3.46}                    & \textbf{6.14}                    \\ \hline
+  \multicolumn{1}{|c|}{10}        & \textbf{1.92}           & 3.03                    & 3.30                    & 1.90                    & \textbf{3.61}                   & \textbf{6.41}                    \\ \hline
+  \multicolumn{1}{|c|}{20}        & \textbf{1.91}           & 3.14                    & 3.27                    & 1.89                    & \textbf{3.73}                   & \textbf{6.71}                    \\ \hline
+  \multicolumn{1}{|c|}{40}        & \textbf{2.00}           & 2.92                    & 3.12                    & 1.95                    & \textbf{3.80}                     & \textbf{7.52}                    \\ \hline
   \end{tabular}
@@
   Dimension & \multicolumn{2}{l}{Increment} & Speedup         & p-value         & Speedup          & p-value         \\
-  2         & 1w            & 2w            & 1.94            & 6E-16           & 1.84             & 2E-16           \\
+  2         & 1w            & 2w            & {\ul 1.94}            & 6E-16           & 1.84             & 2E-16           \\
             & 2w            & 4w            & 1.61            & 2E-15           & 1.83             & 1E-15           \\
             & 4w            & 8w            & 1.24            & 1E-06           & 1.77             & 2E-16           \\
-  3         & 1w            & 2w            & 1.93            & 2E-16           & 1.90             & 2E-16           \\
+  3         & 1w            & 2w            & {\ul 1.93}            & 2E-16           & 1.90             & 2E-16           \\
             & 2w            & 4w            & 1.44            & 8E-15           & 1.71             & 2E-16           \\
@@
             & 2w            & 4w            & 1.58            & 3E-16           & 1.89             & 2E-16           \\
-            & 4w            & 8w            & 1.09            & 1E-03           & 1.78             & 2E-16           \\
+            & 4w            & 8w            & {\ul 1.09}      & 1E-03           & 1.78             & 2E-16           \\
   20        & 1w            & 2w            & 1.91            & 2E-16           & 1.89             & 2E-16           \\
             & 2w            & 4w            & 1.64            & 2E-15           & 1.97             & 2E-16           \\
-            & 4w            & 8w            & 1.04            & 1E-02           & 1.80             & 2E-16           \\
-  40        & 1w            & 2w            & 2.00            & 2E-16           & 1.95             & 2E-16           \\
+            & 4w            & 8w            & {\ul 1.04 }     & 1E-02           & 1.80             & 2E-16           \\
+  40        & 1w            & 2w            & \textbf{2.00}   & 2E-16           & 1.95             & 2E-16           \\
             & 2w            & 4w            & 1.46            & 1E-09           & 1.95             & 2E-16           \\
-            & 4w            & 8w            & 1.07            & 3E-02           & 1.98             & 2E-16          
+            & 4w            & 8w            & {\ul 1.07}            & 3E-02           & \textbf{1.98}    & 2E-16          
   \end{tabular}
@@
 
-Figure \ref{fig:spworker} shows the execution time achieved by a certain number of
+Figure \ref{fig:spworker:time} shows the execution time achieved by a certain number of
 workers, for every problem dimension, which, besides, changes the
@@
 
+To measure the speedup obtained by the addition of workers, we will use the
+number of evaluations per second, instead of the time required to finish an
+instance of a problem. The reason for this is that in some cases reaching the
+desired target can require less work. We are going to measure the total increase
+of work done in a second by adding more workers. The baseline for the comparison
+will be one worker, and the minimum amount of work available will be five
+populations. Five populations are selected because the controller needs at least
+three populations to migrate individuals between them, and we add two more to
+keep even a single worker busy.Figure \ref{fig:spworker:evalspersecond} 
+shows the number of function evaluation per second, achieved by a certain number of
+workers, for every problem dimension, for five and ten dimensions of $f_4$.
+
+As expected, within the same number of populations, there is a linear speedup
+when we add more workers. The speedup is marginal in the case of five
+populations when increasing the number of workers above 4. The reason for this
+is that at any given time, there are at least two populations in the message
+queue, leaving four workers waiting. We can also notice that the spread in the
+amount of work is less when using up to four workers.
+
+Table~\ref{tab:speedup-table} shows the speedup obtained against the baseline of
+one worker. When using eight workers, we achieved the maximum speed up for five
+and ten populations, but interestingly, it was in dimension 40 and dimension 2 
+(these values are shown in bold in the table).
+
+Table~\ref{tab:speedup:test} shows the speedup obtained taking as the baseline
+the previous configuration, for example, when increasing from four to eight
+workers, and the p-rate for the left-tailed Wilcoxon test. There was an
+increment in all cases, but again, for smaller dimensions, five populations with
+fewer workers is a better option (see values underlined in the table) with the best  
+for 40 dimensions (in bold). The worst speed was going from four to eight workers 
+at higher dimensions (underlined in the table). For ten populations, overall 
+has the linear speedup but the best is from four to eight workers at 40 dimensions
+(shown in bold).
+
 % No idea where this fits in. I think it's the second question in 7, 

### 2020-05-16 draft.tex
Mainly deleting addressed comments
Some rehashing of images, and more questions

@@
 
-% Maybe these three sections should be merged; they are all experimental set up and the two next ones start with "experimental"
-% (and he next one too) - JJ
-% done - Mario
-% But also change title? - JJ
-% Experimental design of a container-based application or something like that - JJ 
-% Like it - M
 \section{Experimental Design of a Container-based Application} 
@@
 
-% TODO: connection to cloud subsection  - M
-% Done - M
-
 \subsection{Deploying to a Cloud Provider}
@@
 
-\subsubsection{Setup}
+\subsubsection{Setup} % Even if it's a sub-sub section, flesh this out ... - JJ
 \label{setup:local-cloud}
@@
 % Remember to mention cost at the end - JJ
-
+% Ping - JJ
 We can define an ECS cluster directly on AWS Elastic Cloud VMs or through the AWS Fargate service.
@@
 
+% Mention cost here - JJ
+
 \subsubsection{Experimental Results}
 
-% Write some introduction to this subsubsection.
-% Done - M
 In this experiment, we wanted to test several deployment options for the 
@@
 
-% Mention costs - JJ
-% Done  - M
-
 \begin{figure}[h!tbp]
@@
 
-% Emphasize these results for the purpose of the paper - JJ
-% done - Mario
-
 \section{Experimental Study} 
 \label{setup} % To be renamed
-
-% You should clarify that, before, you were explaining how to set up
-% an experiment. In this section, you are explaining the experiments
-% that you are doing in this paper. But you should relate both
-% sections to the general objectives of the paper. In the prior
-% section, you want to prove that EvoSwarm can be used by the public
-% to create reproductive research (and maybe you should add a small
-% conclusion sub or sub-sub section), in the next you explain the
-% experiment we are doing for this paper. It would be interesting to
-% make references from the next section to the previous one,
-% indicating how the general setup of EvoSwarm helps _us_ to create
-% this experiment - JJ  
-% Connected below - M
 
@@
 Next, we describe the general experimental setup to both experiments.
-
-% Also, point forward to the section (or subsections) that will be devoted to proving each point.
-% BTW, they should probably be subsections.
-% Done - M
 
@@
 algorithm parameters were set. In the GA, we randomly set the mutation and
-crossover probabilities to have more heterogeneous workers; % highlight this and add citation - JJ 
-% done below - M
+crossover probabilities to have more heterogeneous workers; 
 in these parameters, the range of values is specified. We did not change these parameters during the 
@@
 
-% Maybe subsection? Also say the question you're answering - JJ
-% Done, I added the research question at the beginning of the setup question - M
 \subsection{Measuring Scalability}
@@
 
-\subsubsection{Setup}
+\subsubsection{Setup} % Again, this in the index with no context should say something. Please flesh it out.
 
@@
 
+%\subsubsection{Experimental Results} 
+%\label{results}
+
+In this experiment, we wanted to test how the number of workers affected the
+time needed to reach a solution and the speedup that each configuration can
+produce. Adding more populations might add more overhead, so this might have a
+noticeable effect. In principle, the main issue is how to find the right number
+of workers for a particular problem, but in general, we wanted to check the
+communication overhead when the size of the problem increases and how the
+presented framework reacted to it.    % Added some explanation % here - JJ
+We report the comparison regarding the execution time in Section~\ref{sec:exec-time}
+and the summary of the speedup in section Section~ref{sec:speedup}.
+
+\subsubsection{Execution Time}
+\label{sec:exec-time}
+
+Figure \ref{fig:spworker:time} shows the execution time achieved by a certain number of
+workers, for every problem dimension, which, besides, changes the
+population size.
+Each box-plot indicates the time required for the execution of 30 function
+instances, in seconds. Each row has a different scale on the y-axis, to
+compensate for the dimension increments. It is important to notice that the
+execution time can also be affected by the algorithm finding a solution with an
+error smaller than the threshold, before completing all the FEs.
+
+In Table~\ref{tab:time} we can see that as expected, and in most cases, 
+increasing the number of workers reduces the time
+required for the execution of an instance, 
+mainly when using the same parameters
+(number of populations and number of workers), that is, performance
+scales with the number of workers for a fixed problem and number of
+sub-populations used. The exception is the case when using five populations;
+the time does not decrease when progressing from four to eight workers.  
+That is because most of the time, workers are waiting for a population 
+to be available in the message queue.
+
+However, when we compare the two
+options, we find some subtle differences: When using one or two
+workers, using 5 populations takes less time in most cases. 
+In Table~\ref{tab:time} results with p-value of the two-tailed Wilcoxon
+signed-ranks test was grater than the level of significance of 0.05 
+(i.e., accepting the null hypothesis of inequality) are marked in boldface.
+A reason for this could be that the migration of individuals takes more time to
+reach the worker. % how come? - JJ
+For instance, when a single worker has 10 populations, only
+after processing every one of them, it will start receiving mixed populations.  Although, in
+all cases, and even with a single worker, populations are heterogeneous, this
+advantage takes more time to propagate. % which advantage? - JJ
+With four workers, almost the same
+results can be achieved regardless of our selection between five or ten
+populations. Also, there is not much difference between 4 and 8 workers, when
+using five populations, which might indicate that some workers are
+idling while populations are being processed by other workers. In
+general, we can observe that keeping a ratio between the number of workers and populations a bit over 1:1  exploits better the computing
+resources available. This also decreases the relation between the
+overhead needed to migrate and process incoming and outgoing
+population and the total computing time: more populations will need
+more time to do that.
+
 %
@@
 \end{figure}
+%
+\subsubsection{Speedup}
+\label{sec:speedup}
+
+% Introduction saying why we are measuring this, how it's related to the objectives of the paper - JJ
+To measure the speedup obtained by the addition of workers, we will use the
+number of evaluations per second, instead of the time required to finish an
+instance of a problem. The reason for this is that in some cases reaching the
+desired target can require less work. We are going to measure the total increase
+of work done in a second by adding more workers. The baseline for the comparison
+will be the algorithm with a single worker, and the minimum amount of work available will be five
+populations. This number is selected because the controller needs at least
+three populations to migrate individuals between them, and we add two more to
+keep even a single worker busy. Figure \ref{fig:spworker:evalspersecond} 
+shows the number of function evaluation per second, achieved by a certain number of
+workers, for every problem dimension, for five and ten dimensions of $f_4$.  % Why did you choose f4? - JJ
 
@@
 %
+As expected, within the same number of populations, there is a linear speedup
+when we add more workers. The speedup is marginal in the case of five
+populations when increasing the number of workers above 4. The reason for this
+is that at any given time, there are at least two populations in the message
+queue, leaving four workers waiting. We can also notice that the spread in the
+amount of work is less when using up to four workers.
 
@@
   \caption{Speedup achieved with incrementing the number of workers for
-   each dimension showing the p-rate for the Wilcoxon test}
-  % mark significant differences in some way - JJ
+    each dimension showing the p-rate for the Wilcoxon test.}
+  % Say what highlights mean - JJ
   \label{tab:speedup:test}
@@
   \end{table}
-
-%\subsubsection{Experimental Results} 
-%\label{results}
-
-In this experiment, we wanted to test how the number of workers affected the
-time needed to reach a solution and the speedup that each configuration can
-produce. Adding more populations might add more overhead, so this might have a
-noticeable effect. In principle, the main issue is how to find the right number
-of workers for a particular problem, but in general, we wanted to check the
-communication overhead when the size of the problem increases and how the
-presented framework reacted to it.    % Added some explanation % here - JJ
-We report the comparison regarding the execution time in Section~\ref{sec:exec-time}
-and the summary of the speedup in section Section~ref{sec:speedup}.
-
-\subsubsection{Execution Time}
-\label{sec:exec-time}
-
-Figure \ref{fig:spworker:time} shows the execution time achieved by a certain number of
-workers, for every problem dimension, which, besides, changes the
-population size.
-Each box-plot indicates the time required for the execution of 30 function
-instances, in seconds. Each row has a different scale on the y-axis, to
-compensate for the dimension increments. It is important to notice that the
-execution time can also be affected by the algorithm finding a solution with an
-error smaller than the threshold, before completing all the FEs.
-
-In Table~\ref{tab:time} we can see that as expected, and in most cases, 
-increasing the number of workers reduces the time
-required for the execution of an instance, 
-mainly when using the same parameters
-(number of populations and number of workers), that is, performance
-scales with the number of workers for a fixed problem and number of
-sub-populations used. The exception is the case when using five populations;
-the time does not decrease when progressing from four to eight workers.  
-That is because most of the time, workers are waiting for a population 
-to be available in the message queue.
-
-However, when we compare the two
-options, we find some subtle differences: When using one or two
-workers, using 5 populations takes less time in most cases. 
-In Table~\ref{tab:time} results with p-value of the two-tailed Wilcoxon
-signed-ranks test was grater than the level of significance of 0.05 
-(i.e., accepting the null hypothesis of inequality) are marked in boldface.
-A reason for this could be that the migration of individuals takes more time to
-reach the worker. For instance, when a single worker has 10 populations, only
-after processing every one of them, it will start receiving mixed populations.  Although, in
-all cases, and even a single worker, populations are heterogeneous, this
-advantage takes more time to propagate. With four workers, almost the same
-results can be achieved regardless of our selection between five or ten
-populations. Also, there is not much difference between 4 and 8 workers, when
-using five populations, which might indicate that some workers are
-idling while populations are being processed by other workers. In
-general, we can observe that keeping a ratio close to 1:1, but a bit
-better than that, between the number of workers and populations, better exploits the computing
-resources available. This also decreases the relation between the
-overhead needed to migrate and process incoming and outgoing
-population and the total computing time: more populations will need
-more time to do that.
-
-\subsubsection{Speedup}
-\label{sec:speedup}
-
-% TBD? - JJ
-
-To measure the speedup obtained by the addition of workers, we will use the
-number of evaluations per second, instead of the time required to finish an
-instance of a problem. The reason for this is that in some cases reaching the
-desired target can require less work. We are going to measure the total increase
-of work done in a second by adding more workers. The baseline for the comparison
-will be one worker, and the minimum amount of work available will be five
-populations. Five populations are selected because the controller needs at least
-three populations to migrate individuals between them, and we add two more to
-keep even a single worker busy.Figure \ref{fig:spworker:evalspersecond} 
-shows the number of function evaluation per second, achieved by a certain number of
-workers, for every problem dimension, for five and ten dimensions of $f_4$.
-
-As expected, within the same number of populations, there is a linear speedup
-when we add more workers. The speedup is marginal in the case of five
-populations when increasing the number of workers above 4. The reason for this
-is that at any given time, there are at least two populations in the message
-queue, leaving four workers waiting. We can also notice that the spread in the
-amount of work is less when using up to four workers.
-
+%
 Table~\ref{tab:speedup-table} shows the speedup obtained against the baseline of
@@
 increment in all cases, but again, for smaller dimensions, five populations with
-fewer workers is a better option (see values underlined in the table) with the best  
-for 40 dimensions (in bold). The worst speed was going from four to eight workers 
+fewer workers is a better option (see values underlined in the table) with the best speedup, achieved for 40 dimensions, shown in bold. The worst speed was going from four to eight workers 
 at higher dimensions (underlined in the table). For ten populations, overall 
-has the linear speedup but the best is from four to eight workers at 40 dimensions
+has the linear speedup % don't understand this - JJ
+but the best is from four to eight workers at 40 dimensions
 (shown in bold).
@@
 \label{sec:exp2}
+
+% Write an introduction - JJ
 

### 2020-05-16 draft.tex
Update draft.tex

@@
 % Mention cost here - JJ
+% I putted earlier literally at the end of the section :) - M
 
@@
 
+From these parameters, we give more importance in this paper to the number of
+populations, because they are the basic unit of work that can be executed in
+parallel. Together with the size of each population and the number of
+generations, this parameter determines the maximum number of function
+evaluations for a single iteration. We do not have control over each algorithm's
+complexity, but future analyses could focus on balancing populations according
+to this.
+
+In this experiment, we evaluate how the number of populations and the number of
+workers in the system are related to the scalability of the system. In the next
+section, we describe the experimental setup, and a discussion of the results
+follows, with respect to execution time (Section~\ref{sec:exec-time}) 
+and speedup Section~\ref{sec:speedup}.
+
 \subsubsection{Setup} % Again, this in the index with no context should say something. Please flesh it out.
-
-To answer the first research question, we need to evaluate the effect of these parameters on the scalability of the system,   we
+ % I will add more context in the parent section, and continue here
+
+To answer the first research question, we need to evaluate the effect of 
+the number of populations and the number of
+workers on the scalability of the system, we 
 propose  an experiment in which we selected $f_4$ (Skew Rastrigin-Bueche separable) from the
@@
 \cite{nicolau2009application} to solve with the required FEs. As we
-are only comparing in terms of execution time, the results from a single function
+are only comparing in terms of scalability, the results from a single function
 can be better understood, since there are fewer factors involved.
-Following the procedure described in the previous section, we executed 
+Following the procedure described in Section~\ref{setup}, we executed 
 two sets of experiments with five and ten initial populations,
@@
 
+% Added some explanation % here - JJ
+
+%We report the comparison regarding the execution time in Section~\ref{sec:exec-time}
+%and the summary of the speedup in section Section~ref{sec:speedup}.
+% Move this to parent - M
+
+\subsubsection{Results Regarding Execution Time}
+\label{sec:exec-time}
+
+% I think this is now a bit redundant or we need to move it to the parent section - M
+% But is a good intro to the results.
+
 In this experiment, we wanted to test how the number of workers affected the
@@
 produce. Adding more populations might add more overhead, so this might have a
-noticeable effect. In principle, the main issue is how to find the right number
-of workers for a particular problem, but in general, we wanted to check the
-communication overhead when the size of the problem increases and how the
-presented framework reacted to it.    % Added some explanation % here - JJ
-We report the comparison regarding the execution time in Section~\ref{sec:exec-time}
-and the summary of the speedup in section Section~ref{sec:speedup}.
-
-\subsubsection{Execution Time}
-\label{sec:exec-time}
+noticeable effect.
+
+%In principle, the main issue is how to find the right number
+%of workers for a particular problem, but in general, we wanted to check the
+%communication overhead when the size of the problem increases and how the
+%presented framework reacted to it.
+% Commenting because we did not checked the communication over head - M
 
@@
 %
-\subsubsection{Speedup}
+\subsubsection{Results Regarding Speedup}
 \label{sec:speedup}
@@
 % Introduction saying why we are measuring this, how it's related to the objectives of the paper - JJ
+% This is done on the parent section, I renamed the sections to clarify this - M
 To measure the speedup obtained by the addition of workers, we will use the
@@
 shows the number of function evaluation per second, achieved by a certain number of
-workers, for every problem dimension, for five and ten dimensions of $f_4$.  % Why did you choose f4? - JJ
+workers, for every problem dimension, for five and ten dimensions of $f_4$.  % Why did you choose f4? - JJ This is mentioned on the setup - M
 
@@
 fewer workers is a better option (see values underlined in the table) with the best speedup, achieved for 40 dimensions, shown in bold. The worst speed was going from four to eight workers 
-at higher dimensions (underlined in the table). For ten populations, overall 
-has the linear speedup % don't understand this - JJ
-but the best is from four to eight workers at 40 dimensions
+at higher dimensions (underlined in the table). 
+For ten populations, adding more workers resulted in a linear speedup, % don't understand this - JJ % Me either - M, rewritten  
+but the best speedup is obtained when increasing from four to eight workers at 40 dimensions
 (shown in bold).
@@
 % so it should probably be a subsection - JJ
+% done - M
 
@@
 % Write an introduction - JJ
+% I removed the Setup section, because we use the same configuration as before
+% I will mention something about it - M
+% mm but I don't like to have a single subsection, as I remember is not right.
 
@@
 
-\subsubsection{Setup}
+%\subsubsection{Setup}
 
@@
 
-Therefore, in this experiment we compare a multi-population with  GA- and PSO- only populations,
+Therefore, in this experiment we compare a multi-population with GA and PSO only populations,
 versus an ensemble of GA and PSO algorithms. We have tested on the first five functions of the
@@
 same parameters as before.
+
+In order to run this experiment, we just needed to edit a configuration file and run 
+the experiments. The plots presented in the results are generated by the BBOB framework
+from the JOSN log file of each run. There was no additional setup needed.
 

### 2020-05-16 draft.tex
Two last questions

@@
 (i.e., accepting the null hypothesis of inequality) are marked in boldface.
-A reason for this could be that the migration of individuals takes more time to
-reach the worker. % how come? - JJ
+A reason for this could be that when having more populations, the migration
+of individuals takes more time to reach to all of them.
+% how come? - JJ  - Explained a bit more - M
 For instance, when a single worker has 10 populations, only
 after processing every one of them, it will start receiving mixed populations.  Although, in
-all cases, and even with a single worker, populations are heterogeneous, this
-advantage takes more time to propagate. % which advantage? - JJ
+all cases, and even with a single worker, populations are heterogeneous, but
+mixed individuals take more time to propagate. % which advantage? - JJ - Individuals from other pops - M
 With four workers, almost the same

### 2020-05-17 draft.tex
Rounding out and eliminating comments

@@
 A reason for this could be that when having more populations, the migration
-of individuals takes more time to reach to all of them.
-% how come? - JJ  - Explained a bit more - M
+of individuals takes more time from the beginning to reach to all of them.
 For instance, when a single worker has 10 populations, only
 after processing every one of them, it will start receiving mixed populations.  Although, in
-all cases, and even with a single worker, populations are heterogeneous, but
-mixed individuals take more time to propagate. % which advantage? - JJ - Individuals from other pops - M
+all cases, and even with a single worker, populations are heterogeneous, 
+ individuals resulting from the mixing of populations take more time to propagate. 
 With four workers, almost the same

### 2020-05-17 draft.tex
Minor typographic changes

@@
 
-% Mention cost here - JJ
-% I putted earlier literally at the end of the section :) - M
-
 \subsubsection{Experimental Results}
@@
     Population Size  & 50 & 60 & 60 & 70 & 100 & 125 \\ \hline
-    Populations      & 10 & 10 & 10 & 10 & 10  & 10  \\ \hline
-    Iterations       & 10 & 20 & 30 & 30 & 30  & 40  \\ \hline  
+    Populations      & \multicolumn{6}{c|}{10} \\ \hline
+    Iterations       & 10 & 20 & \multicolumn{3}{c|}{30}  & 40  \\ \hline  
   \end{tabular}
@@
     Population Size  & 100 & 120 & 120 & 140 & 200 & 250 \\ \hline
-    Populations      & 5 & 5 & 5 & 5 & 5  & 5  \\ \hline
-    Iterations       & 10 & 20 & 30 & 30 & 30  & 40  \\ \hline
+    Populations      & \multicolumn{6}{c|}{5} \\ \hline
+    Iterations       & 10 & 20 & \multicolumn{3}{c|}{30}  & 40  \\ \hline
   \end{tabular}

### 2020-05-17 draft.tex
Deleting addressed comments and adding some results

@@
 
-From these parameters, we give more importance in this paper to the number of
-populations, because they are the basic unit of work that can be executed in
+From this description it can be deduced that we give more importance in this paper to the number of
+populations, since they are the basic unit of work that can be executed in
 parallel. Together with the size of each population and the number of
@@
 evaluations for a single iteration. We do not have control over each algorithm's
-complexity, but future analyses could focus on balancing populations according
+complexity, but future analyses could focus on balancing populations and number of worker nodes according
 to this.
@@
 % But is a good intro to the results.
+% OK - JJ
 
@@
 shows the number of function evaluation per second, achieved by a certain number of
-workers, for every problem dimension, for five and ten dimensions of $f_4$.  % Why did you choose f4? - JJ This is mentioned on the setup - M
+workers, for every problem dimension, for five and ten dimensions of $f_4$.
 
@@
 at higher dimensions (underlined in the table). 
-For ten populations, adding more workers resulted in a linear speedup, % don't understand this - JJ % Me either - M, rewritten  
+For ten populations, adding more workers resulted in a linear speedup, 
 but the best speedup is obtained when increasing from four to eight workers at 40 dimensions
 (shown in bold).
-
-% No idea where this fits in. I think it's the second question in 7, 
-% so it should probably be a subsection - JJ
-% done - M
 
@@
 
-In order to run this experiment, we just needed to edit a configuration file and run 
-the experiments. The plots presented in the results are generated by the BBOB framework
-from the JOSN log file of each run. There was no additional setup needed.
+In order to run this experiment, we just needed to edit a configuration file and start it up. The plots presented in the results are generated by the BBOB framework
+from the JOSN log file for each run.
+There was no additional setup needed, proving how this framework is able to not only offer an reproducible way to run experiments, but also how easily it interfaces with other common tools in the EC ecosystem.
 

### 2020-05-18 draft.tex
Added conclusions and reflowed

@@
 pattern is an essential component of a reactive architecture and makes
-it 
-highly scalable; one of the reasons for this is the use of functions
+it highly scalable; one of the reasons for this is the use of functions
 that have no secondary effects or access global storage, that is,
@@
 clear and idempotent description of the architecture through {\tt docker-compose} 
-% Be consistent with docker compose. Either {\tt docker-compose} or {\sf docker-compose}, but not docker compose, for instance. - JJ - ok
 are the main reasons.
@@
 In this experiment, we wanted to test several deployment options for the
-cloud-native application, and have an idea of the computing capacity of each option
-measured as the number of function evaluations per second.
-The results are shown in Figure~\ref{fig:aws-results} and Table~\ref{tab:aws}. 
-As we can see, the local environment (see section \ref{setup}) achieved a higher
-ratio of evaluations per second (in bold). Still, we must consider that there is minimum
+cloud-native application, and have an idea of the computing capacity of each
+option measured as the number of function evaluations per second. The results
+are shown in Figure~\ref{fig:aws-results} and Table~\ref{tab:aws}. As we can
+see, the local environment (see section \ref{setup}) achieved a higher ratio of
+evaluations per second (in bold). Still, we must consider that there is minimum
 network overhead in a local workstation and a clock speed of 3.8 GHz. The first
-option (labeled Single) is straightforward to deploy, needing only a single
-{\tt docker-compose} file, and one command to execute. But with the limit of 4 CPUs and
-ten containers, it is the slowest option. The second option (labeled Multi), is
-faster, and not too far from the local execution, especially when solving the
-$f_4$ in 20 dimensions (underlined in table). With the capacity for increasing the number of worker
-containers, it is reasonable to think that we can achieve a higher number of
-evaluations with an appropriate configuration. Nevertheless, the goal of this
-experiment was not to increase the number of evaluations but to test the
-flexibility of a cloud-native application. The total cost of testing, configuration, 
-and execution of the experiments in ECS was less than \$1.00 US Dollar.
+option (labeled Single) is straightforward to deploy, needing only a single {\tt
+docker-compose} file, and one command to execute. But with the limit of 4 CPUs
+and ten containers, it is the slowest option. The second option (labeled Multi),
+is faster, and not too far from the local execution, especially when solving the
+$f_4$ in 20 dimensions (underlined in table). With the capacity for increasing
+the number of worker containers, it is reasonable to think that we can achieve a
+higher number of evaluations with an appropriate configuration. Nevertheless,
+the goal of this experiment was not to increase the number of evaluations but to
+test the flexibility of a cloud-native application. The total cost of testing,
+configuration, and execution of the experiments in ECS was less than \$1.00 US
+Dollar.
 
@@
 
-From this description it can be deduced that we give more importance in this paper to the number of
-populations, since they are the basic unit of work that can be executed in
-parallel. Together with the size of each population and the number of
-generations, this parameter determines the maximum number of function
+From this description it can be deduced that we give more importance in this
+paper to the number of populations, since they are the basic unit of work that
+can be executed in parallel. Together with the size of each population and the
+number of generations, this parameter determines the maximum number of function
 evaluations for a single iteration. We do not have control over each algorithm's
-complexity, but future analyses could focus on balancing populations and number of worker nodes according
-to this.
+complexity, but future analyses could focus on balancing populations and number
+of worker nodes according to this.
 
@@
 
-To answer the first research question, we need to evaluate the effect of 
-the number of populations and the number of
-workers on the scalability of the system, we 
-propose  an experiment in which we selected $f_4$ (Skew Rastrigin-Bueche separable) from the
-BBOB testbed. This function has been used because it is computationally demanding,
-and in higher dimensions has been difficult for PSO \cite{el2009black} and GAs
-\cite{nicolau2009application} to solve with the required FEs. As we
-are only comparing in terms of scalability, the results from a single function
-can be better understood, since there are fewer factors involved.
-Following the procedure described in Section~\ref{setup}, we executed 
-two sets of experiments with five and ten initial populations,
-repeating each experiment 30 times using  1, 2, 4, and 8 workers.  We kept the same 
-maximum number of FEs, changing the relevant parameters for this. 
-See Tables \ref{tab:params:10} and \ref{tab:params:5} 
-for the complete list. 
+To answer the first research question, we need to evaluate the effect of the
+number of populations and the number of workers on the scalability of the
+system, we propose an experiment in which we selected $f_4$ (Skew
+Rastrigin-Bueche separable) from the BBOB testbed. This function has been used
+because it is computationally demanding, and in higher dimensions has been
+difficult for PSO \cite{el2009black} and GAs \cite{nicolau2009application} to
+solve with the required FEs. As we are only comparing in terms of scalability,
+the results from a single function can be better understood, since there are
+fewer factors involved. Following the procedure described in
+Section~\ref{setup}, we executed two sets of experiments with five and ten
+initial populations, repeating each experiment 30 times using 1, 2, 4, and 8
+workers. We kept the same maximum number of FEs, changing the relevant
+parameters for this. See Tables \ref{tab:params:10} and \ref{tab:params:5} for
+the complete list.
+
 %
@@
 
-However, when we compare the two
-options, we find some subtle differences: When using one or two
-workers, using 5 populations takes less time in most cases. 
-In Table~\ref{tab:time} results with p-value of the two-tailed Wilcoxon
-signed-ranks test was grater than the level of significance of 0.05 
-(i.e., accepting the null hypothesis of inequality) are marked in boldface.
-A reason for this could be that when having more populations, the migration
-of individuals takes more time from the beginning to reach to all of them.
-For instance, when a single worker has 10 populations, only
-after processing every one of them, it will start receiving mixed populations.  Although, in
-all cases, and even with a single worker, populations are heterogeneous, 
- individuals resulting from the mixing of populations take more time to propagate. 
-With four workers, almost the same
-results can be achieved regardless of our selection between five or ten
-populations. Also, there is not much difference between 4 and 8 workers, when
-using five populations, which might indicate that some workers are
-idling while populations are being processed by other workers. In
-general, we can observe that keeping a ratio between the number of workers and populations a bit over 1:1  exploits better the computing
-resources available. This also decreases the relation between the
-overhead needed to migrate and process incoming and outgoing
-population and the total computing time: more populations will need
-more time to do that.
+However, when we compare the two options, we find some subtle differences: When
+using one or two workers, using 5 populations takes less time in most cases. In
+Table~\ref{tab:time} results with p-value of the two-tailed Wilcoxon
+signed-ranks test was grater than the level of significance of 0.05 (i.e.,
+accepting the null hypothesis of inequality) are marked in boldface. A reason
+for this could be that when having more populations, the migration of
+individuals takes more time from the beginning to reach to all of them. For
+instance, when a single worker has 10 populations, only after processing every
+one of them, it will start receiving mixed populations. Although, in all cases,
+and even with a single worker, populations are heterogeneous, individuals
+resulting from the mixing of populations take more time to propagate. With four
+workers, almost the same results can be achieved regardless of our selection
+between five or ten populations. Also, there is not much difference between 4
+and 8 workers, when using five populations, which might indicate that some
+workers are idling while populations are being processed by other workers. In
+general, we can observe that keeping a ratio between the number of workers and
+populations a bit over 1:1 exploits better the computing resources available.
+This also decreases the relation between the overhead needed to migrate and
+process incoming and outgoing population and the total computing time: more
+populations will need more time to do that.
 
@@
 increment in all cases, but again, for smaller dimensions, five populations with
-fewer workers is a better option (see values underlined in the table) with the best speedup, achieved for 40 dimensions, shown in bold. The worst speed was going from four to eight workers 
-at higher dimensions (underlined in the table). 
-For ten populations, adding more workers resulted in a linear speedup, 
-but the best speedup is obtained when increasing from four to eight workers at 40 dimensions
-(shown in bold).
+fewer workers is a better option (see values underlined in the table) with the
+best speedup, achieved for 40 dimensions, shown in bold. The worst speed was
+going from four to eight workers at higher dimensions (underlined in the table).
+For ten populations, adding more workers resulted in a linear speedup, but the
+best speedup is obtained when increasing from four to eight workers at 40
+dimensions (shown in bold).
 
@@
 
-In order to run this experiment, we just needed to edit a configuration file and start it up. The plots presented in the results are generated by the BBOB framework
-from the JOSN log file for each run.
-There was no additional setup needed, proving how this framework is able to not only offer an reproducible way to run experiments, but also how easily it interfaces with other common tools in the EC ecosystem.
+In order to run this experiment, we just needed to edit a configuration file and
+start it up. The plots presented in the results are generated by the BBOB
+framework from the JOSN log file for each run. There was no additional setup
+needed, proving how this framework is able to not only offer an reproducible way
+to run experiments, but also how easily it interfaces with other common tools in
+the EC ecosystem.
 
@@
 
-In this paper we have presented the EvoSwarm framework that implements
-an event-based, cloud native architecture suitable for hosting any
-kind of multi-population optimization algorithm. The framework is
-based in industry-standard development tools such as Docker and Docker
-Compose, making easy to develop and deploy multi-algorithm distributed
-experiments.
-
-The set up of this kind of algorithms will have a big influence on the
-results, since diversity is important for this kind of algorithms and
-its creation and destruction play a big part of the algorithm's
-performance. One of the main levers to control it in EvoSwarm is via
-the number of workers that will be allocated from the beginning, which
-is why we have performed an experiment mainly intended as a proof of
-concept, but also devoted to show how two parameters in the framework,
-number of workers and number of populations used to initially seed the
-system, interact with each other. The conclusion of this experiment is
-that there is an interesting interplay, with the most benefit being
-obtained with a number of populations equal to the number of workers
-or slightly inferior or superior, which brings algorithmic benefit
-(experiment taking less time because it needs less evaluations) and
-also performance benefits (experiment taking less time because
-evaluations are being split between different workers.
-
-% this is not a conclusion. The conclusion is an answer to the
-% research question. This works, and works precisely for the reasons
-% we say it works. - JJ
-% Again, revise search questions and explain answers - JJ
-As a conclusion, EvoSwarm is a framework designed and tested according
-to industry standards which is able to bring many benefits to the
-bioinspired algorithm community, by showing the possibility of
-creating new, high performance, mixed algorithms, which can also be
-easily deployed to the cloud, or locally to systems that simply need
-to have Docker and {\tt docker-compose} installed.
-
-It also opens new possibilities. From the design point of view, it
+In this paper, we have presented the EvoSwarm framework that implements an
+event-based, cloud-native architecture suitable for hosting multi-population
+optimization algorithms. The framework is based in industry-standard development
+tools such as Docker and Docker Compose, making it easy to develop and deploy
+multi-algorithm distributed experiments. The model is based on the asynchronous
+exchange of messages between stateless functions that react to a continuous
+stream of data. We compared the message-driven architecture we presented against
+four other proposals for multi-population algorithms found in the literature. We
+found that it is best suited for asynchronous multi-population algorithms.
+
+We proposed a new methodology for the reproducible deployment and execution of
+multiple experiments by specifying the infrastructure as part of an experiment
+definition in both local or cloud environments. We presented the workflow of an
+experiment's entire life-cycle viewed from the perspective of developers and
+researchers. We tested the deployment on both local and Amazon's Elastic
+Container Service.
+
+We did an empirical evaluation of the same framework to learn if the proposed
+solution can efficiently improve the scalability and performance of
+population-based optimization algorithms.
+
+First, we focused on knowing how scaling is related to the number of populations
+and workers in the system. This experiment concludes that there is an
+interesting interplay, with the most benefit obtained with a number of
+populations near-equal to the number of workers, which brings performance gains.
+Also, we found that for lower dimensions, there is an advantage in using fewer
+populations.
+
+A second experiment centered on evaluating the algorithmic benefits (experiments
+taking less time because they need fewer evaluations) of distributed
+multi-algorithm algorithms. We compared the homogeneous and an ensemble of
+multi-populations, using Genetic Algorithms (GAs) and Particle Swarm
+Optimization (PSO) using the COCO benchmark framework for continuous functions.
+We found that the set-up of these algorithms has a significant influence on
+their results since diversity is essential for their performance. The ensemble
+achieved better results and, in most cases, needed fewer evaluations.
+
+% In this paper, we have presented the EvoSwarm framework that implements
+% an event-based, cloud-native architecture suitable for hosting 
+% multi-population optimization algorithms. The framework is
+% based in industry-standard development tools such as Docker and Docker
+% Compose, making it easy to develop and deploy multi-algorithm distributed
+% experiments.
+
+% The set up of this kind of algorithms will have a big influence on the
+% results, since diversity is important for this kind of algorithms and
+% its creation and destruction play a big part of the algorithm's
+% performance. One of the main levers to control it in EvoSwarm is via
+% the number of workers that will be allocated from the beginning, which
+% is why we have performed an experiment mainly intended as a proof of
+% concept, but also devoted to show how two parameters in the framework,
+% number of workers and number of populations used to initially seed the
+% system, interact with each other. The conclusion of this experiment is
+% that there is an interesting interplay, with the most benefit being
+% obtained with a number of populations equal to the number of workers
+% or slightly inferior or superior, which brings algorithmic benefit
+% (experiment taking less time because it needs less evaluations) and
+% also performance benefits (experiment taking less time because
+% evaluations are being split between different workers.
+
+% % this is not a conclusion. The conclusion is an answer to the
+% % research question. This works, and works precisely for the reasons
+% % we say it works. - JJ
+% % Again, revise search questions and explain answers - JJ
+% As a conclusion, EvoSwarm is a framework designed and tested according
+% to industry standards which is able to bring many benefits to the
+% bioinspired algorithm community, by showing the possibility of
+% creating new, high performance, mixed algorithms, which can also be
+% easily deployed to the cloud, or locally to systems that simply need
+% to have Docker and {\tt docker-compose} installed.
+
+This work also opens new possibilities. From the design point of view, it
 would be interesting to try and use automatic scaling, instead of

### 2020-05-18 draft.tex
Revised with several questions

@@
 In this paper, we have presented the EvoSwarm framework that implements an
-event-based, cloud-native architecture suitable for hosting multi-population
-optimization algorithms. The framework is based in industry-standard development
-tools such as Docker and Docker Compose, making it easy to develop and deploy
+event-based, cloud-native architecture suitable for hosting multi-population % multi-algorithm? - JJ
+optimization algorithms. % reproducible? It's in the title - JJ
+The framework is based in industry-standard development
+tools such as Docker and Docker Compose, making it easy for scientists and practitioners to develop and deploy
 multi-algorithm distributed experiments. The model is based on the asynchronous
@@
 stream of data. We compared the message-driven architecture we presented against
-four other proposals for multi-population algorithms found in the literature. We
-found that it is best suited for asynchronous multi-population algorithms.
+four other proposals for multi-population algorithms found in the literature. % But it's not a practical comparison, but a theoretical one. Show the differences instead of compare? - JJ
+We
+found that it is best suited for asynchronous multi-population algorithms. % This is not really proved. You chose, you didn't find - JJ
 
@@
 researchers. We tested the deployment on both local and Amazon's Elastic
-Container Service.
+Container Service. % and we proved... Remember the experimental hypothesis - JJ
 
@@
 interesting interplay, with the most benefit obtained with a number of
-populations near-equal to the number of workers, which brings performance gains.
+populations slightly above the number of workers, which brings performance gains. % of what kind? - JJ
 Also, we found that for lower dimensions, there is an advantage in using fewer
@@
 multi-populations, using Genetic Algorithms (GAs) and Particle Swarm
-Optimization (PSO) using the COCO benchmark framework for continuous functions.
+Optimization (PSO) using the COCO benchmark framework for continuous functions. % as a visualization device? - JJ
 We found that the set-up of these algorithms has a significant influence on

### 2020-05-18 draft.tex
Update Conclusions

--- 2020-05-18 draft.tex
@@
 
-As a proof-of-concept, we present a brief experiment that is deployed in three
-distinct environments, taking advantage that as a container-based application,
 EvoSwarm describes the deployment of loosely coupled processes in a consistent
-and portable way.
-
-The term cloud-native describes these container-based
+and portable way. The term cloud-native describes these container-based
 environments because we can deploy them on an infrastructure that abstracts the
-underlying compute, storage, and networking primitives. This infrastructure does not
-need to be a cloud provider; it could be a developer's laptop, a single server,
-or a production environment. When we deploy a cloud-native application to a
-cloud provider, we have some more options to choose:
+underlying compute, storage, and networking primitives. This infrastructure does
+not need to be a cloud provider; it could be a developer's laptop, a single
+server, or a production environment. In this section, we answer the following
+question:
+
+Can EvoSwarm as a cloud-native, container-based application be deployed locally
+or in the cloud by specifying the configuration of container images and
+resources in an orchestration service?
+
+To prove this question, we present a brief optimization experiment that is
+deployed in three distinct environments, taking advantage of it as a
+container-based application. When we deploy a cloud-native application to a cloud provider, we have some more
+options to choose:
 
@@
 
-In this paper, we have presented the EvoSwarm framework that implements an
-event-based, cloud-native architecture suitable for hosting multi-population
-optimization algorithms. The framework is based in industry-standard development
-tools such as Docker and Docker Compose, making it easy to develop and deploy
+In this paper, we have presented the EvoSwarm framework for 
+the reproducible execution of experiments with an event-based, cloud-native architecture 
+suitable for hosting multi-population, multi-algorithm optimization algorithms. % reproducible? It's in the title - JJ Yes - M
+The framework is based in industry-standard development
+tools such as Docker and Docker Compose, making it easy for scientists and practitioners to develop and deploy
 multi-algorithm distributed experiments. The model is based on the asynchronous
 exchange of messages between stateless functions that react to a continuous
-stream of data. We compared the message-driven architecture we presented against
-four other proposals for multi-population algorithms found in the literature. We
-found that it is best suited for asynchronous multi-population algorithms.
+stream of data. We show the differences of the message-driven architecture we presented against
+four other proposals for multi-population algorithms found in the literature, and highlight the 
+advantages for conducting asynchronous multi-population algorithms.
+% But it's not a practical comparison, but a theoretical one. Show the differences instead of compare? - JJ - Ok 
+% We found that it is best suited for asynchronous multi-population algorithms. % This is not really proved. You chose, you didn't find - JJ
+% Ok, I put it in the same package - M
 
@@
 experiment's entire life-cycle viewed from the perspective of developers and
-researchers. We tested the deployment on both local and Amazon's Elastic
-Container Service.
+researchers.  % and we proved... Remember the experimental hypothesis - JJ - Ok - 
+We tested the deployment of an experiment on Amazon's Elastic
+Container Service, and proved that as cloud-native, container-based application,
+EvoSwarm could be deployed locally or in the cloud by specifying the configuration 
+of container images and resources in an orchestration service.
 
@@
 interesting interplay, with the most benefit obtained with a number of
-populations near-equal to the number of workers, which brings performance gains.
+populations slightly above the number of workers, which brings performance gains 
+in terms of execution speed with experiments taking less time because
+evaluations are shared between different workers. % of what kind? - JJ
 Also, we found that for lower dimensions, there is an advantage in using fewer
@@
 multi-populations, using Genetic Algorithms (GAs) and Particle Swarm
-Optimization (PSO) using the COCO benchmark framework for continuous functions.
+Optimization (PSO) using the COCO benchmark framework 
+for visualization and definition of the optimization problems. % as a visualization device? - JJ  
+% we also used their functions -M
 We found that the set-up of these algorithms has a significant influence on

### 2020-05-18 draft.tex
Still one question to go

@@
 
-In this paper, we have presented the EvoSwarm framework for 
+In this paper we have presented the EvoSwarm framework for 
 the reproducible execution of experiments with an event-based, cloud-native architecture 
-suitable for hosting multi-population, multi-algorithm optimization algorithms. % reproducible? It's in the title - JJ Yes - M
+suitable for hosting multi-population, multi-algorithm optimization
+methods. % Too many algorithms, there's no such thing as a
+         % multi-algorithm algorithm - JJ
 The framework is based in industry-standard development
 tools such as Docker and Docker Compose, making it easy for scientists and practitioners to develop and deploy
-multi-algorithm distributed experiments. The model is based on the asynchronous
+multi-algorithm distributed experiments. The model implemented by the
+framework is based on the asynchronous % Please distinguish model,
+                                % framework, implementation, algorithm
+                                % - JJ
 exchange of messages between stateless functions that react to a continuous
-stream of data. We show the differences of the message-driven architecture we presented against
+stream of data; populations are the messages that flow in this stream. We show the differences of the message-driven architecture we presented against
 four other proposals for multi-population algorithms found in the literature, and highlight the 
 advantages for conducting asynchronous multi-population algorithms.
-% But it's not a practical comparison, but a theoretical one. Show the differences instead of compare? - JJ - Ok 
-% We found that it is best suited for asynchronous multi-population algorithms. % This is not really proved. You chose, you didn't find - JJ
-% Ok, I put it in the same package - M
 
@@
 multiple experiments by specifying the infrastructure as part of an experiment
-definition in both local or cloud environments. We presented the workflow of an
+definition in both local and cloud environments. We presented the workflow of an
 experiment's entire life-cycle viewed from the perspective of developers and
-researchers.  % and we proved... Remember the experimental hypothesis - JJ - Ok - 
+researchers. 
 We tested the deployment of an experiment on Amazon's Elastic
-Container Service, and proved that as cloud-native, container-based application,
+Container Service, and proved that as a cloud-native, container-based application,
 EvoSwarm could be deployed locally or in the cloud by specifying the configuration 
@@
 Optimization (PSO) using the COCO benchmark framework 
-for visualization and definition of the optimization problems. % as a visualization device? - JJ  
-% we also used their functions -M
+for visualization and definition of the optimization problems.
 We found that the set-up of these algorithms has a significant influence on

### 2020-05-18 draft.tex
More questions and fixes

@@
 populations slightly above the number of workers, which brings performance gains 
-in terms of execution speed with experiments taking less time because
+in terms of execution speed with experiments taking less time since
 evaluations are shared between different workers. % of what kind? - JJ
 Also, we found that for lower dimensions, there is an advantage in using fewer
-populations.
+populations. % Specify that.
 

### 2020-05-18 draft.tex
Eliminates question and solves

@@
 in terms of execution speed with experiments taking less time since
-evaluations are shared between different workers. % of what kind? - JJ
+evaluations are performed simultaneously by different workers.
 Also, we found that for lower dimensions, there is an advantage in using fewer

### 2020-05-18 draft.tex
A few corrections

@@
 
-\begin{lstlisting}[language=Python, caption=Fragment of output message, label=code:result ]
+\begin{lstlisting}[language=Python, caption=Fragment of an output message, label=code:result ]
 {
@@
   \caption{Boxplot of the number of evaluations per second, for 15 instances of function $f_4$,
-   with three execution environments Local, Single task and Multi-task.
+   with three execution environments Local, Single task, and Multi-task.
    }
@@
   \caption{ Median number of the evaluations per second, for 15 instances of function $f_4$,
-  with three execution environments Local, Single task and Multi task. The best result is shown in {\bf boldface}.}
+  with three execution environments Local, Single task, and Multi task. The best result is shown in {\bf boldface}.}
   \label{tab:aws}
@@
   \caption{Speedup achieved with incrementing the number of workers for
-    each dimension showing the p-rate for the Wilcoxon test.}
+    each dimension showing the p-value for the Wilcoxon test. Best and Worst are highlighted (bold and underlined) }
   % Say what highlights mean - JJ
@@
 the previous configuration, for example, when increasing from four to eight
-workers, and the p-rate for the left-tailed Wilcoxon test. There was an
+workers, and the p-value for the left-tailed Wilcoxon test. There was an
 increment in all cases, but again, for smaller dimensions, five populations with

### 2020-05-19 draft.tex
Spell check

@@
 by the {\texttt ecs compose command} that uses a {\texttt docker-compose.yml} file to generate 
-the configuration on-the-fly. For example to deploy the controller, setup and redis 
+the configuration on-the-fly. For example to deploy the controller, setup and Redis 
 containers we used the file in Listing ~\ref{code:compose}.
@@
 and another one, where we assigned each worker to a different task and the controller, 
-redis, and setup in another, for a total of nine concurrent Fargate tasks.
+Redis, and setup in another, for a total of nine concurrent Fargate tasks.
 
@@
 (COCO) framework \cite{hansen2016coco}. The testbed includes 24 real-parameter,
-single-objective ben chmark functions, and the capability to provide additional
+single-objective benchmark functions, and the capability to provide additional
 instances of each function. Each instance of a function has a different optimal
@@
 We have deployed the EvoSwarm application locally in the same hardware and
-software described in Section~\ref{setup:local-cloud}, additionally the BOBB
+software described in Section~\ref{setup:local-cloud}, additionally the BBOB
 experiments were performed with COCO \cite{hansen2016coco} version bbob.v15.03
@@
   \small
-  \caption{Speedup achieved with incrementing the number of workers for
+  \caption{Speedup achieved with increasing the number of workers for
     each dimension showing the p-value for the Wilcoxon test. Best and Worst are highlighted (bold and underlined) }
@@
 start it up. The plots presented in the results are generated by the BBOB
-framework from the JOSN log file for each run. There was no additional setup
+framework from the JSON log file for each run. There was no additional setup
 needed, proving how this framework is able to not only offer an reproducible way

### 2020-05-19 draft.tex
Final touches

@@
 suitable for hosting multi-population, multi-algorithm optimization
-methods. % Too many algorithms, there's no such thing as a
-         % multi-algorithm algorithm - JJ
+methods. 
 The framework is based in industry-standard development
@@
 multi-algorithm distributed experiments. The model implemented by the
-framework is based on the asynchronous % Please distinguish model,
-                                % framework, implementation, algorithm
-                                % - JJ
+framework is based on the asynchronous
 exchange of messages between stateless functions that react to a continuous
@@
 We found that the set-up of these algorithms has a significant influence on
-their results since diversity is essential for their performance. The ensemble
+their results since diversity is essential for their performance. The
+hybrid method using both algorithms 
 achieved better results and, in most cases, needed fewer evaluations.

### 2020-08-12 draft.tex
update figures
New Table

@@
 
-Section~\label{sec:exp1} is devoted to answering the first question, and Section~\label{sec:exp2} to the second.
+Section~\ref{sec:exp1} is devoted to answering the first question, and Section~\ref{sec:exp2} to the second.
 Next, we describe the general experimental setup to both experiments.
@@
             20              & 66      & 200                 &      5                 & 30               \\ \hline
-            40              & 80      & 250                 &      5                 & 40                \\ \hline
   \end{tabular}
@@
   \small
-  \caption{Parameters used in the experiments with ten populations
+  \caption{Parameters used in the experiments.
   }
@@
   \small
-  \begin{tabular}{|l|c|c|c|c|c|c|}
+  \begin{tabular}{|c|c|c|c|}
     \hline
-    Dimension        & 2  & 3  & 5  & 10 & 20  & 40  \\ \hline
-    Generations      & 40 & 25 & 28 & 50 & 66  & 80  \\ \hline
-    Population Size  & 50 & 60 & 60 & 70 & 100 & 125 \\ \hline
-    Populations      & \multicolumn{6}{c|}{10} \\ \hline
-    Iterations       & 10 & 20 & \multicolumn{3}{c|}{30}  & 40  \\ \hline  
-  \end{tabular}
-\end{table}
-%
-\begin{table}[h!tbp]
-  \small
-  \caption{Parameters used in the experiments with five populations
-  }
-  \label{tab:params:5}
-  \vspace{0.25cm}
-  \centering
-  \small
-  \begin{tabular}{|l|c|c|c|c|c|c|}
-    \hline
-    Dimension        & 2  & 3  & 5  & 10 & 20  & 40  \\ \hline
-    Generations      & 40 & 25 & 28 & 50 & 66  & 80  \\ \hline
-    Population Size  & 100 & 120 & 120 & 140 & 200 & 250 \\ \hline
-    Populations      & \multicolumn{6}{c|}{5} \\ \hline
-    Iterations       & 10 & 20 & \multicolumn{3}{c|}{30}  & 40  \\ \hline
+      Populations & Population Size & Generations & Maximum Iterations  \\ \hline
+                 \multicolumn{4}{|c|}{10 Dimension } \\ \hline
+         5        &  70             &        50   &   60                \\ \hline
+         10       &  70             &        50   &   60                \\ \hline
+         20       &  70             &        50   &   60                \\ \hline
+                  \multicolumn{4}{|c|}{20 Dimension } \\ \hline
+         5        &  70             &        50   &   60                \\ \hline
+         10       &  70             &        50   &   60                \\ \hline
+         20       &  70             &        50   &   60                \\ \hline
+
   \end{tabular}
@@
 workers. We kept the same maximum number of FEs, changing the relevant
-parameters for this. See Tables \ref{tab:params:10} and \ref{tab:params:5} for
-the complete list.
+parameters for this. See Table \ref{tab:params:10} for the complete list.
 
@@
 %
-\begin{table}[]
+\begin{table}[tbp]
   \caption{The median of the time required to finish an instance of the function $f_4$, 30 runs

### 2020-08-13 draft.tex
Rewrite of section 6
I describe the aws setup used in the following section. @jj  please have a look.

@@
 not need to be a cloud provider; it could be a developer's laptop, a single
-server, or a production environment. In this section, we answer the following
-question:
-
-Can EvoSwarm as a cloud-native, container-based application be deployed locally
-or in the cloud by specifying the configuration of container images and
-resources in an orchestration service?
-
-To prove this question, we present a brief optimization experiment that is
-deployed in three distinct environments, taking advantage of it as a
-container-based application. When we deploy a cloud-native application to a cloud provider, we have some more
+server, or a production environment. In this section, we give specific details 
+about how we deployed  EvoSwarm in a cloud-setting for running the experiments 
+within this paper.
+
+When we deploy a cloud-native application to a cloud provider, we have two
 options to choose:
@@
 \begin{description}
-
-  \item[A virtual machine] with a container-optimized operating system, for example,
-  Google Compute Engine VMs, or Amazon EC2.  We still need to provision the
-  virtual machine and run {\tt docker-compose} commands as we would do in a local 
-  workstation. Salza et al. \cite{salza2019speed} tested this option by using
-  CoreOs cluster.
 
@@
     services are Amazon Elastic Container Service,  Red Hat Openshift, and Marantis Docker Enterprise.
+    Salza et al. \cite{salza2019speed} tested this option by using CoreOs cluster.
 
@@
 
-We chose the third option, that is, a container orchestration service, and
+We chose the second option, that is, a container orchestration service, and
 deployed the EvoSwarm application in Amazon Elastic Container Service (ECS),
@@
 % Ping - JJ
-We can define an ECS cluster directly on AWS Elastic Cloud VMs or through the AWS Fargate service.
-Fargate is a serverless compute engine for containers that provides automatic provision 
-and management of AWS ECS resources. With this service, calculating the cost of execution
-is simplified because users only specify the computing units and memory for all the application,
-instead of every component.
+
+We can define an ECS cluster directly on the AWS Elastic Cloud VMs or through the
+AWS Fargate service. Fargate is a serverless compute engine for containers that
+provides automatic provision and management of AWS ECS resources. With this
+service, calculating the cost of execution is simplified because users only
+specify the computing units and memory for the application, instead of every
+component. However, AWS Fargate is not suited for an experimental setting because
+many of the details are out of user control. For this reason, we chose the
+alternative of using a cluster of EC2 containers. The drawback is that we need to manually
+provision the  EC2 containers, networking, and group permissions for each
+experimental setup, for instance, when going from 2 to 8 workers, we need to increase the
+number of EC2 instances to meet the required amount of CPU units.
 
 Multicontainer applications are deployed to ECS as tasks, and a single cluster can run multiple tasks in
-parallel. When using the Fargate service as a backend, tasks are limited to a
-maximum of 10 containers.
-
-Using the ECS CLI, we can provision a cluster with a configuration file
-specifying the computing resources needed. We can specify the configuration of a task 
+parallel. Using the ECS CLI, we can provision a cluster with a configuration file
+specifying the computing resources needed. We can also specify the configuration of a task 
 by using a task definition file (see Listing~\ref{code:ecs}) in YAML format.
-The main attributes are the {\texttt task\_size} indicating the memory and CPU units
-assigned, and network configuration. The complete configuration is generated
-by the {\texttt ecs compose command} that uses a {\texttt docker-compose.yml} file to generate 
+The main attributes are the {\texttt cpu\_shares} indicating the CPU units
+assigned, in this case 2048; and the network configuration, we have two private 
+subnets and a security group in this configuration. When deploying to EC2 containers, these requirements
+need to be supported by the currently available cluster. For instance, we need to have all 
+the CPU units required by all tasks, if they are not available, the task initiation will fail.
+
+The complete task configuration is generated by the {\texttt ecs compose command} that uses a {\texttt docker-compose.yml} file to generate 
 the configuration on-the-fly. For example to deploy the controller, setup and Redis 
 containers we used the file in Listing ~\ref{code:compose}.
-Each container is defined as a  {\texttt service} , with the following attributes: {\texttt image},
+
+Each container is defined as a {\texttt service} (task in AWS terms) , with the following attributes: {\texttt image},
 that reference the container images we have published in Docker hub
@@
 
-We deployed two ECS configurations, one that included all containers in a single task,
-and another one, where we assigned each worker to a different task and the controller, 
-Redis, and setup in another, for a total of nine concurrent Fargate tasks.
-
-For the first option, we deployed an AWS Fargate cluster and defined a task configuration file with
-8GBs of memory and 4096 computing units, equivalent to 4 CPUs. Because of the
-ten container limit, we only used seven workers and tested function $f_4$ as
-in the previous experiment, and to save resources we only tested for dimensions 10 and 20 for 15 instances.
-
-For the second option, we used the same cluster and additionally
-assigned each worker to a task with 4GBs of memory and 2048 computing units. Then scale the 
-number of worker tasks to eight. We can scale the number of tasks for up to 100 instances in 
-a Fargate configuration.
-
-For local experiments, we have deployed the EvoSwarm application in a
-Desktop PC with an AMD Ryzen 9 3900x 12-core processor with 24 threads with a 3.8 GHz base clock, 
-and 16 GB RAM. We used Docker version 19.03.3, build a872fc2f86, and {\tt docker-compose} version 1.21.0, 
-in Ubuntu Linux 18.04; code is run using Python 3.7.5. Container images and the {\tt docker-compose} file are available at
-\url{https://hub.docker.com/u/mariosky}, and \url{https://github.com/mariosky/EvoSwarm} respectively.
-
-We reproduced the experiment of Listing~\ref{code:exp} in each environment. The experiment consists of
-solving $f_4$ for 10 and 20 dimensions. We give more details about additional parameters of the configuration in Section~\ref{setup}.
+Figure \ref{fig:aws-configuration} shows the basic EC2 configuration used for the deployment
+of experiments. We selected three types of EC2 instances according to our needs, the  
+{\texttt setup}, {\texttt controller}, and {\texttt redis} containers do not demand many resources
+and we do not need to scale them as the amount of parallel workers are increased. They run in a
+single {\texttt C5  High-CPU large} instance, with the available 2048 CPU shares divided as follows:
+{\texttt setup} (128),  {\texttt controller} (896),{\texttt redis} (1024). We assigned 2048 units to 
+each {\texttt worker}   container, by using a more capable {\texttt C5 High-CPU XLarge} instance,  with 
+four vCPUs and 4096 cpu shares. We scaled the number if instances as needed, using one instance
+for every two {\texttt worker} containers. Because all these instances run in a private subnet, we
+run a Bastion instance in a public subnet, from where we executed the experiments and collected the results,
+through a {\texttt ssh} session. This is a {\texttt R5AD Large} instance,  with 75GiB of storage for
+storing the logs, and more memory (16GiB) for processing the logs. The details and cost of these instances at the 
+time of writing are shown in Table \ref{tab:ec2}.
+
+\begin{table}[h!tbp]
+  \small
+  \caption{Details of EC2 instance types used in our deployment}
+  \label{tab:ec2} 
+  \centering
+  \small
+  \begin{tabular}{|l|l|r|l|l|}
+    \hline
+    Name & Memory & vCPUs & Storage  & Cost \\ \hline
+    R5AD Large	&	16.0 GiB	& 2 vCPUs	& 75 GiB 	& \$0.131000 hourly \\ \hline
+    C5 High-CPU Large	&	4.0 GiB	& 2 vCPUs	& EBS Only 	& \$0.102000 hourly \\ \hline
+    C5 High-CPU XLarge	&	8.0 GiB	& 4 vCPUs	& EBS Only 	& \$0.154000 hourly \\ \hline
+  \end{tabular}
+\end{table}
 
@@
   task_definition:
-  task_execution_role: ecsTaskExecutionRole
     ecs_network_mode: awsvpc
-  task_size:
-    mem_limit: 8GB
-    cpu_limit: 4096
+    services:
+      worker:
+        cpu_shares: 2048
+
+  run_params:
+    network_configuration:
+      awsvpc_configuration:
+        subnets:
+          - "subnet-0ca2a542bbd74239e"
+          - "subnet-0dfd367b5ca531b07"
+        security_groups:
+          - "sg-0681c18469874b391"  
 \end{lstlisting}
@@
 
-\subsubsection{Experimental Results}
-
-In this experiment, we wanted to test several deployment options for the
-cloud-native application, and have an idea of the computing capacity of each
-option measured as the number of function evaluations per second. The results
-are shown in Figure~\ref{fig:aws-results} and Table~\ref{tab:aws}. As we can
-see, the local environment (see section \ref{setup}) achieved a higher ratio of
-evaluations per second (in bold). Still, we must consider that there is minimum
-network overhead in a local workstation and a clock speed of 3.8 GHz. The first
-option (labeled Single) is straightforward to deploy, needing only a single {\tt
-docker-compose} file, and one command to execute. But with the limit of 4 CPUs
-and ten containers, it is the slowest option. The second option (labeled Multi),
-is faster, and not too far from the local execution, especially when solving the
-$f_4$ in 20 dimensions (underlined in table). With the capacity for increasing
-the number of worker containers, it is reasonable to think that we can achieve a
-higher number of evaluations with an appropriate configuration. Nevertheless,
-the goal of this experiment was not to increase the number of evaluations but to
-test the flexibility of a cloud-native application. The total cost of testing,
-configuration, and execution of the experiments in ECS was less than \$1.00 US
-Dollar.
-
 \begin{figure}[h!tbp]
   \centering
-  \includegraphics[width=\textwidth]{aws}
-  \caption{Boxplot of the number of evaluations per second, for 15 instances of function $f_4$,
-   with three execution environments Local, Single task, and Multi-task.
+  \includegraphics[width=\textwidth]{aws_ec2}
+  \caption{AWS EC2 configuration for the deployment of EvoSwarm containers.
    }
-  \label{fig:aws-results}
+  \label{fig:aws-configuration}
 \end{figure}
 %
-\begin{table}[h!tbp]
-  \small
-  \caption{ Median number of the evaluations per second, for 15 instances of function $f_4$,
-  with three execution environments Local, Single task, and Multi task. The best result is shown in {\bf boldface}.}
-  \label{tab:aws}
-  \vspace{0.25cm}
-  \centering
-
-  \begin{tabular}{lll}
-  \hline
-  Dim & Env         & Median of ratio \\
-  \hline
-  10  & Local       & \textbf{47,166.29}       \\
-      & Multi task  & 40,466.45    \\
-      & Single task & 18,193.05      \\
-  \hline
-  20  & Local       & \textbf{36,101.86}   \\
-      & Multi task  & {\ul 33,916.02}       \\
-      & Single task & 16,546.95  \\
-  \hline
-  \end{tabular}
-\end{table}
-
-The same workflow and configuration files used by this brief experiment, will be
-used on the next section, to carry a more extensive evaluation of the framework.
-Again, each configuration describe the deployment of containers in a consistent
+
+This configuration files will be
+used on the next section, to carry an extensive evaluation of the framework.
+Again, each configuration describes the deployment of containers in a consistent
 and portable way, an it was used in each phase of the life cycle, from
@@
 
-In the prior section, we wanted to prove that EvoSwarm can be used by 
-the public to create reproductive research, but in this section are
+In the prior section, we describe an option for an EvoSwarm deployment that can be used by 
+the public to create reproductive research, in this section we are
 using the same framework to learn if the proposed solution can efficiently

### 2020-08-13 draft.tex
Section header removed
Update general setup

@@
 
-\subsubsection{Setup} % Even if it's a sub-sub section, flesh this out ... - JJ
-\label{setup:local-cloud}
+%\subsubsection{Setup} % Even if it's a sub-sub section, flesh this out ... - JJ
+%\label{setup:local-cloud}
 
@@
 
-We have deployed the EvoSwarm application locally in the same hardware and
-software described in Section~\ref{setup:local-cloud}, additionally the BBOB
-experiments were performed with COCO \cite{hansen2016coco} version bbob.v15.03
+We have deployed the EvoSwarm application in Amazon Elastic Container Service as described in Section~\ref{cloud-aws}, 
+additionally the BBOB experiments were performed with COCO \cite{hansen2016coco} version bbob.v15.03
 in Python, the plots were produced with version 2.3.2.

### 2020-08-13 draft.tex
Update draft.tex

@@
 and we do not need to scale them as the amount of parallel workers are increased. They run in a
-single {\texttt C5  High-CPU large} instance, with the available 2048 CPU shares divided as follows:
-{\texttt setup} (128),  {\texttt controller} (896),{\texttt redis} (1024). We assigned 2048 units to 
-each {\texttt worker}   container, by using a more capable {\texttt C5 High-CPU XLarge} instance,  with 
-four vCPUs and 4096 cpu shares. We scaled the number if instances as needed, using one instance
-for every two {\texttt worker} containers. Because all these instances run in a private subnet, we
-run a Bastion instance in a public subnet, from where we executed the experiments and collected the results,
-through a {\texttt ssh} session. This is a {\texttt R5AD Large} instance,  with 75GiB of storage for
-storing the logs, and more memory (16GiB) for processing the logs. The details and cost of these instances at the 
-time of writing are shown in Table \ref{tab:ec2}.
+single {\texttt C5  High-CPU large} instance. With the available 2048 CPU shares divided as follows:
+{\texttt setup} (128),  {\texttt controller} (896),{\texttt redis} (1024). With this assignment 
+the {\texttt redis} container gets a dedicated vCPU, needed because it is the most demanding 
+task of the three. We assigned 2048 units to each {\texttt worker} container, and using a more 
+capable {\texttt C5 High-CPU XLarge} instance,  with 
+four vCPUs and 4096 cpu shares, with this configuration each worker gets a pair of vCPUs.  
+We scaled the number if instances as needed, using one instance for every two {\texttt worker} containers.
+
+Because all these instances run in a private subnet, we also
+run a Bastion instance in a public subnet, from where we send the commands to execute the experiments 
+and later collected the results. We conected to this instance through a {\texttt ssh} session from 
+our personal workstation. This is a {\texttt R5AD Large} instance, selected with 75GiB of storage, needed for
+keeping the logs, and more memory (16GiB) for processing the logs in memory. The details these EC2 instabces 
+and their cost at the time of writing are shown in Table \ref{tab:ec2}.
 
@@
 
-This configuration files will be
-used on the next section, to carry an extensive evaluation of the framework.
+The complete configuration files are available at public GitHub repository, and
+they are used on the next section to carry an extensive evaluation of the framework.
 Again, each configuration describes the deployment of containers in a consistent

### 2020-08-13 draft.tex
url added

@@
 
-The complete configuration files are available at public GitHub repository, and
+The complete configuration files are available at public GitHub repository 
+\url{https://github.com/mariosky/EvoSwarm/tree/master/aws-EC2}, and
 they are used on the next section to carry an extensive evaluation of the framework.

### 2020-08-13 draft.tex
more details of EC2

@@
 
-We have deployed the EvoSwarm application in Amazon Elastic Container Service as described in Section~\ref{cloud-aws}, 
-additionally the BBOB experiments were performed with COCO \cite{hansen2016coco} version bbob.v15.03
+We have deployed the EvoSwarm application in Amazon Elastic Container Service as 
+described in Section~\ref{cloud-aws}, each EC2 instance was an Amazon Linux 2 AMI with 
+ECS Agent 1.42.0 and Docker version 19.03.6-ce.
+Additionally the BBOB experiments were performed with COCO \cite{hansen2016coco} version bbob.v15.03
 in Python, the plots were produced with version 2.3.2.

### 2020-08-13 draft.tex
New experiment details updated

@@
 The complete configuration files are available at public GitHub repository 
-\url{https://github.com/mariosky/EvoSwarm/tree/master/aws-EC2}, and
+\footnote{\url{https://github.com/mariosky/EvoSwarm/tree/master/aws-EC2}}, and
 they are used on the next section to carry an extensive evaluation of the framework.
@@
 system, we propose an experiment in which we selected $f_4$ (Skew
-Rastrigin-Bueche separable) from the BBOB testbed. This function has been used
+Rastrigin-Bueche separable) from the BBOB testbed, over 10 and 20 dimensions. This function has been used
 because it is computationally demanding, and in higher dimensions has been
@@
 fewer factors involved. Following the procedure described in
-Section~\ref{setup}, we executed two sets of experiments with five and ten
-initial populations, repeating each experiment 30 times using 1, 2, 4, and 8
+Section~\ref{setup}, we executed two sets of experiments with five, ten, and twenty
+initial populations, repeating each experiment 30 times using 1, 2, 4, 8, and 16
 workers. We kept the same maximum number of FEs, changing the relevant

### 2020-08-14 draft.tex
Updated table
new table and figures

@@
          5        &  70             &        50   &   60                \\ \hline
-         10       &  70             &        50   &   60                \\ \hline
-         20       &  70             &        50   &   60                \\ \hline
+         10       &  30             &        15   &   200                \\ \hline
+         20       &  15             &        10   &   400                \\ \hline
                   \multicolumn{4}{|c|}{20 Dimension } \\ \hline
-         5        &  70             &        50   &   60                \\ \hline
-         10       &  70             &        50   &   60                \\ \hline
-         20       &  70             &        50   &   60                \\ \hline
+         5        &  100            &        66   &   60                \\ \hline
+         10       &  60             &        15   &   200                \\ \hline
+         20       &  30             &        8   &    400                \\ \hline
 
@@
   \centering
-  \includegraphics[width=\textwidth]{time_diff}
+  \includegraphics[width=\textwidth]{secs_per_instance}
   \caption{Boxplot of the median time required for the execution of 30 
-  instances of function $f_4$, in seconds, different number of workers and populations sizes.  }
+  instances of function $f_4$, in seconds, different number of workers and populations sizes. The y-axis is shown in log scale.  }
   \label{fig:spworker:time}
@@
     \centering
-    \begin{tabular}{ccrr}
+    \begin{tabular}{ccrrr}
     \hline
-              &         & 5 Populations    & 10 Populations \\
-     Dimension& Workers & Median of rate & Median of rate  \\
+              &         & 5 Populations    & 10 Populations & 20 Populations \\
+     Dimension& Workers & Median of rate & Median of rate & Median of rate  \\
     \hline
-          2   & 1w      & 2.96   & 3.15   \\
-              & 2w      & 1.84   & 1.72   \\
-              & 4w      & 1.22   & \textbf{0.96}   \\
-              & 8w      & 1.05   & \textbf{0.59}   \\
+          10  & 1w      & 22.24  & 20.48           & 24.55  \\
+              & 2w      & 10.31  & 10.00           & 12.79 \\
+              & 4w      &  6.23  & 4.75            &  5.70 \\
+              & 8w      &  5.86  &  2.33           & 2.74 \\
+              & 16w     &  5.96  &  2.33           & \textbf{1.59}\\
+
     \hline
-          3   & 1w      & 6.53   & \textbf{7.76}   \\
-              & 2w      & 3.64   & 3.83   \\
-              & 4w      & 2.33   & 2.26   \\
-              & 8w      & 2.17   &  \textbf{1.36}   \\
-    \hline
-          5   & 1w      & \textbf{12.33}  & 14.19  \\
-              & 2w      & 6.63   & \textbf{7.41}   \\
-              & 4w      & 4.25   & 4.34   \\
-              & 8w      & 3.67   &  \textbf{2.41}   \\
-    \hline
-          10  & 1w      & \textbf{31.50}  & 38.30  \\
-              & 2w      & \textbf{16.07}  & 19.31  \\
-              & 4w      &  9.99   & 10.35  \\
-              & 8w      & 8.97   &  \textbf{5.92}  \\
-    \hline
-          20  & 1w      &  116.35 & 124.51 \\
-              & 2w      & \textbf{51.18}  & 63.68  \\
-              & 4w      & 32.26  & 33.07  \\
-              & 8w      & 28.64  & \textbf{18.25}  \\
-    \hline
-          40  & 1w      & \textbf{369.89} & 448.89 \\
-              & 2w      & \textbf{198.43} & 234.50 \\
-              & 4w      & 122.39 &  120.62 \\
-              & 8w      & 122.31 & \textbf{63.35} \\
+          20  & 1w   & 18.36   & 65.18  & 80.31 \\
+              & 2w   & 34.18   & 33.46  & 38.87 \\
+              & 4w   & 19.55   & 16.58  & 18.69 \\
+              & 8w   & 18.38   & 7.77   & 8.66 \\
+              & 16w  & 18.36   & 7.27   & \textbf{4.57} \\
     \hline
@@
 %
+
 \begin{figure}[h!tbp]
   \centering
-  \includegraphics[width=\textwidth]{evalspersecond}
+  \includegraphics[width=\textwidth]{evals_per_instance}
+  \caption{Boxplot of the number of function evaluations required to reach the target, for 30 instances of function $f_4$, with different number of workers and populations sizes}
+  \label{fig:spworker:evalspersecond}
+\end{figure}
+
+%
+
+\begin{figure}[h!tbp]
+  \centering
+  \includegraphics[width=\textwidth]{eval_ratio}
   \caption{Boxplot of the number of evaluations per second, for 30 instances of function $f_4$, with different number of workers and populations sizes }
@@
 \end{figure}
-%
+
 \subsubsection{Results Regarding Speedup}
@@
 
-  \begin{tabular}{c|r|r|r|r|r|r|}
-  \cline{2-7}
-  \multicolumn{1}{l|}{}           & \multicolumn{3}{c|}{5 Populations}                                          & \multicolumn{3}{c|}{10 Populations}                                         \\ \hline
-  \multicolumn{1}{|l|}{Dimension} & \multicolumn{1}{c|}{2W} & \multicolumn{1}{c|}{4W} & \multicolumn{1}{c|}{8W} & \multicolumn{1}{c|}{2W} & \multicolumn{1}{c|}{4W} & \multicolumn{1}{c|}{8W} \\ \hline
-  \multicolumn{1}{|c|}{2}         & \textbf{1.94}           & 3.12                    & 3.86                    & 1.84                    & \textbf{3.38}                    & \textbf{5.98}                    \\ \hline
-  \multicolumn{1}{|c|}{3}         & \textbf{1.93}           & 2.77                    & 3.15                    & 1.90                    & \textbf{3.26}                   & \textbf{5.96}                    \\ \hline
-  \multicolumn{1}{|c|}{5}         & 1.91                    & 2.90                    & 3.09                    & 1.91                    & \textbf{3.46}                    & \textbf{6.14}                    \\ \hline
-  \multicolumn{1}{|c|}{10}        & \textbf{1.92}           & 3.03                    & 3.30                    & 1.90                    & \textbf{3.61}                   & \textbf{6.41}                    \\ \hline
-  \multicolumn{1}{|c|}{20}        & \textbf{1.91}           & 3.14                    & 3.27                    & 1.89                    & \textbf{3.73}                   & \textbf{6.71}                    \\ \hline
-  \multicolumn{1}{|c|}{40}        & \textbf{2.00}           & 2.92                    & 3.12                    & 1.95                    & \textbf{3.80}                     & \textbf{7.52}                    \\ \hline
+  \begin{tabular}{c|r|r|r|r|}
+  \hline
+  \multicolumn{1}{|l|}{Dimension} & \multicolumn{1}{c|}{2W} & \multicolumn{1}{c|}{4W} & \multicolumn{1}{c|}{8W} & \multicolumn{1}{c|}{16W}  \\ \hline
+  \multicolumn{5}{| c|}{5 Populations} \\ \hline
+  \multicolumn{1}{|c|}{10}        & 2.15           & 3.61                    & 3.75                    & 3.96     \\ \hline
+  \multicolumn{1}{|c|}{20}        & 2.09           & 3.53                    & 3.83                    & 3.95     \\ \hline
+  \multicolumn{5}{| c|}{10 Populations} \\ \hline
+  \multicolumn{1}{|c|}{10}        & 1.91           & 3.82                    & 7.67                    & 8.25     \\ \hline
+  \multicolumn{1}{|c|}{20}        & 1.97           & 3.97                    & 7.90                    & 8.90     \\ \hline
+  \multicolumn{5}{| c|}{20 Populations} \\ \hline
+  \multicolumn{1}{|c|}{10}        & 1.96           & 4.05                    & 8.29                    & 14.50     \\ \hline
+  \multicolumn{1}{|c|}{20}        & 2.00           & 4.09                    & 8.18                    & 15.16     \\ \hline
   \end{tabular}
@@
   \centering
-  \begin{tabular}{lllllll}
-            &               &               & \multicolumn{2}{l}{5 Populations} & \multicolumn{2}{l}{10 Populations} \\
-  Dimension & \multicolumn{2}{l}{Increment} & Speedup         & p-value         & Speedup          & p-value         \\
-  2         & 1w            & 2w            & {\ul 1.94}            & 6E-16           & 1.84             & 2E-16           \\
-            & 2w            & 4w            & 1.61            & 2E-15           & 1.83             & 1E-15           \\
-            & 4w            & 8w            & 1.24            & 1E-06           & 1.77             & 2E-16           \\
-  3         & 1w            & 2w            & {\ul 1.93}            & 2E-16           & 1.90             & 2E-16           \\
-            & 2w            & 4w            & 1.44            & 8E-15           & 1.71             & 2E-16           \\
-            & 4w            & 8w            & 1.14            & 3E-05           & 1.83             & 2E-16           \\
-  5         & 1w            & 2w            & 1.91            & 9E-16           & 1.91             & 2E-16           \\
-            & 2w            & 4w            & 1.52            & 3E-11           & 1.82             & 2E-16           \\
-            & 4w            & 8w            & 1.06            & 5E-03           & 1.77             & 2E-16           \\
-  10        & 1w            & 2w            & 1.92            & 3E-16           & 1.90             & 2E-16           \\
-            & 2w            & 4w            & 1.58            & 3E-16           & 1.89             & 2E-16           \\
-            & 4w            & 8w            & {\ul 1.09}      & 1E-03           & 1.78             & 2E-16           \\
-  20        & 1w            & 2w            & 1.91            & 2E-16           & 1.89             & 2E-16           \\
-            & 2w            & 4w            & 1.64            & 2E-15           & 1.97             & 2E-16           \\
-            & 4w            & 8w            & {\ul 1.04 }     & 1E-02           & 1.80             & 2E-16           \\
-  40        & 1w            & 2w            & \textbf{2.00}   & 2E-16           & 1.95             & 2E-16           \\
-            & 2w            & 4w            & 1.46            & 1E-09           & 1.95             & 2E-16           \\
-            & 4w            & 8w            & {\ul 1.07}            & 3E-02           & \textbf{1.98}    & 2E-16          
+  \begin{tabular}{llllll}
+  Populations & Dimension  & \multicolumn{2}{l}{Increment}  & Speedup            & p-value\\
+  5  & 10 & 1w & 2w  & 2.15 & 1.51E-11 \\
+  & 10 & 2w & 4w  & 1.68 & 1.51E-11 \\
+  & 10 & 4w & 8w  & 1.04 & 0.004942 \\
+  & 10 & 8w & 16w & 1.05 & 0.072660 \\
+  & 20 & 1w & 2w  & 2.09 & 1.51E-11 \\
+  & 20 & 2w & 4w  & 1.69 & 1.51E-11 \\
+  & 20 & 4w & 8w  & 1.08 & 5.97E-07 \\
+  & 20 & 8w & 16w & 1.03 & 0.214482 \\
+  &    &    &     &      &          \\
+10 & 10 & 1w & 2w  & 1.91 & 1.51E-11 \\
+  & 10 & 2w & 4w  & 2.00 & 1.51E-11 \\
+  & 10 & 4w & 8w  & 2.01 & 1.51E-11 \\
+  & 10 & 8w & 16w & 1.08 & 1.39E-05 \\
+  & 20 & 1w & 2w  & 1.97 & 1.51E-11 \\
+  & 20 & 2w & 4w  & 2.01 & 1.51E-11 \\
+  & 20 & 4w & 8w  & 1.99 & 1.51E-11 \\
+  & 20 & 8w & 16w & 1.13 & 4.24E-09 \\
+  &    &    &     &      &          \\
+ 20 & 10 & 1w & 2w  & 1.96 & 1.51E-11 \\
+  & 10 & 2w & 4w  & 2.07 & 1.51E-11 \\
+  & 10 & 4w & 8w  & 2.05 & 1.51E-11 \\
+  & 10 & 8w & 16w & 1.75 & 1.51E-11 \\
+  & 20 & 1w & 2w  & 2.00 & 1.51E-11 \\
+  & 20 & 2w & 4w  & 2.04 & 1.51E-11 \\
+  & 20 & 4w & 8w  & 2.00 & 1.51E-11 \\
+  & 20 & 8w & 16w & 1.85 & 1.51E-11 \\
  \end{tabular}

### 2020-08-15 draft.tex
Update results

@@
 In the prior section, we describe an option for an EvoSwarm deployment that can be used by 
-the public to create reproductive research, in this section we are
+the public to create reproductive research in the cloud, in this section we are
 using the same framework to learn if the proposed solution can efficiently
@@
 Next, we show the parameters for each algorithm, with Table \ref{tab:GAparams} for
-the GA and Table \ref{tab:PSOparams} for PSO. We obtained these parameters
+the GA and Table \ref{tab:PSOparams} for the PSO. We obtained these parameters
 following the same method as in a previous work \cite{garcia2017benchmarking}.
@@
 in these parameters, the range of values is specified. We did not change these parameters during the 
-experiments, and only the population size and number of generations were
+experiments, and only the number of populations, population size and number of generations were
 provided as parameters. This strategy has been applied successfully for increasing 
@@
 experiments as seen in Section~\ref{sec:evoswarm:config}. 
-As a first step, we must deploy the docker application using a
-{\tt docker-compose.yml} file, as described in Section~\ref{sec:evoswarm:config}. 
+As a first step, we can deploy the docker application using a
+{\tt docker-compose.yml} file, as described in Section~\ref{sec:evoswarm:config} for a local
+test, or by following the steps for a cloud setting as described in Section~\ref{cloud-aws}. 
 Then a JSON file containing the configuration parameters of the
@@
 be tested, and we must select additional parameters for each dimension.
-According to the maximum number of function evaluations (FEs) we mentioned
+According to the maximum number of FEs we mentioned
 earlier, we define for each dimension: the number of populations that will be
@@
 described in Section~\ref{cloud-aws}, each EC2 instance was an Amazon Linux 2 AMI with 
-ECS Agent 1.42.0 and Docker version 19.03.6-ce.
-Additionally the BBOB experiments were performed with COCO \cite{hansen2016coco} version bbob.v15.03
+ECS Agent 1.42.0 and Docker version 19.03.6-ce. Additionally the BBOB experiments 
+were performed with COCO \cite{hansen2016coco} version bbob.v15.03
 in Python, the plots were produced with version 2.3.2.
@@
     \item The size of populations and the number of generations:
-    These parameters naturally increase the execution time.
+    These parameters naturally increase the execution time, because they mean more FEs.
     \item The complexity of the algorithm: For instance, in our case,
@@
 In this experiment, we evaluate how the number of populations and the number of
-workers in the system are related to the scalability of the system. In the next
-section, we describe the experimental setup, and a discussion of the results
-follows, with respect to execution time (Section~\ref{sec:exec-time}) 
-and speedup Section~\ref{sec:speedup}.
+workers in the system are related to the scalability of the system. 
+In the next section, we describe the experimental setup, and a discussion of the results
+follows, with respect to the speedup in terms of evaluations per second (Section~\ref{sec:speedup}), and 
+execution time (Section~\ref{sec:exec-time}).
 
@@
 parameters for this. See Table \ref{tab:params:10} for the complete list.
+
+\subsubsection{Results Regarding Speedup}
+\label{sec:speedup}
+
+\begin{figure}[h!tbp]
+  \centering
+  \includegraphics[width=\textwidth]{eval_ratio}
+  \caption{Boxplot of the number of evaluations per second, for 30 instances of function $f_4$,
+          with different number of workers and number of populations. The 7 axis is in log scale.}
+  \label{fig:spworker:evalspersecond}
+\end{figure}
+
+% Introduction saying why we are measuring this, how it's related to the objectives of the paper - JJ
+% This is done on the parent section, I renamed the sections to clarify this - M
+To measure the speedup obtained by the addition of workers, we will use the
+number of evaluations per second, instead of the time required to finish an
+instance of a problem. The reason for this is that in some cases reaching the
+desired target can require less work, because we are using a stochastic search. 
+%% JJ, can you please check if this is clear or if it needs more explanation?
+We are going to measure the total increase of work done in a second by adding more workers.
+The baseline for the comparison will be the algorithm with a single worker, 
+and the minimum amount of work available will be five populations. 
+This number is selected because the controller needs at least
+three populations to migrate individuals between them, and we add two more to
+keep even a single worker busy all the time. Figure \ref{fig:spworker:evalspersecond} 
+shows the number of function evaluations per second, achieved by a certain number of
+workers, for ten and twenty dimensions of $f_4$.
+
+\begin{table}[h!tbp]
+  \caption{Speedup by worker and dimension, taking one worker as the baseline.}
+  \label{tab:speedup-table}
+  \vspace{0.25cm}
+  \centering
+
+  \begin{tabular}{c|r|r|r|r|}
+  \hline
+  \multicolumn{1}{|l|}{Dimension} & \multicolumn{1}{c|}{2W} & \multicolumn{1}{c|}{4W} & \multicolumn{1}{c|}{8W} & \multicolumn{1}{c|}{16W}  \\ \hline
+  \multicolumn{5}{| c|}{5 Populations} \\ \hline
+  \multicolumn{1}{|c|}{10}        & \textbf{2.15} & 3.61 & 3.75 & 3.96     \\ \hline
+  \multicolumn{1}{|c|}{20}        & \textbf{2.09} & 3.53 & 3.83 & 3.95     \\ \hline
+  \multicolumn{5}{| c|}{10 Populations} \\ \hline
+  \multicolumn{1}{|c|}{10}        & 1.91           & 3.82 & 7.67 & 8.25     \\ \hline
+  \multicolumn{1}{|c|}{20}        & 1.97           & 3.97 & 7.90 & 8.90     \\ \hline
+  \multicolumn{5}{| c|}{20 Populations} \\ \hline
+  \multicolumn{1}{|c|}{10}        & 1.96           & \textbf{4.05} & \textbf{8.29} & \textbf{14.50}     \\ \hline
+  \multicolumn{1}{|c|}{20}        & 2.00           & \textbf{4.09} & \textbf{8.18} & \textbf{15.16}     \\ \hline
+  \end{tabular}
+\end{table}
+%
+As expected, within the same number of populations, there is a linear speedup
+when we add more workers. The speedup is marginal in those cases where there 
+are fewer populations than workers.  The reason for this is that at any given time, 
+there are at least two populations in the message queue, leaving some of the workers 
+waiting for populations to be available in the queue.
+
+\begin{table}[h!tbp]
+  \small
+  \caption{Speedup achieved with increasing the number of workers for
+    each dimension showing the p-value for the Wilcoxon test. 
+    Best and Worst are highlighted (bold and underlined)}
+  % Say what highlights mean - JJ
+  \label{tab:speedup:test}
+  \vspace{0.25cm}
+  \centering
+  \begin{tabular}{llllll}
+  Populations & Dimension  & \multicolumn{2}{l}{Increment}  & Speedup            & p-value\\
+  5  & 10 & 1w & 2w  & 2.15 & 1.51E-11 \\
+  & 10 & 2w & 4w  & 1.68 & 1.51E-11 \\
+  & 10 & 4w & 8w  & 1.04 & 0.004942 \\
+  & 10 & 8w & 16w & 1.05 & 0.072660 \\
+  & 20 & 1w & 2w  & \textbf{2.09} & 1.51E-11 \\
+  & 20 & 2w & 4w  & 1.69 & 1.51E-11 \\
+  & 20 & 4w & 8w  & 1.08 & 5.97E-07 \\
+  & 20 & 8w & 16w & {\ul 1.03} & 0.214482 \\
+  &    &    &     &      &          \\
+10 & 10 & 1w & 2w  & 1.91 & 1.51E-11 \\
+  & 10 & 2w & 4w  & 2.00 & 1.51E-11 \\
+  & 10 & 4w & 8w  & 2.01 & 1.51E-11 \\
+  & 10 & 8w & 16w & 1.08 & 1.39E-05 \\
+  & 20 & 1w & 2w  & 1.97 & 1.51E-11 \\
+  & 20 & 2w & 4w  & 2.01 & 1.51E-11 \\
+  & 20 & 4w & 8w  & 1.99 & 1.51E-11 \\
+  & 20 & 8w & 16w & 1.13 & 4.24E-09 \\
+  &    &    &     &      &          \\
+ 20 & 10 & 1w & 2w  & 1.96 & 1.51E-11 \\
+  & 10 & 2w & 4w  & 2.07 & 1.51E-11 \\
+  & 10 & 4w & 8w  & 2.05 & 1.51E-11 \\
+  & 10 & 8w & 16w & 1.75 & 1.51E-11 \\
+  & 20 & 1w & 2w  & 2.00 & 1.51E-11 \\
+  & 20 & 2w & 4w  & 2.04 & 1.51E-11 \\
+  & 20 & 4w & 8w  & 2.00 & 1.51E-11 \\
+  & 20 & 8w & 16w & 1.85 & 1.51E-11 \\
+ \end{tabular}
+  \end{table}
+
+Table~\ref{tab:speedup-table} shows the speedup obtained against the baseline of
+one worker. We achieved the maximum speedup when using four workers or more with
+twenty populations. On the other hand, the maximum speedup when we consider the
+cost it was with eight workers. The best speedups for each number of workers are
+shown in bold in the table.
+
+Table~\ref{tab:speedup:test} shows the speedup obtained taking as the baseline
+the previous configuration, for example, when increasing from four to eight
+workers, and the p-value for the left-tailed Wilcoxon test. There was an
+increment in all cases. Again, the best effect is achieved when the number of
+populations is higher than the number of workers. Configurations following this
+basic rule doubled the speed or where near it. Using five populations, we
+obtained the best and worst speedup depending on the number of workers (shown in
+underline and bold, respectively).
 
@@
 
-In this experiment, we wanted to test how the number of workers affected the
-time needed to reach a solution and the speedup that each configuration can
-produce. Adding more populations might add more overhead, so this might have a
-noticeable effect.
-
 %In principle, the main issue is how to find the right number
@@
 population size.
+
 Each box-plot indicates the time required for the execution of 30 function
@@
 scales with the number of workers for a fixed problem and number of
-sub-populations used. The exception is the case when using five populations;
-the time does not decrease when progressing from four to eight workers.  
-That is because most of the time, workers are waiting for a population 
-to be available in the message queue.
+sub-populations used.
 
@@
   \caption{Boxplot of the number of function evaluations required to reach the target, for 30 instances of function $f_4$, with different number of workers and populations sizes}
-  \label{fig:spworker:evalspersecond}
+  \label{fig:spworker:evals}
 \end{figure}
@@
 %
-
-\begin{figure}[h!tbp]
-  \centering
-  \includegraphics[width=\textwidth]{eval_ratio}
-  \caption{Boxplot of the number of evaluations per second, for 30 instances of function $f_4$, with different number of workers and populations sizes }
-  \label{fig:spworker:evalspersecond}
-\end{figure}
-
-\subsubsection{Results Regarding Speedup}
-\label{sec:speedup}
-
-% Introduction saying why we are measuring this, how it's related to the objectives of the paper - JJ
-% This is done on the parent section, I renamed the sections to clarify this - M
-To measure the speedup obtained by the addition of workers, we will use the
-number of evaluations per second, instead of the time required to finish an
-instance of a problem. The reason for this is that in some cases reaching the
-desired target can require less work. We are going to measure the total increase
-of work done in a second by adding more workers. The baseline for the comparison
-will be the algorithm with a single worker, and the minimum amount of work available will be five
-populations. This number is selected because the controller needs at least
-three populations to migrate individuals between them, and we add two more to
-keep even a single worker busy. Figure \ref{fig:spworker:evalspersecond} 
-shows the number of function evaluation per second, achieved by a certain number of
-workers, for every problem dimension, for five and ten dimensions of $f_4$.
-
-\begin{table}[h!tbp]
-  \caption{Speedup by worker and dimension, taking one worker as the baseline.}
-  \label{tab:speedup-table}
-  \vspace{0.25cm}
-  \centering
-
-  \begin{tabular}{c|r|r|r|r|}
-  \hline
-  \multicolumn{1}{|l|}{Dimension} & \multicolumn{1}{c|}{2W} & \multicolumn{1}{c|}{4W} & \multicolumn{1}{c|}{8W} & \multicolumn{1}{c|}{16W}  \\ \hline
-  \multicolumn{5}{| c|}{5 Populations} \\ \hline
-  \multicolumn{1}{|c|}{10}        & 2.15           & 3.61                    & 3.75                    & 3.96     \\ \hline
-  \multicolumn{1}{|c|}{20}        & 2.09           & 3.53                    & 3.83                    & 3.95     \\ \hline
-  \multicolumn{5}{| c|}{10 Populations} \\ \hline
-  \multicolumn{1}{|c|}{10}        & 1.91           & 3.82                    & 7.67                    & 8.25     \\ \hline
-  \multicolumn{1}{|c|}{20}        & 1.97           & 3.97                    & 7.90                    & 8.90     \\ \hline
-  \multicolumn{5}{| c|}{20 Populations} \\ \hline
-  \multicolumn{1}{|c|}{10}        & 1.96           & 4.05                    & 8.29                    & 14.50     \\ \hline
-  \multicolumn{1}{|c|}{20}        & 2.00           & 4.09                    & 8.18                    & 15.16     \\ \hline
-  \end{tabular}
-\end{table}
-%
-As expected, within the same number of populations, there is a linear speedup
-when we add more workers. The speedup is marginal in the case of five
-populations when increasing the number of workers above 4. The reason for this
-is that at any given time, there are at least two populations in the message
-queue, leaving four workers waiting. We can also notice that the spread in the
-amount of work is less when using up to four workers.
-
-\begin{table}[h!tbp]
-  \small
-  \caption{Speedup achieved with increasing the number of workers for
-    each dimension showing the p-value for the Wilcoxon test. Best and Worst are highlighted (bold and underlined) }
-  % Say what highlights mean - JJ
-  \label{tab:speedup:test}
-  \vspace{0.25cm}
-  \centering
-  \begin{tabular}{llllll}
-  Populations & Dimension  & \multicolumn{2}{l}{Increment}  & Speedup            & p-value\\
-  5  & 10 & 1w & 2w  & 2.15 & 1.51E-11 \\
-  & 10 & 2w & 4w  & 1.68 & 1.51E-11 \\
-  & 10 & 4w & 8w  & 1.04 & 0.004942 \\
-  & 10 & 8w & 16w & 1.05 & 0.072660 \\
-  & 20 & 1w & 2w  & 2.09 & 1.51E-11 \\
-  & 20 & 2w & 4w  & 1.69 & 1.51E-11 \\
-  & 20 & 4w & 8w  & 1.08 & 5.97E-07 \\
-  & 20 & 8w & 16w & 1.03 & 0.214482 \\
-  &    &    &     &      &          \\
-10 & 10 & 1w & 2w  & 1.91 & 1.51E-11 \\
-  & 10 & 2w & 4w  & 2.00 & 1.51E-11 \\
-  & 10 & 4w & 8w  & 2.01 & 1.51E-11 \\
-  & 10 & 8w & 16w & 1.08 & 1.39E-05 \\
-  & 20 & 1w & 2w  & 1.97 & 1.51E-11 \\
-  & 20 & 2w & 4w  & 2.01 & 1.51E-11 \\
-  & 20 & 4w & 8w  & 1.99 & 1.51E-11 \\
-  & 20 & 8w & 16w & 1.13 & 4.24E-09 \\
-  &    &    &     &      &          \\
- 20 & 10 & 1w & 2w  & 1.96 & 1.51E-11 \\
-  & 10 & 2w & 4w  & 2.07 & 1.51E-11 \\
-  & 10 & 4w & 8w  & 2.05 & 1.51E-11 \\
-  & 10 & 8w & 16w & 1.75 & 1.51E-11 \\
-  & 20 & 1w & 2w  & 2.00 & 1.51E-11 \\
-  & 20 & 2w & 4w  & 2.04 & 1.51E-11 \\
-  & 20 & 4w & 8w  & 2.00 & 1.51E-11 \\
-  & 20 & 8w & 16w & 1.85 & 1.51E-11 \\
- \end{tabular}
-  \end{table}
-%
-Table~\ref{tab:speedup-table} shows the speedup obtained against the baseline of
-one worker. When using eight workers, we achieved the maximum speed up for five
-and ten populations, but interestingly, it was in dimension 40 and dimension 2 
-(these values are shown in bold in the table).
-
-Table~\ref{tab:speedup:test} shows the speedup obtained taking as the baseline
-the previous configuration, for example, when increasing from four to eight
-workers, and the p-value for the left-tailed Wilcoxon test. There was an
-increment in all cases, but again, for smaller dimensions, five populations with
-fewer workers is a better option (see values underlined in the table) with the
-best speedup, achieved for 40 dimensions, shown in bold. The worst speed was
-going from four to eight workers at higher dimensions (underlined in the table).
-For ten populations, adding more workers resulted in a linear speedup, but the
-best speedup is obtained when increasing from four to eight workers at 40
-dimensions (shown in bold).
 

### 2020-08-17 draft.tex
Update draft.tex

@@
 
-% Introduction saying why we are measuring this, how it's related to the objectives of the paper - JJ
-% This is done on the parent section, I renamed the sections to clarify this - M
 To measure the speedup obtained by the addition of workers, we will use the
@@
 desired target can require less work, because we are using a stochastic search. 
-%% JJ, can you please check if this is clear or if it needs more explanation?
 We are going to measure the total increase of work done in a second by adding more workers.
@@
   \caption{The median of the time required to finish an instance of the function $f_4$, 30 runs
-  with different number of workers and populations} % time or rate? - JJ
+  with different number of workers and populations}
   \label{tab:time}
@@
 
-Figure \ref{fig:spworker:time} shows the execution time achieved by a certain number of
-workers, for every problem dimension, which, besides, changes the
-population size.
+Figure \ref{fig:spworker:time} shows the median (30 runs) of the execution time achieved by 2, 4 , 8, and 16  
+workers, to finish an instance of the function $f_4$, for dimensions 10 and 20, with 5, 10, and 20
+population sizes.
 
 Each box-plot indicates the time required for the execution of 30 function
-instances, in seconds. Each row has a different scale on the y-axis, to
+instances, in seconds. Each row has a different log scale on the y-axis, to
 compensate for the dimension increments. It is important to notice that the
@@
 
-In Table~\ref{tab:time} we can see that as expected, and in most cases, 
-increasing the number of workers reduces the time
-required for the execution of an instance, 
-mainly when using the same parameters
-(number of populations and number of workers), that is, performance
+In Table~\ref{tab:time} we can see that as expected, and in all cases, 
+increasing the number of workers reduces the time required for the execution of an instance, 
+mainly when using the same parameters (number of populations and number of workers), that is, performance
 scales with the number of workers for a fixed problem and number of

### 2020-08-21 draft.tex
Update draft.tex

@@
 \usepackage{xcolor}
+\usepackage{multirow}
 \usepackage[normalem]{ulem}
@@
 
+\begin{figure}[h!tbp]
+  \centering
+  \includegraphics[width=\textwidth]{evals_per_instance}
+  \caption{Boxplot of the number of function evaluations required to reach the target, for 30 instances of function $f_4$, with different number of workers and populations sizes}
+  \label{fig:evals_instance}
+\end{figure}
+
+\begin{table}[tbp]
+  \caption{The median of the number of evaluations required to finish an instance of the function $f_4$, 30 runs
+  with different number of workers and populations}
+  \label{tab:evals_instance}
+  \vspace{0.25cm}
+  \centering
+
+  \begin{tabular}{|c|c|r|r|r|}
+  \hline
+  \multicolumn{1}{|c}{} &   \multicolumn{1}{c}{}  & \multicolumn{3}{|c|}{Number of Populations} \\
+  \hline
+  Dimension           & Workers & \multicolumn{1}{c|}{5}       & \multicolumn{1}{c|}{10}      & \multicolumn{1}{c|}{20}          \\
+  \hline
+  \multirow{5}{*}{10} & 1w      & 159,625      & \textbf{142,368*}      & 151,680*     \\
+                      & 2w      & 160,258      & \textbf{137,385}      & 150,445*     \\
+                      & 4w      & 161,144      & \textbf{129,031}      & 144,354     \\
+                      & 8w      & 159,893      & \textbf{{\ul 127,987}}      & 141,896     \\
+                      & 16w     & 167,678      & \textbf{137,446}      & 138,852     \\
+                      \hline
+  \multirow{5}{*}{20} & 1w      & 462,069      & \textbf{396,841}      & 421,249*     \\
+                      & 2w      & 437,952      & \textbf{395,814}      & 407,984     \\
+                      & 4w      & 437,563      & \textbf{396,587}      & 408,461     \\
+                      & 8w      & 440,130      & \textbf{370,264}      & 372,265     \\
+                      & 16w     & 443,108      &         386,272       & \textbf{{\ul 355,999}}\\
+                      \hline  
+  \end{tabular}
+  \end{table}
+
 %
@@
               &         & 5 Populations    & 10 Populations & 20 Populations \\
-     Dimension& Workers & Median of rate & Median of rate & Median of rate  \\
+     Dimension& Workers & Median of Time   & Median of Time & Median of Time  \\
     \hline
           10  & 1w      & 22.24  & 20.48           & 24.55  \\
-              & 2w      & 10.31  & 10.00           & 12.79 \\
-              & 4w      &  6.23  & 4.75            &  5.70 \\
-              & 8w      &  5.86  &  2.33           & 2.74 \\
-              & 16w     &  5.96  &  2.33           & \textbf{1.59}\\
+              & 2w      & 10.31  & \textbf{10.00} (2.05)    & 12.79 \\
+              & 4w      &  6.23  & \textbf{4.75} (2.10)     &  5.70 \\
+              & 8w      &  5.86  &  \textbf{2.33} (2.03)           & 2.74 \\
+              & 16w     &  5.96*  &  2.33*           & {\ul 1.59} \\
 
     \hline
-          20  & 1w   & 18.36   & 65.18  & 80.31 \\
+          20  & 1w   & 73.87   & 65.18                 & 80.31 \\
               & 2w   & 34.18   & 33.46                 & 38.87 \\
-              & 4w   & 19.55   & 16.58  & 18.69 \\
-              & 8w   & 18.38   & 7.77   & 8.66 \\
-              & 16w  & 18.36   & 7.27   & \textbf{4.57} \\
+              & 4w   & 19.55   & \textbf{16.58} (2.02) & 18.69 \\
+              & 8w   & 18.38   & \textbf{7.77} (2.15)  & 8.66 \\
+              & 16w  & 18.36*  & 7.27                  & {\ul 4.57} \\
     \hline
@@
 
-% I think this is now a bit redundant or we need to move it to the parent section - M
-% But is a good intro to the results.
-% OK - JJ
-
-%In principle, the main issue is how to find the right number
-%of workers for a particular problem, but in general, we wanted to check the
-%communication overhead when the size of the problem increases and how the
-%presented framework reacted to it.
-% Commenting because we did not checked the communication over head - M
-
-Figure \ref{fig:spworker:time} shows the median (30 runs) of the execution time achieved by 2, 4 , 8, and 16  
-workers, to finish an instance of the function $f_4$, for dimensions 10 and 20, with 5, 10, and 20
-population sizes.
-
-Each box-plot indicates the time required for the execution of 30 function
-instances, in seconds. Each row has a different log scale on the y-axis, to
-compensate for the dimension increments. It is important to notice that the
-execution time can also be affected by the algorithm finding a solution with an
-error smaller than the threshold, before completing all the FEs.
+In the previous section, we have shown that as we increase the number of
+populations and workers, we increase the ratio of the number of evaluations per
+second in the same proportion. However, it is vital to notice that we can also
+affect the number of evaluations we need to execute to find a solution. This
+reduction means that the multi-population based algorithm can perform better
+when we increase the number of populations, or in some cases, even workers, as the experiments
+show. If we increase the evaluations per second and, at the same time, decrease
+the number of evaluations needed, we can have a superlinear speedup of the
+algorithm, if we measure the time required to complete an instance.
+
+First, we will see how the configuration we selected can reduce the number of
+evaluations needed in an experiment. Figure \ref{fig:evals_instance} shows the
+median (30 runs) of the number of evaluations needed to complete an instance of
+the function $f_4$. For every selection in the number of populations, we could
+expect to have about the same number of FEs needed regardless of the number of
+workers. However, we can see that for some configurations, i.e., 10D and ten
+populations, eight workers will need less FEs then one (p = 0.012). Also, for
+20D and twenty populations, there are differences between one and four workers
+(p = 0.08) and between four and sixteen (p = 0.01). These results indicate that
+there are benefits in the potential change of order in the message queue
+resulting from the parallelization of work.
+
+In Table \ref{tab:evals_instance} we can see some additional insights in more
+detail. When we increase the number of populations from five to ten, or twenty,
+the algorithm needs fewer FEs in both cases. The reduction is more pronounced
+when using ten populations, needing fewer FEs in all but a single case (shown in
+boldface). In some cases, marked with an asterisk, there is no statistical
+difference between the number of evaluations; this happens when using only one
+or two workers. The configurations that required less FEs have an underline in
+the table.
+
+The reduction in the number of FEs has an impact when we measure the speed up
+concerning the time needed to complete an experiment. In this case, we have
+measured the median of the seconds needed to complete an instance of the
+problem. Figure \ref{fig:spworker:time} shows in each box-plot the median (30
+runs) of the execution time achieved by 2, 4, 8, and 16 workers, to finish an
+instance of the function $f_4$, for dimensions 10 and 20, when using 5, 10, and
+20 population sizes. We present each dimension in a row having a log scale on
+the y-axis, to compensate for the dimension increments and scalability.
 
@@
 for this could be that when having more populations, the migration of
-individuals takes more time from the beginning to reach to all of them. For
+individuals takes more time from the bdyeginning to reach to all of them. For
 instance, when a single worker has 10 populations, only after processing every
@@
 %
-
-\begin{figure}[h!tbp]
-  \centering
-  \includegraphics[width=\textwidth]{evals_per_instance}
-  \caption{Boxplot of the number of function evaluations required to reach the target, for 30 instances of function $f_4$, with different number of workers and populations sizes}
-  \label{fig:spworker:evals}
-\end{figure}
 

### 2020-08-21 draft.tex
Update draft.tex

@@
 
-However, when we compare the two options, we find some subtle differences: When
-using one or two workers, using 5 populations takes less time in most cases. In
-Table~\ref{tab:time} results with p-value of the two-tailed Wilcoxon
-signed-ranks test was grater than the level of significance of 0.05 (i.e.,
-accepting the null hypothesis of inequality) are marked in boldface. A reason
-for this could be that when having more populations, the migration of
-individuals takes more time from the bdyeginning to reach to all of them. For
-instance, when a single worker has 10 populations, only after processing every
-one of them, it will start receiving mixed populations. Although, in all cases,
-and even with a single worker, populations are heterogeneous, individuals
-resulting from the mixing of populations take more time to propagate. With four
-workers, almost the same results can be achieved regardless of our selection
-between five or ten populations. Also, there is not much difference between 4
-and 8 workers, when using five populations, which might indicate that some
-workers are idling while populations are being processed by other workers. In
-general, we can observe that keeping a ratio between the number of workers and
-populations a bit over 1:1 exploits better the computing resources available.
-This also decreases the relation between the overhead needed to migrate and
-process incoming and outgoing population and the total computing time: more
-populations will need more time to do that.
+% However, when we compare the two options, we find some subtle differences: When
+% using one or two workers, using 5 populations takes less time in most cases. In
+% Table~\ref{tab:time} results with p-value of the two-tailed Wilcoxon
+% signed-ranks test was grater than the level of significance of 0.05 (i.e.,
+% accepting the null hypothesis of inequality) are marked in boldface. A reason
+% for this could be that when having more populations, the migration of
+% individuals takes more time from the bdyeginning to reach to all of them. For
+% instance, when a single worker has 10 populations, only after processing every
+% one of them, it will start receiving mixed populations. Although, in all cases,
+% and even with a single worker, populations are heterogeneous, individuals
+% resulting from the mixing of populations take more time to propagate. With four
+% workers, almost the same results can be achieved regardless of our selection
+% between five or ten populations. Also, there is not much difference between 4
+% and 8 workers, when using five populations, which might indicate that some
+% workers are idling while populations are being processed by other workers. In
+% general, we can observe that keeping a ratio between the number of workers and
+% populations a bit over 1:1 exploits better the computing resources available.
+% This also decreases the relation between the overhead needed to migrate and
+% process incoming and outgoing population and the total computing time: more
+% populations will need more time to do that.
 

### 2020-08-21 draft.tex
Fixes issue #17

--- 2020-08-21 draft.tex
@@
 realization, from the origins where Salza and Ferrucci
-\cite{salza2016approach,salza2016develop} proposed an architecture
-that carried evolutionary algorithms to the cloud, and which were
-developed further in their next paper \cite{de2017parallel}. These
-papers started to introduce cloud native aspects such as the use of
-messaging queues and CoreOS as an operating system designed from the
-ground up for container utilization in the area of evolutionary
-algorithms. The way these containers are managed, however, is rather
-classical in nature from the point of view of distributed EC: a
-master-worker approach, communicated using RabbitMQ (a messaging
-queue), where replicated workers perform the tasks in parallel. They
-called their approach AMQPGA, inspired by the protocol, AMQP, used by
-RabbitMQ, the messaging system. This classical approach has been also
-followed by other authors by using other EA ``classical'' topologies
-in the cloud: Dziurzanski et al. \cite{dziurzanski2020scalable} adapts an island model to
-a container architecture, in a system that is functionally equivalent
-to classical models, but with the added advantage of being easy to
-deploy and run.
+\cite{salza2016approach,salza2016develop} proposed an architecture that carried
+evolutionary algorithms to the cloud, and which were developed further in their
+next paper \cite{de2017parallel}. These papers started to introduce cloud native
+aspects such as the use of messaging queues and CoreOS as an operating system
+designed from the ground up for container utilization in the area of
+evolutionary algorithms. The way these containers are managed, however, is
+rather classical in nature from the point of view of distributed EC: a
+master-worker approach, communicated using RabbitMQ (a messaging queue), where
+replicated workers perform the tasks in parallel. They called their approach
+AMQPGA, inspired by the protocol, AMQP, used by RabbitMQ, the messaging system.
+Other authors have also followed this classical approach by using other EA
+``classical'' topologies in the cloud: Dziurzanski et al.
+\cite{dziurzanski2020scalable} adapts an island model to a container
+architecture, in a system that is functionally equivalent to classical models,
+with the advantage of the auto-scaling of islands, and being easy to deploy and
+run.
 
@@
 about how we deployed EvoSwarm in a cloud-setting for running the experiments
-within this paper.
+within this paper. Running the cloud-native application in the cloud has the
+benefit of a multi-instance deployment, where we can run containers in several
+instances executing at the same time, not only in a single machine.
 
@@
 
-  \item[A Kubernetes cluster] provides auto-scaling and management of containerized
-  applications, but it needs additional configuration for provisioning. All major
-  cloud-providers have Kubernetes services.
+  \item[A Kubernetes cluster] provides auto-scaling and management of
+  containerized applications, but it needs additional configuration for
+  provisioning. All major cloud-providers have Kubernetes services. This option
+  was used by S. Zhao et al. \cite{zhao2019cloud} to run a multi-island
+  algorithm, and Dziurzanski et al. \cite{dziurzanski2020scalable} later refined
+  it.
 
@@
 variables used by code running inside containers.
+
+\begin{figure}[h!tbp]
+  \centering
+  \includegraphics[width=\textwidth]{aws_ec2}
+  \caption{AWS EC2 configuration for the deployment of EvoSwarm containers.
+   }
+  \label{fig:aws-configuration}
+\end{figure}
 
@@
 
-\begin{figure}[h!tbp]
-  \centering
-  \includegraphics[width=\textwidth]{aws_ec2}
-  \caption{AWS EC2 configuration for the deployment of EvoSwarm containers.
-   }
-  \label{fig:aws-configuration}
-\end{figure}
 %
@@
 
-As we mentioned before in Section~\ref{method}, reactive systems can scale by adding additional copies
-of serverless functions. In our case, we can start additional worker containers
-to have the same effect. Other authors, like Salza et al. \cite{salza2019speed},
-have used a similar architecture, which implies that worker nodes need to have at least a certain level
-of complexity, in terms of execution time,  to effectively scale on multiple
-nodes, tending to linear scalability. In that paper, Salza et al. tested scalability
-by simulating the work of nodes by using a sleep function and messages of different sizes.
-In a population-based algorithm like EvoSwarm, several parameters can increase or 
-decrease the workload of workers:
-
-\begin{itemize}
-    \item The number of populations: if there are more workers than populations, workers
-    must wait for work to arrive. If there are too many populations, they could be
-    standing in the queue for more time.
-    \item The size of populations and the number of generations:
-    These parameters naturally increase the execution time, because they mean more FEs.
-    \item The complexity of the algorithm: For instance, in our case,
-    the PSO implementation has a lower execution time than the GA.
-\end{itemize}
-
 \begin{table}[h!tbp]
@@
 \end{table}
+
+As we mentioned before in Section~\ref{method}, reactive systems can scale by adding additional copies
+of serverless functions. In our case, we can start additional worker containers
+to have the same effect. Other authors, like Salza et al. \cite{salza2019speed},
+have used a similar architecture, which implies that worker nodes need to have at least a certain level
+of complexity, in terms of execution time,  to effectively scale on multiple
+nodes, tending to linear scalability. In that paper, Salza et al. tested scalability
+by simulating the work of nodes by using a sleep function and messages of different sizes.
+In a population-based algorithm like EvoSwarm, several parameters can increase or 
+decrease the workload of workers:
+
+\begin{itemize}
+    \item The number of populations: if there are more workers than populations, workers
+    must wait for work to arrive. If there are too many populations, they could be
+    standing in the queue for more time.
+    \item The size of populations and the number of generations:
+    These parameters naturally increase the execution time, because they mean more FEs.
+    \item The complexity of the algorithm: For instance, in our case,
+    the PSO implementation has a lower execution time than the GA.
+\end{itemize}
 
@@
 
+When comparing the results in this section, we performed the Wilcoxon Signed-Rank test, with a
+significance level of 5\%. We indicate with an asterisk those results that are
+not significant.
+
 To measure the speedup obtained by the addition of workers, we will use the
@@
 We are going to measure the total increase of work done in a second by adding more workers.
-The baseline for the comparison will be the algorithm with a single worker, 
-and the minimum amount of work available will be five populations. 
-This number is selected because the controller needs at least
-three populations to migrate individuals between them, and we add two more to
-keep even a single worker busy all the time. Figure \ref{fig:spworker:evalspersecond} 
-shows the number of function evaluations per second, achieved by a certain number of
-workers, for ten and twenty dimensions of $f_4$.
 
@@
 \end{table}
+
+The baseline for the comparison will be the algorithm with a single worker, 
+and the minimum amount of work available will be five populations. 
+This number is selected because the controller needs at least
+three populations to migrate individuals between them, and we add two more to
+keep even a single worker busy all the time. Figure \ref{fig:spworker:evalspersecond} 
+shows the number of function evaluations per second, achieved by a certain number of
+workers, for ten and twenty dimensions of $f_4$.
+
 %
@@
 the previous configuration, for example, when increasing from four to eight
-workers, and the p-value for the left-tailed Wilcoxon test. There was an
+workers, and the p-value for the Wilcoxon Signed-Rank test. There was an
 increment in all cases. Again, the best effect is achieved when the number of
@@
 \end{figure}
+
+%\subsubsection{Experimental Results} 
+%\label{results}
+
+% Added some explanation % here - JJ
+
+%We report the comparison regarding the execution time in Section~\ref{sec:exec-time}
+%and the summary of the speedup in section Section~ref{sec:speedup}.
+% Move this to parent - M
+
+\subsubsection{Results Regarding Execution Time}
+\label{sec:exec-time}
+
+In the previous section, we have shown that as we increase the number of
+populations and workers, we increase the ratio of the number of evaluations per
+second in the same proportion. However, it is vital to notice that we can also
+affect the number of evaluations we need to execute to find a solution. This
+reduction means that the multi-population based algorithm can perform better
+when we increase the number of populations, or in some cases, even workers, as the experiments
+show. If we increase the evaluations per second and, at the same time, decrease
+the number of evaluations needed, we can have a superlinear speedup of the
+algorithm, if we measure the time required to complete an instance.
 
@@
 
-%
-\begin{figure}[h!tbp]
-  \centering
-  \includegraphics[width=\textwidth]{secs_per_instance}
-  \caption{Boxplot of the median time required for the execution of 30 
-  instances of function $f_4$, in seconds, different number of workers and populations sizes. The y-axis is shown in log scale.  }
-  \label{fig:spworker:time}
-\end{figure}
-%
-\begin{table}[tbp]
-  \caption{The median of the time required to finish an instance of the function $f_4$, 30 runs
-  with different number of workers and populations}
-  \label{tab:time}
-  \vspace{0.25cm}
-    \centering
-    \begin{tabular}{ccrrr}
-    \hline
-              &         & 5 Populations    & 10 Populations & 20 Populations \\
-     Dimension& Workers & Median of Time   & Median of Time & Median of Time  \\
-    \hline
-          10  & 1w      & 22.24  & 20.48           & 24.55  \\
-              & 2w      & 10.31  & \textbf{10.00} (2.05)    & 12.79 \\
-              & 4w      &  6.23  & \textbf{4.75} (2.10)     &  5.70 \\
-              & 8w      &  5.86  &  \textbf{2.33} (2.03)           & 2.74 \\
-              & 16w     &  5.96*  &  2.33*           & {\ul 1.59} \\
-
-    \hline
-          20  & 1w   & 73.87   & 65.18                 & 80.31 \\
-              & 2w   & 34.18   & 33.46                 & 38.87 \\
-              & 4w   & 19.55   & \textbf{16.58} (2.02) & 18.69 \\
-              & 8w   & 18.38   & \textbf{7.77} (2.15)  & 8.66 \\
-              & 16w  & 18.36*  & 7.27                  & {\ul 4.57} \\
-    \hline
-  \end{tabular}
-\end{table}
-
-%\subsubsection{Experimental Results} 
-%\label{results}
-
-% Added some explanation % here - JJ
-
-%We report the comparison regarding the execution time in Section~\ref{sec:exec-time}
-%and the summary of the speedup in section Section~ref{sec:speedup}.
-% Move this to parent - M
-
-\subsubsection{Results Regarding Execution Time}
-\label{sec:exec-time}
-
-In the previous section, we have shown that as we increase the number of
-populations and workers, we increase the ratio of the number of evaluations per
-second in the same proportion. However, it is vital to notice that we can also
-affect the number of evaluations we need to execute to find a solution. This
-reduction means that the multi-population based algorithm can perform better
-when we increase the number of populations, or in some cases, even workers, as the experiments
-show. If we increase the evaluations per second and, at the same time, decrease
-the number of evaluations needed, we can have a superlinear speedup of the
-algorithm, if we measure the time required to complete an instance.
-
 First, we will see how the configuration we selected can reduce the number of
@@
 
+%
+\begin{figure}[h!tbp]
+  \centering
+  \includegraphics[width=\textwidth]{secs_per_instance}
+  \caption{Boxplot of the median time required for the execution of 30 
+  instances of function $f_4$, in seconds, different number of workers and 
+  populations sizes. The y-axis is shown in log scale.}
+  \label{fig:spworker:time}
+\end{figure}
+%
+
 The reduction in the number of FEs has an impact when we measure the speed up
@@
 
-In Table~\ref{tab:time} we can see that as expected, and in all cases, 
-increasing the number of workers reduces the time required for the execution of an instance, 
-mainly when using the same parameters (number of populations and number of workers), that is, performance
-scales with the number of workers for a fixed problem and number of
-sub-populations used.
-
-However, when we compare the two options, we find some subtle differences: When
-using one or two workers, using 5 populations takes less time in most cases. In
-Table~\ref{tab:time} results with p-value of the two-tailed Wilcoxon
-signed-ranks test was grater than the level of significance of 0.05 (i.e.,
-accepting the null hypothesis of inequality) are marked in boldface. A reason
-for this could be that when having more populations, the migration of
-individuals takes more time from the bdyeginning to reach to all of them. For
-instance, when a single worker has 10 populations, only after processing every
-one of them, it will start receiving mixed populations. Although, in all cases,
-and even with a single worker, populations are heterogeneous, individuals
-resulting from the mixing of populations take more time to propagate. With four
-workers, almost the same results can be achieved regardless of our selection
-between five or ten populations. Also, there is not much difference between 4
-and 8 workers, when using five populations, which might indicate that some
-workers are idling while populations are being processed by other workers. In
-general, we can observe that keeping a ratio between the number of workers and
-populations a bit over 1:1 exploits better the computing resources available.
-This also decreases the relation between the overhead needed to migrate and
-process incoming and outgoing population and the total computing time: more
-populations will need more time to do that.
-
-%
-%
-
-%
+\begin{table}[tbp]
+  \caption{The median of the time required to finish an instance of the function $f_4$, 30 runs
+  with different number of workers and populations}
+  \label{tab:time}
+  \vspace{0.25cm}
+    \centering
+    \begin{tabular}{ccrrr}
+    \hline
+              &         & 5 Populations    & 10 Populations & 20 Populations \\
+     Dimension& Workers & Median of Time   & Median of Time & Median of Time  \\
+    \hline
+          10  & 1w      & 22.24  & 20.48           & 24.55  \\
+              & 2w      & 10.31  & \textbf{10.00} (2.05)    & 12.79 \\
+              & 4w      &  6.23  & \textbf{4.75} (2.10)     &  5.70 \\
+              & 8w      &  5.86  &  \textbf{2.33} (2.03)           & 2.74 \\
+              & 16w     &  5.96*  &  2.33*           & {\ul 1.59} \\
+
+    \hline
+          20  & 1w   & 73.87   & 65.18                 & 80.31 \\
+              & 2w   & 34.18   & 33.46                 & 38.87 \\
+              & 4w   & 19.55   & \textbf{16.58} (2.02) & 18.69 \\
+              & 8w   & 18.38   & \textbf{7.77} (2.15)  & 8.66 \\
+              & 16w  & 18.36*  & 7.27                  & {\ul 4.57} \\
+    \hline
+  \end{tabular}
+\end{table}
+
+In Table~\ref{tab:time} we can see that in some cases we have a supralinear
+speedup, we highlighted these in boldface and indicated the speedup value in
+parenthesis. The best overall times are underlined. As expected, increasing the
+number of workers for a fixed dimension and populations reduced the time
+required to find a solution, as long as the population number was higher than
+the number of workers. The values that had no statistical difference (in the
+Wilcoxon test) when increasing the number of workers have an asterisk.
 
@@
 
-% In this paper, we have presented the EvoSwarm framework that implements
-% an event-based, cloud-native architecture suitable for hosting 
-% multi-population optimization algorithms. The framework is
-% based in industry-standard development tools such as Docker and Docker
-% Compose, making it easy to develop and deploy multi-algorithm distributed
-% experiments.
-
-% The set up of this kind of algorithms will have a big influence on the
-% results, since diversity is important for this kind of algorithms and
-% its creation and destruction play a big part of the algorithm's
-% performance. One of the main levers to control it in EvoSwarm is via
-% the number of workers that will be allocated from the beginning, which
-% is why we have performed an experiment mainly intended as a proof of
-% concept, but also devoted to show how two parameters in the framework,
-% number of workers and number of populations used to initially seed the
-% system, interact with each other. The conclusion of this experiment is
-% that there is an interesting interplay, with the most benefit being
-% obtained with a number of populations equal to the number of workers
-% or slightly inferior or superior, which brings algorithmic benefit
-% (experiment taking less time because it needs less evaluations) and
-% also performance benefits (experiment taking less time because
-% evaluations are being split between different workers.
-
-% % this is not a conclusion. The conclusion is an answer to the
-% % research question. This works, and works precisely for the reasons
-% % we say it works. - JJ
-% % Again, revise search questions and explain answers - JJ
-% As a conclusion, EvoSwarm is a framework designed and tested according
-% to industry standards which is able to bring many benefits to the
-% bioinspired algorithm community, by showing the possibility of
-% creating new, high performance, mixed algorithms, which can also be
-% easily deployed to the cloud, or locally to systems that simply need
-% to have Docker and {\tt docker-compose} installed.
-
-This work also opens new possibilities. From the design point of view, it
-would be interesting to try and use automatic scaling, instead of
-setting the number of workers by hand. This would need a certain
-amount of redesign, the most important of which would be using
-container orchestration systems such as Kubernetes or Docker
-Swarm. Instead of setting a fixed number of workers, only the maximum
-amount would need to be established, which could be done directly or
-by setting an evaluation budget.
+This work also opens new possibilities. From the design point of view, it would
+be interesting to use automatic scaling, instead of setting the number of
+workers by hand. This would need a certain amount of configuration, as we would
+be using container orchestration with auto-scaling capabilities such as
+Kubernetes or Docker Swarm. The current orchestration (Amazon ECS) also has
+auto-scaling capabilities, but additional configuration is needed; also, we need
+to establish the auto-scaling of both containers and Amazons EC2 instances.
+Instead of setting a fixed number of workers, only the maximum amount would need
+to be established, which could be done directly or by setting an evaluation
+budget.
 
@@
 
-\section*{References}
+%\section*{References}
 

### 2020-08-21 draft.tex
Please @mariosky check this

@@
 \begin{table}[tbp]
-  \caption{The median of the number of evaluations required to finish an instance of the function $f_4$, 30 runs
+  \caption{Median of the number of evaluations required to finish an instance of the function $f_4$ in 30 runs
   with different number of workers and populations}
@@
 tools such as Docker and Docker Compose, making it easy for scientists and practitioners to develop and deploy
+% The last version still uses Docker Compose? - JJ
 multi-algorithm distributed experiments. The model implemented by the

### 2020-08-21 draft.tex
Again, question for @mariosky

@@
 
-We proposed a new methodology for the reproducible deployment and execution of
+We have proposed a new methodology for the reproducible deployment and execution of
 multiple experiments by specifying the infrastructure as part of an experiment
@@
 researchers. 
-We tested the deployment of an experiment on Amazon's Elastic
+We have tested the deployment of an experiment on Amazon's Elastic
 Container Service, and proved that as a cloud-native, container-based application,
@@
 
-We did an empirical evaluation of the same framework to learn if the proposed
+We have performed an empirical evaluation of the same framework to learn if the proposed
 solution can efficiently improve the scalability and performance of
@@
 and workers in the system. This experiment concludes that there is an
-interesting interplay, with the most benefit obtained with a number of
+interesting interplay between them, with the most benefit obtained with a number of
 populations slightly above the number of workers, which brings performance gains 
-in terms of execution speed with experiments taking less time since
-evaluations are performed simultaneously by different workers.
+in terms of execution speed; these experiments take less time since
+evaluations are performed, as part of its corresponding algorithm, simultaneously by different workers.
 Also, we found that for lower dimensions, there is an advantage in using fewer
@@
 taking less time because they need fewer evaluations) of distributed
-multi-algorithm algorithms. We compared the homogeneous and an ensemble of
+multi-algorithm algorithms such as the ones implemented in this kind of framework.
+We compared the homogeneous and an ensemble of
 multi-populations, using Genetic Algorithms (GAs) and Particle Swarm
@@
 hybrid method using both algorithms 
-achieved better results and, in most cases, needed fewer evaluations.
+achieved better results and, in most cases, needed fewer evaluations. % Clarify if they actually achieve that way superlinear improvements - JJ
 

### 2020-08-21 draft.tex
Spell check

@@
 capable {\texttt C5 High-CPU XLarge} instance,  with 
-four vCPUs and 4096 cpu shares, with this configuration each worker gets a pair of vCPUs.  
+four vCPUs and 4096 CPU shares, with this configuration each worker gets a pair of vCPUs.  
 We scaled the number if instances as needed, using one instance for every two {\texttt worker} containers.
@@
 run a Bastion instance in a public subnet, from where we send the commands to execute the experiments 
-and later collected the results. We conected to this instance through a {\texttt ssh} session from 
+and later collected the results. We connected to this instance through a {\texttt ssh} session from 
 our personal workstation. This is a {\texttt R5AD Large} instance, selected with 75GiB of storage, needed for
-keeping the logs, and more memory (16GiB) for processing the logs in memory. The details these EC2 instabces 
+keeping the logs, and more memory (16GiB) for processing the logs in memory. The details these EC2 instances 
 and their cost at the time of writing are shown in Table \ref{tab:ec2}.
@@
 
+% Probably a final conclusion should be added here - JJ<
+
 This work also opens new possibilities. From the design point of view, it would

### 2020-08-21 draft.tex
6.3 revised; please @mariosky check 1 comment

@@
 about how we deployed EvoSwarm in a cloud-setting for running the experiments
-within this paper. Running the cloud-native application in the cloud has the
+within this paper. Running the cloud-native application in the cloud has the obvious
 benefit of a multi-instance deployment, where we can run containers in several
@@
 
-When we deploy a cloud-native application to a cloud provider, we have two
-options to choose:
+When we deploy a cloud-native application to a cloud provider, there
+are, among others, two options to choose from:
 
@@
   containerized applications, but it needs additional configuration for
-  provisioning. All major cloud-providers have Kubernetes services. This option
+  provisioning and its cost might be higher. All major cloud-providers have Kubernetes services. This option
   was used by S. Zhao et al. \cite{zhao2019cloud} to run a multi-island
@@
     files for configuration and provide transparent provisioning. Examples of these
-    services are Amazon Elastic Container Service,  Red Hat Openshift, and Marantis Docker Enterprise.
+    services are Amazon Elastic Container Service,  Red Hat Openshift
+    (which is also open source), and Marantis Docker Enterprise.
     Salza et al. \cite{salza2019speed} tested this option by using CoreOs cluster.
@@
 alternative.
-
-% Remember to mention cost at the end - JJ 
-% Ping - JJ
 
@@
 provision the  EC2 containers, networking, and group permissions for each
-experimental setup, for instance, when going from 2 to 8 workers, we need to increase the
+experimental setup; for instance, when going from 2 to 8 workers, we need to increase the
 number of EC2 instances to meet the required amount of CPU units.
@@
 The main attributes are the {\texttt cpu\_shares} indicating the CPU units
-assigned, in this case 2048; and the network configuration, we have two private 
+assigned, in this case $2048$, and the network configuration; in this case we have two private 
 subnets and a security group in this configuration. When deploying to EC2 containers, these requirements
@@
 
-Each container is defined as a {\texttt service} (task in AWS terms) , with the following attributes: {\texttt image},
+Each container is defined as a {\texttt service} ({\em task} in AWS terms), with the following attributes: {\texttt image},
 that reference the container images we have published in Docker hub
@@
   }
+% Add authorship if necessary - JJ
   \label{fig:aws-configuration}
@@
 Figure \ref{fig:aws-configuration} shows the basic EC2 configuration used for the deployment
-of experiments. We selected three types of EC2 instances according to our needs, the  
+of experiments. We selected three types of EC2 instances according to our needs; the  
 {\texttt setup}, {\texttt controller}, and {\texttt redis} containers do not demand many resources
@@
 Because all these instances run in a private subnet, we also
-run a Bastion instance in a public subnet, from where we send the commands to execute the experiments 
+run a Bastion instance in a public subnet, from where we send the
+% what is a Bastion instance? - JJ
+commands to execute the experiments
 and later collected the results. We connected to this instance through a {\texttt ssh} session from 

### 2020-08-21 draft.tex
Explain out Bastion

@@
 Because all these instances run in a private subnet, we also
-run a Bastion instance in a public subnet, from where we send the
-% what is a Bastion instance? - JJ
+run a Bastion instance, that is, a specially protected network node, in a public subnet, from where we send the
 commands to execute the experiments
@@
 
-In the prior section, we describe an option for an EvoSwarm deployment that can be used by 
-the public to create reproductive research in the cloud, in this section we are
+In the previous section, we have described an option for an EvoSwarm deployment that can be used by 
+the public to create reproductive research in the cloud; in this section we are
 using the same framework to learn if the proposed solution can efficiently
 improve the scalability and performance of population-based optimization
-algorithms. % If you want to prove it *improves*, you need to provide a baseline. 
-            %What will be your baseline? You need to specify anyway 
-            %in which way it really improves - JJ
-            % The baseline is a single population GA or PSO? - M
-% For instance. But I was actually looking for another word that does not imply a comparison - JJ
-Hence, in the following sections, we verify the following questions:
+algorithms. 
+Hence, in the following sections, we provide answers to the following questions:
 
@@
 Section~\ref{sec:exp1} is devoted to answering the first question, and Section~\ref{sec:exp2} to the second.
-Next, we describe the general experimental setup to both experiments.
+Next, we describe the general experimental setup designed to answer
+both questions.
 
@@
 crossover probabilities to have more heterogeneous workers;
-in these parameters, the range of values is specified. We did not change these parameters during the 
+for these parameters, the range of values is specified. We did not change these parameters during the 
 experiments, and only the number of populations, population size and number of generations were

### 2020-08-21 draft.tex
Remove request for authorship
Another question for @mariosky

@@
   }
-% Add authorship if necessary - JJ
   \label{fig:aws-configuration}
@@
 specify a value of $0$, this will give us an algorithm with only PSO
-populations, and finally, a value of  $1$ means that all populations will run
+populations, and finally, a value of  $1$ means that every population will run
 the GA algorithm.
@@
 population size. Finally, the {\em Iterations} parameter indicates how many
-complete loops will be performed. All these parameters give us the maximum
+complete loops will be performed. % These are generations, right ? - jj
+All these parameters give us the maximum
 number of FEs that will be performed. For instance, for $D = 2$, the maximum

### 2020-08-21 draft.tex
Some fixes

@@
 From this description it can be deduced that we give more importance in this
-paper to the number of populations, since they are the basic unit of work that
+paper to the number of populations, since they are the basic unit of
+work, the message, that
 can be executed in parallel. Together with the size of each population and the
@@
 number of populations and the number of workers on the scalability of the
-system, we propose an experiment in which we selected $f_4$ (Skew
+system, and in order to achieve this we propose an experiment for
+which we have selected $f_4$ (Skew
 Rastrigin-Bueche separable) from the BBOB testbed, over 10 and 20 dimensions. This function has been used
@@
   \caption{Boxplot of the number of evaluations per second, for 30 instances of function $f_4$,
-          with different number of workers and number of populations. The 7 axis is in log scale.}
+          with different number of workers and number of populations. The $y$ axis is in log scale.}
   \label{fig:spworker:evalspersecond}

### 2020-08-21 draft.tex
Clarify this @mariosky
Another question for @mariosky
Final comment
Check out this explanation @mariosky

@@
 population size. Finally, the {\em Iterations} parameter indicates how many
-complete loops will be performed. % These are generations, right ? - jj
+complete loops, that is, complete message processings, will be performed.
 All these parameters give us the maximum
@@
 significance level of 5\%. We indicate with an asterisk those results that are
-not significant.
+not significant. % Where? - JJ
 
@@
 \begin{table}[h!tbp]
-  \caption{Speedup by worker and dimension, taking one worker as the baseline.}
+  \caption{Speedup by worker and dimension, taking one worker as the
+    baseline.}
+% Clarify here what you are measuring. Number of evaluations for
+% whatever divided by whatever - JJ
   \label{tab:speedup-table}
@@
     with different number of workers and populations}
+  % But this is total evaluation for all workers, right? - JJ
   \label{tab:evals_instance}
@@
 
-In Table \ref{tab:evals_instance} we can see some additional insights in more
-detail. When we increase the number of populations from five to ten, or twenty,
+In Table \ref{tab:evals_instance} we try to acquire more insights on
+how this works. When we increase the number of populations from five to ten, or twenty,
 the algorithm needs fewer FEs in both cases. The reduction is more pronounced
@@
 
-% Probably a final conclusion should be added here - JJ<
+% Probably a final conclusion should be added here, including BBOB
+% benchmark and the overall performance - JJ
 

### 2020-08-21 draft.tex
Answers
Response to a12e4f1
Carification
References comment in 58d851e
Clarification
refernces 071d207

@@
 significance level of 5\%. We indicate with an asterisk those results that are
-not significant. % Where? - JJ
-
+not significant throughout the tables of this section.  % Where? - JJ
+% Answered - Mario
 To measure the speedup obtained by the addition of workers, we will use the
@@
   \caption{Speedup by worker and dimension, taking one worker as the
-    baseline.}
+    baseline. Speedup is obtained by dividing the evaluation ratio of the baseline by the 
+    ratio of the increased number of workers.}
 % Clarify here what you are measuring. Number of evaluations for
 % whatever divided by whatever - JJ
+% Done - Mario
   \label{tab:speedup-table}
@@
 % The last version still uses Docker Compose? - JJ
+% Well it uses docker-compose files, and these files are  
+% converted to tasks and services by the ECS-CLI. -Mario
+% It still can be used to run the application locally - Mario 
 multi-algorithm distributed experiments. The model implemented by the
@@
 hybrid method using both algorithms 
-achieved better results and, in most cases, needed fewer evaluations. % Clarify if they actually achieve that way superlinear improvements - JJ
+achieved better results and, in most cases, needed fewer evaluations.
+
+% Clarify if they actually achieve that way superlinear improvements - JJ
+% They should, in this experiment we only tested the #FEs, but they imply 
+% a superlinear improvements as we established in the previous section - Mario
 

### 2020-08-21 draft.tex
Update draft.tex
Conclusions

@@
   % But this is total evaluation for all workers, right? - JJ
+  %  Yes - Mario
   \label{tab:evals_instance}
@@
 Also, we found that for lower dimensions, there is an advantage in using fewer
-populations. % Specify that.
+populations.
 
@@
 
+We conclude that superlinear improvements are achievable by selecting an
+appropriate configuration consisting of the number of populations, worker
+containers, and algorithm selection. Further research should focus on learning
+more about how these parameters interact with each other. The BBOB benchmark
+results show that the selection and mixing of algorithms can lower the number of
+FEs needed to find a solution or find solutions that single algorithms could not
+find.
+
 % Probably a final conclusion should be added here, including BBOB

### 2020-08-22 draft.tex
Update draft.tex

@@
   \includegraphics[width=\textwidth]{evals_per_instance}
-  \caption{Boxplot of the number of function evaluations required to reach the target, for 30 instances of function $f_4$, with different number of workers and populations sizes}
+  \caption{Boxplot of the number of function evaluations required to reach the target, for 30 instances of function $f_4$, with a different number of workers and populations sizes}
   \label{fig:evals_instance}
@@
   \caption{Median of the number of evaluations required to finish an instance of the function $f_4$ in 30 runs
-    with different number of workers and populations}
+    with a different number of workers and populations}
   % But this is total evaluation for all workers, right? - JJ
@@
   \caption{Boxplot of the median time required for the execution of 30 
-  instances of function $f_4$, in seconds, different number of workers and 
+  instances of function $f_4$, in seconds, a different number of workers and 
   populations sizes. The y-axis is shown in log scale.}
@@
   \caption{The median of the time required to finish an instance of the function $f_4$, 30 runs
-  with different number of workers and populations}
+  with a different number of workers and populations}
   \label{tab:time}

### 2020-08-22 draft.tex
Update draft.tex

@@
 benefit of a multi-instance deployment, where we can run containers in several
-instances executing at the same time, not only in a single machine.
+instances executing at the same time, not only in a single machine, not only 
+in a single machine as it is the case when using docker-compose.
 
@@
   \item[A Kubernetes cluster] provides auto-scaling and management of
-  containerized applications, but it needs additional configuration for
-  provisioning and its cost might be higher. All major cloud-providers have Kubernetes services. This option
+  containerized applications, and it supports other container technologies 
+  but it needs additional configuration for provisioning, 
+  and its cost might be higher. All major cloud-providers have Kubernetes services. This option
   was used by S. Zhao et al. \cite{zhao2019cloud} to run a multi-island
@@
     services are Amazon Elastic Container Service,  Red Hat Openshift
-    (which is also open source), and Marantis Docker Enterprise.
-    Salza et al. \cite{salza2019speed} tested this option by using CoreOs cluster.
+    (which is also open source), and Marantis Docker Enterprise using Docker Swarm.
+    Salza et al. \cite{salza2019speed} tested a variant of this option by using a CoreOs cluster.
 

### 2020-08-22 draft.tex
Minor reformatting and fix bib warning

@@
 
-A second experiment centered on evaluating the algorithmic benefits (experiments
-taking less time because they need fewer evaluations) of distributed
-multi-algorithm algorithms such as the ones implemented in this kind of framework.
-We compared the homogeneous and an ensemble of
-multi-populations, using Genetic Algorithms (GAs) and Particle Swarm
-Optimization (PSO) using the COCO benchmark framework 
-for visualization and definition of the optimization problems.
-We found that the set-up of these algorithms has a significant influence on
+A second experiment centered on evaluating the algorithmic benefits
+(experiments taking less time because they need fewer evaluations) of
+distributed multi-algorithm algorithms such as the ones implemented in
+this kind of framework.  We compared the homogeneous and an ensemble
+of multi-populations, using Genetic Algorithms (GAs) and Particle
+Swarm Optimization (PSO) using the COCO benchmark framework for
+visualization and definition of the optimization problems.  We found
+that the set-up of these algorithms has a significant influence on
 their results since diversity is essential for their performance. The
-hybrid method using both algorithms 
-achieved better results and, in most cases, needed fewer evaluations.
-
-% Clarify if they actually achieve that way superlinear improvements - JJ
-% They should, in this experiment we only tested the #FEs, but they imply 
-% a superlinear improvements as we established in the previous section - Mario
+hybrid method using both algorithms achieved better results and, in
+most cases, needed fewer evaluations.
 
@@
 appropriate configuration consisting of the number of populations, worker
-containers, and algorithm selection. Further research should focus on learning
+containers, and algorithm combination. Further research should focus on learning
 more about how these parameters interact with each other. The BBOB benchmark
@@
 find.
-
-% Probably a final conclusion should be added here, including BBOB
-% benchmark and the overall performance - JJ
 

### 2020-08-22 draft.tex
Reformatting and rewriting

@@
 
-On the other hand, there are several caveats as well:
-It is more costly to move entire populations as messages than only passing certain individuals 
-from a process to another. Designers must consider this cost when working with large
-individuals, and if possible send populations using several messages, use compression or 
-adjust the size of populations. Pool-based algorithms also suffer from this drawback, but 
-web based implementations have been working with continuous optimization problems of 1000 dimensions.
-But, in other cases this is not a viable approach.
+On the other hand, there are several caveats as well: It is more
+costly to move entire populations as messages than only passing
+certain individuals from a process to another. Designers must consider
+this cost when working with large individuals, and if possible send
+populations using several messages, use compression or adjust the size
+of populations. Pool-based algorithms also suffer from this drawback,
+but web based implementations have been working with continuous
+optimization problems with 1000 dimensions.  However, in other cases
+this is not a viable approach.
 
@@
 
-Finally, the proposed PSO\&GA mixed algorithm has reached the $10^{-8}$ target
+Finally, the proposed PSO\&GA hybrid algorithm has reached the $10^{-8}$ target
 value on all functions ($f_1-f_5$) and scales well to higher dimensions, even on

### 2020-10-14 draft.tex
Title channged
Closes #26
Removed paragraph about credit cads
Closes #25

@@
 
-\title{EvoSwarm: A Cloud-Native Architecture for the Reproducible Execution of Multi-population Optimization Algorithms}
+\title{A Container-Based Cloud-Native Architecture for the Reproducible Execution of Multi-population Optimization Algorithms}
 
@@
 
-We have mentioned several advantages of a cloud-native application for running
-multi-population based algorithms, but there are in our experience, several
-issues that make this approach more practical. % several issues that make it practical or useful instead? - JJ 
-In general, cloud platforms have free-tiers    
-for their services, and many offer academic discounts to students. But once we
-pass to the pay-as-you-go tier, computing resources need to be
-managed, and, where possible, its cost minimized.  First,
-in many cases,  a credit-card is needed to guarantee the payment of the
-resources consumed and someone responsible for assigning quotas and, more
-importantly, special keys for using the cloud's API.  If these keys are
-compromised or shared by their holders,  additional costs or administrative
-procedures are in order. For those cases when a prototype platform is needed,
-for instance, students developing a new algorithm variant or experimenting with
-a published work, a container-based platform can be more practical. Container
-technology enables researchers the local deployment of all the infrastructure
-needed to run the experiments. Later, if needed, the same system could be
-deployed to the cloud.
+% We have mentioned several advantages of a cloud-native application for running
+% multi-population based algorithms, but there are in our experience, several
+% issues that make this approach more practical. % several issues that make it practical or useful instead? - JJ 
+% In general, cloud platforms have free-tiers    
+% for their services, and many offer academic discounts to students. But once we
+% pass to the pay-as-you-go tier, computing resources need to be
+% managed, and, where possible, its cost minimized.  First,
+% in many cases,  a credit-card is needed to guarantee the payment of the
+% resources consumed and someone responsible for assigning quotas and, more
+% importantly, special keys for using the cloud's API.  If these keys are
+% compromised or shared by their holders,  additional costs or administrative
+% procedures are in order. For those cases when a prototype platform is needed,
+% for instance, students developing a new algorithm variant or experimenting with
+% a published work, a container-based platform can be more practical. Container
+% technology enables researchers the local deployment of all the infrastructure
+% needed to run the experiments. Later, if needed, the same system could be
+% deployed to the cloud.
 

### 2020-10-14 draft.tex
No in-text URLs
Closes #23

@@
 heterogeneous populations. The platform uses the {\tt docker-compose} tool for
-defining and running experiments as multi-container applications. EvoSwarm is also released with a free licence in \url{https://github.com/mariosky/EvoSwarm}.
+defining and running experiments as multi-container applications. EvoSwarm is also released with a free licence in
+\footnote{\url{https://github.com/mariosky/EvoSwarm}}.
 We preferred the use of a container-based application for this prototype because
@@
 The open-source library is called Reactive Extensions (ReactiveX) for Python 
-\url{https://github.com/ReactiveX/RxPY} an it is based on the Observer 
+\footnote{\url{https://github.com/ReactiveX/RxPY}} an it is based on the Observer 
 pattern \cite{gamma1995design} and functional programming.
@@
 that reference the container images we have published in Docker hub
-\url{https://hub.docker.com/u/mariosky}, open {\texttt ports}, a {\texttt command} to be executed at
+\footnote{\url{https://hub.docker.com/u/mariosky}}, open {\texttt ports}, a {\texttt command} to be executed at
 startup, configuration for the AWS {\texttt logging} service, and finally {\texttt environment}

### 2020-10-14 draft.tex
Fixed grammar issues
These fixes were highlithed by reviewer #4 it closes #19

@@
 
-Researchers 
-and algorithm designers willing to  adapt to the cloud their
-multi-population solutions face the additional challenges:
+Researchers  and algorithm designers willing to adapt to the cloud; however, their
+multi-population solutions face additional challenges:
 \begin{itemize}
@@
 %%%%%% to method - JJ
-Some parameters, specially the population size, are
-kept fixed in order to control more easily the execution of the algorithm. For
+Some parameters, especially the population size, are
+kept fixed in order to control the execution of the algorithm more easily. For
 instance, by having the size of populations fixed, it is easier to control
@@
 % introduction - JJ 
-Based on these general principles we define next the proposed model.
+Based on these general principles, we next define the proposed model.
 
@@
 benefit of a multi-instance deployment, where we can run containers in several
-instances executing at the same time, not only in a single machine, not only 
-in a single machine as it is the case when using docker-compose.
+instances executing at the same time, not only in a single machine as it is the case when using docker-compose.
 
@@
 single {\texttt C5  High-CPU large} instance. With the available 2048 CPU shares divided as follows:
-{\texttt setup} (128),  {\texttt controller} (896),{\texttt redis} (1024). With this assignment 
+{\texttt setup} (128),  {\texttt controller} (896), {\texttt redis} (1024). With this assignment 
 the {\texttt redis} container gets a dedicated vCPU, needed because it is the most demanding 
@@
 four vCPUs and 4096 CPU shares, with this configuration each worker gets a pair of vCPUs.  
-We scaled the number if instances as needed, using one instance for every two {\texttt worker} containers.
+We scaled the number of instances as needed, using one instance for every two {\texttt worker} containers.
 

### 2020-10-14 draft.tex
Not a challenge
Fixes #22
Introduction
Proof read and deleted a repetiive paragraph. @JJ  please check.

@@
 
-Besides being all population-based, they also share the common characteristic
-of creating an initial set of random candidate solutions that are later used to
-generate a new set of candidates, using a nature-inspired heuristic.
-Population-based algorithms are intrinsically parallel, and can be implemented to run asynchronously.
+Besides being all population-based, they also share the common characteristic of 
+creating an initial set of random candidate solutions later used by a 
+nature-inspired heuristic to generate a new set of candidates. 
+Population-based algorithms are intrinsically parallel.
 The fitness of each individual can be (in general) evaluated independently of 
 others, and similarly, each population could evolve in isolation. Researchers use the term
-multi-population based methods when generally referring to techniques using
+multi-population based methods when referring to techniques using
 many populations as part of their optimization strategy. 
-Since 
-earlier works, researchers have been proposing some form of parallelization
-\cite{muhlenbein1988evolution} to increase the scalability of these
-algorithms that adds to this simple parallel execution some form of
-interaction between the independently running populations.
+Since earlier works, researchers have been proposing some form of parallelization
+\cite{muhlenbein1988evolution} to increase these algorithms' scalability
+by adding some form of interaction between populations running in isolation.
 The island model was one of the first techniques proposed for parallelization,
-which led to an early proof of speedup   \cite{gorges1990explicit,grosso1985computer}. 
+which led to early proofs of speed-up \cite{gorges1990explicit,grosso1985computer}. 
 The concept was to divide a large population into communicated subpopulations. 
-Since then, other population-based algorithms have adopted the concept, 
-and researchers have found additional benefits
+Since then, other population-based algorithms have adopted the idea. 
+Researchers have found additional benefits
 besides the execution speed; these include avoiding a premature convergence and
@@
 Moreover, parallel implementations can run in two different ways: synchronous and
-asynchronous. When operations run in parallel but every node must maintain synchrony with
-the other operations,  all operations need to finish to continue to the next
-step. For instance, in a  master/worker model, the master needs to wait for all
-workers to finish their operations before moving to the next iteration.  In
+asynchronous. When operations run in parallel, but every node must maintain synchrony with
+the other operations, they all need to finish before moving to the next
+step. For instance, in a  master/worker model, masters need to wait for all
+workers to complete their operations before moving to the next iteration.  In
 contrast, in an asynchronous execution,  operations are not synchronized with
 each other, so there's no time lost waiting for other populations to
-arrive to the synchronization point; results arriving from other workers are 
+arrive at the synchronization point; results coming from other workers are 
 incorporated as soon as they are received. Following the previous example, now the master operation can
-continue to the next iteration, even if a worker has not finished its work.  In
-the literature, we can find many asynchronous algorithms
-\cite{coleman89,baugh2003asynchronous}, reporting benefits in execution time,
+continue to the next iteration, even if a worker has not finished its work.  We can find many asynchronous algorithms
+\cite{coleman89,baugh2003asynchronous}, reporting benefits in execution time
 and scalability. In the particular case of asynchronous and cloud-native
@@
 through a central repository \cite{sofea:cec2012, JSON} or message queues
-\cite{salza2019speed, guervos2018introducing}, in this work we follow this
+\cite{salza2019speed, guervos2018introducing}. In this work, we follow this
 practice.
@@
 multi-core systems \cite{Serrano2008,lai2019adaptive} to multi-threaded \cite{merelo2019scaling} ones.
-And recently the focus has been on exploiting a higher number of processing
+Moreover, recently, the focus has been on exploiting a higher number of processing
 units by using GPUs \cite{tan2015survey,li2007efficient}, or distributed
@@
 and, more importantly, cloud computing
-\cite{GValdez2015,salza2019speed,valenzuela2015implementing,FlexGP}. Cloud computing has become the standard way of running
-enterprise applications, not only because of the convenience of the
+\cite{GValdez2015,salza2019speed,valenzuela2015implementing,FlexGP}.
+
+%% JJ Can you improve the redability of the paragraph below? 
+%% Grammarly suggests that is hard to read, verly long sentences.
+Cloud computing has become the standard way of running
+enterprise applications.  not only because of the convenience of the
 pay-as-you-go model or the non-existent sunk cost of physical facilities, but also because
@@
 built on a single stack of services that communicate by layers, to microservices
-\cite{microservices} and serverless architectures \cite{varghese2018next,Varghese2018849} that favor the parallel and asynchronous communication of
-heterogeneous resources; in the process, a new methodology for designing {\em cloud
-native applications} has been created, with brand new methods and
-techniques \cite{Baldini2016287}. In these architectures, services or even
+\cite{microservices} and serverless architectures \cite{varghese2018next,Varghese2018849} 
+that favor the parallel and asynchronous communication of
+heterogeneous resources; in the process, a new methodology for designing 
+{\em cloud-native applications} has been created, with brand new methods and
+techniques \cite{Baldini2016287}. 
+In these architectures, services or even
 functions are seen as independent processing nodes, departing from the monolithic
@@
 Systems developed with the patterns outlined above have the
-properties of a reactive system \cite{boner2014reactive} and they are commonly
+properties of a reactive system \cite{boner2014reactive}, and they are commonly
 more flexible, fault-tolerant, and scalable. % Some properties and reference - Mario  
@@
 
-    \item To scale the system, when it is possible, there are several options, including the  use of stateless functions for
-    processing.
+  %  \item To scale the system, when it is possible, there are several options, including the  use of stateless functions for
+  %  processing.
 
@@
 the interaction between them, the search area of each population, and the search
-strategy and parametrization of each population. We must take into account
+strategy and parametrization for each population. We must take into account
 these high-level issues when designing a parallel architecture.
 
-Taking these factors into account, in this work, we present EvoSwarm, 
+Considering these factors, in this work, we present EvoSwarm, 
 a container-based application that reactively processes
-isolated and
-heterogeneous populations. The platform uses the {\tt docker-compose} tool for
-defining and running experiments as multi-container applications. EvoSwarm is also released with a free licence in
+isolated and heterogeneous populations. The platform uses the {\tt docker-compose} tool for
+defining and running experiments as multi-container applications. EvoSwarm is also released with a free license
+and available in a public 
+GitHub repository
 \footnote{\url{https://github.com/mariosky/EvoSwarm}}.
-We preferred the use of a container-based application for this prototype because
-it is more suitable for development and replication without the need for
+We preferred the use of a container-based implementation for this prototype because
+it is more suitable for development and replication without needing
 additional components. Nevertheless, the application is compatible with other
@@
 to a Docker Swarm or a Kubernetes system.
-
-Docker container images host the stateless functions responsible for 
-running a search strategy on populations they receive as messages. 
-A single container can pull many messages on its lifetime, processing one at a time. 
-% clarify "on several populations" Sequentially? At the same time? - JJ 
-% done - M
-These functions could also be ported
-directly to a Function as a Service (FaaS) \cite{Roberts2016} implementation.
-% Need to improve the next paragraph or maybe move it to another place - Mario
-Researchers then define the configuration of services used by
-a multi-population algorithm using a YAML file. Any researcher can deploy and
-start all the services from this configuration file with a single command. Once
-the services are up, multiple instances of an experiment can be sent to a queue
-for their execution. The platform includes containerized services that are
-compatible with other container orchestration technologies like Kubernetes and
-Docker Swarm. The services included are a Redis Server used for the Message
-Queue implementation and logging and an Experiment Controller implemented in
-Python. Developers of population processing functions have the freedom to
-implement new population-based algorithms or operators in the language they
-choose inside a container, as they only need to subscribe to a channel of the
-Message Queue and have the ability to read the population and parameters as JSON  
-file.
+%% Here we could talk about Cloud-based platform for executing the application
+%% and the cloud-lock risk - Mario
+
+%% This is explained later in detail, we do not needed here.
+%% JJ Please check this - Mario
+
+% Docker container images host the stateless functions responsible for 
+% running a search strategy on populations they receive as messages. 
+% A single container can pull many messages in its lifetime, processing one at a time. 
+% % clarify "on several populations" Sequentially? At the same time? - JJ 
+% % done - M
+% These functions could also be ported
+% directly to a Function as a Service (FaaS) \cite{Roberts2016} implementation.
+% % Need to improve the next paragraph or maybe move it to another place - Mario
+% Researchers then define the configuration of services used by
+% a multi-population algorithm using a YAML file. Any researcher can deploy and
+% start all the services from this configuration file with a single command. Once
+% the services are up, multiple instances of an experiment can be sent to a queue
+% for their execution. The platform includes containerized services that are
+% compatible with other container orchestration technologies like Kubernetes and
+% Docker Swarm. The services included are a Redis Server used for the Message
+% Queue implementation and logging and an Experiment Controller implemented in
+% Python. Developers of population processing functions have the freedom to
+% implement new population-based algorithms or operators in the language they
+% choose inside a container, as they only need to subscribe to a channel of the
+% Message Queue and have the ability to read the population and parameters as JSON  
+% file.
 
@@
 follows:
-We have proved that cloud-native applications are
-able to easily accommodate multi-paradigm bioinspired algorithms
+We have proved that cloud-native applications can easily accommodate multi-paradigm bioinspired algorithms
 with little efforts of parametrization and
-that a combination of algorithms has better speedup and performance
+that a combination of algorithms has better speed-up and performance
 than any single one taken independently.
@@
     \item Third, the application is compatible with the COCO benchmark 
-    framework, allowing researchers to compare the performance of their 
-    algorithms with other works. % Not really a contribution, but an
+    framework, allowing researchers to compare their 
+    algorithms' performance against other works. % Not really a contribution, but an
                                 % implementation detail. Convenient,
@@
                                 % the art - JJ
-    \item Fourth, we present an empirical study to validate the 
-    applicability of our application,  measuring the execution time and  
-    speedup.
+    \item Fourth, we present an empirical study to validate 
+    our application's applicability, measuring the execution time and speed-up.
 
@@
     multi-populations, using Genetic Algorithms (GAs) and Particle 
-    Swarm Optimization (PSO) in a benchmark for the optimization of 
-    continuous functions. % "Validate the applicability" falls short
+    Swarm Optimization (PSO) using a benchmark for 
+    continuous function optimization. % "Validate the applicability" falls short
                           % of being also a contribution. It must
@@
 
-The organization of the paper is as follows: First, in Section \ref{multi}, we present a
+We organized the paper as follows: First, in Section \ref{multi}, we present a
 background of the fundamental issues of integrating nature-inspired optimization
@@
 of the method,  with respect to scalability and % This is a subsection, fix it. 
-the performance of heterogeneous populations.  % No references to sections 9 and 10
+the heterogeneous populations' performance.  % No references to sections 9 and 10
 Finally, we offer the conclusions of this paper and suggestions on future work in Section \ref{conclusions}.
@@
 services, and also automation of the whole workflow. This research
-bumped into some problems, since speedup is limited, mainly due to the
+bumped into some problems, since speed-up is limited, mainly due to the
 non-automatic scaling of services and also the presence of a single
@@
 In the next section, we describe the experimental setup, and a discussion of the results
-follows, with respect to the speedup in terms of evaluations per second (Section~\ref{sec:speedup}), and 
+follows, with respect to the speed-up in terms of evaluations per second (Section~\ref{sec:speedup}), and 
 execution time (Section~\ref{sec:exec-time}).
@@
 
-\subsubsection{Results Regarding Speedup}
+\subsubsection{Results Regarding Speed-up}
 \label{sec:speedup}
@@
 % Answered - Mario
-To measure the speedup obtained by the addition of workers, we will use the
+To measure the speed-up obtained by the addition of workers, we will use the
 number of evaluations per second, instead of the time required to finish an
@@
 \begin{table}[h!tbp]
-  \caption{Speedup by worker and dimension, taking one worker as the
-    baseline. Speedup is obtained by dividing the evaluation ratio of the baseline by the 
+  \caption{Speed-up by worker and dimension, taking one worker as the
+    baseline. Speed-up is obtained by dividing the evaluation ratio of the baseline by the 
     ratio of the increased number of workers.}
@@
 %
-As expected, within the same number of populations, there is a linear speedup
-when we add more workers. The speedup is marginal in those cases where there 
+As expected, within the same number of populations, there is a linear speed-up
+when we add more workers. The speed-up is marginal in those cases where there 
 are fewer populations than workers.  The reason for this is that at any given time, 
@@
   \small
-  \caption{Speedup achieved with increasing the number of workers for
+  \caption{Speed-up achieved with increasing the number of workers for
     each dimension showing the p-value for the Wilcoxon test. 
@@
   \begin{tabular}{llllll}
-  Populations & Dimension  & \multicolumn{2}{l}{Increment}  & Speedup            & p-value\\
+  Populations & Dimension  & \multicolumn{2}{l}{Increment}  & Speed-up            & p-value\\
   5  & 10 & 1w & 2w  & 2.15 & 1.51E-11 \\
@@
 
-Table~\ref{tab:speedup-table} shows the speedup obtained against the baseline of
-one worker. We achieved the maximum speedup when using four workers or more with
-twenty populations. On the other hand, the maximum speedup when we consider the
-cost it was with eight workers. The best speedups for each number of workers are
+Table~\ref{tab:speedup-table} shows the speed-up obtained against the baseline of
+one worker. We achieved the maximum speed-up when using four workers or more with
+twenty populations. On the other hand, the maximum speed-up when we consider the
+cost it was with eight workers. The best speed-ups for each number of workers are
 shown in bold in the table.
 
-Table~\ref{tab:speedup:test} shows the speedup obtained taking as the baseline
+Table~\ref{tab:speedup:test} shows the speed-up obtained taking as the baseline
 the previous configuration, for example, when increasing from four to eight
@@
 basic rule doubled the speed or where near it. Using five populations, we
-obtained the best and worst speedup depending on the number of workers (shown in
+obtained the best and worst speed-up depending on the number of workers (shown in
 underline and bold, respectively).
@@
 show. If we increase the evaluations per second and, at the same time, decrease
-the number of evaluations needed, we can have a superlinear speedup of the
+the number of evaluations needed, we can have a superlinear speed-up of the
 algorithm, if we measure the time required to complete an instance.
@@
 In Table~\ref{tab:time} we can see that in some cases we have a supralinear
-speedup, we highlighted these in boldface and indicated the speedup value in
+speed-up, we highlighted these in boldface and indicated the speed-up value in
 parenthesis. The best overall times are underlined. As expected, increasing the

### 2020-10-14 draft.tex
Update draft.tex

@@
 application in Section \ref{docker} and the workflow for reproducible experiment design in Section \ref{experiment_flow}. % No reference to section 6.
-Section \ref{setup} describes the design of the empirical evaluation to assess the effectiveness
-of the method,  with respect to scalability and % This is a subsection, fix it. 
+Section \ref{setup} describes the empirical evaluation design to assess the effectiveness
+of the method, concerning scalability and % This is a subsection, fix it. 
 the heterogeneous populations' performance.  % No references to sections 9 and 10
-Finally, we offer the conclusions of this paper and suggestions on future work in Section \ref{conclusions}.
+Finally, we offer the conclusions of this paper and suggestions for future work in Section \ref{conclusions}.
 

### 2020-10-15 draft.tex
Chapter 2
Related to #27 and #21 I commented some  sentences that were out of place.

@@
 which led to early proofs of speed-up \cite{gorges1990explicit,grosso1985computer}. 
-The concept was to divide a large population into communicated subpopulations. 
+% Repetitive - M
+%The concept was to divide a large population into communicated subpopulations. 
 Since then, other population-based algorithms have adopted the idea. 
@@
 incorporated as soon as they are received. Following the previous example, now the master operation can
-continue to the next iteration, even if a worker has not finished its work.  We can find many asynchronous algorithms
-\cite{coleman89,baugh2003asynchronous}, reporting benefits in execution time
+continue to the next iteration, even if a worker has not yet finished its work. 
+We can find many asynchronous algorithms
+\cite{coleman89,baugh2003asynchronous} reporting benefits in execution time
 and scalability. In the particular case of asynchronous and cloud-native
@@
 %% Grammarly suggests that is hard to read, verly long sentences.
+%% Edited a little - Mario
 Cloud computing has become the standard way of running
-enterprise applications.  not only because of the convenience of the
+enterprise applications.  Not only because of the convenience of the
 pay-as-you-go model or the non-existent sunk cost of physical facilities, but also because
@@
 is much easier to reproduce results and this has been a boon for scientific
-computing.  However,  cloud computing has also been evolving, going from simply
-putting on virtual machines elsewhere than on premises old-style monolithic applications, that is, applications
-built on a single stack of services that communicate by layers, to microservices
-\cite{microservices} and serverless architectures \cite{varghese2018next,Varghese2018849} 
+computing.  
+However,  cloud computing has also been evolving, going from simply
+putting old-style monolithic applications on virtual machines executing on an external 
+data center to microservices \cite{microservices} and serverless architectures \cite{varghese2018next,Varghese2018849} 
 that favor the parallel and asynchronous communication of
-heterogeneous resources; in the process, a new methodology for designing 
+heterogeneous resources. In the process, a new methodology for designing 
 {\em cloud-native applications} has been created, with brand new methods and
 techniques \cite{Baldini2016287}. 
-In these architectures, services or even
-functions are seen as independent processing nodes, departing from the monolithic
+In these architectures, services or even functions are seen as independent 
+processing nodes, departing from the monolithic
 and even distributed paradigms, and becoming a loosely coupled collection of {\em
-stateless functions} \cite{malawski2017serverless}, that react to events and
-only come into existence while they are doing some processing.
+stateless functions} \cite{malawski2017serverless}, that react to events. 
+%and only come into existence while they are doing some processing.
 Systems developed with the patterns outlined above have the
-properties of a reactive system \cite{boner2014reactive}, and they are commonly
+properties of a reactive system \cite{boner2014reactive}, and they are generally 
 more flexible, fault-tolerant, and scalable. % Some properties and reference - Mario  
@@
 
-Researchers  and algorithm designers willing to adapt to the cloud; however, their
-multi-population solutions face additional challenges:
+Researchers and algorithm designers are willing to adapt to the cloud; 
+however, when developing their multi-population algorithms, they face additional challenges:
+
 \begin{itemize}
@@
 Beyond the choice of design methodology and implementation, researchers need to
-consider additional issues \cite{Ma2019} when designing efficient multi-population algorithms. These include the number and size of populations,
+consider additional issues \cite{Ma2019} when designing efficient multi-population algorithms. 
+These include the number and size of populations,
 the interaction between them, the search area of each population, and the search
@@
 Considering these factors, in this work, we present EvoSwarm, 
-a container-based application that reactively processes
-isolated and heterogeneous populations. The platform uses the {\tt docker-compose} tool for
-defining and running experiments as multi-container applications. EvoSwarm is also released with a free license
+a cloud-native, container-based application that reactively processes
+isolated and heterogeneous populations. 
+% The platform uses the {\tt docker-compose} tool for
+% defining and running experiments as multi-container applications. 
+% docker-compose is no longer the only tool used for deployment
+EvoSwarm is released with a free license
 and available in a public 
@@
 \footnote{\url{https://github.com/mariosky/EvoSwarm}}.
+
 We preferred the use of a container-based implementation for this prototype because
@@
                      % using here - JJ
-This relative isolation helps in maintaining an overall
+This relative isolation maintains an overall
 diversity since each population will search in a particular area, at least
@@
 However, in most cases, algorithms applied to each population are 
-homogeneous, or at any rate, the same variant of the algorithm. As long as this
-parallel operation is not synchronous, other population-based algorithms, or, as
-a matter of fact, any algorithm, could be easily integrated. That is why several
+homogeneous; but as long as this parallel operation is event-based, 
+other population-based algorithms could be easily integrated.  That is why several
 works based on multi-population are heterogeneous, integrating various
@@
 
-Heterogeneous algorithms add another degree of freedom to the problem of finding
-the correct parameter settings for an algorithm; because some parameters affect
+Heterogeneous algorithms add another degree of freedom to the parameter
+tuning problem; because some parameters affect 
 the accuracy of the solution and the convergence speed of the algorithms as they
@@
 other hand, current studies show that by having a high number of populations 
-interacting in parallel, the effect of the individual parameters of each
-population is compensated by those selected in other
-populations \cite{li2016multi,tanabe2013evaluation}. % reference - JJ
+interacting in parallel, the effect of each population's parameters is compensated 
+by those selected in other populations \cite{li2016multi,tanabe2013evaluation}. % reference - JJ
 
 %%%%% ----------------- This gets lost in this section -----------------------------
-In this work, we will use random settings within a specific range as results have shown
-this is a valid solution to this problem \cite{garcia2014randomized}. % This should have been
+%In this work, we will use random settings within a specific range as results have shown
+%this is a valid solution to this problem \cite{garcia2014randomized}. % This should have been
                                 % mentioned in the introduction - JJ
@@
 %%%%%% to method - JJ
-Some parameters, especially the population size, are
-kept fixed in order to control the execution of the algorithm more easily. For
-instance, by having the size of populations fixed, it is easier to control
-the number of evaluations and the communication costs, when the algorithm is in
-operation. % And to keep them more or less synchronized, which can be
+%Some parameters, especially the population size, are
+%kept fixed in order to control the execution of the algorithm more easily. For
+%instance, by having the size of populations fixed, it is easier to control
+%the number of evaluations and the communication costs, when the algorithm is in
+%operation. % And to keep them more or less synchronized, which can be
            % interesting, too... But this should go to the
@@
 %%%% ---------------------- Also to the introduction
-Combining multiple algorithms, with different parameters, interacting with each
+Combining multiple algorithms with different parameters, interacting with each
 other at the same time, can benefit from the strengths of each. For instance, a

### 2020-10-15 draft.tex
Update SOA

@@
 
-Although at first sight containers look like simply a method that
-allows easy shipping of applications, they are much more than
+Although containers look like merely a method for easy shipping 
+of applications, at first sight, they are much more than
 that. One of the first features that make them interesting is their
 ephemerality: fast startup and teardown time make them usable for
-machine-independent and cloud deployment of simple, ephemeral and
-maybe stateless functions. On the other hand, containerization also
-allows, through the use of orchestration engines like Docker Swarm or
-Kubernetes, or through simple, non-self-scaling specifications using
-{\tt docker-compose}, a definition of all the relationships between
-different parts of the application, or services. The full realization
+machine-independent and cloud deployment of simple, ephemeral, and
+maybe stateless functions. 
+%On the other hand, containerization also
+%allows, through the use of orchestration engines like Docker Swarm or
+%Kubernetes, or through simple, non-self-scaling specifications using
+%{\tt docker-compose}, a definition of all the relationships between
+%different parts of the application, or services. 
+% We said this earlier - Mario
+The full realization
 of Docker-based architectures has led to the creation of a series of
-patterns that, together, conform cloud-native application development
+patterns that together define the cloud-native application development
 \cite{kratzke2017understanding}.
 
-However, in the field of evolutionary algorithms, this has been a slow
-realization, from the origins where Salza and Ferrucci
-\cite{salza2016approach,salza2016develop} proposed an architecture that carried
-evolutionary algorithms to the cloud, and which were developed further in their
-next paper \cite{de2017parallel}. These papers started to introduce cloud native
+However, this has been a slow realization in the evolutionary algorithms 
+field, from the origins where Salza and Ferrucci \cite{salza2016approach,salza2016develop} 
+proposed an architecture that carried evolutionary algorithms to the cloud, 
+and which were developed further in their next paper \cite{de2017parallel}.
+These papers started to introduce cloud-native
 aspects such as the use of messaging queues and CoreOS as an operating system
 designed from the ground up for container utilization in the area of
-evolutionary algorithms. The way these containers are managed, however, is
-rather classical in nature from the point of view of distributed EC: a
+evolutionary algorithms. However, the way these containers are managed, is
+rather classical from the distributed EC point of view: a
 master-worker approach, communicated using RabbitMQ (a messaging queue), where
 replicated workers perform the tasks in parallel. They called their approach
-AMQPGA, inspired by the protocol, AMQP, used by RabbitMQ, the messaging system.
-Other authors have also followed this classical approach by using other EA
-``classical'' topologies in the cloud: Dziurzanski et al.
+AMQPGA, inspired by the protocol, AMQP, used by RabbitMQ.
+Other authors have also followed this classical approach by using other 
+``classical'' EA topologies in the cloud: Dziurzanski et al.
 \cite{dziurzanski2020scalable} adapts an island model to a container
@@
 
-Later approaches use microservices \cite{khalloof2018generic}, or were
-systematized \cite{salza2019speed} achieving speed-up through the
-creation of an abstraction layer over the evolutionary algorithm
-services, and also automation of the whole workflow. This research
-bumped into some problems, since speed-up is limited, mainly due to the
-non-automatic scaling of services and also the presence of a single
-master which is the one that runs the evolutionary
-algorithms. However, this master-worker architecture can be turned
-around with other kind of evolutionary architectures, such
+Later approaches used microservices \cite{khalloof2018generic} and 
+provided the automation of the whole workflow\cite{salza2019speed}, 
+achieving speed-up by creating an abstraction layer over the 
+evolutionary algorithm services.
+%, and also automation of the whole workflow. 
+This research bumped into some problems, since speed-up is limited, mainly due to the
+non-automatic scaling of services and the presence of a single
+master, which is the one that runs the evolutionary
+algorithm. However, this master-worker architecture can be turned
+around with other kinds of evolutionary architectures, such
 as pool-based systems
-\cite{valenzuela2015implementing,merelo2012sofea,sofea:cec2012} where
+\cite{valenzuela2015implementing,merelo2012sofea,sofea:cec2012} that were 
 a better match to the inherently heterogeneous nature of cloud-based
 architectures. Pool-based systems are closer to serverless systems
-\cite{malawski2017serverless}, and they pull from the pool and run
-whole algorithms, instead of just evaluation. The pool from where they
-pull is the common store of results, and can be accessed in an
-asynchronous way, which suits better the nature of cloud systems.
+\cite{malawski2017serverless}, as they pull populations from the pool and run
+whole algorithms, instead of just doing the evaluation. The pool 
+is a shared data storage for results, and can be accessed in an
+asynchronous way, which suits better the nature of cloud-native systems.
 

### 2020-10-15 draft.tex
Proposed method intro

@@
 whole algorithms, instead of just doing the evaluation. The pool 
-is a shared data storage for results, and can be accessed in an
+is a shared data storage for results and can be accessed in an
 asynchronous way, which suits better the nature of cloud-native systems.
@@
 In this section, we present the design
-and a container-based implementation of the model we propose for the execution
-of multi-population based algorithms. We follow the requirements outlined in
+and a container-based implementation of the model we propose to execute
+multi-population based algorithms. We follow the requirements outlined in
 Section \ref{multi} and the general design principles of cloud-native
-applications we have mentioned earlier; our intention is to steer
+applications we have mentioned earlier; we intend to steer
 clear of a direct translation of ``classical'' GA architectures to the
 cloud and design from the ground up a cloud-native method. As a first step, we describe the general
-architecture, and then  every component in detail.
-
-The most elemental data structure used throughout the system is the population; and at a higher
+architecture and then every component in detail.
+
+The most basic data structure used throughout the system is the population, and at a higher
 level, we can view the system as a continuous stream of populations
@@
 to communicate a population to another, it must pack the population into
-a message, and then send the message 
+a message and then send the message 
 to the queuing system without waiting for a response; this
@@
 without waiting for a result. On the other end, the recipient subscribes to the
-message queue by defining an event handler method that is going to be triggered when a
-message is pushed to it. The two components at each end of the queue are
+message queue by defining an event handler method that will be triggered when a
+message is pushed to it. At each end of the queue, the two components are 
 Producers that push messages to the queue and Consumers that pull them. This
 pattern is an essential component of a reactive architecture and makes
-it highly scalable; one of the reasons for this is the use of functions
-that have no secondary effects or access global storage, that is,
+it highly scalable; one reason for this is the use of functions
+with no secondary effects, that is,
 stateless functions. These functions do not need to read, keep, or modify data outside of the method. If
-Consumers are implemented using stateless functions, and thus, there is no difference
-between having one or many copies of the same function pulling work from the queue all at the
-same time, because no side effects, including unwanted resource
-locking problems,  are going to result from this concurrency.
+Consumers are implemented using stateless functions, there is no difference
+between having one or many copies of the same function pulling work from the queue 
+simultaneously, because there are no side effects, including unwanted resource
+locking problems resulting from this concurrency.
 % This is important and we should probably mention that in the

### 2020-10-15 draft.tex
Update draft.tex

@@
 Figure \ref{fig:kafkEO}. Again, we can explain the model using the analogy
-of producers and consumers of messages. First, we can see that
-there are two queues, one labeled \texttt{Input}, and the other one \texttt{Output}. In the diagram, 
+of producers and consumers of messages. First, we can see two queues, 
+one labeled \texttt{Input} and the other one \texttt{Output}. In the diagram, 
 push operations on a queue are represented by solid arrows connecting to the left side
 of the queue box, and pull or pop operations as solid arrows leaving from the right side.
-Also, the architecture has at least four processes indicated in the diagram as
+The architecture has at least four processes indicated in the diagram as
 swimlanes: First, the \texttt{Setup} process, responsible for the reading a configuration
-file, and creating the initial populations. Second, the \texttt{Controller} process,
-responsible for the migration between populations, and keeping track of the
-iterations of the algorithm. The \texttt{Message Queue} process runs the \texttt{Input} and \texttt{Output}
+file and creating the initial populations. Second, the \texttt{Controller} process,
+responsible for the migration between populations and keeping track of the algorithm's
+iterations. The \texttt{Message Queue} process runs the \texttt{Input} and \texttt{Output}
 queues. Finally, there is at least one \texttt{Stateless Function} process responsible
@@
 The algorithm starts with the \texttt{Setup} process pulling a configuration message from
-the \texttt{Experiments} queue, this is not shown in the diagram. The configuration message 
-includes all the parameters needed to execute the algorithm, among other things,
-the number of populations, the number of individuals, and the number of iterations
+the \texttt{Experiments} queue (not shown in the diagram). The configuration message 
+includes all the parameters needed to execute the algorithm, the number of populations,
+the number of individuals, and the number of iterations 
 of the algorithm. We give more details about the configuration data structure
@@
 
-\item Each population is then pushed to the \texttt{Input} queue, so they can be consumed 
-by Stateless functions responsible for the execution of the search.
+\item Each population is then pushed to the \texttt{Input} queue, so that they can be consumed 
+by stateless functions responsible for the search's execution.
 
 \item One or more \texttt{Stateless functions} are constantly pulling population messages
-from the \texttt{Input} queue. Taking the population data and metadata as parameters, these 
-functions are responsible for the actual execution of the optimization algorithm. 
-They take the current state of the population and run the algorithm for a certain 
-number of iterations.
+from the \texttt{Input} queue. They take the current state of the population and 
+run the algorithm for a certain number of iterations.
 
 \item Once they finish the execution, the population state
-is packed again along with additional metadata about the execution of the algorithm.
+is packed again with additional metadata about the algorithm's execution.
 The resulting populations are now pushed to the Output queue. Once finished, another 
@@
 \item The \texttt{Controller} process is responsible for keeping track of the progress of 
-the search. It pulls current populations from the \texttt{Output} queue, inspects the metadata 
+the search. It pulls current populations from the \texttt{Output} queue, inspects the metadata, 
 and if an optimal solution has been found or the maximum number of iteration has been
-reached it signals the stop of the execution.
+reached, it signals the stop of the execution.
 
 \item Otherwise, it passes the stream of messages 
-to a migration process, where populations are mixed with each other. New populations 
-are generated from this migration, and they are again pushed to the Input queue to continue 
-in a loop. The \texttt{Controller} is also responsible of logging the metadata received with the
+to a migration process, where populations are mixed. New populations 
+are generated from this migration and again pushed to the Input queue to continue 
+in a loop. The \texttt{Controller} is also responsible for logging the metadata received with the
 messages.  

### 2020-10-15 draft.tex
Update draft.tex

@@
 \item 
-An important aspect of proposal is the decoupling of the population and 
+An important aspect of the proposal is the decoupling of the population and 
 the population-based algorithm. It is common that in a classic island-based algorithm
-each island is executed in a separate processing node, i.e. in a virtual machine, CPU,
-or thread. In this case we can have just one processing node, or many nodes, each 
-running a different search strategy, or using its own parameters. 
+each island is executed in a separate processing node, i.e., in a virtual machine, CPU,
+or thread. In this case, we can have just one processing node or many nodes, each 
+running with a different search strategy or parameters. 
 \item 
@@
 Another advantage is that algorithm designers have many options for implementing this simple architecture. 
-The same basic components can be implemented as a single multi-threaded program, or 
+The same components can be implemented as a single multi-threaded program or 
 as a highly scalable serverless cloud application. Most modern languages include 
-constructs for asynchronous programming using queues or channels, for a multi-threaded
+constructs for asynchronous programming using queues or channels for multi-threaded
 execution.
@@
 
-On the other hand, there are several caveats as well: It is more
+On the other hand, there are several caveats: It is more
 costly to move entire populations as messages than only passing
-certain individuals from a process to another. Designers must consider
-this cost when working with large individuals, and if possible send
-populations using several messages, use compression or adjust the size
+specific individuals from a process to another. Designers must consider
+this cost when working with large individuals, and if possible, send
+populations using several messages, use compression, or adjust the size
 of populations. Pool-based algorithms also suffer from this drawback,
-but web based implementations have been working with continuous
-optimization problems with 1000 dimensions.  However, in other cases
+but web-based implementations have been working with continuous
+optimization problems with 1000 dimensions.  However, in other cases, % Reference that paper - Mario
 this is not a viable approach.
@@
 comparison on the coupling and communication between the main components:
-populations, processing nodes, and algorithms.  In Figure \ref{fig:population_message} we show the
+populations, processing nodes, and algorithms.  In Figure \ref{fig:population_message}, we show the
 main components of a message-driven architecture. We can see that the algorithm
@@
 only pass specific individuals between them.  In this model, adding additional
-processing nodes, on-runtime, can be more difficult, because they need to have
+processing nodes, on-runtime, can be more difficult because they need to have
 an addressing mechanism or simply be aware of these new nodes. This is

### 2020-10-15 draft.tex
Section 6

@@
 
-A typical design pattern used to alleviate the above drawbacks is the use of a
-central repository or pool of individuals that is available to all processing
+A typical design pattern used to alleviate the above drawbacks is using a
+central repository or pool of individuals available to all processing
 nodes.  In Figure \ref{fig:pool_island} we show the main components of a pool-based
-multi-population algorithm. Although algorithms and population state remains
+multi-population algorithm. Although algorithms and population state remain
 coupled, now, processing nodes do not communicate directly with each other.
@@
 found is that when a processing node returns a population to the pool, the
-state of the population is lost. But having the algorithm decoupled opens the
+population's state is lost. Nevertheless, having the algorithm decoupled opens the
 possibility of implementing the system using serverless functions.
@@
 
-There are many advantages of container-based architectures over one
-based on virtual machines. The main ones are architectural: it makes
-the design of an application gravitate towards a set of decoupled
-tasks, that use reactive, event based programming, to interact with
-the rest of the application; containers hold micro-services that are
-much easier to test, integrate and deploy. It also simplifies
-the interaction APIs. Of course, that makes them more economical in a
-pay-per-use cloud provider, but that is not the main reason why we
-have chosen them in this paper; reproducibility of experiments and a
-clear and idempotent description of the architecture through {\tt docker-compose} 
-are the main reasons.
-
-In this section, we propose a design of a reactive container-based application
-for executing multi-population based optimization experiments.  For the design,
-we followed the design patterns highlighted in the previous section, and again
+% There are many advantages of container-based architectures over one
+% based on virtual machines. The main ones are architectural: it makes
+% the design of an application gravitate towards a set of decoupled
+% tasks, that use reactive, event based programming, to interact with
+% the rest of the application; containers hold micro-services that are
+% much easier to test, integrate and deploy. It also simplifies
+% the interaction APIs. Of course, that makes them more economical in a
+% pay-per-use cloud provider, but that is not the main reason why we
+% have chosen them in this paper; reproducibility of experiments and a
+% clear and idempotent description of the architecture through {\tt docker-compose} 
+% are the main reasons.
+
+% Was mentioned before - M
+
+This section proposes a design of a reactive container-based application 
+for executing multi-population-based optimization experiments. We followed the
+cloud-native design patterns highlighted in previous sections, and again
 we can go back to Figure \ref{fig:kafkEO} and use it as a guide for explaining
-the main components and their interactions. The containers used in this design will be explained next.
+the main components and their interactions. 
+The containers used in this design will be explained next.
 
@@
 \cite{de2017parallel} used queues for communication, although in that
-case RabbitMQ was the chosen tool. Redis is faster, and also can act
-as a data store from where we can control the state of the application.
+case, RabbitMQ was the chosen tool. Redis is faster and can act as a 
+data store from where we can control the application's state. 
 Each queue is a Redis List object, and we use
-the \texttt{LPOP} (left pop), and \texttt{RPOP} (right pop) commands for a queue like behavior.
+the \texttt{LPOP} (left pop) and \texttt{RPOP} (right pop) commands for a queue like behavior.
 In those cases where we needed a blocking behavior, i.e., when a process needs
 to wait until a message is available, we used the blocking versions \texttt{BLPOP} and
-\texttt{BRPOP}. Redis in-memory operation are very fast, and all of the above 
+\texttt{BRPOP}. Redis in-memory operations are very fast, and all of the above 
 operations have a time complexity of $\mathcal{O}(1)$. We use the official 
@@
 The open-source library is called Reactive Extensions (ReactiveX) for Python 
-\footnote{\url{https://github.com/ReactiveX/RxPY}} an it is based on the Observer 
+\footnote{\url{https://github.com/ReactiveX/RxPY}} and it is based on the Observer 
 pattern \cite{gamma1995design} and functional programming.
@@
 emitted by the \texttt{Observable}. Reactive operators are represented as text boxes,
-showing the transformation to be applied. Normally, a transformation results in
-another \texttt{Observable}, again emitting new results.  For asynchronous programming, 
-this approach is more elegant than nested callbacks, that are more difficult 
-to code and debug.
+showing the transformation to be applied. Typically, a transformation results in
+another \texttt{Observable}, again emitting new results. This approach is more 
+elegant for asynchronous programming than nested callbacks, which are more 
+difficult to code and debug.
 
@@
 \texttt{Observable}. In this particular period, we can see that the stream receives 
-two populations from another experiment, and are shown as two red marbles.  
+two populations from another experiment and are shown as two red marbles.  
 The filter operator removes all messages that belong to a different experiment; 
@@
 end the loop. In this example, \texttt{max\_iterations = 6},  the take operator 
-assures that only 6 messages are received. After the \texttt{Observable} emits 
+assures that only six messages are received. After the \texttt{Observable} emits 
 the sixth message, it triggers the completed event.
@@
 \item Many observers can be subscribed to the \texttt{valid\_messages} \texttt{Observable} because it
-emits all the valid items. At the moment, there are to additional methods
+emits all the valid items. At the moment, there are two additional methods
 subscribed that we are not showing, one for logging and the other for monitoring
@@
 
-\item The   \texttt{buffer\_with\_count(3)} operator, waits until it receives 
+\item The \texttt{buffer\_with\_count(3)} operator waits until it receives 
 three valid messages to emit a single message that contains a list with
@@
 In the next section, we follow the algorithm flow and describe the 
-implementation details of the \texttt{worker} containers.
+\texttt{worker} containers' implementation details.
 
@@
 worker object responsible for initializing and running the specified stateless
-function (i.e., GA or PSO) with the population and configuration in the message.
+function (i.e., GA or PSO) with the population and configuration inside the message.
 Once the function returns the evolved population, the
 the main script pushes the results to the \texttt{Output} queue, and continues
-the execution to the infinite loop. This process is shown in Figure
-\ref{fig:worker}.
+to execute the infinite loop. This process is shown in Figure \ref{fig:worker}.
 
@@
 perspective of the two roles users can play. On one side is the user role that
-designs and runs experiments and, on the other, the role of a developer of
+designs and runs experiments and, on the other, a developer of
 algorithms. The scenario presented in Figure~\ref{fig:experiment_flow} shows the
-workflow of an experiment in Evo\-Swarm the two roles users can play. The first
+workflow of an experiment in EvoSwarm and the two roles users can play. The first
 step is to develop the search strategy algorithms that are going to be used by
@@
 experiments, monitor the execution, and analyze the results. Depending on the
-type of experiment executed, users could have additional needs, in this case
+type of experiment executed, users could have additional needs. In this case,
 there are additional scripts for processing the results of experiments to be
 used with the COCO framework. In the next sections, we give more details about
-these roles and their workflows, and at the end we present a proof-of-concept example of a
+these roles and their workflows, followed by a proof-of-concept example of a
 reproducible experiment in EvoSwarm.
@@
 
-Developers can add new algorithm as stateless functions in
+Developers can add new algorithms as stateless functions in
 several ways and with two levels of integration. In the first level of
@@
 configuration created in the setup, and it only needs to take messages from the
-input queue and return the modified population to the output queue. On the
+input queue and return the modified population to the output queue. At the
 second level, the algorithm receives a configuration from the initial setup,
@@
 file received by workers in Listing \ref{code:config}  in which we abbreviated
-the population key (the ellipsis is used instead). The message includes the current state
-of the population, together with data about the problem, algorithm,
-and its parameters.
+the population key (the ellipsis is used instead). The message includes the 
+current state of the population and data about the problem, algorithm, and parameters.
 
@@
 
-In the case of the first level of integration, developers
+In the first level of integration, developers
 can implement a new stateless Docker container following
 the description in the previous section (\ref{workers}). The worker
-container executes a daemon that continuously pulls messages from the input queue
-on the Redis host specified as an environment variable {\texttt REDIS\_HOST}. If the new
+container executes a daemon that continuously pulls messages from the Redis host's input queue
+specified as an environment variable {\texttt REDIS\_HOST}. If the new
 population-based algorithm is written in Python, developers can base the new
@@
 The worker must include a constructor with the configuration dictionary 
-(see Listing \ref{code:config}) as the only parameter, and call a the new function 
+(see Listing \ref{code:config}) as the only parameter and call a the new function 
 with the required parameters, unpacking them from the configuration dictionary if necessary.
-All the code of the EvoSwarm Application is published with the open source MIT license.
+All the code of the EvoSwarm Application is published with an open-source MIT license.
 
 If developers use an existing library for population-based algorithms, they need
-to change the standard model of execution, that creates first a random
+to change the standard model of execution, which creates a random
 population because, in this case, the algorithm needs to start with the
@@
 JSON message and replace the population object created by the default library implementation. For
-example, the DEAP library we used for the GA algorithm, creates the type of the
+example, the DEAP library we used for the GA algorithm creates the type of the
 population at runtime (i.e., bit, float), so we have to replace the population
@@
 iteration, so we just read it and put it in the format required.
-Again, the state of the population must be encoded as a JSON message to be
+Again, the population's state must be encoded as a JSON message to be
 returned as a result, with additional data as
-depicted in Listing \ref{code:result}. The output message contains the id of the
-worker container together with records giving information about each iteration
-of the algorithm, including the best solution found. The controller uses this
-information to track the execution of the algorithm and keep a log in the Redis
-container for posterior analyses.
+depicted in Listing \ref{code:result}. The output message contains worker container's id 
+together with records giving information about each iteration
+of the algorithm, including the best solution found. The controller uses 
+this information to track the algorithm's execution and keep a 
+log in the Redis container for posterior analyses.
 
@@
 container to specify additional configuration options. For instance,
-in the current configuration message (see Listing \ref{code:config})
-workers executing a GA, initialize the mutation and crossover 
+in the current configuration message (see Listing \ref{code:config}),
+workers executing a GA initialize the mutation and crossover 
 probabilities randomly from a specific range of values. Users specify
@@
 Migration between populations is independent of the type of algorithm, as the
-controller treats messages as the same type of objects. But developers could
-also change the controller to specify new rules of operation, for instance, to
+controller treats messages as the same type of objects. However, developers could
+also change the controller to specify new operation rules, for instance, to
 change the type of migration depending on the algorithm. 
-For example, in the case of swarm-based algorithms,  migration between  
+For example, in swarm-based algorithms,  migration between  
 populations (swarms) can follow a topology, indicating valid connections 
@@
 can apply these rules selectively only to swarm-based algorithms. 
-In general, these configuration options will be algorithm and library specific.
+In general, these configuration options will be algorithm and library-specific.
 
@@
 Moreover, having more algorithms available can benefit researchers trying to
-compare against other algorithms or increase the diversity of their algorithms
-by adding more search methods. Finally, in some cloud services, having hosted
+compare against other algorithms or increase their algorithms' diversity by 
+adding more search methods. Finally, in some cloud services, having hosted
 images is a requirement, as it is the case of Amazon ECS.
 
-Moreover, public repositories have the functionality for documenting, sharing,
-and tracking different versions of both images and image definitions by using
-tags. This infrastructure supports the objective of the paper, which is to
+Moreover, public repositories can document, share, 
+and track different versions of both images and image definitions by using
+tags. This infrastructure supports this paper's objective, which is to
 contribute to the reproducibility of experiments and platform independence.
@@
 docker-compose.yml} file to specify the type and number of containers required
-for the experiments. Then, the {\tt docker-compose} application is responsible for
-starting all the containers, and downloading new versions if it is necessary,
+for the experiments. The {\tt docker-compose} application is then responsible for
+starting all the containers and downloading new versions if necessary
 by executing the command {\tt docker-compose up}  inside the EvoSwarm directory.
@@
 After starting the application, users can push several Experiment Definition 
-files at a time and start monitoring the current execution in the terminal.
+files and start monitoring the current execution in the terminal.
 After an experiment is finished, the user must execute another script
@@
 scripts are included in the repository to plot the running times for the
-experiments, and all other plots used in this paper;  there is also a script to
+experiments and all other plots used in this paper;  there is also a script to
 generate the files needed by the COCO framework \cite{hansen2016coco}, which can generate standard

### 2020-10-15 draft.tex
Update draft.tex

@@
 not need to be a cloud provider; it could be a developer's laptop, a single
-server, or a production environment. In this section, we give specific details
+server, or a production environment. This section gives specific details
 about how we deployed EvoSwarm in a cloud-setting for running the experiments
@@
   \item[A Kubernetes cluster] provides auto-scaling and management of
-  containerized applications, and it supports other container technologies 
+  containerized applications, and it supports other container technologies, 
   but it needs additional configuration for provisioning, 
@@
     services are Amazon Elastic Container Service,  Red Hat Openshift
-    (which is also open source), and Marantis Docker Enterprise using Docker Swarm.
+    (which is also open-source), and Marantis Docker Enterprise using Docker Swarm.
     Salza et al. \cite{salza2019speed} tested a variant of this option by using a CoreOs cluster.
@@
 
-We chose the second option, that is, a container orchestration service, and
-deployed the EvoSwarm application in Amazon Elastic Container Service (ECS),
+We chose the second option, a container orchestration service. We
+deployed the EvoSwarm application in Amazon Elastic Container Service (ECS)
 since it needs fewer changes in configuration (from deploying it locally using
-{\tt docker-compose}) and it is has a lower cost than the Kubernetes
+{\tt docker-compose}) and it has a lower cost than the Kubernetes
 alternative.
@@
 The main attributes are the {\texttt cpu\_shares} indicating the CPU units
-assigned, in this case $2048$, and the network configuration; in this case we have two private 
+assigned, in this case, $2048$, and the network configuration; in this case, we have two private 
 subnets and a security group in this configuration. When deploying to EC2 containers, these requirements
 need to be supported by the currently available cluster. For instance, we need to have all 
-the CPU units required by all tasks, if they are not available, the task initiation will fail.
+the CPU units required by all tasks because if they are not available, the task initiation will fail.
 
 The complete task configuration is generated by the {\texttt ecs compose command} that uses a {\texttt docker-compose.yml} file to generate 
-the configuration on-the-fly. For example to deploy the controller, setup and Redis 
+the configuration on-the-fly. For example, to deploy the controller, setup, and Redis 
 containers we used the file in Listing ~\ref{code:compose}.
@@
 \footnote{\url{https://hub.docker.com/u/mariosky}}, open {\texttt ports}, a {\texttt command} to be executed at
-startup, configuration for the AWS {\texttt logging} service, and finally {\texttt environment}
+startup, configuration for the AWS {\texttt logging} service, and finally, {\texttt environment}
 variables used by code running inside containers.

### 2020-10-16 draft.tex
Proofreading abstract refs #27

@@
 cloud-native computing.  However, the design of these cloud-native,
-distributed, multi-population algorithms is not a trivial task. Coming
-from monolithic (single-instance) solutions, adaptations at several
+distributed, multi-population algorithms is not a trivial task. Using
+as a foundation monolithic (single-instance) solutions, adaptations at several
 levels, from the algorithmic to the functional, must be made to
@@
 same time, conserving the intended functionality. Instead of an
-evolutive approach, in this paper we propose to create from scratch a
-cloud-native optimization framework that is able to include multiple
-algorithms and use few parameters to work. This solution goes beyond
-current state of the art via a framework that is able to support
+evolutive approach, in this paper we propose a
+cloud-native optimization framework created from scratch, that is able to include multiple
+(population-based) algorithms without increasing the number of
+parameters that need tuning. This solution goes beyond
+current state of the art, since is able to support
 different algorithms at the same time, work asynchronously, and also
@@
 algorithms perform better in their ``natural'' environment than other
-kinds of algorithms, % Did we compare it to other kind of algorithms?
-                     % If we didn't, this should go - JJ 
-                     % Compared against GA and PSO and others from BBOB - Mario
-                     % on the algorithm, not speedup 
-                     % I like the sentence, If we did not prove it, 
-                     % at least we build a tool that will allow
-                     % algorithms to run in their natural environment.
-                     % The focus of the Journal after all is in computer systems - M
-
-and setting a new baseline for scaling and
-performance in the cloud.
+kinds of algorithms, and setting a new baseline for scaling and
+performance of this kind of algorithms in the cloud.
 \end{abstract}

### 2020-10-16 draft.tex
Minor changes

@@
 applied to solve many complex real-world problems
-\cite{yang2014nature}; these algorithms can be broadly grouped into evolutionary algorithms (EAs)
+\cite{yang2014nature}. These algorithms can be broadly grouped into evolutionary algorithms (EAs)
 \cite{back1996evolutionary} and swarm intelligence (SI)
@@
 
-Besides being all population-based, they also share the common characteristic of 
+Besides being, all of them, population-based, they also share the common characteristic of
 creating an initial set of random candidate solutions later used by a 

### 2020-10-16 draft.tex
Grammar correction according to GDrive write refs #27

@@
 
-The trend concerning the parallel execution multi-population algorithms goes from
+The trend concerning parallel execution multi-population algorithms goes from
 earlier hardware-based implementations using transputers
@@
 and even distributed paradigms, and becoming a loosely coupled collection of {\em
-stateless functions} \cite{malawski2017serverless}, that react to events. 
+stateless functions}\cite{malawski2017serverless}, that react to events. 
 %and only come into existence while they are doing some processing.
@@
 The architecture has at least four processes indicated in the diagram as
-swimlanes: First, the \texttt{Setup} process, responsible for the reading a configuration
+swimlanes: First, the \texttt{Setup} process, responsible for reading a configuration
 file and creating the initial populations. Second, the \texttt{Controller} process,
@@
 function (i.e., GA or PSO) with the population and configuration inside the message.
-Once the function returns the evolved population, the
+Once the function returns the evolved population,
 the main script pushes the results to the \texttt{Output} queue, and continues
@@
 The worker must include a constructor with the configuration dictionary 
-(see Listing \ref{code:config}) as the only parameter and call a the new function 
+(see Listing \ref{code:config}) as the only parameter and call the new function 
 with the required parameters, unpacking them from the configuration dictionary if necessary.
@@
 population at runtime (i.e., bit, float), so we have to replace the population
-once DEAP created it. In the case of EvoloPy, the PSO method directly uses a
+once DEAP creates it. In the case of EvoloPy, the PSO method directly uses a
 list of floats for storing the swarm, so we just added a parameter to receive
@@
 returned as a result, with additional data as
-depicted in Listing \ref{code:result}. The output message contains worker container's id 
+depicted in Listing \ref{code:result}. The output message contains the
+worker container's id 
 together with records giving information about each iteration
@@
 
-  \item[A container orchestration service] these services are compatible with {\tt docker-compose}
+  \item[A container orchestration service]: These services are compatible with {\tt docker-compose}
     files for configuration and provide transparent provisioning. Examples of these
@@
 
-Because all these instances run in a private subnet, we also
-run a Bastion instance, that is, a specially protected network node, in a public subnet, from where we send the
-commands to execute the experiments
-and later collected the results. We connected to this instance through a {\texttt ssh} session from 
-our personal workstation. This is a {\texttt R5AD Large} instance, selected with 75GiB of storage, needed for
-keeping the logs, and more memory (16GiB) for processing the logs in memory. The details these EC2 instances 
-and their cost at the time of writing are shown in Table \ref{tab:ec2}.
+Because all these instances run in a private subnet, we also run a
+Bastion instance, that is, a specially protected network node, in a
+public subnet, from where we send the commands to execute the
+experiments and later collect the results. We connected to this
+instance through a {\texttt ssh} session from our personal
+workstation. This is a {\texttt R5AD Large} instance, selected with
+75GiB of storage, needed for keeping the logs, and more memory (16GiB)
+for processing the logs in memory. The details these EC2 instances and
+their cost at the time of writing are shown in Table \ref{tab:ec2}.
 
@@
 Again, each configuration describes the deployment of containers in a consistent
-and portable way, an it was used in each phase of the life cycle, from
+and portable way, and it was used in each phase of the life cycle, from
 development through testing to production independently on the kind of
@@
   \caption{Boxplot of the number of evaluations per second, for 30 instances of function $f_4$,
-          with different number of workers and number of populations. The $y$ axis is in log scale.}
+          with different number of workers and  populations. The $y$ axis is in log scale.}
   \label{fig:spworker:evalspersecond}
@@
 twenty populations. On the other hand, the maximum speed-up when we consider the
-cost it was with eight workers. The best speed-ups for each number of workers are
+cost was with eight workers. The best speed-ups for each number of workers are
 shown in bold in the table.
@@
   \includegraphics[width=\textwidth]{evals_per_instance}
-  \caption{Boxplot of the number of function evaluations required to reach the target, for 30 instances of function $f_4$, with a different number of workers and populations sizes}
+  \caption{Boxplot of the number of function evaluations required to reach the target, for 30 instances of function $f_4$, with a different number of workers and population sizes}
   \label{fig:evals_instance}
@@
 it might enhance evolutionary search and avoid unneeded
-evaluations to a greater extent than homogeneous systems systems do; heterogeneous settings, if
+evaluations to a greater extent than homogeneous systems do; heterogeneous settings, if
 done right, increase the diversity of the whole population\cite{araujo2008multikulti}.
@@
 framework from the JSON log file for each run. There was no additional setup
-needed, proving how this framework is able to not only offer an reproducible way
+needed, proving how this framework is able to not only offer a reproducible way
 to run experiments, but also how easily it interfaces with other common tools in
@@
 
-In this paper we have presented the EvoSwarm framework for 
-the reproducible execution of experiments with an event-based, cloud-native architecture 
-suitable for hosting multi-population, multi-algorithm optimization
-methods. 
-The framework is based in industry-standard development
-tools such as Docker and Docker Compose, making it easy for scientists and practitioners to develop and deploy
-% The last version still uses Docker Compose? - JJ
-% Well it uses docker-compose files, and these files are  
-% converted to tasks and services by the ECS-CLI. -Mario
-% It still can be used to run the application locally - Mario 
+In this paper we have presented the EvoSwarm framework for the
+reproducible execution of experiments with an event-based,
+cloud-native architecture suitable for hosting multi-population,
+multi-algorithm optimization methods.  The framework is based on
+industry-standard development tools such as Docker and Docker Compose,
+making it easy for scientists and practitioners to develop and deploy
 multi-algorithm distributed experiments. The model implemented by the
-framework is based on the asynchronous
-exchange of messages between stateless functions that react to a continuous
-stream of data; populations are the messages that flow in this stream. We show the differences of the message-driven architecture we presented against
-four other proposals for multi-population algorithms found in the literature, and highlight the 
-advantages for conducting asynchronous multi-population algorithms.
+framework is based on the asynchronous exchange of messages between
+stateless functions that react to a continuous stream of data;
+populations are the messages that flow in this stream. We show the
+differences of the message-driven architecture we presented against
+four other proposals for multi-population algorithms found in the
+literature, and highlight the advantages for conducting asynchronous
+multi-population algorithms.
 
@@
 auto-scaling capabilities, but additional configuration is needed; also, we need
-to establish the auto-scaling of both containers and Amazons EC2 instances.
+to establish the auto-scaling of both containers and Amazon's EC2 instances.
 Instead of setting a fixed number of workers, only the maximum amount would need

### 2020-10-16 draft.tex
Passed through grammarly, closes #27

@@
 techniques \cite{Baldini2016287}. 
-In these architectures, services or even functions are seen as independent 
+In these architectures, services, or even functions are seen as independent 
 processing nodes, departing from the monolithic
@@
 
-    \item They need to establish a workflow of local development and prototyping while having the 
-    option of deployment in a cloud provider. It is desirable to use modern development methodologies and technologies.
+    \item They need to establish a workflow of local development and
+prototyping while having the option of deployment to services provided
+by a cloud provider, thus using modern methodologies and technologies.
 
@@
 
-\subsection{Event Driven Model} 
+\subsection{Event-Driven Model} 
 \label{edm}
@@
 
-\item Each population is then pushed to the \texttt{Input} queue, so that they can be consumed 
+\item Each population is then pushed to the \texttt{Input} queue so that they can be consumed 
 by stateless functions responsible for the search's execution.
@@
 node keeps both the algorithm process and the state of the population.  Other
-processing nodes follow the same configuration and form of execution, and they
+processing nodes follow the same configuration and form of execution and they
 only pass specific individuals between them.  In this model, adding additional
@@
 Once the function returns the evolved population,
-the main script pushes the results to the \texttt{Output} queue, and continues
+the main script pushes the results to the \texttt{Output} queue and continues
 to execute the infinite loop. This process is shown in Figure \ref{fig:worker}.
@@
 
-Figure \ref{fig:aws-configuration} shows the basic EC2 configuration used for the deployment
-of experiments. We selected three types of EC2 instances according to our needs; the  
-{\texttt setup}, {\texttt controller}, and {\texttt redis} containers do not demand many resources
-and we do not need to scale them as the amount of parallel workers are increased. They run in a
-single {\texttt C5  High-CPU large} instance. With the available 2048 CPU shares divided as follows:
-{\texttt setup} (128),  {\texttt controller} (896), {\texttt redis} (1024). With this assignment 
-the {\texttt redis} container gets a dedicated vCPU, needed because it is the most demanding 
-task of the three. We assigned 2048 units to each {\texttt worker} container, and using a more 
-capable {\texttt C5 High-CPU XLarge} instance,  with 
-four vCPUs and 4096 CPU shares, with this configuration each worker gets a pair of vCPUs.  
-We scaled the number of instances as needed, using one instance for every two {\texttt worker} containers.
+Figure \ref{fig:aws-configuration} shows the basic EC2 configuration
+used for the deployment of experiments. We selected three types of EC2
+instances according to our needs; the {\texttt setup}, {\texttt
+controller}, and {\texttt redis} containers do not demand many
+resources and we do not need to scale them as the amount of parallel
+workers are increased. They run in a single {\texttt C5 High-CPU
+large} instance. With the available 2048 CPU shares divided as
+follows: {\texttt setup} (128), {\texttt controller} (896), {\texttt
+redis} (1024). With this assignment, the {\texttt redis} container
+gets a dedicated vCPU, needed because it is the most demanding task of
+the three. We assigned 2048 units to each {\texttt worker} container,
+and using a more capable {\texttt C5 High-CPU XLarge} instance, with
+four vCPUs and 4096 CPU shares, with this configuration each worker
+gets a pair of vCPUs.  We scaled the number of instances as needed,
+using one instance for every two {\texttt worker} containers.
 
@@
 \footnote{\url{https://github.com/mariosky/EvoSwarm/tree/master/aws-EC2}}, and
-they are used on the next section to carry an extensive evaluation of the framework.
+they are used in the next section to carry an extensive evaluation of the framework.
 Again, each configuration describes the deployment of containers in a consistent
@@
 
-In the previous section, we have described an option for an EvoSwarm deployment that can be used by 
-the public to create reproductive research in the cloud; in this section we are
-using the same framework to learn if the proposed solution can efficiently
-improve the scalability and performance of population-based optimization
-algorithms. 
-Hence, in the following sections, we provide answers to the following questions:
+In the previous section, we have described an option for an EvoSwarm
+deployment that can be used by the public to create reproductive
+research in the cloud; in this section, we are using the same
+framework to learn if the proposed solution can efficiently improve
+the scalability and performance of population-based optimization
+algorithms.  Hence, in the following sections, we provide answers to
+the following questions:
 
@@
 
-Next, we show the parameters for each algorithm, with Table \ref{tab:GAparams} for
-the GA and Table \ref{tab:PSOparams} for the PSO. We obtained these parameters
-following the same method as in a previous work \cite{garcia2017benchmarking}.
-To obtain the parameters, we tested first on the Rastrigin separable function
-with five dimensions. After about fifteen experiments, the most challenging
-targets were achieved for this particular function. We tested again with
-functions one to three, and after obtaining favorable results, the PSO and GA
-algorithm parameters were set. In the GA, we randomly set the mutation and
-crossover probabilities to have more heterogeneous workers;
-for these parameters, the range of values is specified. We did not change these parameters during the 
-experiments, and only the number of populations, population size and number of generations were
-provided as parameters. This strategy has been applied successfully for increasing 
-the performance of multi-population algorithms without requiring the setting 
-of initial parameters \cite{garcia2014randomized}.
+Next, we show the parameters for each algorithm, with Table
+\ref{tab:GAparams} for the GA and Table \ref{tab:PSOparams} for the
+PSO. We obtained these parameters following the same method as in a
+previous work \cite{garcia2017benchmarking}.  To obtain the
+parameters, we tested first on the Rastrigin separable function with
+five dimensions. After about fifteen experiments, the most challenging
+targets were achieved for this particular function. We tested again
+with functions one to three, and after obtaining favorable results,
+the PSO and GA algorithm parameters were set. In the GA, we randomly
+set the mutation and crossover probabilities to have more
+heterogeneous workers; for these parameters, the range of values is
+specified. We did not change these parameters during the experiments,
+and only the number of populations, population size, and number of
+generations were provided as parameters. This strategy has been
+applied successfully for increasing the performance of
+multi-population algorithms without requiring the setting of initial
+parameters \cite{garcia2014randomized}.
 
@@
 
-From this description it can be deduced that we give more importance in this
-paper to the number of populations, since they are the basic unit of
+From this description, it can be deduced that we give more importance in this
+paper to the number of populations since they are the basic unit of
 work, the message, that
@@
 
-In this experiment, we evaluate how the number of populations and the number of
-workers in the system are related to the scalability of the system.
-In the next section, we describe the experimental setup, and a discussion of the results
-follows, with respect to the speed-up in terms of evaluations per second (Section~\ref{sec:speedup}), and 
+In this experiment, we evaluate how the number of populations and the
+number of workers in the system are related to the scalability of the
+system.  In the next section, we describe the experimental setup and a
+discussion of the results follows, with respect to the speed-up in
+terms of evaluations per second (Section~\ref{sec:speedup}), and
 execution time (Section~\ref{sec:exec-time}).
@@
 which we have selected $f_4$ (Skew
-Rastrigin-Bueche separable) from the BBOB testbed, over 10 and 20 dimensions. This function has been used
+Rastrigin-Bueche separable) from the BBOB testbed, over 10, and 20 dimensions. This function has been used
 because it is computationally demanding, and in higher dimensions has been
@@
 instance of a problem. The reason for this is that in some cases reaching the
-desired target can require less work, because we are using a stochastic search. 
+desired target can require less work because we are using a stochastic search. 
 We are going to measure the total increase of work done in a second by adding more workers.
@@
 populations is higher than the number of workers. Configurations following this
-basic rule doubled the speed or where near it. Using five populations, we
+basic rule doubled the speed or were near that. Using five populations, we
 obtained the best and worst speed-up depending on the number of workers (shown in
@@
 From the algorithmic point of view, there are many possible lines of
-research. Mixing population-based algorithms with other algorithms
+research. Mixing population-based algorithms with other kinds
 would be a possibility, but also using different instances of
@@
 which could be tested, trying to find out which sets of algorithms are
-a better match to each other. All these will be explored as future
+a better match for each other. All these will be explored as future
 lines of work.

### 2020-10-16 draft.tex
Fleshes out cloud possibilities, closes #24

@@
     services are Amazon Elastic Container Service,  Red Hat Openshift
-    (which is also open-source), and Marantis Docker Enterprise using Docker Swarm.
+    (which is also open source), and Marantis Docker Enterprise using Docker Swarm.
     Salza et al. \cite{salza2019speed} tested a variant of this option by using a CoreOs cluster.
@@
 \end{description}
-
-%\subsubsection{Setup} % Even if it's a sub-sub section, flesh this out ... - JJ
-%\label{setup:local-cloud}
 
@@
 {\tt docker-compose}) and it has a lower cost than the Kubernetes
-alternative.
-
-We can define an ECS cluster directly on the AWS Elastic Cloud VMs or through the
+alternative. This does not imply any loss of generality; first,
+because it's {\tt docker-compose} compatible, and it can be deployed
+in approximately the same way to any of the services, mentioned above,
+compatible with it. Since {\tt docker-compose} is open source, it can
+be ported to any computing node, and can thus be deployed to any other
+cloud service like Azure or Google Cloud Platform. At any rate, please
+note that there are several levels of infrastructure programming
+here. At the lowest level, we are using docker containers, which are
+compatible across a myriad of cloud providers in one way or other. At
+the next level, we are using a container composition technology that
+is also standard, {\tt docker compose}, once again compatible across
+any platform that can host that kind of service. There's, however, a
+final, platform-specific, infrastructure definition that will have to
+vary across any public or private cloud platform that we are using,
+and that will tap the cloud provider API. These changes, however, are
+usually minimal, and can also be done through the CLI or using the web
+console.
+
+For instance, we can define an ECS cluster directly on the AWS Elastic Cloud VMs or through the
 AWS Fargate service. Fargate is a serverless compute engine for containers that

### 2020-10-16 draft.tex
Revised sentences, closes #21

@@
 
-Besides being, all of them, population-based, they also share the common characteristic of
-creating an initial set of random candidate solutions later used by a 
-nature-inspired heuristic to generate a new set of candidates. 
-Population-based algorithms are intrinsically parallel.
-The fitness of each individual can be (in general) evaluated independently of 
-others, and similarly, each population could evolve in isolation. Researchers use the term
-multi-population based methods when referring to techniques using
-many populations as part of their optimization strategy. 
-Since earlier works, researchers have been proposing some form of parallelization
-\cite{muhlenbein1988evolution} to increase these algorithms' scalability
-by adding some form of interaction between populations running in isolation.
-The island model was one of the first techniques proposed for parallelization,
-which led to early proofs of speed-up \cite{gorges1990explicit,grosso1985computer}. 
-% Repetitive - M
-%The concept was to divide a large population into communicated subpopulations. 
+Besides being, all of them, population-based, they also share the
+common characteristic of creating an initial set of random candidate
+solutions which is later used by a nature-inspired metaheuristic to
+generate a new set of candidates; there are parts of this process that
+can be easily done in parallel. For instance, the fitness of each
+individual can be (in general) evaluated independently of others, and
+similarly, each population could evolve in isolation. Researchers use
+the term multi-population based methods when referring to techniques
+using many populations as part of their optimization strategy.  Since
+earlier works, researchers have been proposing some form of
+parallelization \cite{muhlenbein1988evolution} to increase these
+algorithms' scalability by adding some form of interaction between
+populations running in isolation.  The island model was one of the
+first techniques proposed for parallelization, which led to early
+proofs of speed-up \cite{gorges1990explicit,grosso1985computer}.
 Since then, other population-based algorithms have adopted the idea.
-Researchers have found additional benefits
-besides the execution speed; these include avoiding a premature convergence and
-maintaining the diversity of the global population \cite{li2015multi}.
+Researchers have found additional benefits besides the execution
+speed; these include avoiding a premature convergence and maintaining
+the diversity of the global population \cite{li2015multi}.
 
@@
 
-%% JJ Can you improve the redability of the paragraph below? 
-%% Grammarly suggests that is hard to read, verly long sentences.
-%% Edited a little - Mario
 Cloud computing has become the standard way of running

### 2020-10-16 draft.tex
Eliminates repetitive sentence

@@
 
-We preferred the use of a container-based implementation for this prototype because
-it is more suitable for development and replication without needing
-additional components. Nevertheless, the application is compatible with other
-orchestration and deployment technologies if more scalability is needed. 
-With additional configuration settings, applications can be deployed 
-to a Docker Swarm or a Kubernetes system.
+% ------------------------ Commented this out since this is said
+% elsewhere ----------------------------------------------------
+% We preferred the use of a container-based implementation for this prototype because
+% it is more suitable for development and replication without needing
+% additional components. Nevertheless, the application is compatible with other
+% orchestration and deployment technologies if more scalability is needed. 
+% With additional configuration settings, applications can be deployed 
+% to a Docker Swarm or a Kubernetes system.
 %% Here we could talk about Cloud-based platform for executing the application

### 2020-10-16 draft.tex
Proofreading 6.2 refs #20

@@
 
-\subsection{EvoSwarm from the User Point of View} 
+\subsection{EvoSwarm from the user's point of view} 
 \label{sec:evoswarm:config}
@@
 data contained in the file in the next section. Users can use these files
-to analyze and plot experiment data.  Python
-scripts are included in the repository to plot the running times for the
-experiments and all other plots used in this paper;  there is also a script to
+to analyze and plot experiment data. We include Python
+scripts  in the repository to plot the running times for the
+experiments and all other charts used in this paper;  there is also a script to
 generate the files needed by the COCO framework \cite{hansen2016coco}, which can generate standard

### 2020-10-16 draft.tex
More changes

--- 2020-10-16 draft.tex
@@
 
-Besides being, all of them, population-based, they also share the
-common characteristic of creating an initial set of random candidate
-solutions which is later used by a nature-inspired metaheuristic to
-generate a new set of candidates; there are parts of this process that
-can be easily done in parallel. For instance, the fitness of each
-individual can be (in general) evaluated independently of others, and
-similarly, each population could evolve in isolation. Researchers use
-the term multi-population based methods when referring to techniques
-using many populations as part of their optimization strategy.  Since
-earlier works, researchers have been proposing some form of
-parallelization \cite{muhlenbein1988evolution} to increase these
-algorithms' scalability by adding some form of interaction between
-populations running in isolation.  The island model was one of the
-first techniques proposed for parallelization, which led to early
-proofs of speed-up \cite{gorges1990explicit,grosso1985computer}.
+Besides being, all of them, population-based, they also share the common characteristic of
+creating an initial set of random candidate solutions later used by a 
+nature-inspired heuristic to generate a new set of candidates. 
+Population-based algorithms are intrinsically parallel.
+The fitness of each individual can be (in general) evaluated independently of 
+others, and similarly, each population could evolve in isolation. Researchers use the term
+multi-population based methods when referring to techniques using
+many populations as part of their optimization strategy. 
+Since earlier works, researchers have been proposing some form of parallelization
+\cite{muhlenbein1988evolution} to increase these algorithms' scalability
+by adding some form of interaction between populations running in isolation.
+The island model was one of the first techniques proposed for parallelization,
+which led to early proofs of speed-up \cite{gorges1990explicit,grosso1985computer}. 
+% Repetitive - M
+%The concept was to divide a large population into communicated subpopulations. 
 Since then, other population-based algorithms have adopted the idea. 
-Researchers have found additional benefits besides the execution
-speed; these include avoiding a premature convergence and maintaining
-the diversity of the global population \cite{li2015multi}.
+Researchers have found additional benefits
+besides the execution speed; these include avoiding a premature convergence and
+maintaining the diversity of the global population \cite{li2015multi}.
 
@@
 
+%% JJ Can you improve the redability of the paragraph below? 
+%% Grammarly suggests that is hard to read, verly long sentences.
+%% Edited a little - Mario
 Cloud computing has become the standard way of running
@@
 techniques \cite{Baldini2016287}. 
-In these architectures, services, or even functions are seen as independent 
-processing nodes, departing from the monolithic
-and even distributed paradigms, and becoming a loosely coupled collection of {\em
+In these architectures, services are seen as independent processing nodes, 
+departing from monolithic or distributed paradigms to becoming a loosely 
+coupled collection of {\em
 stateless functions}\cite{malawski2017serverless}, that react to events. 
@@
     \item They need to establish a workflow of local development and
-prototyping while having the option of deployment to services provided
-by a cloud provider, thus using modern methodologies and technologies.
+prototyping while having the option of deployment to cloud services,
+ thus using modern methodologies and technologies.
 
@@
 
-    \item They must consider additional parameters that can affect the monetary cost or the performance of the
+    \item They must consider additional parameters that can affect the monetary cost 
+    or the performance of the
     execution — for instance, the format and size of messages, the number of
@@
 application in Section \ref{docker} and the workflow for reproducible experiment design in Section \ref{experiment_flow}. % No reference to section 6.
-Section \ref{setup} describes the empirical evaluation design to assess the effectiveness
-of the method, concerning scalability and % This is a subsection, fix it. 
+Section \ref{setup} describes the empirical evaluation design to assess the 
+method's effectiveness, concerning scalability and % This is a subsection, fix it. 
 the heterogeneous populations' performance.  % No references to sections 9 and 10
-Finally, we offer the conclusions of this paper and suggestions for future work in Section \ref{conclusions}.
+Finally, we offer the conclusions of this paper and suggestions for future work in 
+Section \ref{conclusions}.
 
@@
 and a container-based implementation of the model we propose to execute
-multi-population based algorithms. We follow the requirements outlined in
+multi-population-based algorithms. We follow the requirements outlined in
 Section \ref{multi} and the general design principles of cloud-native
@@
 Consumers are implemented using stateless functions, there is no difference
-between having one or many copies of the same function pulling work from the queue 
-simultaneously, because there are no side effects, including unwanted resource
+between having one or many copies of the same function simultaneously pulling work
+from the queue. There are no side effects, including unwanted resource
 locking problems resulting from this concurrency.
@@
 \label{edm}
-Based on a reactive architecture we proposed in a previous work
+Based on a reactive architecture we proposed in our previous work
 \cite{guervos2018introducing}, we now describe the general architecture shown in
@@
 \item Once they finish the execution, the population state
-is packed again with additional metadata about the algorithm's execution.
+is again packed with additional metadata about the algorithm's execution.
 The resulting populations are now pushed to the Output queue. Once finished, another 
@@
 \item The \texttt{Controller} process is responsible for keeping track of the progress of 
-the search. It pulls current populations from the \texttt{Output} queue, inspects the metadata, 
-and if an optimal solution has been found or the maximum number of iteration has been
-reached, it signals the stop of the execution.
+the search. It pulls current populations from the \texttt{Output} queue, inspects the metadata. 
+If an optimal solution has been found or the maximum number of iteration has been
+reached, it signals the execution to stop.
 
@@
 An important aspect of the proposal is the decoupling of the population and 
-the population-based algorithm. It is common that in a classic island-based algorithm
+the population-based algorithm. It is common that in a classic island-based algorithm,
 each island is executed in a separate processing node, i.e., in a virtual machine, CPU,
@@
 
-On the other hand, there are several caveats: It is more
-costly to move entire populations as messages than only passing
+On the other hand, we have several caveats: It is more
+costly to move entire populations as messages than passing only
 specific individuals from a process to another. Designers must consider
 this cost when working with large individuals, and if possible, send
-populations using several messages, use compression, or adjust the size
-of populations. Pool-based algorithms also suffer from this drawback,
-but web-based implementations have been working with continuous
-optimization problems with 1000 dimensions.  However, in other cases, % Reference that paper - Mario
-this is not a viable approach.
+populations using several messages, use compression, or adjust the populations'
+size. Pool-based algorithms also suffer from this drawback.
+%,but web-based implementations have been working with continuous
+%optimization problems with 1000 dimensions.  However, in other cases, % Reference that paper - Mario
+%this is not a viable approach.
 
@@
 
-Next, we compare the message-driven architecture we presented against four other
-proposals for multi-population algorithms found in the literature. We center the
+Next, we compare the message-driven architecture we presented against four
+multi-population-based algorithms found in the literature. We center the
 comparison on the coupling and communication between the main components:
@@
 node keeps both the algorithm process and the state of the population.  Other
-processing nodes follow the same configuration and form of execution and they
+processing nodes follow the same configuration and execution form, and they
 only pass specific individuals between them.  In this model, adding additional
@@
 solved in part in Peer to Peer (P2P) evolutionary algorithms
-\cite{juanlu:europar}, but coupling still exists, meaning that every
+\cite{juanlu:europar}. However, coupling still exists, meaning that every
 node must run the whole evolutionary algorithm and hold full copies of
@@
 message queue. These messages are instantly emitted by the  \texttt{consumed\_messages}
-\texttt{Observable}. In this particular period, we can see that the stream receives 
-two populations from another experiment and are shown as two red marbles.  
+\texttt{Observable}.This particular example shows that the stream receives 
+two populations from another experiment, and these are shown as two red marbles.  
 The filter operator removes all messages that belong to a different experiment; 
@@
 the population key (the ellipsis is used instead). The message includes the 
-current state of the population and data about the problem, algorithm, and parameters.
+population's current state and data about the problem, algorithm, and parameters.
 
@@
 \end{description}
+
+%\subsubsection{Setup} % Even if it's a sub-sub section, flesh this out ... - JJ
+%\label{setup:local-cloud}
 
@@
 {\tt docker-compose}) and it has a lower cost than the Kubernetes
-alternative. This does not imply any loss of generality; first,
-because it's {\tt docker-compose} compatible, and it can be deployed
-in approximately the same way to any of the services, mentioned above,
-compatible with it. Since {\tt docker-compose} is open source, it can
-be ported to any computing node, and can thus be deployed to any other
-cloud service like Azure or Google Cloud Platform. At any rate, please
-note that there are several levels of infrastructure programming
-here. At the lowest level, we are using docker containers, which are
-compatible across a myriad of cloud providers in one way or other. At
-the next level, we are using a container composition technology that
-is also standard, {\tt docker compose}, once again compatible across
-any platform that can host that kind of service. There's, however, a
-final, platform-specific, infrastructure definition that will have to
-vary across any public or private cloud platform that we are using,
-and that will tap the cloud provider API. These changes, however, are
-usually minimal, and can also be done through the CLI or using the web
-console.
-
-For instance, we can define an ECS cluster directly on the AWS Elastic Cloud VMs or through the
+alternative.
+
+We can define an ECS cluster directly on the AWS Elastic Cloud VMs or through the
 AWS Fargate service. Fargate is a serverless compute engine for containers that

### 2020-10-16 draft.tex
Fill 6.2 refs #20

@@
 
-After starting the application, users can push several Experiment Definition
-files and start monitoring the current execution in the terminal.
-After an experiment is finished, the user must execute another script
-that takes as argument the experiment id to generate a collection of files containing all the
-data generated by the experiment, again in a JSON format. An example of an
-experiment definition is shown in Listing \ref{code:exp}. We give more details about the
-data contained in the file in the next section. Users can use these files
-to analyze and plot experiment data. We include Python
+After starting the application, users can push several Experiment
+Definition files and start monitoring the current execution in the
+terminal.  After an experiment is finished, the user must execute
+another script that takes as argument the experiment id to generate a
+collection of files containing all the data generated by the
+experiment, again in a JSON format. An example of an experiment
+definition is shown in Listing \ref{code:exp}. We give more details
+about the data contained in the file in the next section. Users can
+use these files to analyze and plot experiment data. We include Python
 scripts in the repository to plot the running times for the
-experiments and all other charts used in this paper;  there is also a script to
-generate the files needed by the COCO framework \cite{hansen2016coco}, which can generate standard
-comparisons against other methods.
+experiments and all other charts used in this paper; there is also a
+script to generate the files needed by the COCO framework
+\cite{hansen2016coco}, which can generate standard comparisons against
+other methods.
 

### 2020-10-16 draft.tex
Revised 6.3, refs #20, please @mariosky re-revise and close

@@
 
-\subsection{Deploying to a Cloud Provider}
+\subsection{Deploying to a cloud provider}
 \label{cloud-aws}
@@
 server, or a production environment. This section gives specific details
-about how we deployed EvoSwarm in a cloud-setting for running the experiments
+about how we deployed EvoSwarm in a cloud setting for running the experiments
 within this paper. Running the cloud-native application in the cloud has the obvious
 benefit of a multi-instance deployment, where we can run containers in several
-instances executing at the same time, not only in a single machine as it is the case when using docker-compose.
+computing nodes at the same time, not only in a single machine as it is the case when using docker-compose.
 
@@
 
-Multicontainer applications are deployed to ECS as tasks, and a single cluster can run multiple tasks in
+Multicontainer applications are deployed to ECS as a series of tasks {\em tasks}, and a single cluster can run multiple tasks in
 parallel. Using the ECS CLI, we can provision a cluster with a configuration file
@@
 that reference the container images we have published in Docker hub
-\footnote{\url{https://hub.docker.com/u/mariosky}}, open {\texttt ports}, a {\texttt command} to be executed at
+\footnote{They are shown as different repositories for user {\tt
+    mariosky} here: \url{https://hub.docker.com/u/mariosky}}, open {\texttt ports}, a {\texttt command} to be executed at
 startup, configuration for the AWS {\texttt logging} service, and finally, {\texttt environment}

### 2020-10-16 draft.tex
Update draft.tex

@@
 After starting the application, users can push several Experiment
-Definition files and start monitoring the current execution in the
-terminal.  After an experiment is finished, the user must execute
+Definition files and monitor the current execution using the PC console.
+After an experiment is finished, the user must execute
 another script that takes as argument the experiment id to generate a
@@
 within this paper. Running the cloud-native application in the cloud has the obvious
-benefit of a multi-instance deployment, where we can run containers in several
+benefit of a multi-instance deployment. We can run containers in several
 computing nodes at the same time, not only in a single machine as it is the case when using docker-compose.
@@
 deployed the EvoSwarm application in Amazon Elastic Container Service (ECS)
-since it needs fewer changes in configuration (from deploying it locally using
+since it needs fewer configuration changes (from deploying it locally using
 {\tt docker-compose}) and it has a lower cost than the Kubernetes
@@
 the configuration on-the-fly. For example, to deploy the controller, setup, and Redis 
-containers we used the file in Listing ~\ref{code:compose}.
-
-Each container is defined as a {\texttt service} ({\em task} in AWS terms), with the following attributes: {\texttt image},
+containers, we used the file in Listing ~\ref{code:compose}.
+
+Each container is defined as a {\texttt service} (a {\em task} in AWS terms), with the following attributes: {\texttt image},
 that reference the container images we have published in Docker hub
@@
 controller}, and {\texttt redis} containers do not demand many
-resources and we do not need to scale them as the amount of parallel
-workers are increased. They run in a single {\texttt C5 High-CPU
+resources and we do not need to scale them as the number of parallel
+workers is increased. They run in a single {\texttt C5 High-CPU
 large} instance. With the available 2048 CPU shares divided as
@@
 redis} (1024). With this assignment, the {\texttt redis} container
-gets a dedicated vCPU, needed because it is the most demanding task of
-the three. We assigned 2048 units to each {\texttt worker} container,
-and using a more capable {\texttt C5 High-CPU XLarge} instance, with
-four vCPUs and 4096 CPU shares, with this configuration each worker
+gets a dedicated vCPU because it is the most demanding task.
+We assigned 2048 units to each {\texttt worker} container
+and used a more capable {\texttt C5 High-CPU XLarge} instance, with
+four vCPUs and 4096 CPU shares. With this configuration, each worker
 gets a pair of vCPUs.  We scaled the number of instances as needed,
@@
 Because all these instances run in a private subnet, we also run a
-Bastion instance, that is, a specially protected network node, in a
+Bastion instance, a specially protected network node, in a
 public subnet, from where we send the commands to execute the
 experiments and later collect the results. We connected to this
-instance through a {\texttt ssh} session from our personal
-workstation. This is a {\texttt R5AD Large} instance, selected with
+instance through an {\texttt ssh} session from our workstation.
+This is an {\texttt R5AD Large} instance, selected with
 75GiB of storage, needed for keeping the logs, and more memory (16GiB)
-for processing the logs in memory. The details these EC2 instances and
+for processing the logs in memory. The details of these EC2 instances and
 their cost at the time of writing are shown in Table \ref{tab:ec2}.
@@
 
-The complete configuration files are available at public GitHub repository 
+The complete configuration files are available at public the GitHub repository 
 \footnote{\url{https://github.com/mariosky/EvoSwarm/tree/master/aws-EC2}}, and
@@
 Again, each configuration describes the deployment of containers in a consistent
-and portable way, and it was used in each phase of the life cycle, from
+and portable way. It was used in each phase of the life cycle, from
 development through testing to production independently on the kind of
@@
 framework to learn if the proposed solution can efficiently improve
-the scalability and performance of population-based optimization
-algorithms.  Hence, in the following sections, we provide answers to
+population-based optimization algorithms' scalability and performance.
+Hence, in the following sections, we provide answers to
 the following questions:
@@
 \begin{itemize}
-\item Can we improve the execution time of the algorithm by adding populations
+\item Can we improve the algorithm's execution time by adding populations
  and serverless functions (workers)  to the system and, in particular,
@@
 \item Does having a multi-population enabled platform, with heterogeneous populations and
-the support for mixing search strategies,  increase the performance of the
-search by needing fewer function evaluations than a homogeneous setting? 
+the support for mixing search strategies,  increase the search's
+performance by needing fewer function evaluations than a homogeneous setting? 
 % Moved reproducibility question from here - M

### 2020-10-17 draft.tex
More proof reading

@@
 parallel. Using the ECS CLI, we can provision a cluster with a configuration file
-specifying the computing resources needed. We can also specify the configuration of a task 
+specifying the computing resources needed. We can also specify a task's configuration 
 by using a task definition file (see Listing~\ref{code:ecs}) in YAML format.
@@
 
-The complete task configuration is generated by the {\texttt ecs compose command} that uses a {\texttt docker-compose.yml} file to generate 
-the configuration on-the-fly. For example, to deploy the controller, setup, and Redis 
+The complete task configuration is generated on-the-fly by the {\texttt ecs compose command} 
+that uses a {\texttt docker-compose.yml} file to generate 
+the configuration. For example, to deploy the controller, setup, and Redis 
 containers, we used the file in Listing ~\ref{code:compose}.
@@
 Bastion instance, a specially protected network node, in a
-public subnet, from where we send the commands to execute the
+public subnet. From this instance, we send the commands to execute the
 experiments and later collect the results. We connected to this
@@
 deployment that can be used by the public to create reproductive
-research in the cloud; in this section, we are using the same
+research in the cloud. In this section, we are using the same
 framework to learn if the proposed solution can efficiently improve
@@
  and serverless functions (workers)  to the system and, in particular,
- what is its effect on the scalability of the system? % Improve with respect to what? Single population? 
+ what is its effect on the system's scalability? % Improve with respect to what? Single population? 
                          %Please explain carefully the issues and how scaling is related to the number of populations. - JJ
@@
 instances of each function. Each instance of a function has a different optimal
-value. The standard benchmark of the testbed utilizes 15 instances per function
+value. The standard benchmark runs 15 instances per function
 over 2, 3, 5, 10, 20, and 40 dimensions. The maximum number of Function
@@
 
-The COCO framework offers several tools to compare the performance of
-algorithms, generating data sets, tables, and reports for an experiment. There
+The COCO framework offers several tools to compare algorithms' performance,
+generating data sets, tables, and reports for an experiment. There
 is a repository\footnote{\url{https://coco.gforge.inria.fr/doku.php?id=algorithms-bbob}} 
@@
 compatible with the scripts of the framework, to compare with other algorithms
-in the repository.
+stored in the repository.
 
@@
 performance between a homogeneous and an ensemble of multi-populations, using
-Genetic Algorithms (GAs) and Particle Swarm Optimization (PSO). For the GA
-implementation, we used the DEAP library \cite{fortin2012deap} and for the PSO,
-the EvoloPy library \cite{faris2016evolopy}. Both Python libraries are
-open-source. In EvoSwarm these two algorithms are implemented as stateless
+Genetic Algorithms (GAs) and Particle Swarm Optimization (PSO). 
+We used the DEAP library \cite{fortin2012deap}, for the GA implementation. 
+For the PSO algorithm, the EvoloPy library \cite{faris2016evolopy}. Both Python libraries are
+open-source. In EvoSwarm, we implemented these two algorithms as stateless
 functions.
@@
 
-Next, we show the parameters for each algorithm, with Table
+Next, we show each algorithm's parameters, with Table
 \ref{tab:GAparams} for the GA and Table \ref{tab:PSOparams} for the
-PSO. We obtained these parameters following the same method as in a
+PSO. We obtained these parameters following the same method as in 
 previous work \cite{garcia2017benchmarking}.  To obtain the
 parameters, we tested first on the Rastrigin separable function with
-five dimensions. After about fifteen experiments, the most challenging
-targets were achieved for this particular function. We tested again
+five dimensions.  After about fifteen experiments, the algorithm 
+achieved the most challenging targets for this particular function. We tested again
 with functions one to three, and after obtaining favorable results, 
-the PSO and GA algorithm parameters were set. In the GA, we randomly
+we set the PSO and GA parameters to these values. In the GA, we randomly
 set the mutation and crossover probabilities to have more
-heterogeneous workers; for these parameters, the range of values is
-specified. We did not change these parameters during the experiments,
-and only the number of populations, population size, and number of
-generations were provided as parameters. This strategy has been
+heterogeneous workers; for these parameters, we specify the range 
+of values in Table \ref{tab:GAparams}. We did not change these 
+parameters during the experiments and only modified the number of
+populations, population size, and generations. This strategy has been
 applied successfully for increasing the performance of
@@
 The experiments can be easily reproduced by using the configuration
-files the EvoSwarm application uses to run the noiseless BBOB testbed
-experiments as seen in Section~\ref{sec:evoswarm:config}. 
+files described in Section~\ref{sec:evoswarm:config}. 
 As a first step, we can deploy the docker application using a
@@
 populations that will use the GA algorithm.  In the example, with a value of
-$0.50$ there will be about the same proportion of GA and PSO populations. If we
+$0.50$, there will be about the same proportion of GA and PSO populations. If we
 specify a value of $0$, this will give us an algorithm with only PSO
@@
 Next, we specify a list indicating which of the 24 benchmark
-functions will be tested. In the example, the experiment will use the first five
+functions we will test. In the example, the experiment will use the first five
 functions.  The {\em Instances} parameter indicates how many instances of each
-function will be tested. {\em Instances}  has a default value of 15. We used this
+function we will teste. {\em Instances}  have a default value of 15. We used this
 value because it is the standard for the BBOB benchmark \cite{hansen2016coco}.
@@
 According to the maximum number of FEs we mentioned
-earlier, we define for each dimension: the number of populations that will be
+earlier, for each dimension, we define the number of populations that will be
 generated in the setup, and for each population, the number of generations, and
 population size. Finally, the {\em Iterations} parameter indicates how many
-complete loops, that is, complete message processings, will be performed.
+complete loops, that is, the maximum number of messages taken from the queue.
 All these parameters give us the maximum
-number of FEs that will be performed. For instance, for $D = 2$, the maximum
+number of FEs we will have as budget. For instance, for $D = 2$, the maximum
 number of FEs is $200,000$, which is the same as $40*50*10*10$.
@@
 described in Section~\ref{cloud-aws}, each EC2 instance was an Amazon Linux 2 AMI with 
-ECS Agent 1.42.0 and Docker version 19.03.6-ce. Additionally the BBOB experiments 
+ECS Agent 1.42.0 and Docker version 19.03.6-ce. Additionally, the BBOB experiments 
 were performed with COCO \cite{hansen2016coco} version bbob.v15.03
-in Python, the plots were produced with version 2.3.2.
+in Python, and the plots were produced with version 2.3.2.
 
@@
 
-As we mentioned before in Section~\ref{method}, reactive systems can scale by adding additional copies
+As we mentioned before, in Section~\ref{method}, reactive systems can scale by adding additional copies
 of serverless functions. In our case, we can start additional worker containers
@@
 have used a similar architecture, which implies that worker nodes need to have at least a certain level
-of complexity, in terms of execution time,  to effectively scale on multiple
+of complexity, in terms of execution time, to scale on multiple
 nodes, tending to linear scalability. In that paper, Salza et al. tested scalability
-by simulating the work of nodes by using a sleep function and messages of different sizes.
+by simulating the nodes' work by using a sleep function and messages of different sizes.
 In a population-based algorithm like EvoSwarm, several parameters can increase or 
@@
 
-From this description, it can be deduced that we give more importance in this
-paper to the number of populations since they are the basic unit of
-work, the message, that
-can be executed in parallel. Together with the size of each population and the
+From this description, we can deduce that we give more importance 
+to the number of populations since they are the basic unit of
+work, the message, that can be executed in parallel. Together with each population's size and the
 number of generations, this parameter determines the maximum number of function
-evaluations for a single iteration. We do not have control over each algorithm's
-complexity, but future analyses could focus on balancing populations and number
-of worker nodes according to this.
-
-In this experiment, we evaluate how the number of populations and the
-number of workers in the system are related to the scalability of the
-system.  In the next section, we describe the experimental setup and a
-discussion of the results follows, with respect to the speed-up in
-terms of evaluations per second (Section~\ref{sec:speedup}), and
+evaluations for a single iteration. According to this, we do not have control 
+over each algorithm's complexity, but future analyses could focus on balancing 
+populations and the number of worker nodes.
+
+In this experiment, we evaluate how the number of populations and workers 
+are related to the system's scalability.
+In the next section, we describe the experimental setup.A
+discussion of the results follows, concerning the speed-up in
+terms of evaluations per second (Section~\ref{sec:speedup}) and
 execution time (Section~\ref{sec:exec-time}).
@@
 To answer the first research question, we need to evaluate the effect of the
-number of populations and the number of workers on the scalability of the
-system, and in order to achieve this we propose an experiment for
-which we have selected $f_4$ (Skew
-Rastrigin-Bueche separable) from the BBOB testbed, over 10, and 20 dimensions. This function has been used
-because it is computationally demanding, and in higher dimensions has been
+number of populations and the number of workers on the system's scalability.
+To achieve this, we propose an experiment for
+which we have selected $f_4$ (Skew Rastrigin-Bueche separable) from the BBOB testbed,
+over 10 and 20 dimensions. This function has been used
+because it is computationally demanding, and in higher dimensions, it has been
 difficult for PSO \cite{el2009black} and GAs \cite{nicolau2009application} to
-solve with the required FEs. As we are only comparing in terms of scalability,
-the results from a single function can be better understood, since there are
-fewer factors involved. Following the procedure described in
-Section~\ref{setup}, we executed two sets of experiments with five, ten, and twenty
-initial populations, repeating each experiment 30 times using 1, 2, 4, 8, and 16
+solve with the required number of FEs. As we are only comparing in terms of scalability,
+the results from a single function can be better understood since fewer factors are involved.
+Following the procedure described in Section~\ref{setup}, we executed two sets of 
+experiments with five, ten, and twenty initial populations, repeating each 
+experiment 30 times using 1, 2, 4, 8, and 16
 workers. We kept the same maximum number of FEs, changing the relevant
@@
   \caption{Boxplot of the number of evaluations per second, for 30 instances of function $f_4$,
-          with different number of workers and  populations. The $y$ axis is in log scale.}
+          with a different number of workers and populations. The $y$-axis is in log scale.}
   \label{fig:spworker:evalspersecond}
@@
 To measure the speed-up obtained by the addition of workers, we will use the
-number of evaluations per second, instead of the time required to finish an
-instance of a problem. The reason for this is that in some cases reaching the
+number of evaluations per second instead of the time required to finish an
+instance of a problem. This is because in some cases reaching the
 desired target can require less work because we are using a stochastic search. 

### 2020-10-17 draft.tex
Update draft.tex

@@
   \caption{Speed-up by worker and dimension, taking one worker as the
-    baseline. Speed-up is obtained by dividing the evaluation ratio of the baseline by the 
+    baseline. Speed-up is obtained by dividing the baseline's evaluation ratio by the 
     ratio of the increased number of workers.}
@@
 %
-As expected, within the same number of populations, there is a linear speed-up
-when we add more workers. The speed-up is marginal in those cases where there 
-are fewer populations than workers.  The reason for this is that at any given time, 
+As expected, there is a linear speed-up when we add more workers 
+within the same number of populations. The speed-up is marginal in those cases where there 
+are fewer populations than workers. This is because, at any given time, 
 there are at least two populations in the message queue, leaving some of the workers 
@@
 Table~\ref{tab:speedup:test} shows the speed-up obtained taking as the baseline
-the previous configuration, for example, when increasing from four to eight
+the previous configuration, such as increasing from four to eight
 workers, and the p-value for the Wilcoxon Signed-Rank test. There was an
@@
 populations is higher than the number of workers. Configurations following this
-basic rule doubled the speed or were near that. Using five populations, we
+basic rule doubled the speed or where near that. Using five populations, we
 obtained the best and worst speed-up depending on the number of workers (shown in
@@
 In the previous section, we have shown that as we increase the number of
-populations and workers, we increase the ratio of the number of evaluations per
+populations and workers, we increase the number of evaluations per
 second in the same proportion. However, it is vital to notice that we can also
@@
 workers. However, we can see that for some configurations, i.e., 10D and ten
-populations, eight workers will need less FEs then one (p = 0.012). Also, for
+populations, eight workers will need fewer FEs than one (p = 0.012). Also, for
 20D and twenty populations, there are differences between one and four workers
@@
 there are benefits in the potential change of order in the message queue
-resulting from the parallelization of work.
+resulting from work parallelization.
 
@@
 
-The reduction in the number of FEs has an impact when we measure the speed up
+The reduction in the number of FEs impacts when we measure the speed up
 concerning the time needed to complete an experiment. In this case, we have
 measured the median of the seconds needed to complete an instance of the
-problem. Figure \ref{fig:spworker:time} shows in each box-plot the median (30
+problem. Figure \ref{fig:spworker:time} shows in each boxplot the median (30
 runs) of the execution time achieved by 2, 4, 8, and 16 workers, to finish an
@@
 20 population sizes. We present each dimension in a row having a log scale on
-the y-axis, to compensate for the dimension increments and scalability.
+the y-axis to compensate for the dimension increments and scalability.
 
@@
 
-In Table~\ref{tab:time} we can see that in some cases we have a supralinear
-speed-up, we highlighted these in boldface and indicated the speed-up value in
+In Table~\ref{tab:time} we can see that we have a supralinear
+speed-up in some cases; we highlighted these in boldface and indicated the speed-up value in
 parenthesis. The best overall times are underlined. As expected, increasing the

### 2020-10-19 draft.tex
Finished proof reading

@@
 A multi-population based algorithm can decrease the execution time
-thanks to running the algorithm on several populations at the same
-time. But, if the populations are heterogeneous 
-it might enhance evolutionary search and avoid unneeded
+thanks to running the algorithm on several populations simultaneously.
+However, if the populations are heterogeneous 
+it might enhance the evolutionary search and avoid unneeded
 evaluations to a greater extent than homogeneous systems do; heterogeneous settings, if
@@
 
-But this is a rule of thumb, and it will depend on the degree of heterogeneity,
-as well as on the algorithm itself. Some level of heterogeneity can be
-implemented by just changing the configuration parameters of each population,
-as in the previous experiment. But in this case, we are interested in
-answering the second research question, and we want to validate if by using
-heterogeneous populations and mixing search strategies,  we can increase the performance of the
-search.
-
-Therefore, in this experiment we compare a multi-population with GA and PSO only populations,
+Nevertheless, this is a rule of thumb, and it will depend on the degree of heterogeneity and 
+the algorithm itself. Some level of heterogeneity can be 
+achieved by just changing the configuration parameters of each population, 
+as in the previous experiment.However, in this case, we are interested in answering 
+the second research question. We want to validate if we can increase the search's performance 
+using heterogeneous populations and mixing search strategies.
+
+Therefore, in this experiment, we compare a multi-population with GA and PSO only populations,
 versus an ensemble of GA and PSO algorithms. We have tested on the first five functions of the
@@
 
-In order to run this experiment, we just needed to edit a configuration file and
+To run this experiment, we just needed to edit a configuration file and
 start it up. The plots presented in the results are generated by the BBOB
 framework from the JSON log file for each run. There was no additional setup
-needed, proving how this framework is able to not only offer a reproducible way
-to run experiments, but also how easily it interfaces with other common tools in
+needed, proving how this framework can offer a reproducible way
+to run experiments and how easily it interfaces with other common tools in
 the EC ecosystem.
@@
 the runtime scales with dimension to reach certain target values $\Delta f$;
-The color of the lines indicates the average runtime, for each target reached
+The color of the lines indicates the average runtime, for each target, reached
 at least once. A red circle without a number on top indicates the algorithm 
-has reached the $10^{-8}$ target on all instances for that dimension. If there
-is that number, it indicates how many times the algorithm has reached
+reached the $10^{-8}$ target on all instances for that dimension. If there
+is a number, it indicates how many times the algorithm has reached
 the target. Fixed values of targets $\Delta f = 10^{k}$ with $k$ colors are
-given in the legend of $f_1$ in the top row, $k$ values are $[-8,-5,-3,-2,-1,0,1]$. Results
+given in the $f_1$ legend in the top row, $k$ values are $[-8,-5,-3,-2,-1,0,1]$. Results
 from each configuration are presented in each column, GA, PSO, and GA\&PSO
@@
 We can see that the GA algorithm (on the leftmost column) has a hard time reaching even the
-$10^{-5}$ target in functions ($f_1-f_4$) for all dimensions, and has the worst
+$10^{-5}$ target in functions ($f_1-f_4$) for all dimensions and has the worst
 performance in 20 and 40 dimensions. Finally, it performs better on $f_5$
@@
 
-On the other hand, the multi-population PSO (taking the column on the middle) 
+On the other hand, the multi-population PSO (taking the column in the middle) 
 has better performance; its results are spread over a smaller range,
@@
 runtime than the multi-strategy configuration. It is not able to reach the
-$10^{-8}$ target on the higher dimensions of $f_3$ and $f_4$. Moreover, the PSO
+$10^{-8}$ target on higher dimensions of $f_3$ and $f_4$. Moreover, the PSO
 multi-population has the best runtime for the lower dimensions of $f_5$.
@@
 Finally, the proposed PSO\&GA hybrid algorithm has reached the $10^{-8}$ target
-value on all functions ($f_1-f_5$) and scales well to higher dimensions, even on
-$f_3$ and $f_4$ (charts on the lower rows), which are usually considered
+value on all functions ($f_1-f_5$). It scales nicely to higher dimensions, even on
+$f_3$ and $f_4$ (charts on the lower rows), usually considered
 difficult functions. This experiment also exemplifies the type of analyses that
-we can perform with the results of the EvoSwarm application by reusing the COCO framework.
-We also proved that multi-population algorithms can be compared under the same 
+we can perform with the EvoSwarm application results by reusing the COCO framework.
+We also proved that multi-population algorithms could be compared under the same 
 conditions, including single and multi-algorithm versions, and results show 
-that the later have better performance.
+that the latter has the best performance.
 
@@
 \label{conclusions}
-
-In this paper we have presented the EvoSwarm framework for the
+This paper has presented the EvoSwarm framework for the
 reproducible execution of experiments with an event-based,
@@
 four other proposals for multi-population algorithms found in the
-literature, and highlight the advantages for conducting asynchronous
+literature and highlight the advantages of conducting asynchronous
 multi-population algorithms.
@@
 We have tested the deployment of an experiment on Amazon's Elastic
-Container Service, and proved that as a cloud-native, container-based application,
+Container Service and proved that as a cloud-native, container-based application,
 EvoSwarm could be deployed locally or in the cloud by specifying the configuration 
@@
 We have performed an empirical evaluation of the same framework to learn if the proposed
-solution can efficiently improve the scalability and performance of
-population-based optimization algorithms.
+solution can efficiently improve population-based optimization algorithms' 
+scalability and performance.
 
@@
 A second experiment centered on evaluating the algorithmic benefits
-(experiments taking less time because they need fewer evaluations) of
+(experiments are taking less time because they need fewer evaluations) of
 distributed multi-algorithm algorithms such as the ones implemented in
@@
 of multi-populations, using Genetic Algorithms (GAs) and Particle
-Swarm Optimization (PSO) using the COCO benchmark framework for
-visualization and definition of the optimization problems.  We found
-that the set-up of these algorithms has a significant influence on
+Swarm Optimization (PSO), using the COCO benchmark framework to visualize and define 
+optimization problems.  We found
+that these algorithms' set-up has a significant influence on
 their results since diversity is essential for their performance. The
@@
 appropriate configuration consisting of the number of populations, worker
-containers, and algorithm combination. Further research should focus on learning
+containers, and algorithm combinations. Further research should focus on learning
 more about how these parameters interact with each other. The BBOB benchmark
@@
 Kubernetes or Docker Swarm. The current orchestration (Amazon ECS) also has
-auto-scaling capabilities, but additional configuration is needed; also, we need
+auto-scaling capabilities, but additional configuration is needed; we also need
 to establish the auto-scaling of both containers and Amazon's EC2 instances.
@@
 research. Mixing population-based algorithms with other kinds
-would be a possibility, but also using different instances of
-population based algorithms, such as checking how Estimation of
+would be a possibility and using different instances of
+population-based algorithms, such as checking how Estimation of
 Distribution Algorithms would work together with evolutionary
-algorithms. EvoloPy also includes other population based algorithms,
-which could be tested, trying to find out which sets of algorithms are
-a better match for each other. All these will be explored as future
-lines of work.
+algorithms. EvoloPy also includes other population-based algorithms,
+which could be tested, trying to determine which sets of algorithms are
+a better match for each other. We will explore all this as future lines of work.
 
